### 2021-12-27 dedication/dedication.tex
finish dedication

\cleardoublepage

\begin{dedication}
  I hereby declare that the content of this thesis is the product of my own research work. The main parts of this thesis have been published in the following list of my publications. I made the major contribution in design, implementation and experiments for these work under the guidance of Prof. Anil Anthony Bharath and some others which have been acknowledged appropriately in text. The information, codes, and former ideas guided from other sources have been cited in a list of references is in the bibliography.
\end{dedication}
% List the publications used in the thesis
\textbf{Publication List}:
\begin{enumerate}
	\item \underline{Tianhong Dai}, Magda Dubois, Kai Arulkumaran, Jonathan Campbell, Cher Bass, Benjamin Billot, Fatmatulzehra Uslu, Vincenzo de Paola, Claudia Clopath, and Anil Anthony Bharath, ``Deep Reinforcement Learning for Subpixel Neural Tracking", International Conference on Medical Imaging with Deep Learning, 2019 (Spotlight).
	\item \underline{Tianhong Dai}, Hengyan Liu, Anil Anthony Bharath, ``Episodic Self-Imitation Learning with Hindsight", Electronics, Special Issue on Deep Reinforcement Learning: Methods and Applications, 2020.
	\item \underline{Tianhong Dai}, Hengyan Liu, Kai Arulkumaran, Guangyu Ren, and Anil Anthony Bharath, ``Diversity‐based Trajectory and Goal Selection with Hindsight Experience Replay", Pacific Rim International Conference on Artificial Intelligence, 2021.
	\item \underline{Tianhong Dai}, Yali Du, Meng Fang, and Anil Anthony Bharath, ``Diversity-Augmented Intrinsic Motivation for Deep Reinforcement Learning", Neurocomputing, 2021.
	\item \underline{Tianhong Dai}, Kai Arulkumaran, Tamara Gerbert, Samyakh Tukra, Feryal Behbahani, and Anil Anthony Bharath, ``Analysing Deep Reinforcement Learning Agents Trained with Domain Randomisation", Neurocomputing, 2022.
\end{enumerate}

% List extra publications
In addition to the above papers, these following publications that I authored/co-authored during my PhD are out of the main scope of this thesis, and are briefly mentioned in the background or future directions.
\begin{enumerate}
	\item Shafa Balaram, Kai Arulkumaran, \underline{Tianhong Dai}, and Anil Anthony Bharath, ``A Maximum Entropy Deep Reinforcement Learning Neural Tracker", International Workshop on Machine Learning in Medical Imaging, 2019.
	\item Yali Du, Lei Han, Meng Fang, \underline{Tianhong Dai}, Ji Liu, and Dacheng Tao, ``LIIR: Learning Individual Intrinsic Reward in Multi-Agent Reinforcement Learning", Advances in Neural Information Processing Systems, 2019.
	\item \underline{Tianhong Dai}, Wei Li, Xilei Cao, Jianzhuang Liu, Xu Jia, Ales Leonardis, Youliang Yan, and Shanxin Yuan, ``Wavelet-Based Network For High Dynamic Range Imaging", Submitted to IEEE Transactions on Imaging Processing, 2021.
\end{enumerate}

### 2021-12-31 chapters/chapter3.tex
init draft of chapter3

\chapter{Episodic Self-Imitation Learning with Hindsight}
\section{Introduction}
Reinforcement learning (RL) has has been shown to be very effective in training agents within gaming environments~\cite{mnih2015human,silver2016mastering}, particularly when combined with deep neural networks~\cite{lecun2015deep,silver2016mastering,liu2017survey}. In most tasks settings that are solved by RL algorithms, reward shaping is an essential requirement for guiding the learning of the  agent. Reward shaping, however, often requires significant quantities of domain knowledge that {are} highly task-specific~\cite{ng1999policy} and, even with careful design, can lead to undesired policies. Moreover, for complex robotic manipulation tasks, manually designing  reward shaping functions to guide the learning agent becomes intractable~\cite{arulkumaran2017deep,florensa2017reverse} if even minor variations to the task are introduced. For such settings, the application of deep reinforcement learning requires algorithms that can learn from unshaped, and usually sparse, reward signals. The complicated dynamics of robot manipulation exacerbate the difficulty posed by sparse rewards, especially for on-policy RL algorithms. For example, achieving goals that require successfully executing multiple steps over a long horizon involves high dimensional control {that must also generalise to work across variations in the environment for each step}. These aspects of robot control result in a situation where a naive RL agent so rarely receives a reward at the start of training that it is not able to learn at all. A common solution in the robotics community is to collect a sufficient quantity of expert demonstrations, then use imitation learning to train the agent. However, in some scenarios, demonstrations are expensive to collect and the achievable performance of a trained agent is restricted by their quantity. One solution is to use the valuable past experiences of the agent to enhance training, and this is particularly useful in sparse reward environments.

To alleviate the problems associated with having sparse rewards, there are two kinds of approaches: imitation learning and hindsight experience replay (HER). First, the standard approach of imitation learning is to use supervised learning algorithms and minimise a surrogate loss with respect to an oracle. The most common form is learning from demonstrations~\cite{hester2018deep,gao2018reinforcement}. Similar techniques are applied to robot manipulation tasks~\cite{rajeswaran2017learning,vevcerik2017leveraging,nair2018overcoming,james2018task}. When the demonstrations are not attainable, self-imitation learning (SIL)~\cite{oh2018self}, which uses past good experiences (episodes in which the goal is achieved), can~be used to enhance exploration or speed up the training of the agent. {Self-imitation learning} works well in discrete control environments, such as Atari Games. Whilst being able to learn policies for continuous control tasks with dense or delayed rewards~\cite{oh2018self}, {the present} experiments suggest that SIL struggles when rewards are sparse. Recently, {hindsight experience replay} has been proposed to solve such goal-conditional, sparse reward problems. The main idea of HER~\cite{andrychowicz2017hindsight} is that during replay, the selected transitions are sampled from state--action pairs derived from {achieved goals} that are substituted for the real goals of the task; this increases the frequency of positive rewards. {Hindsight experience replay} is used with off-policy RL algorithms, such as DQN~\cite{mnih2015human} and DDPG~\cite{lillicrap2015continuous}, for experience replay and has several extensions~\cite{schaul2015prioritized,liu2018competitive}. The present experiments show that simply applying HER with SIL does not lead to an agent capable of performing tasks from the Fetch robot environment. In summary, self-imitation learning with on-policy algorithms for tasks that require continuous control, and for which rewards are sparse, remains unsolved.

In this paper, {episodic self-imitation learning (ESIL) for goal-oriented problems that provide only sparse rewards is proposed and combined with a state-of-the-art on-policy RL algorithm}: proximal policy optimization (PPO). In contrast to standard SIL, which samples past good transitions from the replay buffer for imitation learning, {the proposed} ESIL adopts entire current episodes {(successful or not)}, and modifies them into ``expert'' trajectories based on HER. An extra trajectory selection module is also introduced to relieve the effects of sample correlation~\cite{lee2019sample} in updating the network. Figure~\ref{fig:esil} shows the difference between naive SIL+HER and ESIL. During training by SIL+HER, a batch of transitions is sampled from the replay buffer; these are modified into ``hindsight experiences'' and used directly in self-imitation learning. In contrast, ESIL utilises entire current collected episodes and converts them into hindsight episodes. The trajectory selection module removes undesired transitions in the hindsight episodes. Using tasks from the Open AI Fetch environment, this paper demonstrates that the proposed ESIL approach is effective in training agents which are required to solve continuous control problems, and shows that it achieves state-of-the-art results on several tasks.

\begin{figure}[t]
\centering
\includegraphics[width=0.95\textwidth]{figures/chapter3/buffer.png}
\caption{Illustration of difference between self-imitation learning (SIL)+hindsight experience replay (HER) and episodic self-imitation learning (ESIL).}
%% PPO+SIL+HER samples a batch of transitions from the replay buffer and modify them into hindsight experiences and conduct self-imitation learning directly. PPO+ESIL utilizes entire current collected episodes and convert them into hindsight episodes. Another trajectory selection module is used to remove undesired transitions in each hindsight episodes.
\label{fig:esil}
\end{figure}

The primary contribution of this paper is a novel episodic self-imitation learning (ESIL) algorithm that can solve continuous control problems {in environments providing only sparse rewards}; in doing so, it also empirically answers an open question posed by~\cite{plappert2018multi}. {The proposed ESIL approach also provides} a more efficient way to perform exploration in goal-conditional settings than the standard self-imitation learning algorithm. Finally, this approach achieves, to our knowledge, the best results for four moderately complex robot control tasks in simulation.

% Related work
\section{Related Work}
% exploration
Imitation learning (IL) can be divided into two main categories: behavioural cloning and inverse reinforcement learning~\cite{hussein2017imitation}. Behavioural cloning involves the learning of behaviours from demonstrations~\cite{bojarski2016end,xu2017end,torabi2018behavioral}. Other extensions have an expert in the loop, such as DAgger~\cite{ross2011reduction}, or use an adversarial paradigm for the behavioural cloning method~\cite{ho2016generative,wang2017robust}. The inverse reinforcement learning estimates a reward model from expert trajectories~\cite{ng2000algorithms,abbeel2004apprenticeship,ziebart2008maximum}. Learning from demonstrations is powerful for complex robotic manipulation tasks~\cite{finn2016guided,zhang2018deep,pmlr-v78-finn17a,rajeswaran2017learning,fang2019survey}. ~\cite{ho2016generative} propose generative adversarial imitation learning (GAIL), which employs generative adversarial training to match the distribution of state--action pairs of demonstrations. Compared with behavioural cloning, the GAIL framework shows strong improvements in continuous control tasks. In the work of~\cite{ding2019goal}, goalGAIL is proposed to speed up the training in goal-conditional environments; goalGAIL was also shown to be able to learn from demonstrations without action information. Prior work has used demonstrations to accelerate  learning~\cite{rajeswaran2017learning,vevcerik2017leveraging,nair2018overcoming}. Demonstrations are often collected by an expert policy or human actions. In contrast to these approaches, {episodic self-imitation learning (ESIL) does not need demonstrations.}

% self-learning and self-imitation learning
Self-imitation learning (SIL)~\cite{oh2018self} is used for exploiting past experiences for parametric policies. It~has a similar flavor to ~\cite{gangwani2018learning, pmlr-v97-wu19a}, in that the agent learns from imperfect demonstrations. During~training, past good experiences are stored in the replay buffer. When SIL starts, transitions are sampled from the replay buffer according to the advantage values. {In the work of~\cite{tang2020self}, generalised~SIL was proposed as an extension of SIL. It uses an $n$-bound $Q$-learning approach to generalise the original SIL technique, and shows robustness to a wide range of continuous control tasks. Generalised SIL can also be combined with both deterministic and stochastic RL algorithms. ~\cite{guo2019self} points out that using imitation learning with past good experience could lead to a sub-optimal policy. Instead of imitating past good trajectories, a trajectory-conditioned policy~\cite{guo2019self} is proposed to imitate trajectories in diverse directions, encouraging exploration in environments where exploration is otherwise difficult.} Unlike SIL, episodic self imitation learning (ESIL) applies HER to the current episodes to create ``imperfect'' demonstrations for imitation learning; this also requires introducing a trajectory-selection module to reject undesired samples from the hindsight experiences. In the work of~\cite{lee2019sample}, it was shown that the agent benefits from using whole episodes in updates, rather~than uniformly sampling the sparse or delayed reward environments. {The present} experiments suggest that episodic self-imitation learning {achieves better performance in an agent that must learn to perform continuous control in environments delivering sparse rewards}.

% Hindsight
Recently, the technique known as hindsight learning was developed. Hindsight experience replay (HER)~\cite{andrychowicz2017hindsight} is an algorithm that can overcome the exploration problems in multi-goal environments, delivering sparse rewards. Hindsight policy gradient (HPG)~\cite{rauber2018hindsight} introduces techniques that enable the learning of goal-conditional policies using hindsight experiences. However, the current implementation of HPG has only been evaluated for agents that need to perform discrete actions, and one drawback of hindsight policy gradient estimators is the computational cost because of the goal-oriented sampling. An extension of HER, called dynamic hindsight experience replay (DHER)~\cite{fang2018dher}, was proposed to deal with dynamic goals. ~\cite{liu2019hindsight} uses the GAIL framework~\cite{ho2016generative} to generate trajectories that are similar to hindsight experiences; it then applies imitation learning, using these trajectories. Competitive Experience Replay (CER) complements HER by introducing a competition between two agents for exploration~\cite{liu2018competitive}. ~\cite{zhao2018energy} point out that the hindsight trajectories which contain higher energy are more valuable during training, leading to a more efficient learning system. ~\cite{fang2019curriculum} proposed curriculum-guided HER, which~ incorporates curriculum learning in the work. During training, the agent focuses on the closest goals in the initial stage, then focuses on the expanding the diversity of goals. This approach accelerates training compared with other baseline methods. Unlike these works, episodic self-imitation learning (ESIL) combines episodic hindsight experiences with imitation learning, which aids learning at the start of training. Furthermore, ESIL can be applied to continuous control, making it more suitable for control problems that demand greater precision.

% Methodology
\section{Methodology}
\label{sec:method}
The proposed method combines PPO and episodic self-imitation learning to maximally use hindsight experiences for exploration to improve learning. {Recent advantages in episodic backward update~\cite{lee2019sample} and hindsight experiences~\cite{andrychowicz2017hindsight} are also leveraged} to guide exploration for on-policy RL.

\subsection{Episodic Self-Imitation Learning}
The present method aims to use episodic hindsight experiences to guide the exploration of the PPO algorithm. To this end, hindsight experiences are created from current episodes. For an episode $i$, let there be $T$ time steps; after $T$, a series of transitions $\tau_i =\{ (s_{t}|\langle g, g^{ac}_{t} \rangle, a_{t}, r_{t}, s_{t+1}|\langle g, g^{ac}_{t+1}\rangle )\}_{t=0:T-1}$ is collected. If at time step $t=T-1$, in $s_{T}$, $g^{ac}_{T} \neq g$, it implies that in this episode, the agent failed to achieve the original goal. Simply, to create hindsight experiences, {the achieved goal $g^{ac}_{T}$ in the last state $s_{T}$ is selected and considered as the modified desired goal $g^{\prime}$, i.e., $g^\prime = g^{ac}_{T}$.} Next, a new reward $r_{t}^{\prime}$ is computed under the new goal $g^\prime$. Then, {a new ``imagined'' episode is achieved}, and a new series of transitions $\tau_i^{\prime} = \{(s_{t}|\langle g^\prime, g^{ac}_{t} \rangle, a_{t}, r_{t}^{\prime},s_{t+1}| \langle g^\prime, g^{ac}_{t+1}\rangle )\}_{t=0:T-1}$ is collected.
Then, an approach to self-imitation learning based on episodic hindsight experiences is proposed, which applies the policy updates to both hindsight and in-environment episodes. Proximal policy optimization (PPO) is used as the base RL algorithm, which is a state-of-the-art on-policy RL algorithm. With current and corresponding hindsight experiences, a new objective function is introduced and defined as
\begin{equation}
  \centering
  \mathcal{L} = \alpha\cdot\mathcal{L}_{PPO} + \beta\cdot\mathcal{L}_{ESIL},
  \label{eq:loss}
\end{equation}
where $\alpha$ is the weight coefficient of $\mathcal{L}_{PPO}$. In the experiments, we set $\alpha=1$ as default {to balance the contribution of $\mathcal{L}_{PPO}$ and $\mathcal{L}_{ESIL}$}. $\mathcal{L}_{PPO}$ is the loss of PPO which can be written as
\begin{equation}
  \mathcal{L}_{PPO} = \mathcal{L}_{policy} - c\cdot \mathcal{L}_{value}
  \label{eq:ppo_loss}
\end{equation}
where $\mathcal{L}_{policy}$ is the policy loss which is parameterised by $\theta$, $\mathcal{L}_{value}$ is the value loss which is parameterised by $\eta$, and $c$ is the weight coefficient of the $\mathcal{L}_{value}$, which is set to 1 to match the default PPO setting~\cite{schulman2017proximal}. The policy loss, $\mathcal{L}_{policy}$, can be represented as
\begin{equation}
  \centering
  \mathcal{L}_{policy} = \mathbb{E}_{s_{t}, a_{t}, g\in\Tau}\biggr[\min \left (\frac{\pi_{\theta}(a_{t}|s_{t}, g)}{\pi_{\theta_{old}}(a_{t}|s_{t}, g)}A_{t}, \clip \left (\frac{\pi_{\theta}(a_{t}|s_{t}, g)}{\pi_{\theta_{old}}(a_{t}|s_{t}, g)}, 1-\epsilon, 1+\epsilon \right )A_{t}\right )\biggr],
  \label{eq:policy_loss}
\end{equation}
here, $A_{t}$ is the advantage value, and can be computed as $R_{t} - V_{\eta}(s_{t}, g)$. $V_{\eta}(s_{t}, g)$ is the state value at time step $t$ which is predicted by the critic network. $R_{t}$ is the return at time step $t$. $\epsilon$ is the clip ratio. $\Tau$~indicates original trajectories. {The value loss is an squared error loss $\mathcal{L}_{value}=(V_{\eta}(s_{t}, g) - R_{t})^{2}$.}

For the $\mathcal{L}_{ESIL}$ term, $\beta$ is an adaptive weight coefficient of $L_{ESIL}$; it can be defined as the ratio of samples which are selected for self-imitation learning
\begin{equation}
  \beta=\frac{N_{ESIL}}{N_{Total}},
\end{equation}
where $N_{ESIL}$ is the number of samples used for self-imitation learning and $N_{Total}$ is the total number of collected samples.
The episodic self-imitation learning loss $\mathcal{L}_{ESIL}$ can be written as 
\begin{equation}
  \centering
  \mathcal{L}_{ESIL}=\mathbb{E}_{s_{t}, a_{t}, g^\prime\in\Tau^{'}}\left[\log\pi_{\theta}\left(a_{t}|s_{t}, g^\prime\right)\cdot\mathcal{F}_t\right],
  \label{eq:imitaion}
\end{equation}
where $\Tau^{'}$ indicates hindsight trajectories and $\mathcal{F}_t$ is the trajectory selection module which is based on returns of the current episodes, $R$, and the returns of corresponding hindsight experiences, $R^{\prime}$.

\subsection{Episodic Update with Hindsight}
Two important issues of ESIL are: (1) hindsight experiences are sub-optimal, and (2) the {detrimental} effect of updating networks with correlated trajectories. Although episodic self-imitation learning makes exploration more effective, hindsight experiences are not from experts and not ``perfect'' demonstrations.  With the training process continuing, if the agent is always learning these imperfect demonstrations, the policy will be stuck at the sub-optimal, or experience overfitting.

To prevent the agent learning from imperfect hindsight experiences, {hindsight experiences are actively selected based on returns.} With the same action, different goals may lead to different results. The proposed method only selects hindsight experiences that can achieve higher returns. The illustration of the trajectory selection module is in Figure~\ref{fig:hs_princple}. For an episodic experience and its hindsight experience, the returns of the episodic experience and its hindsight experience can be calculated, respectively. 
In a trajectory, at time step $t$, the return $R_{t}$ can be calculated by $R_{t} = R(s_t, g) = r_{t} + \gamma \cdot r_{t+1} + \gamma^{2} \cdot r_{t+2} + ... + \gamma^{T-t-1} \cdot r_{T-1}$. Then, for a trajectory $\tau_{i}$, we have $\left\{R_{0}^{i}, R_{1}^{i}, R_{2}^{i}, ..., R_{T-1}^{i}\right\}$.

For the hindsight experiences, similarly, the return $R_{t}'$ for each time step, $t$,~with respect to the hindsight goals $g^\prime$, can be calculated. Based on the modified trajectory $\tau_{i}^{\prime}$ with the same length of $\tau_{i}$, we therefore have the returns $\left\{R_{0}^{i^\prime}, R_{1}^{i^\prime}, R_{2}^{i^\prime}, ..., R_{T-1}^{i^\prime}\right\}$. During training, the~hindsight experiences with higher returns are used for self-imitation learning. The rest of the hindsight experiences will be supposed to be worthless samples and ignored.
Then, Equation (\ref{eq:imitaion}) can be rewritten as
\begin{equation}
  \centering
  \mathcal{L}_{ESIL}=\mathbb{E}_{s_{t}, a_{t}, g^\prime\in\Tau^{'},g\in\Tau}\biggr[\log\pi_{\theta}\left(a_{t}|s_{t}, g^\prime\right)\cdot\mathcal{F}\left(s_{t}, g, g^\prime\right)\biggr],
  \label{eq:final_loss}
\end{equation}
where $\mathcal{F}\left(s_{t}, g, g^{\prime}\right)$ is the trajectory selection module. The selection function can be expressed as
\begin{equation}
  \centering
  \mathcal{F}\left(s_{t}, g, g^\prime \right)=\mathds{1}\left[R(s_{t}, g^\prime) > R(s_{t}, g)\right],
\end{equation}
here, {$\mathds{1}(\cdot)$ is the unit step function}. Consider the OpenAI FetchReach environment as an example. For~a failed trajectory, the rewards $r_{t}$ are $\{-1,-1,\cdots, -1\}$. {The desired goal is modified} to construct a new hindsight trajectory and the new rewards $r_{t}^{\prime}$ become $\{-1,-1,\cdots, 0\}$. Then,  {$R$ and $R^{\prime}$ can be calculated~separately.}
\begin{figure}[t!]
  \centering
  \includegraphics[width=\columnwidth]{figures/chapter3/hgs.png}
  \caption{A simplified illustration of trajectory selection. Blue trajectories indicate original experiences. Orange trajectories indicate hindsight experiences. Solid trajectories in the hindsight experiences are selected by the trajectory selection module of ESIL with new ``imagined'' goals.}
  \label{fig:hs_princple}
\end{figure}

From a goal perspective, episodic self-imitation learning (ESIL) tries to explore (desired) goals to get positive returns.~It can be viewed as a form of multi-task learning, {because ESIL has two objective functions to be optimised jointly.}~It is also related to self-imitation learning (SIL)~\cite{oh2018self}.~However,~ the~difference is that SIL uses $\left(R-V_{\theta}(s)\right)_{+}$ on past experiences to learn to choose the action chosen in the past in a given state, rather than goals. The full description of ESIL can be found in Algorithm~\ref{alg}.
% the algorithm
\begin{algorithm}[t!]
  \caption{Proximal policy optimization (PPO) with Episodic Self-Imitation Learning (ESIL)}
  \label{alg}
  \begin{algorithmic}[1]
    \REQUIRE an actor network $\pi(s, g|\theta)$, a critic network $V(s, g|\eta)$, the maximum steps $T$ of an episode, a reward function $r$
    \FOR{$\mathrm{iteration} = 1, 2, \cdots $}
    \STATE $\Tau = \varnothing$, $\Tau^{\prime} = \varnothing$ 
    \FOR{$\mathrm{episode} = 1, 2, ..., N$}
    \STATE $\tau = \varnothing$
    \FOR{$t = 0, 1, \cdots, T-1$}
    \STATE Sample an action $a_{t}$ using the actor network $\pi(s_{t}, g|\theta)$
    \STATE Execute the action $a_{t}$ and observe a new state $s_{t+1}$
    \STATE Store the transition $\left(s_{t}|\langle g, g^{ac}_{t} \rangle, a_{t}, r_{t}, s_{t+1}|\langle g, g^{ac}_{t+1} \rangle \right)$ in $\tau$
    \ENDFOR

    \FOR{\textbf{each} transition $\left(s_{t}, a_{t}, r_{t}, g, g^{ac}_t \right)$ in $\tau$}
    \STATE Clone the transition and replace $g$ with $g^{\prime}$, where $g^{\prime}=g^{ac}_{T}$
    \STATE $r_{t}^{\prime}:=r\left(s_{t}, a_{t}, g^{\prime}\right)$
    \STATE Store the transition $\left(s_{t}, a_{t}, r_{t}^{\prime}, g^{\prime}, g^{ac}_t\right)$ in $\tau^{\prime}$
    \ENDFOR
    % next loop
    \STATE Store the trajectory $\tau$ and the hindsight trajectory $\tau^{\prime}$ in $\Tau$ and $\Tau^{\prime}$, respectively 
    \ENDFOR

    \STATE Calculate the Return $R$ and $R^{\prime}$ for all transitions in $\Tau$ and $\Tau^{\prime}$, respectively
    \STATE Calculate the PPO loss: $\mathcal{L}_{PPO} = \mathcal{L}_{policy}(\theta) - c\cdot \mathcal{L}_{value}(\eta)$ using $\Tau$ \eqref{eq:ppo_loss}
    \STATE Calculate the ESIL loss: $\mathcal{L}_{ESIL}(\theta)$ using $\Tau^{\prime}$, $R$ and $R^{\prime}$ \eqref{eq:imitaion}
    \STATE Update the parameters $\theta$ and $\eta$ using loss $\mathcal{L}=\alpha\cdot\mathcal{L}_{PPO} + \beta\cdot\mathcal{L}_{ESIL}$ \eqref{eq:loss}
    \ENDFOR 
\end{algorithmic}
\end{algorithm}

\section{Experiments and Results}
\label{sec:experiment}
The proposed method is evaluated on several multi-goal environments, including the Empty Room environment and the OpenAI Fetch environments (see Figure~\ref{fig:env}). The Empty Room environment is a toy example, and has discrete action spaces. In the Fetch environments, there are four robot tasks with continuous action spaces. To obtain a comprehensive comparison between the proposed method and other baseline approaches, suitable baseline approaches are selected for different environments. Ablation studies of the trajectory selection module are also performed.
\begin{figure}[t]
\minipage{0.19\textwidth}
  \centering
  \includegraphics[width=\linewidth]{figures/chapter3/empty_room.png}
  ({a}) Empty Room \hspace{3em}          
\endminipage\hfill
\minipage{0.19\textwidth}
  \centering
  \includegraphics[width=\linewidth]{figures/chapter3/reach.png}
  ({b}) FetchReach \hspace{3em}  
\endminipage\hfill
\minipage{0.19\textwidth}%
  \centering
  \includegraphics[width=\linewidth]{figures/chapter3/push.png}
  ({c}) FetchPush \hspace{3em}        
\endminipage\hfill
\minipage{0.19\textwidth}%
  \centering
  \includegraphics[width=\linewidth]{figures/chapter3/pick.png}
  ({d}) FetchPickPlace
\endminipage\hfill
\minipage{0.19\textwidth}%
  \centering
  \includegraphics[width=\linewidth]{figures/chapter3/slide.png}
  ({e}) FetchSlide \hspace{3em}       
\endminipage
\caption{Evaluation environments. ({a}) is the Empty Room environment, in which a yellow circle indicates the position of the agent and a red star represents a target position. ({b}--{e}) are the Fetch robotic environments. The red spot represents a target position.}
\label{fig:env}
\end{figure}

\subsection{Setup}
\textbf{{Empty Room (grid-world) environment}}:
The Empty Room environment is a simple grid-world environment. The agent is placed in an $11\times11$ grid, representing the room. The goal of the agent is to reach a target position in the room. The start position of the agent is at the left upper corner of the room, and the target position is randomly selected  within the room. When the agent chooses an action that would lead it to fall outside the grid area, the agent stays at the current position. The length of each episode is 32. 
The desired goal, $g$, is a two-dimensional grid coordinate which represents the target position. The achieved goal, $g^{ac}_{t}$, is also a two-dimensional coordinate which represents the current position of the agent at time step $t$, and finally, the observation is a two-dimensional coordinate which represents the current position of the agent. 
The agent has five actions: left, right, up, down and stay; the agent executes a random action with probability 0.2. The agent can get $+1$ as a reward only when $g^{ac}_{t}=g$, otherwise, it gets a reward of $0$.

The agent is trained with 1 CPU core. In each epoch, 100 episodes are collected for the training. After each epoch, the agent is evaluated for 10 episodes. During training, the actions are sampled from the categorical distribution. During evaluation, the action with the highest probability will be chosen.

\textbf{{Fetch robotic (continuous) environments}%.
}~\cite{plappert2018multi}:~The Fetch robotic environments are physically plausible simulations based on the real Fetch robot. The purpose of these environments is to provide a platform to tackle problems which are close to practical challenging robot manipulation tasks. Fetch is a 7-DoF robot arm with a two finger gripper. The Fetch environments include four~tasks: \emph{FetchReach}, \emph{FetchPush}, \emph{FetchPickAndPlace} and \emph{FetchSlide}. 
For all Fetch tasks, the length of each episode is 50. The~desired goal, $g$, is a three-dimensional coordinate which represents the target position. If~a task has an object, the achieved goal $g^{ac}_{t}$ is a three-dimensional coordinate represents the position of the object. Otherwise, $g^{ac}_{t}$ is a three-dimensional coordinate represents the position of the gripper.
Observations include the following information: position, velocity and state of the gripper. If a task has an object, the position, velocity and rotation information of the object is included. Therefore, the~observation of FetchReach is a 10-dimensional vector.~The observation of other tasks is a 25-dimensional vector.~The~action is a four-dimensional vector. The first three dimensions represent the relative position that the gripper needs to move in the next step. The last dimension indicates the distance between the fingers of the gripper.
The reward function can be written as $r_{t}=-\mathds{1}\left(\left\|g^{ac}_{t} - g\right\|>\epsilon\right)$, where $\epsilon =0.05$.

In the Fetch environments, for FetchReach, FetchPush and FetchPickAndPlace tasks, the agent is trained using 16 CPU cores. In each epoch, 50 episodes are collected for training. The FetchSlide task is more complex, so 32 CPU cores are used. In each epoch, 100 episodes are collected for training. The~{Message Passing Interface (MPI) framework is used to perform}  synchronization when updating the network. After each epoch, the agent is evaluated for 10 episodes by each MPI worker. Finally,~the~success rate of each MPI worker is averaged. During training, actions are sampled from {multivariate normal} distributions. In the evaluation phase, the mean vector of the distribution is used as an action.

The proposed method, termed PPO+ESIL, is compared with different baselines on different environments. All experiments are plotted based on five runs with different seeds. The solid line is the median value. The upper bound is the {75th} percentile and the lower bound is the {25th} percentile.

\subsection{Network Structure and Hyperparameters}
\textbf{Network structure}: Both the actor network and the critic network have three hidden layers with 256~neurons. ReLu is selected as the activation function for the hidden layers. In the grid-world environment, the actor network builds a categorical distribution. In the Fetch environment, the actor network builds normal {distributions by producing mean vectors and the standard deviations of the independent variables}.

\textbf{Hyperparameters}: For all experiments, the learning rate is 0.0003 for both the actor and critic networks. The discount factor $\gamma$ is 0.98. Adam is chosen as an optimiser with $\epsilon=0.00001$. For each epoch, the~actor network and critic network are updated 10 times. The clip ratio of the PPO algorithm is 0.2. For the grid-world environment, it trains networks for 100 epochs with batch size equals 160. Each~epoch consists of 100 episodes. For the Fetch environments, in FetchReach task, it trains networks for 100 epochs and other tasks for 1000 epochs with batch size equals to 125. For FetchReach, FetchPush and FetchPickAndPlace tasks, each epoch consists of 50 episodes. For FetchSlide task, each epoch consists of 100 episodes. {In designing the experiments, the number of episodes within an epoch is a balance between being able to train, the length of time required to run experiments and the maximum number of time steps that would be required to a achieve a goal. All environments have a fixed maximum number of time-steps , but this maximum differs depending on the problem or environment. This means that the number of state--action pairs can differ between two environments that have the same number of episodes and the same number of epochs. We arrange the episodes to try to compensate for the number of state--action pairs collected during training to make experiments easier to compare.} {The models are trained on a machine with an Intel i7-5960X CPU and 64GB RAM.}

%\section{Results and Discussions}
\subsection{Grid-World Environments}
To understand the basic properties of the proposed method, the toy Empty Room environment is used to evaluate ESIL.
%We compare the proposed method, named as PPO+HSL, with other baselines on the grid-world environment.
The following baselines are considered:
\begin{itemize}
    \item PPO: vanilla PPO~\cite{schulman2017proximal} for discrete action spaces;
    \item PPO+SIL/PPO+SIL+HER: Self-imitation learning (SIL) is used with PPO to solve hard exploration environments by imitating past good experiences~\cite{oh2018self}. In order to solve sparse rewards tasks, hindsight experience replay (HER) is applied to sampled transitions;
    \item DQN+HER: Hindsight experience replay (HER), designed for sparse reward problems, is~combined with a deep Q-learning network (DQN)~\cite{andrychowicz2017hindsight}; this is an off policy algorithm;
    \item Hindsight Policy Gradients (HPG): the vanilla implementation of HPG that is only suitable for discrete action spaces~\cite{rauber2018hindsight}.
\end{itemize}
More specifically, {PPO+ESIL is compared with above baseline methods} in Figure~\ref{fig:toy_results}a. This shows that PPO+ESIL converges faster than the other four baselines, and PPO+SIL converges faster than vanilla PPO, because PPO+SIL reuses past good experiences to help exploration and training. Hindsight~Policy Gradient (HPG) is slower than the others because goal sampling is not efficient and also unstable. 
\begin{figure}[H]
\minipage{0.5\textwidth}
  \centering
  \includegraphics[width=\linewidth]{figures/chapter3/empty_room_baseline.pdf}
  ({a}) Comparison with on-policy baselines\hspace{3em} 
\endminipage
\minipage{0.5\textwidth}
  \centering
  \includegraphics[width=\linewidth]{figures/chapter3/empty_room_hs.pdf}
  ({b}) {Ablation study of selection module}
\endminipage\hfill
\minipage{0.5\textwidth}%
  \centering
  \includegraphics[width=\linewidth]{figures/chapter3/empty_room_samples.pdf}
  ({c}) Variation of adaptive weight coefficient $\beta$
\endminipage
\minipage{0.5\textwidth}%
  \centering
  \includegraphics[width=\linewidth]{figures/chapter3/empty_room_her.pdf}
  ({d}) Comparison with off-policy baselines
\endminipage\hfill
\caption{Results of the grid-world environment. ({a}) Comparing the performance of PPO+ESIL between the on-policy approaches. ({b}) An ablation study on the trajectory selection module. ({c}) The variation of adaptive weight coefficient $\beta$ through training. ({d}) Comparison of the performance of PPO+ESIL to an off-policy approach: DQN+HER.}
\label{fig:toy_results}
\end{figure}

Further, {the performance of the trajectory selection module is evaluated} in Figure~\ref{fig:toy_results}b. This shows that the selection strategy helps improve the performance. Hindsight experiences are not always perfect; the trajectory selection module filters some undesirable, modified experiences. Through adopting this selection strategy, the chance of agents learning from poor trajectories is reduced. The adaptive weight coefficient $\beta$ is also investigated in these experiments. {In Figure~\ref{fig:toy_results}c, it can be seen that at the initial stages of training, $\beta$ is high}. {This is because at this stage, the agent very seldom achieves the original goals}. The hindsight experiences can yield higher returns than the original experiences. {Therefore,~a~large proportion of hindsight experiences are selected to conduct self-imitation learning, helping the agent learn a policy for moving through the room}. In the later stages of training, the agent can achieve success frequently, and the hindsight experiences might be redundant (e.g., $R(s_{t}, g)\geq R(s_{t}, g^{\prime})$). In~this case, undesired hindsight experiences are removed by using the trajectory selection module and $L_{PPO}$ leads the training. However, when the trajectory selection module is not employed, all hindsight experiences are used through the entire training process which includes the redundant hindsight experiences. This leads to overfitting and makes training unstable. Thus, the $L_{ESIL}$ can provide the agent with a better initial policy, and the adaptive weight coefficient $\beta$ can balance the contributions of $L_{PPO}$ and $L_{ESIL}$ properly during training.

\begin{figure}[t!]
\centering
\minipage{0.5\textwidth}
  \centering
  \includegraphics[width=\linewidth]{figures/chapter3/reach_baseline.pdf}
  ({a}) FetchReach
\endminipage
\minipage{0.5\textwidth}%
  \centering
  \includegraphics[width=\linewidth]{figures/chapter3/push_baseline.pdf}
  ({b}) FetchPush
\endminipage\hfill
\minipage{0.5\textwidth}%
  \centering
  \includegraphics[width=\linewidth]{figures/chapter3/pick_baseline.pdf}
  ({c}) FetchPickAndPlace
\endminipage
\minipage{0.5\textwidth}%
  \centering
  \includegraphics[width=\linewidth]{figures/chapter3/slide_baseline.pdf}
  ({d}) FetchSlide
\endminipage\hfill
\caption{Results of comparison between ESIL and on-policy baselines on all Fetch environments.}
\label{fig:baseline_compare}
\end{figure}
Finally, {the combination of PPO+ESIL is also compared with DQN+HER,} which is an off-policy RL algorithm, in Figure~\ref{fig:toy_results}d. This shows that DQN+HER works a little better than ESIL at the {start of training}. However, the proposed method achieves similar results to DQN+HER later in training.

\subsection{Continuous Environments}
{Continuous control problems are generally more challenging for reinforcement learning. In the experiments of this section, the aim is to investigate how useful the proposed method is for several hard exploration OpenAI Gym Fetch tasks. These environments are commonly used to assess the performance of RL methods for continuous control.} The following baselines are considered: 
\begin{itemize}
    \item PPO: the vanilla PPO~\cite{schulman2017proximal} for continuous action spaces;
    \item PPO+SIL/PPO+SIL+HER: Self-imitation learning is used with PPO to solve hard exploration environments by imitating past good experiences~\cite{oh2018self}. For sparse rewards tasks, hindsight experience replay (HER) is applied to sampled transitions;
    \item DDPG+HER: this is the state-of-the-art off-policy RL algorithm for the Fetch tasks. Deep~deterministic policy gradient (DDPG) is trained with HER to deal with the sparse reward problem~\cite{andrychowicz2017hindsight}.
\end{itemize}
\begin{figure}[H]
\centering
\minipage{0.5\textwidth}
  \centering
  \includegraphics[width=\linewidth]{figures/chapter3/reach_hs.pdf}
  ({a}) FetchReach
\endminipage
\minipage{0.5\textwidth}%
  \centering
  \includegraphics[width=\linewidth]{figures/chapter3/push_hs.pdf}
  ({b}) FetchPush
\endminipage\hfill
\minipage{0.5\textwidth}%
  \centering
  \includegraphics[width=\linewidth]{figures/chapter3/pick_hs.pdf}
  ({c}) FetchPickAndPlace
\endminipage
\minipage{0.5\textwidth}%
  \centering
  \includegraphics[width=\linewidth]{figures/chapter3/slide_hs.pdf}
  ({d}) FetchSlide
\endminipage\hfill
\caption{Results of ablation studies with or without using trajectory selection module on all Fetch environments.}
\label{fig:hs_compare}
\end{figure}

\subsubsection{Comparison to On-Policy Baselines}
{Figure~\ref{fig:baseline_compare}, PPO+ESIL} achieves reasonable results on all Fetch environments. {In contrast}, PPO, PPO+SIL and PPO+SIL+HER do not work on all tasks, with the exception of FetchReach. In comparison with the other selected tasks from the Fetch environments, FetchReach is relatively simple, because there is no object to be manipulated. For other tasks, it is quite difficult for the agent to achieve sufficient positive rewards during exploration, because of their rare occurrence. Although PPO+SIL utilises past good experiences to help exploration, it is still faced with the difficulty that past experiences do not easily achieve positive rewards. From the experiments ({see Figure~\ref{fig:baseline_compare}}), PPO+SIL (no hindsight) converges much more slowly than using PPO only. Attempting to use only the original trajectories for self-imitation learning leads to unsatisfactory performance. For PPO+SIL+HER (with no episodic update), the sampled transitions are modified into hindsight experiences, achieving better performance in the FetchReach and FetchSlide tasks. However, this transition-based method still cannot solve the other two manipulation tasks. In contrast, the proposed PPO+ESIL, through utilizing episodic hindsight experiences from failed trajectories, can achieve positive rewards quickly at the start of~training.

\subsubsection{Ablation Study of Trajectory Selection Module}
In order to investigate the effect of trajectory selection, {ablation studies are performed} to validate the selection strategy of our approach. {Figure~\ref{fig:hs_compare}}, when the trajectory selection module is not used, the~ performance of the agent increases at first, and then starts to decrease. This suggests that the agent starts to converge to a sub-optimal location. However, {Figure~\ref{fig:hs_compare}d}, for the FetchSlide task, the agent converges faster without the trajectory selection module, and has better performance. {This is likely to be because FetchSlide is the most difficult of the Fetch environments. During training, the agent is very unlikely to achieve positive rewards. Figure~\ref{fig:her_samples_compare} also indicates that the value of $\beta$ in FetchSlide is higher than values in other environments, which means the majority of hindsight experiences have higher returns than original experiences. Thus, using {\em more} hindsight experiences (without filtering) accelerates training at this stage.} Nonetheless, the trajectory selection module prevents the agent overfitting the hindsight experience in the other three tasks. Figure~\ref{fig:her_samples_compare}, shows the adaptive weight coefficient $\beta$ on all Fetch environments. When the trajectory selection module is used, the value of $\beta$ decreases with the increase in training epochs. This implies that the agent can achieve a greater proportion of the original goals in the latter stages of training, and fewer hindsight experiences are required for  self-imitation learning.
\begin{figure}[H]
\centering
\minipage{0.5\textwidth}
  \centering
  \includegraphics[width=\linewidth]{figures/chapter3/reach_samples.pdf}
  ({a}) FetchReach
\endminipage
\minipage{0.5\textwidth}%
  \centering
  \includegraphics[width=\linewidth]{figures/chapter3/push_samples.pdf}
  ({b}) FetchPush
\endminipage\hfill
\minipage{0.5\textwidth}%
  \centering
  \includegraphics[width=\linewidth]{figures/chapter3/pick_samples.pdf}
  ({c}) FetchPickAndPlace
\endminipage
\minipage{0.5\textwidth}%
  \centering
  \includegraphics[width=\linewidth]{figures/chapter3/slide_samples.pdf}
  ({d}) FetchSlide
\endminipage\hfill
\caption{Variation in adaptive weight coefficient $\beta$ through the training on all Fetch environments.}
\label{fig:her_samples_compare}
\end{figure}
\subsubsection{Comparison to Off-Policy Baselines}
Finally, {the proposed method is also compared} with a state-of-the-art off-policy algorithm: DDPG+HER. From {Figure~\ref{fig:her_compare}, it may be seen that} DDPG+HER converges faster than PPO+ESIL in all tasks. {However, PPO+ESIL obtains a similar performance to DDPG+HER. This is because DDPG+HER is an off-policy algorithm and uses a large number of hindsight experiences. A replay buffer is also employed to store  samples collected in the past. This approach has better sample efficiency than on-policy algorithms such as PPO.} Even so, {Figure~\ref{fig:her_compare}c} shows that PPO+ESIL still outperforms DDPG+HER in the FetchPickAndPlace task and the success rate is close to 1. {This suggests that PPO+ESIL approximates the characteristics of on-policy algorithms, which have low sample efficiency, but are able to obtain a comparable performance to off-policy algorithms in continuous control tasks~\cite{schulman2017proximal}.}
\begin{figure}[H]
\minipage{0.5\textwidth}
  \centering
  \includegraphics[width=\linewidth]{figures/chapter3/reach_her.pdf}
  ({a}) FetchReach
\endminipage
\minipage{0.5\textwidth}%
  \centering
  \includegraphics[width=\linewidth]{figures/chapter3/push_her.pdf}
  ({b}) FetchPush
\endminipage\hfill
\minipage{0.5\textwidth}%
  \centering
  \includegraphics[width=\linewidth]{figures/chapter3/pick_her.pdf}
  ({c}) FetchPickAndPlace
\endminipage
\minipage{0.5\textwidth}%
  \centering
  \includegraphics[width=\linewidth]{figures/chapter3/slide_her.pdf}
  ({d}) FetchSlide
\endminipage\hfill
\caption{Results of comparison between PPO+ESIL and DDPG+HER on all Fetch environments.}
\label{fig:her_compare}
\end{figure}

\subsection{Overall Performance}
{Table~\ref{tb:last}} shows the average success rate of the last 10 epochs during training of baseline methods and PPO+ESIL. The proposed ESIL achieves the best performance in four out of five tasks. However~PPO and PPO+SIL  only obtain reasonable results for the Empty Room and FetchReach tasks. With the assistance of HER, PPO+SIL+HER obtains a better performance in the FetchSlide task. {For the off-policy methods of DDPG+HER, all five tasks are achieved, but a better performance is  obtained than PPO+ESIL only in the FetchPush task}.

\begin{table}[H]
  \caption{Average success rate $\pm$ standard error in the last 10 epochs over five random seeds on all~environments {(\textbf{bold} indicates the best result among all methods).}}
  \label{tb:last}
  \centering
  \resizebox{\textwidth}{!}{
  \begin{tabular}{llllll}
    \toprule
           & {Empty Room} & {Reach} & {Push} & {Pick} & {Slide}  \\
    \midrule
    PPO & 1.000 $\pm$ 0.000 & 1.000 $\pm$ 0.000& 0.070 $\pm$ 0.001 & 0.033 $\pm$ 0.001 & 0.077 $\pm$ 0.001 \\
    PPO + SIL & 0.998 $\pm$ 0.002 & 0.225 $\pm$ 0.016 & 0.071 $\pm$ 0.001 & 0.036 $\pm$ 0.002 & 0.011 $\pm$ 0.001 \\
    PPO + SIL + HER &0.996 $\pm$ 0.013 & 1.000 $\pm$ 0.000 & 0.066 $\pm$ 0.011 & 0.035 $\pm$ 0.004 & 0.276 $\pm$ 0.011 \\
    DQN + HER &1.000 $\pm$ 0.000 & - & - & - & - \\
    DDPG + HER & - & 1.000 $\pm$ 0.000 & \textbf{0.996 $\pm$ 0.001} & 0.888 $\pm$ 0.008 & 0.733 $\pm$ 0.013\\
    HPG &0.964 $\pm$ 0.012 & - & - & - & - \\
    \midrule
    PPO + ESIL (Ours) & \textbf{1.000 $\pm$ 0.000} & \textbf{1.000 $\pm$ 0.000}& 0.984 $\pm$ 0.003& \textbf{0.986 $\pm$ 0.002}& \textbf{0.812 $\pm$ 0.015}\\
    \bottomrule
  \end{tabular}}
\end{table}
% Conclusion

\section{Conclusions}
\label{sec:conclusion}
This paper proposed a novel method for self-imitation learning (SIL), in which an on-policy RL algorithm uses episodic modified past trajectories, i.e., hindsight experiences, to update policies. Compared with standard self-imitation learning, episodic self-imitation learning (ESIL) has a better performance in continuous control tasks where rewards are sparse. As far as we know, it is also the first time that hindsight experiences have been combined with state-of-the-art on-policy RL algorithms, such as PPO, to solve relatively hard exploration environments in continuous action spaces.

The experiments that we have conducted suggest that simply using self-imitation learning with the PPO algorithm, even with hindsight experience, leads to disappointing performance in continuous control Fetch tasks. In contrast, the episodic approach we take with ESIL is able to learn in these sparse reward settings. The auxiliary trajectory selection module and the adaptive weight $\beta$ help the training process to remove undesired experiences and balance the contributions to learning between the PPO term and the ESIL term automatically, and also increase the stability of training.

Our experiments suggest that the selection module is useful to prevent overfitting to sub-optimal hindsight experiences, but also that it does not always lead to learning a better policy faster. Despite~this, selection filtering appears to support learning a useful policy in challenging environments. The experiments we have conducted to date have utilised relatively small networks, and it would be appropriate to extend the experiments to consider more complex observation spaces, and to actor/critic networks, which are consequently more elaborate.

Future work includes extending {the proposed} method to support hierarchical reinforcement learning (HRL) algorithms for more complex manipulation control tasks, such as in-hand manipulation. Episodic self-imitation learning (ESIL) can also be applied to simultaneously learn sub-goal policies.

### 2021-12-31 dedication/dedication.tex
remove redudant index
init draft of chapter3

@@
 \begin{enumerate}
+	\item Cher Bass, \underline{Tianhong Dai}, Benjamin Billot, Kai Arulkumaran, Antonia Creswell, Claudia Clopath, Vincenzo de Paola, and Anil Anthony Bharath, ``Image Synthesis with a Convolutional Capsule Generative Adversarial Network", International Conference on Medical Imaging with Deep Learning, 2019 (Oral).
 	\item Shafa Balaram, Kai Arulkumaran, \underline{Tianhong Dai}, and Anil Anthony Bharath, ``A Maximum Entropy Deep Reinforcement Learning Neural Tracker", International Workshop on Machine Learning in Medical Imaging, 2019.
 	\item Yali Du, Lei Han, Meng Fang, \underline{Tianhong Dai}, Ji Liu, and Dacheng Tao, ``LIIR: Learning Individual Intrinsic Reward in Multi-Agent Reinforcement Learning", Advances in Neural Information Processing Systems, 2019.
-	\item \underline{Tianhong Dai}, Wei Li, Xilei Cao, Jianzhuang Liu, Xu Jia, Ales Leonardis, Youliang Yan, and Shanxin Yuan, ``Wavelet-Based Network For High Dynamic Range Imaging", Submitted to IEEE Transactions on Imaging Processing, 2021.
 \end{enumerate}

### 2022-01-01 chapters/chapter4.tex
init chapter4

\chapter{Diversity-based Trajectory and Goal Selection with Hindsight Experience Replay}
\section{Introduction}
Deep reinforcement learning (DRL)~\cite{arulkumaran2017deep}, in which neural networks are used as function approximators for reinforcement learning (RL), has been shown to be capable of solving complex control problems in several environments, including board games~\cite{schrittwieser2020mastering,silver2017mastering}, video games~\cite{berner2019dota,mnih2015human,vinyals2019grandmaster}, simulated and real robotic manipulation~\cite{andrychowicz2020learning,gu2017deep,levine2016end} and simulated autonomous driving~\cite{kiran2021deep}.

However, learning from a \textit{sparse} reward signal, where the only reward is provided upon the completion of a task, still remains difficult. An agent may rarely or never encounter positive examples from which to learn in a sparse-reward environment. Many domains therefore provide dense reward signals~\cite{brockman2016openai}, or practitioners may turn to reward shaping~\cite{ng1999theory}. Designing dense reward functions typically requires prior domain knowledge, making this approach difficult to generalise across different environments.

Fortunately, a common scenario is goal-oriented RL, where the RL agent is tasked with solving different goals within the same environment~\cite{kaelbling1993learning,schaul2015universal}. Even if each task has a sparse reward, the agent ideally \textit{generalises} across goals, making the learning process easier. For example, in a robotic manipulation task, the goal during a single episode would be to achieve a specific position of a target object.

Hindsight experience replay (HER)~\cite{NIPS2017_453fadbd} was proposed to improve the learning efficiency of goal-oriented RL agents in sparse reward settings: when past experience is replayed to train the agent, the desired goal is replaced (in ``hindsight'') with the achieved goal, generating many positive experiences. In the above example, the desired target position would be overwritten with the achieved target position, with the achieved reward also being overwritten correspondingly.

We note that HER, whilst it enabled solutions to previously unsolved tasks, can be somewhat inefficient in its use of uniformly sampling transitions during training. In the same way that prioritised experience replay~\cite{schaul2016prioritized} has significantly improved over the standard experience replay in RL, several approaches have improved upon HER by using data-dependent sampling~\cite{fang2019curriculum,zhao2018energy}. HER with energy-based prioritisation (HEBP)~\cite{zhao2018energy} assumes semantic knowledge about the goal-space and uses the energy of the target objects to sample trajectories with high energies, and then samples transitions uniformly. Curriculum-guided HER (CHER)~\cite{fang2019curriculum} samples trajectories uniformly, and then samples transitions based on a mixture of proximity to the desired goal and the diversity of the samples; CHER adapts the weighting of these factors over time. In this work, we introduce diversity-based trajectory and goal selection with HER (DTGSH; See Fig.~\ref{fig:illustration}), which samples trajectories based on the diversity of the goals achieved within the trajectory, and then samples transitions based on the diversity of the set of samples. In this paper, DTGSH is evaluated on five challenging robotic manipulation tasks. From extensive experiments, our proposed method converges faster and reaches higher rewards than prior work, without requiring domain knowledge~\cite{zhao2018energy} or tuning a curriculum~\cite{fang2019curriculum}, and is based on a single concept---determinantal point processes (DPPs)~\cite{kulesza2012determinantal}.
\begin{figure}[h]
    \centering
    \includegraphics[width=\textwidth]{figures/chapter4/illustration_DTGSH_latest.pdf}
    \caption{Overview of DTGSH. Every time a new episode is completed, its diversity is calculated, and it is stored in the episodic replay buffer. During training, $m$ episodes are sampled according to their diversity-based priority, and then $k$ diverse, hindsight-relabelled transitions are sampled using a $k$-DPP~\cite{kulesza2011k}.}
    \label{fig:illustration}
\end{figure}

\section{Methodology}
We now formally describe the two main components of our method, DTGSH: 1) a diversity-based trajectory selection module to sample valuable trajectories for the further goal selection; 2) a diversity-based goal selection module to select transitions with diverse goal states from the previously selected trajectories. Together, these select informative transitions from a large area of the goal space, improving the agent's ability to learn and generalise.

\subsection{Diversity-based Trajectory Selection}
We propose a diversity-based prioritization method to select valuable trajectories for efficient training. Related to HEBP's prioritisation of high-energy trajectories~\cite{zhao2018energy}, we hypothesise that trajectories that achieve diverse goal states $g^{ac}_{t}$ are more valuable for training; however, unlike HEBP, we do not require knowledge of the goal space semantics.

In a robotic manipulation task, the agent needs to move a target object from its initial position, $g^{ac}_{1}$, to the target position, $g$. If the agent never moves the object, despite hindsight relabelling it will not be learning information that would directly help in task completion. On the other hand, if the object moves a lot, hindsight relabelling will help the agent learn about meaningful interactions.

In our approach, DPPs are used to model the diversity of achieved goal states $g^{ac}_{t}$ in an episode, or subsets thereof. For a single trajectory $\Tau$ of length $T$, we divide it into several partial trajectories $\tau_{j}$ of length $b$, with achieved goal states $\{g^{ac}_{t}\}_{t=n:n+b-1}$. That is, with a sliding window of $b = 2$, a trajectory $\Tau$ can be divided into $N_p$ partial trajectories:
\begin{equation}
    \Tau_{i} = \{\{\underbrace{g^{ac}_{1}, g^{ac}_{2}}_{\tau_{1}}\}, \{\underbrace{g^{ac}_{2}, g^{ac}_{3}}_{\tau_{2}}\}, \{\underbrace{g^{ac}_{3}, g^{ac}_{4}}_{\tau_{3}}\}, \cdots, \{\underbrace{g^{ac}_{T-1}, g^{ac}_{T}}_{\tau_{N_{p}}}\}\}.
\end{equation}
The diversity $d_{\tau_{j}}$ of each partial trajectory $\tau_{j}$ can be computed as:
\begin{equation}
    d_{\tau_{j}} = \text{det}(L_{\tau_{j}}), 
\label{eq:partial_div}
\end{equation}
where $L_{\tau_{j}}$ is the kernel matrix of partial trajectory $\tau_{j}$:
\begin{equation}
    L_{\tau_{j}} = M^{T}M,
\end{equation}
and $M=[\hat{g}^{ac}_{n}, \hat{g}^{ac}_{n+1}, \cdots, \hat{g}^{ac}_{n+b-1}]$, where each $\hat{g}^{ac}$ is the $\ell_2$-normalised version of the achieved goal $g^{ac}$~\cite{kulesza2011k}. Finally, the diversity $d_\Tau$ of trajectory $\Tau$ is the sum of the diversity of its $N_p$ constituent partial trajectories:
\begin{equation}
    d_\Tau = \sum_{j=1}^{N_{p}} d_{\tau_{j}}.
\label{eq:total_div}
\end{equation}

Similarly to HEBP~\cite{zhao2018energy}, we use a non-uniform episode sampling strategy. During training, we prioritise sampling episodes proportionally to their diversity; the probability $p(\Tau_{i})$ of sampling trajectory $\Tau_{i}$ from a replay buffer of size $N_{e}$ is:
\begin{equation}
    p(\Tau_{i}) = \frac{d_{\Tau_{i}}}{\sum_{n=1}^{N_{e}} d_{\Tau_{n}}}.
\label{eq:diversity_priority}
\end{equation}

\subsection{Diversity-based Goal Selection}
In prior work~\cite{NIPS2017_453fadbd,zhao2018energy}, after selecting the trajectories from the replay buffer, one transition from each selected trajectory is sampled uniformly to construct a minibatch for training. However, the modified goals $g^{\prime}$ in the minibatch might be similar, resulting in redundant information. In order to form a minibatch with diverse goals for more efficient learning, we use $k$-DPPs~\cite{kulesza2011k} for sampling goals. Compared to the standard DPP, a $k$-DPP is a conditional DPP where the subset $Y$ has a fixed size $k$, with the probability distribution function:
\begin{equation}
    \mathcal{P}_{L}^{k}(\mathbf{Y}=Y) = \frac{\text{det}(L_{Y})}{\sum_{|Y^{\prime}|=k} \text{det}(L_{Y^{\prime}})}.
\end{equation}
$k$-DPPs are more appropriate for goal selection with a minibatch of fixed size $k$. Given $m > k$ trajectories sampled from the replay buffer, we first uniformly sample a transition from each of the $m$ trajectories. Finally, a $k$-DPP is used to sample a diverse set of transitions based on the relabelled goals $g'$ (which, in this context, we denote as ``candidate goals''). Fig.~\ref{fig:goal_sampling} gives an example of uniform vs. $k$-DPP sampling, demonstrating the increased coverage of the latter. Fig.~\ref{fig:kde} provides corresponding estimated density plots; note that the density of the $k$-DPP samples is actually more uniform over the \textit{support} of the candidate goal distribution.
\begin{figure}[h]
    \begin{subfigure}{\textwidth}
        \centering
        \includegraphics[width=0.9\textwidth]{figures/chapter4/goals_distribution_2d.pdf}
        \caption{Plot of candidate goals and selected goals. $k$-DPP sampling is more likely to sample points from the full span of the goal space.}
        \label{fig:goal_sampling}
    \end{subfigure}
    \begin{subfigure}{\textwidth}
        \centering
        \includegraphics[width=0.9\textwidth]{figures/chapter4/kde.pdf}
        \caption{Kernel density estimation of the distributions of goals. $k$-DPP leads to a more uniform selection of goals over the support of the goal space.}
        \label{fig:kde}
    \end{subfigure}
    \caption{Visualisation of $k=32$ goals selected from $m=100$ candidate goals of the Push task using either uniform sampling or $k$-DPP sampling, respectively. The candidate goals are distributed over a 2D ($xy$) space. Note that  $k$-DPP sampling (right hand plots) results in a broader span of selected goals in $xy$ space compared to uniform sampling (left hand plots).}
\end{figure}

% pseudo code of dpp-sampling
\begin{algorithm}[t!]
\caption{Diversity-based Goal Selection using $k$-DPP}
\label{alg:goal_sample}
\begin{algorithmic}[1]
 \REQUIRE set of $m$ candidate goal states $\mathcal{G}:=\{g_{i}\}_{i=1:m}$, minibatch size $k$
 \STATE $J \leftarrow\varnothing$, $M \leftarrow  [g_{1}, g_{2}, \cdots, g_{m}]$
 \STATE Calculate the DPP kernel matrix $L_M$
 \STATE $\{\boldsymbol{v}_{n}, \lambda_{n}\} \leftarrow \text{EigenDecomposition}(L_M)$
 %\STATE $e_{k}(\lambda_{1}, \cdots, \lambda_{m}): = \sum\limits_{|J'|=k}\prod_{n\in J'}\lambda_{n}$ \hfill{$\triangleright$ elementary symmetric polynomial: $e^{m}_{k}$}
 \STATE $e_k(\lambda_1,\lambda_2,\dots,\lambda_m) := 
  \sum\limits_{J^{\prime} \subseteq \{1,2,\dots,m\}\atop|J^{\prime}|=k} \prod\limits_{n \in J^{\prime}} \lambda_n$ \hfill{$\triangleright$ elementary symmetric polynomial: $e^{m}_{k}$}
 \FOR{$\mathrm{n} = m, m-1, \cdots, 1$}
 \IF{$u\sim \text{Uniform}[0, 1] < \lambda_{n}\frac{e^{n-1}_{k-1}}{e_{k}^{n}}$}
 \STATE $J \leftarrow J \cup \{n\}$, $k \leftarrow k - 1$
 \IF{$k = 0$}
 \STATE \textbf{break}
 \ENDIF
 \ENDIF
 \ENDFOR
 \STATE $V \leftarrow \{\boldsymbol{v}_{n}\}_{n \in J}$, $B \leftarrow \varnothing$
 \WHILE{$|V| > 0$}
 \STATE Select $g_{i}$ from $\mathcal{G}$ with $p(g_{i}):=\frac{1}{|V|}\sum_{\boldsymbol{v}\in V}(\boldsymbol{v}^{T}\boldsymbol{b}_{i})^{2}$\hfill{$\triangleright$ $\boldsymbol{b}_{i}$ is the $i^{\text{th}}$ standard basis}
 \STATE $B \leftarrow B \cup \{g_{i}\}$
 \STATE $V \leftarrow V_{\perp}$ \hfill{$\triangleright$ an orthonormal basis for the subspace of $V$ orthogonal to $\boldsymbol{b}_{i}$}
 \ENDWHILE
 \RETURN minibatch $B$ with size $k$
\end{algorithmic}
\end{algorithm}

\begin{algorithm}[t!]
\caption{Diversity-based Trajectory and Goal Selection with HER}
\label{alg:main_alg}
\begin{algorithmic}[1]
 \REQUIRE RL environment with episodes of length $T$, number of episodes $N$, off-policy RL algorithm $\mathbb{A}$, episodic replay buffer $\mathcal{B}$, number of algorithm updates $U$, candidate transitions size $m$, minibatch size $k$
 \STATE Initialize the parameters $\theta$ of all models in $\mathbb{A}$
 \STATE $\mathcal{B}\leftarrow\varnothing$
 \FOR{$\mathrm{i} = 1, 2, \cdots, N$}
 \STATE Sample a desired goal $g$ and an initial state $s_{0}$\hfill{$\triangleright$ start a new episode}
 \FOR{$\mathrm{t} = 1, 2, \cdots, T$}
 \STATE Sample an action $a_{t}$ using the policy $\pi(s_{t}, g;\theta)$
 \STATE Execute action $a_{t}$ and get the next state $s_{t+1}$ and achieved goal state $g^{ac}_{t+1}$
 \STATE Calculate $r_{t}$ according to Eq.~\eqref{eq:sparse_reward}
 \STATE Store transition $(s_t, a_t, r_t, s_{t+1}, g, g^{ac}_{t+1})$ in $\mathcal{B}$
 \ENDFOR
 \STATE Calculate the diversity score of current episode $d_{\Tau_{i}}$ using Eq.~\eqref{eq:partial_div} and Eq.~\eqref{eq:total_div}
 \STATE Calculate the diversity-based priority $p(\Tau)$ of each episode in $\mathcal{B}$ using Eq.~\eqref{eq:diversity_priority}
 \FOR{$\mathrm{iteration} = 1, 2, \cdots, U$}
 \STATE Sample $m$ trajectories from $\mathcal{B}$ according to priority $p(\Tau)$
 \STATE Uniformly sample one transition from each of the $m$ trajectories
 \STATE Relabel goals in each transition and recompute the reward to get $m$ candidate transitions $\{(s_{t}, a_{t}, r_{t}^{\prime}, s_{t+1}, g^{\prime})_{n}\}_{n=1:m}$
 \STATE Sample minibatch $B$ with size $k$ from the $m$ candidates using Alg.~\ref{alg:goal_sample}
 \STATE Optimise $\theta$ with minibatch $B$
 \ENDFOR
 \ENDFOR 
\end{algorithmic}
\end{algorithm}

Alg.~\ref{alg:goal_sample} shows the details of the goal selection subroutine, and Alg.~\ref{alg:main_alg} gives the overall algorithm for our method, DTGSH.

\section{Experiments}
We evaluate our proposed method, and compare it with current state-of-the-art HER-based algorithms~\cite{NIPS2017_453fadbd,fang2019curriculum,zhao2018energy} on challenging robotic manipulation tasks~\cite{plappert2018multi}, pictured in Fig.~\ref{fig:env}. Furthermore, we perform ablation studies on our diversity-based trajectory and goal selection modules. Our code is based on OpenAI Baselines\footnote[1]{\url{https://github.com/openai/baselines}}, and is available at: \url{https://github.com/TianhongDai/div-hindsight}.

\begin{figure}
  \begin{subfigure}[t]{0.19\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter4/push_resize.png}
    \caption{Push}
    \label{subfig:env_push}
  \end{subfigure}\hfill
  \begin{subfigure}[t]{0.19\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter4/pick_resize.png}
    \caption{Pick$\&$Place}
    \label{subfig:env_pick}
  \end{subfigure}\hfill
  \begin{subfigure}[t]{0.19\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter4/handegg_resize.png}
    \caption{EggFull}
    \label{subfig:env_handegg}
  \end{subfigure}\hfill
  \begin{subfigure}[t]{0.19\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter4/handblock_resize.png}
    \caption{BlockRotate}
    \label{subfig:env_handblock}
  \end{subfigure}\hfill
  \begin{subfigure}[t]{0.19\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter4/handpen_resize.png}
    \caption{PenRotate}
    \label{subfig:env_handpen}
  \end{subfigure}\hfill
  \caption{Robotic manipulation environments. (a-b) use the Fetch robot, and (c-e) use the Shadow Dexterous Hand.} 
  \label{fig:env}
\end{figure}

\subsection{Environments}
The robotic manipulation environments used for training and evaluation include five different tasks. Two tasks use the 7-DoF Fetch robotic arm with two-fingers parallel gripper: Push, and Pick\&Place, which both require the agent to move a cube to the target position. The remaining three tasks use a 24-DoF Shadow Dexterous Hand to manipulate an egg, a block and a pen, respectively. The sparse reward function is given by Eq.~\eqref{eq:sparse_reward}.

In the Fetch environments, the state $s_{t}$ contains the position and velocity of the joints, and the position and rotation of the cube. Each action $a_{t}$ is a 4-dimensional vector, with three dimensions specifying the relative position of the gripper, and the final dimension specifying the state of the gripper (i.e., open or closed). The desired goal $g$ is the target position, and the achieved goal $g^{ac}_{t}$ is the position of the cube. Each episode is of length $T = 50$.

In the Shadow Dexterous Hand environments, the state $s_{t}$ contains the position and velocity of the joints. Each action $a_{t}$ is a 20-dimensional vector which specifies the absolute position of 20 non-coupled joints in the hand. The desired goal $g$ and achieved goal $g^{ac}_t$ specify the rotation of the object for the block and pen tasks, and the position + rotation of the object for the egg task. Each episode is of length $T = 200$.

\subsection{Training Settings}
We base our training setup on CHER~\cite{fang2019curriculum}. We train all agents on minibatches of size $k = 64$ for 50 epochs using MPI for parallelisation over 16 CPU cores; each epoch consists of 1600 ($16 \times 100$) episodes, with evaluation over 160 ($16 \times 10$) episodes at the end of each epoch. Remaining hyperparameters for the baselines are taken from the original work~\cite{NIPS2017_453fadbd,fang2019curriculum,zhao2018energy}. Our method, DTGSH, uses partial trajectories of length $b = 2$ and $m = 100$ as the number of candidate goals.

\subsection{Benchmark Results}
We compare DTGSH to DDPG~\cite{lillicrapContinuous}, DDPG+HER~\cite{NIPS2017_453fadbd}, DDPG+HEBP~\cite{zhao2018energy} and DDPG+CHER~\cite{fang2019curriculum}. Evaluation results are given based on repeated runs with 5 different seeds; we plot the median success rate with upper and lower bounds given by the \nth{75} and \nth{25} percentiles, respectively.

\begin{figure}
  \begin{subfigure}[t]{0.33\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter4/FetchPush-v1.pdf}
    \caption{Push}
    \label{subfig:baseline_push}
  \end{subfigure}\hfill
  \begin{subfigure}[t]{0.33\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter4/FetchPickAndPlace-v1.pdf}
    \caption{Pick\&Place}
    \label{subfig:baseline_pick}
  \end{subfigure}\hfill
  \begin{subfigure}[t]{0.33\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter4/HandManipulateEggFull-v0.pdf}
    \caption{EggFull}
    \label{subfig:baseline_handegg}
  \end{subfigure}\hfill
  \begin{subfigure}[t]{0.33\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter4/HandManipulateBlockRotateXYZ-v0.pdf}
    \caption{BlockRotate}
    \label{subfig:baseline_handblock}
  \end{subfigure}\hfill
  \begin{subfigure}[t]{0.33\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter4/HandManipulatePenRotate-v0.pdf}
    \caption{PenRotate}
    \label{subfig:baseline_handpen}
  \end{subfigure}\hfill
  \begin{subfigure}[t]{0.33\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter4/blank.png}
  \end{subfigure}\hfill
  \caption{Success rate of DTGSH and baseline approaches.} 
  \label{fig:baseline_compare}
\end{figure}

Fig.~\ref{fig:baseline_compare} and Tab.~\ref{tab:benchmark} show the performance of DDPG+DTGSH and baseline approaches on all five tasks. In the Fetch tasks, DDPG+DTGSH and DDPG+HEBP both learn significantly faster than the other methods, while in the Shadow Dexterous Hand tasks DDPG+DTGSH learns the fastest and achieves higher success rates than all other methods. In particular, DDPG cannot solve any tasks without using HER, and CHER performs worse in the Fetch tasks. We believe the results highlight the importance of sampling both diverse trajectories \textit{and} goals, as in our proposed method, DTGSH.
\begin{table}[h]
    \centering
    \resizebox{\textwidth}{!}{
    \begin{tabular}{c|c@{\hskip 0.1in}|c@{\hskip 0.1in}|c|c|c}
    \toprule
         & Push & Pick\&Place & EggFull & BlockRotate & PenRotate \\
        \midrule
        DDPG & 0.09$\pm$0.01 & 0.04$\pm$0.00 & 0.00$\pm$0.00 & 0.01$\pm$0.00 & 0.00$\pm$0.00 \\
        DDPG+HER & \textbf{1.00$\pm$0.00} & 0.89$\pm$0.03 & 0.11$\pm$0.01 & 0.55$\pm$0.04 & 0.15$\pm$0.02 \\
        DDPG+HEBP & \textbf{1.00$\pm$0.00} & 0.91$\pm$0.03 & 0.14$\pm$0.02 & 0.59$\pm$0.02 & 0.20$\pm$0.03\\
        DDPG+CHER & \textbf{1.00$\pm$0.00} & 0.91$\pm$0.04 & 0.15$\pm$0.01 & 0.54$\pm$0.04 & 0.17$\pm$0.03 \\
        \midrule
        DDPG+DTGSH & \textbf{1.00$\pm$0.00} & \textbf{0.94$\pm$0.01} & \textbf{0.17$\pm$0.03} & \textbf{0.62$\pm$0.02} & \textbf{0.21$\pm$0.02}\\
        \bottomrule
    \end{tabular}
    }
    \vspace{0.2em}
    \caption{Final mean success rate $\pm$ standard deviation, with best results in \textbf{bold}.}
    %\vspace{-14mm}
    \label{tab:benchmark}
\end{table}
%\vspace{-15mm}
\subsection{Ablation Studies}
In this section, we perform the following experiments to investigate the effectiveness of each component in DTGSH: 1) diversity-based trajectory selection with HER (DTSH) and diversity-based goal selection with HER (DGSH) are evaluated independently to assess the contribution of each stage; 2) the performance using different partial trajectory lengths $b$; 3) the performance of using different candidate goal set sizes $m$.
\begin{figure}[h]
  \begin{subfigure}[t]{0.33\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter4/FetchPush-v1_ab1.pdf}
    \caption{Push}
    \label{subfig:baseline_push_ab1}
  \end{subfigure}\hfill
  \begin{subfigure}[t]{0.33\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter4/FetchPickAndPlace-v1_ab1.pdf}
    \caption{Pick\&Place}
    \label{subfig:baseline_pick_ab1}
  \end{subfigure}\hfill
  \begin{subfigure}[t]{0.33\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter4/HandManipulateEggFull-v0_ab1.pdf}
    \caption{EggFull}
    \label{subfig:baseline_handegg_ab1}
  \end{subfigure}\hfill
  \begin{subfigure}[t]{0.33\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter4/HandManipulateBlockRotateXYZ-v0_ab1.pdf}
    \caption{BlockRotate}
    \label{subfig:baseline_handblock_ab1}
  \end{subfigure}\hfill
  \begin{subfigure}[t]{0.33\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter4/HandManipulatePenRotate-v0_ab1.pdf}
    \caption{PenRotate}
    \label{subfig:baseline_handpen_ab1}
  \end{subfigure}\hfill
  \begin{subfigure}[t]{0.33\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter4/blank.png}
  \end{subfigure}\hfill
  \caption{Success rate of HER, DTGSH, and ablations DTSH and DGSH.} 
  \label{fig:ablation1}
\end{figure}

Fig.~\ref{fig:ablation1} shows the performance of using DTSH and DGSH independently. DDPG+DTSH outperforms DDPG+HER substantially in all tasks, which supports the view that sampling trajectories with diverse achieved goals can substantially improve performance. Furthermore, unlike DDPG+HEBP, DTSH does not require knowing the structure of the goal space in order to calculate the energy of the target object; DDPG+DGSH achieves better performance than DDPG+HER in three environments, and is only worse in one environment. DGSH performs better in environments where it is easier to solve the task (e.g., Fetch tasks), and hence the trajectories selected are more likely to contain useful transitions. However, DTGSH, which is the combination of both modules, performs the best overall.

\begin{figure}[t]
  \begin{subfigure}[t]{0.33\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter4/FetchPush-v1_ab2.pdf}
    \caption{Push}
    \label{subfig:baseline_push_ab2}
  \end{subfigure}\hfill
  \begin{subfigure}[t]{0.33\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter4/FetchPickAndPlace-v1_ab2.pdf}
    \caption{Pick\&Place}
    \label{subfig:baseline_pick_ab2}
  \end{subfigure}\hfill
  \begin{subfigure}[t]{0.33\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter4/HandManipulateEggFull-v0_ab2.pdf}
    \caption{EggFull}
    \label{subfig:baseline_handegg_ab2}
  \end{subfigure}\hfill
  \begin{subfigure}[t]{0.33\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter4/HandManipulateBlockRotateXYZ-v0_ab2.pdf}
    \caption{BlockRotate}
    \label{subfig:baseline_handblock_ab2}
  \end{subfigure}\hfill
  \begin{subfigure}[t]{0.33\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter4/HandManipulatePenRotate-v0_ab2.pdf}
    \caption{PenRotate}
    \label{subfig:baseline_handpen_ab2}
  \end{subfigure}\hfill
  \begin{subfigure}[t]{0.33\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter4/blank.png}
  \end{subfigure}\hfill
  \caption{Success rate of DTGSH with different partial trajectory lengths $b$ and different candidate goal set sizes $m$.} 
  \label{fig:ablation2}
\end{figure}

Fig.~\ref{fig:ablation2} shows the performance of DDPG+DTGSH with different partial trajectory lengths $b$ and different candidate goal set sizes $m$. In this work, we use $b = 2$ and $m = 100$ as the defaults. Performance degrades with $b \gg 2$, indicating that pairwise diversity is best for learning in our method. $m \gg 100$ does not affect performance in the Fetch environments, but degrades performance in the Shadow Dexterous Hand environments.

\subsection{Time Complexity}
Table~\ref{tab:time} gives example training times of all of the HER-based algorithms. DTGSH requires an additional calculation of the diversity score of $\mathcal{O}(N_{p}b^3)$ at the end of every training episode, and sampling of $\mathcal{O}(mk^2)$ for each minibatch.
\begin{table}[h]
    \centering
    \resizebox{\textwidth}{!}{
    \begin{tabular}{c|c|c|c|c}
    \toprule
           & DDPG+HER & DDPG+HEBP & DDPG+CHER & DDPG+DTGSH\\
        \midrule
        Time  & 00:55:08 & 00:56:32 & 03:02:18 & 01:52:30 \\
        \bottomrule
    \end{tabular}
    }
    \vspace{0.2em}
    \caption{Training time (hours:minutes:seconds) of DTGSH and baseline approaches on the Push task for 50 epochs.}
    \label{tab:time}
\end{table}

\section{Conclusion}
In this paper, we introduced diversity-based trajectory and goal selection with hindsight experience replay (DTGSH) to improve the learning efficiency of goal-orientated RL agents in the sparse reward setting. Our method can be divided into two stages: 1) valuable trajectories are selected according to diversity-based priority, as modelled by determinantal point processes (DPPs)~\cite{kulesza2012determinantal}; 2) $k$-DPPs~\cite{kulesza2011k} are leveraged to sample transitions with diverse goal states from previously selected trajectories for training. Our experiments empirically show that DTGSH achieves faster learning and higher final performance in five challenging robotic manipulation tasks, compared to previous state-of-the-art approaches~\cite{NIPS2017_453fadbd,fang2019curriculum,zhao2018energy}. Furthermore, unlike prior extensions of hindsight experience replay, DTGSH does not require semantic knowledge of the goal space~\cite{zhao2018energy}, and does not require tuning a curriculum~\cite{fang2019curriculum}.

### 2022-01-01 chapters/chapter4.tex
switch conclusion to discussion

@@
 
-\section{Conclusion}
+\section{Discussion}
 In this paper, we introduced diversity-based trajectory and goal selection with hindsight experience replay (DTGSH) to improve the learning efficiency of goal-orientated RL agents in the sparse reward setting.

### 2022-04-23 abstract/abstract.tex
update

\cleardoublepage
\addcontentsline{toc}{chapter}{Abstract}

\begin{abstract}
Deep reinforcement learning (DRL) has made great progress in dealing with complex control problems in \td{various testing scenarios}, such as playing video games, playing board games, and dexterous robotic manipulation; \td{and real-world applications, such as controlling plasmas for nuclear fusion.} However, DRL requires large amounts of interactions with an environment to find an optimal policy to solve the task, limiting its application in real-world problems. In this thesis, we focus on two aspects to improve sample efficiency in DRL: 1) solving sparse reward tasks and 2) improving general exploration strategies.

First, we propose two methods to overcome exploration difficulties and improve learning efficiency in goal-oriented RL with the sparse reward setting, where an agent can rarely achieve positive feedback. In the first method, to provide sufficient positive samples for training an agent, \td{hindsight goal relabelling is used to replace goals in original samples with intermediate goals, and these augmented positive samples are leveraged to accelerate the training via a self-imitation learning paradigm.} An additional selection module is also designed to remove undesirable modified samples and stabilise training. In the second method, to alleviate the inefficiency of hindsight experience replay (HER) caused by its uniform sampling strategy, a diversity-based sampling method is employed to select valuable and diverse experiences for efficient training.

Second, a diversity-augmented intrinsic motivation is introduced to encourage the agent to explore novel states in an environment with sparse or delayed rewards. During training, the diversity of adjacent state sequences is measured under the framework of determinantal point processes (DPPs) and this measurement is used as an auxiliary reward to facilitate the exploration of the agent, thus improving the final performance.

Furthermore, we analyse the trained agents with and without domain randomisation (DR), a technique that can reduce the reality gap between a simulator and real-world scenarios. Through evaluating their robustness to previous unseen environments and applying both qualitative and quantitative interpretability methods, we provide the insight into the behaviour of the trained agents. Finally, some suggestions are also given to researchers who intend to adopt interpretability methods to analyse DRL agents.

\end{abstract}

### 2022-04-23 acknowledgements/acknowledgements.tex
update

\cleardoublepage

\addcontentsline{toc}{chapter}{Acknowledgements}

\begin{acknowledgements}

I would like to express my deep gratitude to:

\begin{itemize}
 \item Anil Anthony Bharath: Anil is my supervisor and brings me into an academic research career. During my Ph.D. study, he provided me with plentiful guidance, help and trust in my research. In addition, he has also provided me with valuable suggestions on my career planning. Without him, I would not have been able to achieve the current results.
 \vspace*{3mm}
 \item Kai Arulkumaran: Kai is my closest colleague/collaborator in the BICI group. He has helped me a lot in my research. In addition, he also gives me many suggestions in daily life. He is also my ``leader" in the field of reinforcement learning.
 \vspace*{3mm}
 \item BICI Group Members: I would like to express my gratitude for their daily support.
 \vspace*{3mm}
 \item My parents: From the day I was born, my parents have provided me with unfailing love and support. Without them, I wouldn't be here today.
 \vspace*{3mm}
 \item My wife: From the day we meet, she continually inspires me to chase my dreams and gives me much love and care. During our two years together in the UK, she has been there to guide me through difficult times, to help me with my worries and to allow me to focus on my work. She makes my life so full of hope and joy that no words can express my gratitude and love to her.
 % 从我们认识的那天起，她就给予我充足的关怀以及爱。我们一起在英国的这两年间，她每天为我准备美味的食物以及提供舒适的生活环境，让我能够专注于我自己的工作。也是她让我的生活充满了希望与快乐，再多的文字也表达不出我对她的感激与爱。
\end{itemize}

\end{acknowledgements}

### 2022-04-23 background/background.tex
update

\chapter{Background Theory}

\label{ch:background}
In this chapter, the background covered in this thesis is presented according to the following structure: in Section~\ref{ch3:dnn} and Section~\ref{ch3:bp}, we first briefly discuss deep neural network (DNN), back-propagation and optimisation methods. Then, in Section~\ref{ch3:rl}, some background to reinforcement learning (RL) is introduced. Next, in Section~\ref{ch3:policy-optim}, we discuss more details about policy optimisation in deep reinforcement learning (DRL) and related DRL algorithms that are covered in this thesis. Finally, in Section~\ref{ch3:dpp}, we introduce the fundamentals of determinantal point processes (DPPs).
%Finally, in Section~\ref{ch3:env}, we introduce the benchmark environments which are adopted to evaluate the proposed methods.

\section{Deep Neural Network in DRL}
\begin{figure}[h]
    \centering
    \includegraphics[width=0.7\textwidth]{figures/background/mlp_background.pdf}
    \caption{A simple structure of a multi-layer perceptron (MLP). It consists of an input layer, two hidden layers and an output layer. The highlighted lines (\textcolor{red}{red}) show the connection between a single neuron and its previous layer.}
    \label{fig:nn}
\end{figure}
\label{ch3:dnn}
\subsection{Multi-layer Perceptron}
Deep learning is based on research of artificial neural networks (ANNs). The structure of ANNs is inspired by animal brains, and mimics signal propagation between biological neurons. A feed-forward neural network (FFNN) or multi-layer perceptron (MLP) is a standard structure of ANN, and it consists of several neuron layers, which include an input layer, one or several hidden layers and an output layer (see Figure~\ref{fig:nn}). In each layer, there are multiple neurons, and each neuron is connected to those in the previous layer with corresponding weights $w_{i}$ and a bias $b$. For example, in Figure~\ref{fig:nn}, the transmission of the signal between the highlighted neuron and the previous layer can be expressed as:
\begin{equation}
    f = \sigma (w_{0}x_{0} + w_{1}x_{1} + w_{2}x_{2} + b)
\label{eq:nn_compute}
\end{equation}
here, $\sigma(\cdot)$ represents an activation function, such as a rectifier linear unit (ReLU)~\cite{nair2010rectified}, sigmoid$(\cdot)$ and tanh$(\cdot)$, which are used to add non-linearity to the network, because most real-world problems are nonlinear. The hidden layers in a neural network are used to learn various non-linear combinations of the input data, and the learned non-linear combinations are also called features. These features are the key factors to decide the prediction of the network.

The general forward propagation in an MLP that can be described as the following way. Firstly, the input layer receives the input data of a specific task. The input data can be a flattened image or the property of a house (e.g., the number of rooms, location, and area). Then, in each hidden layer, the output of each neuron is computed according to Equation~\eqref{eq:nn_compute}, and this computation proceeds from the first hidden layer to the last hidden layer. Finally, in the output layer, the output of the network can be the probability of each class in a classification task, the predicted value in a regression task, or the policy/predicted state value in a DRL task.

However, the initial parameters (i.e., weights and bias) of a FFNN may not fit a function for solving a specific task. Thus, the parameters need to be updated through a back-propagation method which is discussed in Section~\ref{ch3:bp}.
% CNN
\subsection{Convolutional Neural Network}
Multi-layer perceptrons (MLPs) are somewhat inefficient in dealing with computer vision tasks because of its fully connected structure. As the input size increases, the number of parameters (weight and bias) also increases. Convolutional neural network (CNN)~\cite{lecun1998gradient} alleviates this problem by using learnable kernels/filters, where each filter traverses the entire image. \td{CNNs are a strong inductive bias that reflects translation invariance of the physical world.} Compared with an MLP, the parameters of these kernels are shared, less, and easy to train. In previous works~\cite{krizhevsky2012imagenet,girshick2014rich,detone2016deep}, CNN has demonstrated its better feature representation ability than other hand-crafted methods, such as histogram of oriented gradient (HOG)~\cite{dalal2005histograms} and scale-invariant feature transform (SIFT)~\cite{lowe2004distinctive}. Nowadays, an increased number of powerful network structures have been proposed, such as the VGG network~\cite{simonyan2014very}, the residual network~\cite{he2016deep}, the dense network~\cite{huang2017densely}, and the capsule network~\cite{sabour2017dynamic}. These network structures improve state-of-the-art performance on various benchmark tasks~\cite{deng2009imagenet,everingham2010pascal}. In DRL, CNNs are regularly used to extract features from high-dimensional inputs (e.g., images) for further processing. In this thesis, as in many DRL studies, we put less effort into the design of the network structure.
%\td{[shifting and scaling in-variance makes CNN important]}

\section{Back-propagation and Optimisation}
\label{ch3:bp}
Gradient descent~\cite{bottou2010large} is a common technique used to update the parameters of the deep neural network (DNN) to fit a model, and it relies on the gradient of an objective function/loss function with respect to the parameters of the network. The gradient can be calculated efficiently through using the back-propagation algorithm~\cite{kelley1960gradient}. In the forward propagation of an MLP with $N$ layers, the output of each hidden layer can be formulated as:
\begin{equation}
    \mathbf{f}_{i} = \sigma(\mathbf{W}_{i-1}\mathbf{f}_{i-1} + \mathbf{b}_{i-1}), \quad i = 1, 2, \cdots, N,
\end{equation}
where $\mathbf{f}_{i}$ is the output of the current hidden layer, $\mathbf{f}_{i-1}$ is the output of the previous hidden layer, and the input is $\mathbf{f}_{0}$; $\mathbf{W}_{i-1}$ and $\mathbf{b}_{i-1}$ are the weights and bias of the current layer, respectively. Then, in this example, we use the squared error as the loss function to calculate the gradient and update the parameters of the network:
\begin{equation}
    \min_{\bm{\theta}}\mathcal{L} = \norm{\mathbf{y} - \mathbf{f}_{N}(\mathbf{x};\bm{\theta})}^{2},
\end{equation}
where $\bm{\theta}$ is the collection of parameters of the network: $\{\mathbf{W_{0}}, \mathbf{b_{0}}, \mathbf{W_{1}}, \mathbf{b_{1}}, \cdots, \mathbf{W_{N-1}}, \mathbf{b_{N-1}}\}$, $\mathbf{y}$ is the ground truth value and $\mathbf{x}$ is the input data. To minimise the error between the ground truth and the prediction, we need to compute the partial derivatives of $\mathcal{L}$ with respect to the parameters $\bm{\theta}_{i}$ of each layer using the chain rule:
\begin{equation}
   \nabla_{\bm{\theta}_{i}}\mathcal{L} = \frac{\partial \mathcal{L}}{\partial \bm{\theta_{i}}} = \frac{\partial \mathcal{L}}{\partial \mathbf{f}_{N}} \frac{\partial\mathbf{f}_{N}}{\partial \mathbf{f}_{N-1}}\cdots \frac{\partial\mathbf{f}_{i+2}}{\partial \mathbf{f}_{i+1}} \frac{\partial\mathbf{f}_{i+1}}{\partial \bm{\theta}_{i}}.
\end{equation}

Once get the gradient $\nabla_{\bm{\theta}}\mathcal{L}$, the parameters $\bm{\theta}$ can be updated in the opposite direction of the gradient with a step size $\eta$ (i.e. learning rate) via gradient descent:
\begin{equation}
    \bm{\theta} \leftarrow \bm{\theta} - \eta\nabla_{\bm{\theta}}\mathcal{L}.
\end{equation}
However, gradient descent requires the calculation of the gradient over all samples, and it is inefficient when the number of samples is large. Stochastic gradient descent (SGD) reduces computational complexity by randomly splitting the entire samples into several minibatches, and each minibatch is comprised of one or more samples. Then, the gradient is computed over each minibatch samples to update the parameters. 
%Furthermore, SGD with momentum~\cite{qian1999momentum} can accelerate the convergence speed on top of standard SGD.
In this thesis, two variants of the SGD method are used to update the parameters of the network: root mean square propagation (RMSProp)~\cite{tieleman2012lecture} and Adam~\cite{kingma2014adam}.

Root mean square propagation (RMSProp) can adaptively adjust the learning rate for different parameters (i.e., per-parameter learning rate) to accelerate the convergence, the update rule can be written as:
\begin{align}
    \mathbf{v} &\leftarrow \beta \mathbf{v} + (1 - \beta)\nabla_{\bm{\theta}}\mathcal{L}\odot\nabla_{\bm{\theta}}\mathcal{L} \label{eq:squared_grad}\\
     \bm{\theta} &\leftarrow \bm{\theta} - \frac{\eta}{\epsilon+\sqrt{\mathbf{v}}}\odot\nabla_{\bm{\theta}}\mathcal{L} \label{eq:rmsprop_update}.
\end{align}
In Equation~\eqref{eq:squared_grad}, $\mathbf{v}$ is the moving average of the squared gradient, $\odot$ is the element-wise multiplication, and $\beta$ is the decay coefficient which indicates the ratio of history information used to compute the average value. In Equation~\eqref{eq:rmsprop_update}, $\epsilon$ is a small value to prevent division by zero, $\frac{1}{(\epsilon + \sqrt{\mathbf{v}})}$ is the weight to balance the learning rate for individual parameter, in doing so, the learning rate for large gradient will decrease to avoid exploding, and the learning rate for small gradient will increase to avoid vanishing.

Adam is a combination of momentum~\cite{qian1999momentum} and RMSProp, and it is the most popular optimiser in DRL community. The update rule of Adam is:
\begin{align}
    \mathbf{m} &\leftarrow \beta_{1} \mathbf{m} + (1 - \beta_{1})\nabla_{\bm{\theta}}\mathcal{L} \label{eq:momentum_adam}\\
    \mathbf{v} &\leftarrow \beta_{2} \mathbf{v} + (1 - \beta_{2})\nabla_{\bm{\theta}}\mathcal{L}\odot\nabla_{\bm{\theta}}\mathcal{L} \label{eq:squared_grad_adam}\\
     \bm{\theta} &\leftarrow \bm{\theta} - \frac{\eta}{\epsilon+\sqrt{\mathbf{v}}}\mathbf{m} \label{eq:adam_update}.
\end{align}
Equation~\eqref{eq:momentum_adam} is the momentum term, $\mathbf{m}$ is the running average of the gradient, and $\beta_{1}$ is the decay coefficient. The momentum term can correct the gradient in the right direction and thus accelerate the convergence. The other parts are similar to RMSProp.

\section{Reinforcement Learning}
\label{ch3:rl}
\subsection{Markov Decision Process}
\begin{figure}[h]
    \centering
    \includegraphics[width=0.8\textwidth]{figures/background/mdp_background.pdf}
    \caption{An overview of reinforcement learning (RL).}
    \label{fig:rl_desc}
\end{figure}
Reinforcement learning (RL) can be formulated under the framework of a Markov Decision Process (MDP); it is used to learn an optimal policy to solve sequential decision-making problems. The MDP is a mathematical framework which describes a fully observable environment in RL, and it is composed of a 5-tuple $<\mathcal{S}, \mathcal{A}, \mathcal{P}, \mathcal{R}, \gamma>$. $\mathcal{S}$ is a set of states, $\mathcal{A}$ is a set of actions, $\mathcal{P}$ is a transition probability/dynamics function: $p(s_{t+1}|s_{t},a_{t})$ which only depends on the current state and action, $\mathcal{R}$ is a reward function, and $\gamma$ is a discount factor. At each timestep $t$, a state $s_{t}\in\mathcal{S}$ is received by an agent from the environment (see Figure~\ref{fig:rl_desc}). An action $a_{t}\in\mathcal{A}$ is sampled by the agent according to its policy $\pi(a_{t}|s_{t})$. Then, the next state $s_{t+1}$ and reward $r_{t}$ are provided to the agent according to the transition function and the reward function. The goal of the RL is to have the agent learn a policy that maximises the expected return $\mathbb{E}_{\pi}[G_{t}]$~\cite{sutton2018reinforcement}:
\begin{equation}
\centering
    G_{t} =\sum_{k=0}^{T-1}\gamma^{k}r_{t+k}.
\label{eq:acc_reward_func}
\end{equation}
Here, the discount factor $\gamma$ exponentially downplays the influence of future rewards. {In a robot manipulation task, the state $s_{t}$ can be the velocity and position of each joint of a robot arm. The action $a_{t}$ can be the velocities of actuators (control signals)} and the reward $r_{t}$ might be calculated based on the distance between the gripper of the robot arm and the target position.

\subsection{Goal-Oriented Reinforcement Learning}
\label{ch3:goal_rl}
The purpose of goal-oriented RL or multi-goal RL is to acquire a policy that can achieve different goals within the environment. 
%Universal Value Function Approximators (UVFA) is proposed to solve such 
In this thesis, the experiments in Chapter~\ref{ch:esil} and Chapter~\ref{ch:dtgsh} follow the terminology suggested by OpenAI~\cite{plappert2018multi}, in which the possible goals are drawn from a goal space $\mathcal{G}$, and the goal being pursued does not influence the environment dynamics. Two types of goal are recognised. One is the desired goal $g\in \mathcal{G}$, which is the target position {or state}, and may be different for different episodes. Within a single episode, $g$ is constant.  The second type of goal is the achieved goal $g^{ac}_{t}$, which is the achieved state in the environment (e.g., the position of an object), and this is considered to be different at each timestep in an episode. In an episode, each transition can be represented as $(s_t|\langle g, g^{ac}_{t} \rangle, a_t, r_t, s_{t+1}|\langle g, g^{ac}_{t+1} \rangle)$, where $s_t$ indicates a state, $a_t$ indicates an action and $r_t$ indicates a~reward; $\langle \, , \rangle$ is simply used to represent the grouping of goals.

\begin{figure}[h]
    \centering
    \includegraphics[width=0.85\textwidth]{figures/background/her_background.pdf}
    \caption{A brief illustration of the hindsight goal relabelling in a robotic manipulation task~\cite{plappert2018multi}. In this task, the agent needs to control the gripper to hit the puck and make it slide to a specific position, where the achieved goal is the position of the puck and the desired goal is the target position (the red spot).}
    \label{fig:her_back}
\end{figure}
In sparse reward settings, an agent will only get non-negative rewards when the desired goal $g$ is achieved. The sparse reward function can be defined as:
\begin{equation}
r_{t}\left(g^{ac}_{t+1}, g\right):=
\begin{cases}
0& \text{if $\left\|g^{ac}_{t+1} - g\right\|\leq\epsilon$}\\
-1& \text{otherwise},
\end{cases}
\label{eq:sparse_reward}
\end{equation}
where $\epsilon$ is a threshold value, used to identify if the agent has achieved the goal. {However, the desired goal, $g$, might be difficult to reach during training. Thus, hindsight experiences~\cite{andrychowicz2017hindsight} (see Figure~\ref{fig:her_back}) are created through replacing the original desired goal $g$ with the current achieved goal $g_{t+1}^{ac}$ to augment successful samples, and then reward $r_{t}$ can be recomputed according to Equation~\eqref{eq:sparse_reward}. The modification of the desired goal can be denoted as $g^\prime$ and transitions from hindsight experiences can be represented as $(s_t|\langle g^\prime, g^{ac}_{t} \rangle, a_t, r_t, s_{t+1}|\langle g^\prime, g^{ac}_{t+1} \rangle)$.} {Intuitively, introducing $g^\prime$ serves a useful purpose in the early stages of training; taking, for example, a robot reaching task, the agent has no prior ``concept" of how to move its effector to a specific location in space. Thus, even these original failed episodes contain valuable information for ultimately learning a useful control policy for the original, desired goal $g$.}

\section{Policy Optimisation in DRL}
\label{ch3:policy-optim}
Policy optimisation in DRL can be roughly classified into three categories: 1) value-based methods (see Section~\ref{ch3:value-based}), 2) policy-based methods (see Section~\ref{ch3:policy-based}), and 3) the actor-critic methods, which are the combination of value-based and policy-based methods (see Section~\ref{ch3:actor-critic}).

\subsection{Value-based Methods}
\label{ch3:value-based}
% value based
Value-based methods aim to learn a state or state--action value function instead of an explicit policy. The value function can evaluate the quality of a given state or the quality of an action performed at a given state. Then, the agent can select an action according to the corresponding estimate. This section will discuss how to derive a policy through the value function.

%and for a given state, the agent selects an action according to the prediction of the value function. When the values of states are predicted (i.e. state value function), the agent can select the action towards the next state with the highest state value. When the values of each action are estimated (i.e. state-value function), the agent selects the action with the highest value.

The state value function $V_{\pi}(s_{t})$ under a policy $\pi$ is the expected return when starting from the state $s_{t}$, and can be defined as:
\begin{equation}
\centering
    V_{\pi}(s_{t}) = \mathbb{E}_{\pi}[G_{t}] = \mathbb{E}_{\pi}\biggr[\sum_{k=0}^{T-1} \gamma^{k}r_{t+k}\biggr].
\end{equation}
Through the Bellman equation, the state value function can be rewritten as the combination of the immediate reward, $r_{t}$, and the discounted state value of the next state, $V_{\pi}(s_{t+1})$:
\begin{equation}
\centering
     V_{\pi}(s_{t}) = \mathbb{E}_{\pi}[r_{t} + \gamma V_{\pi}(s_{t+1})].
\end{equation}
% state-action value function
\td{Apart from estimating the state value,} we can also predict the value of a particular state--action pair, denoted as $Q_{\pi}(s_{t},a_{t})$, and this has a similar definition to the state value function:
\begin{equation}
\centering
     Q_{\pi}(s_{t},a_{t}) = \mathbb{E}_{\pi}[G_{t}] = \mathbb{E}_{\pi}\biggr[\sum_{k=0}^{T-1} \gamma^{k}r_{t+k}\biggr].
\end{equation}
The state--action value function $Q_{\pi}(s_{t},a_{t})$ is also called the Q-value function. Similarly, the Q-value function can also be expressed in a recursive form:
\begin{equation}
\centering
     Q_{\pi}(s_{t},a_{t}) = \mathbb{E}_{\pi}[r_{t} + \gamma Q_{\pi}(s_{t+1}, a_{t+1})].
\end{equation}
Next, the optimal policy can be derived using these value functions. In RL, the agent aims to follow the policy that can achieve more accumulated rewards from the environment. Thus, an optimal policy $\pi_{*}$ is defined as the policy that achieves the maximum/optimal state value function, $V_{*}(s_{t})$:
\begin{equation}
    V_{*}(s_{t}) = \max_{\pi}V_{\pi}(s_{t}).
\end{equation}
Similarly, an optimal policy is also defined as the policy that achieves the maximum/optimal Q-value function $Q_{*}(s_{t},a_{t})$:
\begin{equation}
    Q_{*}(s_{t},a_{t}) = \max_{\pi}Q_{\pi}(s_{t},a_{t}).
\end{equation}
Particularly, the optimal policy $\pi_{*}$ can be achieved through this optimal Q-value function directly:
\begin{equation}
    \pi_{*}(s_{t}) = \argmax_{a_{t}} Q_{\pi}(s_{t}, a_{t}).
\end{equation}
Therefore, finding the optimal Q-value function is equivalent to finding the optimal policy. The optimal Q-value function can be reorganised using the Bellman optimality equation:
% bellman optimality equation
\begin{equation}
\centering
     Q_{*}(s_{t},a_{t}) = \mathbb{E}[r_{t} + \gamma\max_{a_{t+1}} Q_{*}(s_{t+1}, a_{t+1})].
\end{equation}
Q-learning~\cite{watkins1992q} is a tabular and iterative update method to solve this equation. In Q-learning, the state--action pairs and corresponding Q-values are stored in a table. For a given state, the action is selected with an exploration strategy (e.g., $\epsilon$-greedy) to observe the next state and the reward, and the Q-value of the current state is updated via:
\begin{equation}
    Q_{\pi}(s_{t}, a_{t}) \leftarrow Q_{\pi}(s_{t}, a_{t}) + \alpha(r_{t} + \gamma\max_{a_{t+1}}Q_{*}(s_{t+1}, a_{t+1}) - Q_{\pi}(s_{t}, a_{t})),
\end{equation}
where $\alpha$ is the learning rate, and this process is repeated until the estimated Q-value converges to the optimal Q-value.

\subsubsection{Function Approximator and Deep Q-Network}
\td{Q-learning in its tabular form}, is impractical to scale to tasks with high-dimensional states, such as playing a video game using images as inputs. The Deep Q-Network (DQN)~\cite{mnih2015human} overcomes this difficulty by combining deep learning and Q-learning. In DQN, instead of updating the Q-value function in a table, a convolutional neural network (CNN) is leveraged to approximate the Q-value function from a stack of images directly:
\begin{equation}
    Q(s_{t},a_{t};\theta) \approx Q^{*}(s_{t},a_{t}),
\end{equation}
here, $\theta$ represents the parameters of a main network. During training, the agent collect samples from the environment with the $\epsilon$-greedy exploration strategy, and each transition ($s_{t}$, $a_{t}$, $r_{t}$, $s_{t+1}$) is stored in a replay buffer $\mathcal{D}$. The replay buffer is designed to reuse past experiences. When updating the parameters $\theta$, a minibatch of transitions is sampled from the replay buffer to minimise the mean squared error (MSE) between the target Q-value and the estimated Q-value using gradient descent methods (e.g., SGD):
\begin{equation}
    \min_{\theta}\mathcal{L} = \mathbb{E}_{(s_{t}, a_{t}, r_{t}, s_{t+1})\sim\mathcal{D}}[(Q(s_{t}, a_{t};\theta) - (r_{t} + \gamma\max_{a_{t+1}}Q(s_{t+1}, a_{t+1};\theta^{-})))^{2}],
\end{equation}
here, $\theta^{-}$ is the parameters of a target network, which is used to estimate the target Q-value. The target network has the same structure as the main network, and its parameters are copied from the main network at each fixed step. The target network is also used to stabilise the training.

\subsection{Policy-based Methods}
\label{ch3:policy-based}
% policy based
Policy-based methods learn a policy function $\pi(a_{t}|s_{t};{\theta})$, parameterised by $\theta$, and it directly maps a given state $s_{t}$ to an action $a_{t}$. In discrete control tasks, the output of the policy function is a categorical distribution over the action space. In continuous control tasks, the output of the policy function is often the parameters of a Gaussian distribution. During training, the actions are sampled from the corresponding distribution for exploration. In policy-based methods, the optimal parameters $\theta$ can be found by maximising expected returns in the environment:
\begin{align}
    \max_{\theta}J & = \mathbb{E}_{\pi_{\theta}}[G_{t}] \\
                   & = \sum_{s_{t}\in\mathcal{S}}d(s_{t})\sum_{a_{t}\in\mathcal{A}}\pi(a_{t}|s_{t};\theta)G_{t},
\end{align}
where $d(s_{t})$ is the stationary distribution in the Markov chain under the policy $\pi_{\theta}$. In order to maximise the expected returns, we need to increase the probability of the actions which lead to higher returns, and the gradient of the objective function can be expressed as:
\begin{align}
    \nabla_{\theta} J &= \sum_{s_{t}\in\mathcal{S}}d(s_{t})\sum_{a_{t}\in\mathcal{A}}\nabla_{\theta}\pi(a_{t}|s_{t};\theta)G_{t} \\
    &= \sum_{s_{t}\in\mathcal{S}}d(s_{t})\sum_{a_{t}\in\mathcal{A}}\pi(a_{t}|s_{t};\theta)\frac{\nabla_{\theta}\pi(a_{t}|s_{t};\theta)}{\pi(a_{t}|s_{t};\theta)}G_{t}\\
    &= \sum_{s_{t}\in\mathcal{S}}d(s_{t})\sum_{a_{t}\in\mathcal{A}}\pi(a_{t}|s_{t};\theta)\nabla_{\theta}\log\pi(a_{t}|s_{t};\theta)G_{t} \\
    &=\mathbb{E}_{\pi_{\theta}}[\nabla_{\theta}\log\pi(a_{t}|s_{t};\theta)G_{t}], \label{eq:reinforce}
\end{align}
where Equation~\eqref{eq:reinforce} is also known as REINFORCE~\cite{williams1992simple}, and the parameters are updated via gradient ascent. In practice, the return value $G_{t}$ is calculated using Monte Carlo methods, leads to a high variance in the estimated gradient, and slows the convergence of the policy function. In order to alleviate the above shortcomings, actor-critic methods are introduced.

\subsection{Actor-Critic Methods}
\label{ch3:actor-critic}
Actor-critic methods combine both policy-based methods and value-based methods. The actor is the policy function $\pi(a_{t}|s_{t};\theta)$ to generate an action for a given state. The critic is the value function $V(s_{t};\psi)$ or Q-value function $Q(s_{t},a_{t};\psi)$ to evaluate the quality of a state or a state--action pair and thus guides the training of the actor. To reduce the variance in policy-based methods, a common practice is to subtract a baseline from the estimated return $G_{t}$~\cite{sutton2018reinforcement}. In REINFORCE, the state value predicted by the critic $V(s_{t};\psi)$ is selected as the baseline:
\begin{equation}
    A(s_{t}, a_{t}) = G_{t} - V(s_{t};\psi),
\end{equation}
where $A(s_{t}, a_{t})$ is an advantage function, and the gradient of the policy function can be rewritten as:
\begin{equation}
    \nabla_{\theta} J = \mathbb{E}_{\pi_{\theta}}[\nabla_{\theta}\log\pi(s_{t},a_{t};\theta)A_{t}].
\end{equation}
The critic can also be updated through minimising the MSE error between the return value $R_{t}$ and the estimated state value $V(s_{t};\psi)$:
\begin{equation}
    \min_{\psi}\mathcal{L} = \mathbb{E}[(G_{t} - V(s_{t};\psi))^2].
    \label{eq:mse}
\end{equation}
This actor-critic method is called advantage actor-critic (A2C). For more efficient training, A2C also utilises multiple CPU workers to collect samples from separate environments to update the actor and the critic~\cite{mnih2016asynchronous}. However, A2C does not perform well in continuous control tasks. In this paper, two actor-critic DRL methods -- deep deterministic policy gradient (DDPG)~\cite{lillicrap2015continuous} and proximal policy gradient (PPO)~\cite{schulman2017proximal} -- are adopted as the base algorithms to address the continuous control tasks.

\subsubsection{Deep Deterministic Policy Gradient}
Unlike the previous policy gradient methods (e.g., A2C) which output a probability distribution over the action space, DDPG outputs an action from the actor directly:
\begin{equation}
    a_{t} = \pi(s_{t}; \theta),
\end{equation}
and the critic is a Q-value function $Q(s_{t}, a_{t};\psi)$ to guide the learning of the actor. During training, noise is added to the action $a_{t}$ to help exploration, and the collected transitions are stored in a replay buffer $\mathcal{D}$. For each update iteration, a minibatch sized transitions is sampled to update the actor and the critic. The actor aims to maximise the expected returns $\mathbb{E}[Q(s_{t}, a_{t};\psi)|_{a_{t}=\pi(s_{t};\theta)}]$ and the gradient of the objective function can be estimated using chain rule:

\begin{align}
    \nabla_{\theta}J &\approx\mathbb{E}_{s_{t}\sim\mathcal{D}}[\nabla_{\theta}Q(s_{t}, a_{t};\psi)|_{a_{t}=\pi(s_{t};\theta)}]\\
    & = \mathbb{E}_{s_{t}\sim\mathcal{D}}[\nabla_{a_{t}}Q(s_{t}, a_{t};\psi)|_{a_{t}=\pi(s_{t};\theta)} \nabla_{\theta}\pi(s_{t};\theta)].
\end{align}
The update of the critic is similar to DQN which minimises the MSE between the target Q-vaule and the estimated Q-value:
\begin{equation}
    \min_{\psi}\mathcal{L} = \mathbb{E}_{(s_{t}, a_{t}, r_{t}, s_{t+1})\sim\mathcal{D}}[(Q(s_{t}, a_{t};\psi) - (r_{t} + \gamma Q(s_{t+1}, \pi(s_{t+1};\theta^{-});\psi^{-})))^{2}],
\end{equation}
where $\theta^{-}$ and $\psi^{-}$ are the parameters of the target actor network and the target critic network, respectively. These two target networks are updated toward the parameters of the two main networks slowly using a moving average scheme:
\begin{align}
    \theta^{-} &= \tau\theta^{-} + (1-\tau)\theta\\
    \psi^{-} &= \tau\psi^{-} + (1-\tau)\psi
\end{align}
here, $\tau\in$(0,1) is a coefficient that decides the updating speed of the target networks.

\subsubsection{Proximal Policy Optimisation}
Before introducing proximal policy optimisation (PPO), we first review its forerunner -- trust region policy optimisation (TRPO)~\cite{schulman2015trust}. When updating the policy network using gradient-based methods, an inappropriate step size (i.e. learning rate) may lead to a worse policy, and the samples collected by this updated policy may interfere with the further policy improvements. In TRPO, when maximising the expected return of a policy network, an extra Kullback-Leibler (KL) divergence term is used to guarantee a promising step size and to restrict the policy update in a specific (trust) region. This constrained objective function can be expressed as:
\begin{align}
   & \max_{\theta} J = \mathbb{E}_{\pi_{\theta_{old}}}\biggr[\frac{\pi(a_{t}|s_{t};\theta)}{\pi(a_{t}|s_{t};\theta_{old})}A(s_{t},a_{t})\biggr] \\
   & \text{subject to} \enspace \mathbb{E}_{\pi_{\theta_{old}}}[\text{KL}[\pi(\cdot|s_{t};\theta_{old}), \pi(\cdot|s_{t};\theta)]] \leq \delta,
\end{align}
where $\theta$ and $\theta_{old}$ are the parameters of the current policy network and the old policy network, respectively, and $\delta$ is a value to limit changes between policies. This objective function can be solved via the method of Lagrange multipliers. In addition, a line search method can be used to estimate an appropriate step size, with the conjugate gradient method being used to estimate the gradient. Although TRPO achieves great performance in continuous control tasks, it is not flexible to be implemented and it performs poorly on tasks with high-dimensional inputs, such as playing Atari games.

Proximal policy optimisation (PPO) provides a simplified solution of TRPO with a clipped objective function to constrain changes of policy during an update:
\begin{equation}
\centering
\max_{\theta}J = \mathbb{E}_{\theta}\biggr[\min\bigg (\frac{\pi(a_{t}|s_{t};\theta)}{\pi(a_{t}|s_{t};\theta_{old})},
	\clip \bigg(\frac{\pi(a_{t}|s_{t};\theta)}{\pi(a_{t}|s_{t};\theta_{old})}, 1-\epsilon, 1+\epsilon \bigg)\bigg )A(s_t, a_t)\biggr],
\label{eq:standard_ppo}
\end{equation}
where $\epsilon$ represents the clipping ratio. In practice, PPO makes the implementation more accessible and improves the performance in tasks with high-dimensional inputs.

\section{Determinantal Point Processes}
\label{ch3:dpp}
A determinantal point process (DPP)~\cite{kulesza2012determinantal} is a stochastic process that characterises a probability distribution over sets of points using the determinant of some function. In machine learning, it is often used to quantify the diversity of a subset, with applications such as video~\cite{mahasseni2017unsupervised} and document summarisation~\cite{hong2014improving}.

Formally, for a discrete set of points $\mathcal{Y}=\{\mathbf{x}_{1}, \mathbf{x}_{2}, \cdots, \mathbf{x}_{N}\}$, a point process $\mathcal{P}$ is a probability measure over all $2^{|\mathcal{Y}|}$ subsets. $\mathcal{P}$ is a DPP if a random subset $\mathbf{Y}$ is sampled with probability:
\begin{equation}
    \mathcal{P}_{L}(\mathbf{Y}=Y) = \frac{\text{det}(\mathbf{L}_{Y})}{\sum_{Y'\subseteq \mathcal{Y}} \text{det}(\mathbf{L}_{Y'})} = \frac{\text{det}(\mathbf{L}_{Y})}{\text{det}(\mathbf{L}+\mathbf{I})},
\end{equation}
where $Y\subseteq \mathcal{Y}$, $\mathbf{I}$ is the identity matrix, $\mathbf{L} \in \mathbb{R}^{N\times N}$ is the positive semi-definite DPP kernel matrix, and $\mathbf{L}_{Y}$ is the sub-matrix with rows and columns indexed by the elements of the subset $Y$.

The kernel matrix $\mathbf{L}$ can be represented as the Gram matrix $\mathbf{L} = \mathbf{X}^{T}\mathbf{X}$, where each column of $\mathbf{X}$ is the feature vector of an item in $\mathcal{Y}$. The determinant, $\text{det}(\mathbf{L}_{Y})$, represents the (squared) volume spanned by vectors $\mathbf{x}_{i}\in Y$. From a geometric perspective, feature vectors that are closer to being orthogonal to each other will have a larger determinant, and vectors in the spanned subspace are more likely to be sampled: $\mathcal{P}_{L}(\mathbf{Y}=Y) \propto \text{det}(\mathbf{L}_{Y})$. Using orthogonality as a measure of diversity, we leverage DPPs to sample diverse trajectories and goals in Chapter~\ref{ch:dtgsh}. In Chapter~\ref{ch:daim}, we use DPPs to model an intrinsic motivation, which aims to encourage the agent to explore novel states in the environment.

### 2022-04-23 chapters/chapter3.tex
switch conclusion to discussion
update

@@
 \chapter{Episodic Self-Imitation Learning with Hindsight}
+\label{ch:esil}
 \section{Introduction}
-Reinforcement learning (RL) has has been shown to be very effective in training agents within gaming environments~\cite{mnih2015human,silver2016mastering}, particularly when combined with deep neural networks~\cite{lecun2015deep,silver2016mastering,liu2017survey}.
+% \td{In this chapter, we propose episodic self-imitation learning (ESIL), which aims to address }
+%------------------------------ From Main Paper ----------------------------
+Reinforcement learning (RL) has been shown to be very effective in training agents within gaming environments~\cite{mnih2015human,silver2016mastering}, particularly when combined with deep neural networks~\cite{lecun2015deep,silver2016mastering,liu2017survey}.
 In most tasks settings that are solved by RL algorithms, reward shaping is an essential requirement for guiding the learning of the agent.
-Reward shaping, however, often requires significant quantities of domain knowledge that {are} highly task-specific~\cite{ng1999policy} and, even with careful design, can lead to undesired policies.
-Moreover, for complex robotic manipulation tasks, manually designing  reward shaping functions to guide the learning agent becomes intractable~\cite{arulkumaran2017deep,florensa2017reverse} if even minor variations to the task are introduced.
+However, reward shaping often requires significant quantities of domain knowledge that {are} highly task-specific~\cite{ng1999policy} and, even with careful design, can lead to undesired policies.
+Moreover, for complex robot manipulation tasks, the manual design of reward shaping functions to guide the learning agent becomes intractable~\cite{arulkumaran2017deep,florensa2017reverse} if even minor variations are introduced in the task.
 For such settings, the application of deep reinforcement learning requires algorithms that can learn from unshaped, and usually sparse, reward signals.
-The complicated dynamics of robot manipulation exacerbate the difficulty posed by sparse rewards, especially for on-policy RL algorithms.
-For example, achieving goals that require successfully executing multiple steps over a long horizon involves high dimensional control {that must also generalise to work across variations in the environment for each step}.
-These aspects of robot control result in a situation where a naive RL agent so rarely receives a reward at the start of training that it is not able to learn at all.
-A common solution in the robotics community is to collect a sufficient quantity of expert demonstrations, then use imitation learning to train the agent.
-However, in some scenarios, demonstrations are expensive to collect and the achievable performance of a trained agent is restricted by their quantity.
-One solution is to use the valuable past experiences of the agent to enhance training, and this is particularly useful in sparse reward environments.
+The complicated dynamics of robot manipulation exacerbates the difficulty posed by sparse rewards, especially for on-policy RL algorithms.
+For example, achieving goals that require the successful execution of multiple steps over a long horizon involves high-dimensional control that must also be generalised to work across variations in the environment for each step.
+These aspects of robot control result in a situation where a naive RL agent so rarely receives a reward at the start of training that it cannot learn at all.
+A common solution in the \td{DRL} community is collecting a sufficient number of expert demonstrations, then use imitation learning to train the agent.
+However, in some scenarios, demonstrations are expensive to collect and the performance of a trained agent is restricted by their quantity.
+One solution is to use the valuable past experiences of the agent to improve training, which is particularly useful in sparse reward environments.
 
-To alleviate the problems associated with having sparse rewards, there are two kinds of approaches: imitation learning and hindsight experience replay (HER).
+There are two kinds of approaches to alleviate the problems associated with sparse rewards: imitation learning and hindsight experience replay (HER)~\cite{andrychowicz2017hindsight}.
 First, the standard approach of imitation learning is to use supervised learning algorithms and minimise a surrogate loss with respect to an oracle.
@@
 Similar techniques are applied to robot manipulation tasks~\cite{rajeswaran2017learning,vevcerik2017leveraging,nair2018overcoming,james2018task}.
-When the demonstrations are not attainable, self-imitation learning (SIL)~\cite{oh2018self}, which uses past good experiences (episodes in which the goal is achieved), can~be used to enhance exploration or speed up the training of the agent.
-{Self-imitation learning} works well in discrete control environments, such as Atari Games.
-Whilst being able to learn policies for continuous control tasks with dense or delayed rewards~\cite{oh2018self}, {the present} experiments suggest that SIL struggles when rewards are sparse.
-Recently, {hindsight experience replay} has been proposed to solve such goal-conditional, sparse reward problems.
-The main idea of HER~\cite{andrychowicz2017hindsight} is that during replay, the selected transitions are sampled from state--action pairs derived from {achieved goals} that are substituted for the real goals of the task; this increases the frequency of positive rewards.
+When the demonstrations are not attainable, self-imitation learning (SIL)~\cite{oh2018self}, which uses past good experiences (episodes in which the goal is achieved), can~be used to enhance the exploration or speed up the training of the agent.
+{Self-imitation learning} works well in discrete control environments, such as Atari games.
+While being able to learn policies for continuous control tasks with dense or delayed rewards~\cite{oh2018self}, the present experiments suggest that SIL struggles when the rewards are sparse.
+Recently, {hindsight experience replay} has been proposed to solve such goal-oriented, sparse reward problems.
+The main idea of HER is that during replay, the selected transitions are sampled from state--action pairs derived from {achieved goals} that are substituted for the real goals of the task; this increases the frequency of positive rewards.
 {Hindsight experience replay} is used with off-policy RL algorithms, such as DQN~\cite{mnih2015human} and DDPG~\cite{lillicrap2015continuous}, for experience replay and has several extensions~\cite{schaul2015prioritized,liu2018competitive}.
@@
 
-In this paper, {episodic self-imitation learning (ESIL) for goal-oriented problems that provide only sparse rewards is proposed and combined with a state-of-the-art on-policy RL algorithm}: proximal policy optimization (PPO).
-In contrast to standard SIL, which samples past good transitions from the replay buffer for imitation learning, {the proposed} ESIL adopts entire current episodes {(successful or not)}, and modifies them into ``expert'' trajectories based on HER.
+In this chapter, {episodic self-imitation learning (ESIL) for goal-oriented problems that provide only sparse rewards is proposed and combined with a state-of-the-art on-policy RL algorithm:} proximal policy optimisation (PPO).
+In contrast to standard SIL, which samples past good transitions from the replay buffer for imitation learning, {the proposed} ESIL adopts the entire set of current episodes {(successful or not)}, and modifies them into ``expert'' trajectories based on HER.
 An extra trajectory selection module is also introduced to relieve the effects of sample correlation~\cite{lee2019sample} in updating the network.
@@
 During training by SIL+HER, a batch of transitions is sampled from the replay buffer; these are modified into ``hindsight experiences'' and used directly in self-imitation learning.
-In contrast, ESIL utilises entire current collected episodes and converts them into hindsight episodes.
+In contrast, ESIL utilises the entire currently collected episodes and converts them into hindsight episodes.
 The trajectory selection module removes undesired transitions in the hindsight episodes.
-Using tasks from the Open AI Fetch environment, this paper demonstrates that the proposed ESIL approach is effective in training agents which are required to solve continuous control problems, and shows that it achieves state-of-the-art results on several tasks.
+Using tasks from the OpenAI Fetch environment~\cite{plappert2018multi}, this chapter demonstrates that the proposed ESIL approach effectively trains the agent which is required to solve continuous control problems, and shows that it achieves state-of-the-art results on several tasks.
 
@@
 \includegraphics[width=0.95\textwidth]{figures/chapter3/buffer.png}
-\caption{Illustration of difference between self-imitation learning (SIL)+hindsight experience replay (HER) and episodic self-imitation learning (ESIL).}
+\caption{Illustration of the difference between self-imitation learning (SIL)+hindsight experience replay (HER) and episodic self-imitation learning (ESIL).}
 %% PPO+SIL+HER samples a batch of transitions from the replay buffer and modify them into hindsight experiences and conduct self-imitation learning directly.
@@
 
-The primary contribution of this paper is a novel episodic self-imitation learning (ESIL) algorithm that can solve continuous control problems {in environments providing only sparse rewards}; in doing so, it also empirically answers an open question posed by~\cite{plappert2018multi}.
-{The proposed ESIL approach also provides} a more efficient way to perform exploration in goal-conditional settings than the standard self-imitation learning algorithm.
-Finally, this approach achieves, to our knowledge, the best results for four moderately complex robot control tasks in simulation.
+The primary contribution of this chapter is a novel episodic self-imitation learning (ESIL) algorithm that can solve continuous control problems {in environments providing only sparse rewards}; in doing so, it also empirically answers an open question posed by Plappert \textit{et al.}~\cite{plappert2018multi} -- ``\textit{How to combine HER with state-of-the-art on-policy RL algorithms like
+PPO?}". {The proposed ESIL approach also provides} a more efficient way to perform exploration in goal-oriented settings than the standard self-imitation learning algorithm. 
+% Finally, to our knowledge, this approach achieves the best results for four moderately complex robot control tasks in simulation. 
+Finally, this approach outperforms other on-policy RL algorithms in all simulated robot manipulation tasks.
 
@@
 The inverse reinforcement learning estimates a reward model from expert trajectories~\cite{ng2000algorithms,abbeel2004apprenticeship,ziebart2008maximum}.
-Learning from demonstrations is powerful for complex robotic manipulation tasks~\cite{finn2016guided,zhang2018deep,pmlr-v78-finn17a,rajeswaran2017learning,fang2019survey}.
-~\cite{ho2016generative} propose generative adversarial imitation learning (GAIL), which employs generative adversarial training to match the distribution of state--action pairs of demonstrations.
-Compared with behavioural cloning, the GAIL framework shows strong improvements in continuous control tasks.
-In the work of~\cite{ding2019goal}, goalGAIL is proposed to speed up the training in goal-conditional environments; goalGAIL was also shown to be able to learn from demonstrations without action information.
-Prior work has used demonstrations to accelerate  learning~\cite{rajeswaran2017learning,vevcerik2017leveraging,nair2018overcoming}.
-Demonstrations are often collected by an expert policy or human actions.
+Learning from demonstrations is powerful for complex robot manipulation tasks~\cite{finn2016guided,zhang2018deep,pmlr-v78-finn17a,rajeswaran2017learning,fang2019survey}.
+Prior work has used demonstrations to accelerate  learning~\cite{rajeswaran2017learning,vevcerik2017leveraging,nair2018overcoming},  nevertheless demonstrations are often collected by an expert policy or human actions.
 In contrast to these approaches, {episodic self-imitation learning (ESIL) does not need demonstrations.}
@@
 % self-learning and self-imitation learning
-Self-imitation learning (SIL)~\cite{oh2018self} is used for exploiting past experiences for parametric policies.
+Self-imitation learning (SIL)~\cite{oh2018self} can be used for exploiting past experiences for parametric policies.
 It~has a similar flavor to ~\cite{gangwani2018learning, pmlr-v97-wu19a}, in that the agent learns from imperfect demonstrations.
@@
 Generalised SIL can also be combined with both deterministic and stochastic RL algorithms.
-~\cite{guo2019self} points out that using imitation learning with past good experience could lead to a sub-optimal policy.
+Guo \textit{et al.}~\cite{guo2019self} points out that using imitation learning with past successful experiences could lead to a sub-optimal policy.
 Instead of imitating past good trajectories, a trajectory-conditioned policy~\cite{guo2019self} is proposed to imitate trajectories in diverse directions, encouraging exploration in environments where exploration is otherwise difficult.} Unlike SIL, episodic self imitation learning (ESIL) applies HER to the current episodes to create ``imperfect'' demonstrations for imitation learning; this also requires introducing a trajectory-selection module to reject undesired samples from the hindsight experiences.
-In the work of~\cite{lee2019sample}, it was shown that the agent benefits from using whole episodes in updates, rather~than uniformly sampling the sparse or delayed reward environments.
-{The present} experiments suggest that episodic self-imitation learning {achieves better performance in an agent that must learn to perform continuous control in environments delivering sparse rewards}.
+In the work of~\cite{lee2019sample}, it was shown that the agent benefits from using whole episodes in updates, rather~than uniformly sampling in the sparse or delayed reward environments.
+The present experiments suggest that episodic self-imitation learning {achieves better performance in an agent that must learn to perform continuous control in environments delivering sparse rewards}.
 
 % Hindsight
-Recently, the technique known as hindsight learning was developed.
+Recently, the technique known as hindsight learning has been developed.
 Hindsight experience replay (HER)~\cite{andrychowicz2017hindsight} is an algorithm that can overcome the exploration problems in multi-goal environments, delivering sparse rewards.
-Hindsight policy gradient (HPG)~\cite{rauber2018hindsight} introduces techniques that enable the learning of goal-conditional policies using hindsight experiences.
+Hindsight policy gradient (HPG)~\cite{rauber2018hindsight} introduces techniques that enable the learning of goal-oriented policies using hindsight experiences with on-policy RL algorithms.
 However, the current implementation of HPG has only been evaluated for agents that need to perform discrete actions, and one drawback of hindsight policy gradient estimators is the computational cost because of the goal-oriented sampling.
-An extension of HER, called dynamic hindsight experience replay (DHER)~\cite{fang2018dher}, was proposed to deal with dynamic goals.
-~\cite{liu2019hindsight} uses the GAIL framework~\cite{ho2016generative} to generate trajectories that are similar to hindsight experiences; it then applies imitation learning, using these trajectories.
-Competitive Experience Replay (CER) complements HER by introducing a competition between two agents for exploration~\cite{liu2018competitive}.
-~\cite{zhao2018energy} point out that the hindsight trajectories which contain higher energy are more valuable during training, leading to a more efficient learning system.
-~\cite{fang2019curriculum} proposed curriculum-guided HER, which~ incorporates curriculum learning in the work.
-During training, the agent focuses on the closest goals in the initial stage, then focuses on the expanding the diversity of goals.
-This approach accelerates training compared with other baseline methods.
-Unlike these works, episodic self-imitation learning (ESIL) combines episodic hindsight experiences with imitation learning, which aids learning at the start of training.
+Unlike HPG, episodic self-imitation learning (ESIL) combines episodic hindsight experiences with imitation learning, which aids learning at the beginning of training.
 Furthermore, ESIL can be applied to continuous control, making it more suitable for control problems that demand greater precision.
@@
 The proposed method combines PPO and episodic self-imitation learning to maximally use hindsight experiences for exploration to improve learning.
-{Recent advantages in episodic backward update~\cite{lee2019sample} and hindsight experiences~\cite{andrychowicz2017hindsight} are also leveraged} to guide exploration for on-policy RL.
+{Recent advantages in episodic update~\cite{lee2019sample} and hindsight experiences~\cite{andrychowicz2017hindsight} are also leveraged} to guide exploration for on-policy RL.
 
@@
 To this end, hindsight experiences are created from current episodes.
-For an episode $i$, let there be $T$ time steps; after $T$, a series of transitions $\tau_i =\{ (s_{t}|\langle g, g^{ac}_{t} \rangle, a_{t}, r_{t}, s_{t+1}|\langle g, g^{ac}_{t+1}\rangle )\}_{t=0:T-1}$ is collected.
-If at time step $t=T-1$, in $s_{T}$, $g^{ac}_{T} \neq g$, it implies that in this episode, the agent failed to achieve the original goal.
+For an episode $i$, let there be $T$ timesteps; after $T$ timesteps, a series of transitions $\tau_i =\{ (s_{t}, a_{t}, r_{t}, s_{t+1}, g, g^{ac}_{t+1})\}_{t=0:T-1}$ is collected.
+If at timestep $t=T-1$, in $s_{T}$, $g^{ac}_{T} \neq g$, it implies that in this episode, the agent failed to achieve the original goal.
 Simply, to create hindsight experiences, {the achieved goal $g^{ac}_{T}$ in the last state $s_{T}$ is selected and considered as the modified desired goal $g^{\prime}$, i.e., $g^\prime = g^{ac}_{T}$.} Next, a new reward $r_{t}^{\prime}$ is computed under the new goal $g^\prime$.
-Then, {a new ``imagined'' episode is achieved}, and a new series of transitions $\tau_i^{\prime} = \{(s_{t}|\langle g^\prime, g^{ac}_{t} \rangle, a_{t}, r_{t}^{\prime},s_{t+1}| \langle g^\prime, g^{ac}_{t+1}\rangle )\}_{t=0:T-1}$ is collected.
+Then, {a new ``imagined'' episode is achieved}, and a new series of transitions $\tau_i^{\prime} = \{(s_{t}, a_{t}, r_{t}^{\prime},s_{t+1}, g^\prime, g^{ac}_{t+1})\}_{t=0:T-1}$ is collected.
 Then, an approach to self-imitation learning based on episodic hindsight experiences is proposed, which applies the policy updates to both hindsight and in-environment episodes.
-Proximal policy optimization (PPO) is used as the base RL algorithm, which is a state-of-the-art on-policy RL algorithm.
-With current and corresponding hindsight experiences, a new objective function is introduced and defined as
+Proximal policy optimisation (PPO) is used as the base RL algorithm, which is a state-of-the-art on-policy RL algorithm.
+With current and corresponding hindsight experiences, a new objective function is introduced and defined as:
 \begin{equation}
   \centering
-  \mathcal{L} = \alpha\cdot\mathcal{L}_{PPO} + \beta\cdot\mathcal{L}_{ESIL},
+  \mathcal{L} = \alpha\mathcal{L}_{PPO} + \beta\mathcal{L}_{ESIL},
   \label{eq:loss}
@@
 In the experiments, we set $\alpha=1$ as default {to balance the contribution of $\mathcal{L}_{PPO}$ and $\mathcal{L}_{ESIL}$}.
-$\mathcal{L}_{PPO}$ is the loss of PPO which can be written as
+$\mathcal{L}_{PPO}$ is the loss of PPO which can be written as:
 \begin{equation}
-  \mathcal{L}_{PPO} = \mathcal{L}_{policy} - c\cdot \mathcal{L}_{value}
+  \mathcal{L}_{PPO} = \mathcal{L}_{policy} - c\mathcal{L}_{value}
   \label{eq:ppo_loss}
 \end{equation}
-where $\mathcal{L}_{policy}$ is the policy loss which is parameterised by $\theta$, $\mathcal{L}_{value}$ is the value loss which is parameterised by $\eta$, and $c$ is the weight coefficient of the $\mathcal{L}_{value}$, which is set to 1 to match the default PPO setting~\cite{schulman2017proximal}.
-The policy loss, $\mathcal{L}_{policy}$, can be represented as
+where $\mathcal{L}_{policy}$ is the policy loss which is parameterised by $\theta$, $\mathcal{L}_{value}$ is the value loss which is parameterised by $\psi$, and $c$ is the weight coefficient of the $\mathcal{L}_{value}$, which is set to 1 to match the default PPO setting~\cite{schulman2017proximal}.
+The policy loss, $\mathcal{L}_{policy}$, can be represented as:
 \begin{equation}
   \centering
-  \mathcal{L}_{policy} = \mathbb{E}_{s_{t}, a_{t}, g\in\Tau}\biggr[\min \left (\frac{\pi_{\theta}(a_{t}|s_{t}, g)}{\pi_{\theta_{old}}(a_{t}|s_{t}, g)}A_{t}, \clip \left (\frac{\pi_{\theta}(a_{t}|s_{t}, g)}{\pi_{\theta_{old}}(a_{t}|s_{t}, g)}, 1-\epsilon, 1+\epsilon \right )A_{t}\right )\biggr],
+  \mathcal{L}_{policy} = \mathbb{E}_{s_{t}, a_{t}, g\in\Tau}\biggr[\min \left (\frac{\pi(a_{t}|s_{t}, g;\theta)}{\pi(a_{t}|s_{t}, g;\theta_{old})}A_{t}, \clip \left (\frac{\pi(a_{t}|s_{t}, g;\theta)}{\pi(a_{t}|s_{t}, g;\theta_{old})}, 1-\epsilon, 1+\epsilon \right )A_{t}\right )\biggr],
   \label{eq:policy_loss}
 \end{equation}
-here, $A_{t}$ is the advantage value, and can be computed as $R_{t} - V_{\eta}(s_{t}, g)$.
-$V_{\eta}(s_{t}, g)$ is the state value at time step $t$ which is predicted by the critic network.
-$R_{t}$ is the return at time step $t$.
+here, $A_{t}$ is the advantage value, and can be computed as $G_{t} - V(s_{t}, g;\psi)$.
+$V(s_{t}, g;\psi)$ is the state value at timestep $t$ which is predicted by the critic network.
+$G_{t}$ is the return at timestep $t$.
 $\epsilon$ is the clip ratio.
 $\Tau$~indicates original trajectories.
-{The value loss is an squared error loss $\mathcal{L}_{value}=(V_{\eta}(s_{t}, g) - R_{t})^{2}$.}
+{The value loss is a squared error loss $\mathcal{L}_{value}=\mathbb{E}[(V(s_{t}, g;\psi) - G_{t})^{2}]$.}
 
-For the $\mathcal{L}_{ESIL}$ term, $\beta$ is an adaptive weight coefficient of $L_{ESIL}$; it can be defined as the ratio of samples which are selected for self-imitation learning
+For the $\mathcal{L}_{ESIL}$ term, $\beta$ is an adaptive weight coefficient of $\mathcal{L}_{ESIL}$; it can be defined as the ratio of samples which are selected for self-imitation learning:
 \begin{equation}
@@
 where $N_{ESIL}$ is the number of samples used for self-imitation learning and $N_{Total}$ is the total number of collected samples.
-The episodic self-imitation learning loss $\mathcal{L}_{ESIL}$ can be written as 
+The episodic self-imitation learning loss $\mathcal{L}_{ESIL}$ can be written as: 
 \begin{equation}
   \centering
-  \mathcal{L}_{ESIL}=\mathbb{E}_{s_{t}, a_{t}, g^\prime\in\Tau^{'}}\left[\log\pi_{\theta}\left(a_{t}|s_{t}, g^\prime\right)\cdot\mathcal{F}_t\right],
+  \mathcal{L}_{ESIL}=\mathbb{E}_{s_{t}, a_{t}, g^\prime\in\Tau^{'}}\left[\log\pi\left(a_{t}|s_{t}, g^\prime;\theta\right)\mathcal{F}_t\right],
   \label{eq:imitaion}
 \end{equation}
-where $\Tau^{'}$ indicates hindsight trajectories and $\mathcal{F}_t$ is the trajectory selection module which is based on returns of the current episodes, $R$, and the returns of corresponding hindsight experiences, $R^{\prime}$.
+where $\Tau^{'}$ denotes hindsight trajectories and $\mathcal{F}_t$ is the trajectory selection module which is based on returns of the current episodes, $G$, and the returns of corresponding hindsight experiences, $G^{\prime}$.
 
 \subsection{Episodic Update with Hindsight}
-Two important issues of ESIL are: (1) hindsight experiences are sub-optimal, and (2) the {detrimental} effect of updating networks with correlated trajectories.
+Two crucial issues of ESIL are: 1) hindsight experiences are sub-optimal, and 2) the {detrimental} effect of updating networks with correlated trajectories. 
 Although episodic self-imitation learning makes exploration more effective, hindsight experiences are not from experts and not ``perfect'' demonstrations.
@@
 
-To prevent the agent learning from imperfect hindsight experiences, {hindsight experiences are actively selected based on returns.} With the same action, different goals may lead to different results.
+To prevent the agent from learning imperfect hindsight experiences, {hindsight experiences are actively selected based on returns.} With the same action, different goals may lead to different results.
 The proposed method only selects hindsight experiences that can achieve higher returns.
@@
 For an episodic experience and its hindsight experience, the returns of the episodic experience and its hindsight experience can be calculated, respectively.
-In a trajectory, at time step $t$, the return $R_{t}$ can be calculated by $R_{t} = R(s_t, g) = r_{t} + \gamma \cdot r_{t+1} + \gamma^{2} \cdot r_{t+2} + ...
-+ \gamma^{T-t-1} \cdot r_{T-1}$.
-Then, for a trajectory $\tau_{i}$, we have $\left\{R_{0}^{i}, R_{1}^{i}, R_{2}^{i}, ..., R_{T-1}^{i}\right\}$.
+In a trajectory, at each timestep $t$, the return $G_{t}$ can be calculated by $G_{t} = G(s_t, g) = r_{t} + \gamma r_{t+1} + \gamma^{2} r_{t+2} + ...
++ \gamma^{T-t-1} r_{T-1}$.
+Then, for a trajectory $\tau_{i}$, we have $\left\{G_{0}^{i}, G_{1}^{i}, G_{2}^{i}, ..., G_{T-1}^{i}\right\}$.
 
-For the hindsight experiences, similarly, the return $R_{t}'$ for each time step, $t$,~with respect to the hindsight goals $g^\prime$, can be calculated.
-Based on the modified trajectory $\tau_{i}^{\prime}$ with the same length of $\tau_{i}$, we therefore have the returns $\left\{R_{0}^{i^\prime}, R_{1}^{i^\prime}, R_{2}^{i^\prime}, ..., R_{T-1}^{i^\prime}\right\}$.
+For the hindsight experiences, similarly, the return $G_{t}'$ for each timestep $t$,~with respect to the hindsight goal $g^\prime$, can be calculated.
+Based on the modified trajectory $\tau_{i}^{\prime}$ with the same length of $\tau_{i}$, we therefore have the returns $\left\{G_{0}^{i^\prime}, G_{1}^{i^\prime}, G_{2}^{i^\prime}, ..., G_{T-1}^{i^\prime}\right\}$.
 During training, the~hindsight experiences with higher returns are used for self-imitation learning.
-The rest of the hindsight experiences will be supposed to be worthless samples and ignored.
-Then, Equation (\ref{eq:imitaion}) can be rewritten as
+The rest of the hindsight experiences will be \td{assumed to be of lower importance} and ignored.
+Then, Equation~\eqref{eq:imitaion} can be rewritten as:
 \begin{equation}
   \centering
-  \mathcal{L}_{ESIL}=\mathbb{E}_{s_{t}, a_{t}, g^\prime\in\Tau^{'},g\in\Tau}\biggr[\log\pi_{\theta}\left(a_{t}|s_{t}, g^\prime\right)\cdot\mathcal{F}\left(s_{t}, g, g^\prime\right)\biggr],
+  \mathcal{L}_{ESIL}=\mathbb{E}_{s_{t}, a_{t}, g^\prime\in\Tau^{'},g\in\Tau}\biggr[\log\pi_{\theta}\left(a_{t}|s_{t}, g^\prime\right)\mathcal{F}\left(s_{t}, g, g^\prime\right)\biggr],
   \label{eq:final_loss}
@@
   \centering
-  \mathcal{F}\left(s_{t}, g, g^\prime \right)=\mathds{1}\left[R(s_{t}, g^\prime) > R(s_{t}, g)\right],
+  \mathcal{F}\left(s_{t}, g, g^\prime \right)=\mathds{1}\left[G(s_{t}, g^\prime) > G(s_{t}, g)\right],
 \end{equation}
 here, {$\mathds{1}(\cdot)$ is the unit step function}.
-Consider the OpenAI FetchReach environment as an example.
+Consider the OpenAI Fetch environment as an example.
 For~a failed trajectory, the rewards $r_{t}$ are $\{-1,-1,\cdots, -1\}$.
-{The desired goal is modified} to construct a new hindsight trajectory and the new rewards $r_{t}^{\prime}$ become $\{-1,-1,\cdots, 0\}$.
-Then,  {$R$ and $R^{\prime}$ can be calculated~separately.}
+{The desired goal is modified} to construct a new hindsight trajectory, and the new rewards $r_{t}^{\prime}$ become $\{-1,-1,\cdots, 0\}$.
+Then,  {$G$ and $G^{\prime}$ can be calculated~separately.}
 \begin{figure}[t!]
@@
 
-From a goal perspective, episodic self-imitation learning (ESIL) tries to explore (desired) goals to get positive returns.~It can be viewed as a form of multi-task learning, {because ESIL has two objective functions to be optimised jointly.}~It is also related to self-imitation learning (SIL)~\cite{oh2018self}.~However,~ the~difference is that SIL uses $\left(R-V_{\theta}(s)\right)_{+}$ on past experiences to learn to choose the action chosen in the past in a given state, rather than goals.
+From a goal perspective, episodic self-imitation learning (ESIL) tries to explore (desired) goals to get positive returns.~It can be viewed as a form of multi-task learning, {because ESIL has two objective functions to be optimised jointly.}~It is also related to self-imitation learning (SIL)~\cite{oh2018self}.~However,~ the~difference is that SIL uses $\left(G-V(s;\theta)\right)_{+}$ on past experiences to learn to choose the action chosen in the past in a given state, rather than goals.
 The full description of ESIL can be found in Algorithm~\ref{alg}.
@@
 \begin{algorithm}[t!]
-  \caption{Proximal policy optimization (PPO) with Episodic Self-Imitation Learning (ESIL)}
+  \caption{Proximal Policy Optimisation (PPO) with Episodic Self-Imitation Learning (ESIL)}
   \label{alg}
   \begin{algorithmic}[1]
-    \REQUIRE an actor network $\pi(s, g|\theta)$, a critic network $V(s, g|\eta)$, the maximum steps $T$ of an episode, a reward function $r$
-    \FOR{$\mathrm{iteration} = 1, 2, \cdots $}
-    \STATE $\Tau = \varnothing$, $\Tau^{\prime} = \varnothing$ 
-    \FOR{$\mathrm{episode} = 1, 2, ..., N$}
-    \STATE $\tau = \varnothing$
+    \REQUIRE an actor network $\pi(s, g;\theta)$, a critic network $V(s, g;\psi)$, the maximum steps $T$ of an episode, a reward function $\mathcal{R}$, number of iterations $N$, number of episodes for each iteration $M$, weight coefficient of critic loss term $c$, weight coefficient of PPO term $\alpha$, adaptive coefficient $\beta$
+    %\STATE Initialize the parameters $\theta$ and $\psi$ of both actor and critic networks
+    \FOR{$\mathrm{iteration} = 1, 2, \cdots N$}
+    \STATE $\Tau \leftarrow \varnothing$, $\Tau^{\prime} \leftarrow \varnothing$ 
+    \FOR{$\mathrm{episode} = 1, 2, ..., M$}
+    \STATE $\tau \leftarrow \varnothing$
     \FOR{$t = 0, 1, \cdots, T-1$}
-    \STATE Sample an action $a_{t}$ using the actor network $\pi(s_{t}, g|\theta)$
+    \STATE Sample an action $a_{t}$ using the actor network $\pi(s_{t}, g;\theta)$
     \STATE Execute the action $a_{t}$ and observe a new state $s_{t+1}$
-    \STATE Store the transition $\left(s_{t}|\langle g, g^{ac}_{t} \rangle, a_{t}, r_{t}, s_{t+1}|\langle g, g^{ac}_{t+1} \rangle \right)$ in $\tau$
+    \STATE Store the transition $\left(s_{t}, a_{t}, r_{t}, s_{t+1}, g^{ac}_{t+1}\right)$ in $\tau$
     \ENDFOR
 
-    \FOR{\textbf{each} transition $\left(s_{t}, a_{t}, r_{t}, g, g^{ac}_t \right)$ in $\tau$}
+    \FOR{\textbf{each} transition $\left(s_{t}, a_{t}, r_{t}, g, g^{ac}_{t+1} \right)$ in $\tau$}
     \STATE Clone the transition and replace $g$ with $g^{\prime}$, where $g^{\prime}=g^{ac}_{T}$
-    \STATE $r_{t}^{\prime}:=r\left(s_{t}, a_{t}, g^{\prime}\right)$
-    \STATE Store the transition $\left(s_{t}, a_{t}, r_{t}^{\prime}, g^{\prime}, g^{ac}_t\right)$ in $\tau^{\prime}$
+    \STATE $r_{t}^{\prime}:=\mathcal{R}\left(s_{t}, a_{t}, g^{\prime}\right)$
+    \STATE Store the transition $\left(s_{t}, a_{t}, r_{t}^{\prime}, g^{\prime}, g^{ac}_{t+1}\right)$ in $\tau^{\prime}$
     \ENDFOR
@@
 
-    \STATE Calculate the Return $R$ and $R^{\prime}$ for all transitions in $\Tau$ and $\Tau^{\prime}$, respectively
-    \STATE Calculate the PPO loss: $\mathcal{L}_{PPO} = \mathcal{L}_{policy}(\theta) - c\cdot \mathcal{L}_{value}(\eta)$ using $\Tau$ \eqref{eq:ppo_loss}
-    \STATE Calculate the ESIL loss: $\mathcal{L}_{ESIL}(\theta)$ using $\Tau^{\prime}$, $R$ and $R^{\prime}$ \eqref{eq:imitaion}
-    \STATE Update the parameters $\theta$ and $\eta$ using loss $\mathcal{L}=\alpha\cdot\mathcal{L}_{PPO} + \beta\cdot\mathcal{L}_{ESIL}$ \eqref{eq:loss}
+    \STATE Calculate the Returns $G$ and $G^{\prime}$ for all transitions in $\Tau$ and $\Tau^{\prime}$, respectively
+    \STATE Calculate the PPO loss: $\mathcal{L}_{PPO} = \mathcal{L}_{policy} - c\mathcal{L}_{value}$
+    \STATE Calculate the ESIL loss: $\mathcal{L}_{ESIL}$ using Eq.\eqref{eq:imitaion}
+    \STATE Update the parameters $\theta$ and $\psi$ using loss $\mathcal{L}=\alpha\mathcal{L}_{PPO} + \beta\mathcal{L}_{ESIL}$
     \ENDFOR 
@@
 
-\section{Experiments and Results}
+\section{Experiments}
 \label{sec:experiment}
-The proposed method is evaluated on several multi-goal environments, including the Empty Room environment and the OpenAI Fetch environments (see Figure~\ref{fig:env}).
+The proposed method is evaluated on several multi-goal environments, including the Empty Room environment and the OpenAI Fetch environment (see Figure~\ref{fig:env}).
 The Empty Room environment is a toy example, and has discrete action spaces.
-In the Fetch environments, there are four robot tasks with continuous action spaces.
-To obtain a comprehensive comparison between the proposed method and other baseline approaches, suitable baseline approaches are selected for different environments.
-Ablation studies of the trajectory selection module are also performed.
+In the Fetch environment, there are four robot tasks with continuous action spaces.
+To obtain a comprehensive comparison between the proposed method and other baselines, suitable baselines are selected for different environments.
+In addition, ablation studies of the trajectory selection module are also performed.
 \begin{figure}[t]
@@
   \includegraphics[width=\linewidth]{figures/chapter3/empty_room.png}
-  ({a}) Empty Room \hspace{3em}          
+  ({a}) Empty Room         
 \endminipage\hfill
@@
   \includegraphics[width=\linewidth]{figures/chapter3/reach.png}
-  ({b}) FetchReach \hspace{3em}  
+  ({b}) Reach 
 \endminipage\hfill
@@
   \includegraphics[width=\linewidth]{figures/chapter3/push.png}
-  ({c}) FetchPush \hspace{3em}        
+  ({c}) Push      
 \endminipage\hfill
@@
   \includegraphics[width=\linewidth]{figures/chapter3/pick.png}
-  ({d}) FetchPickPlace
+  ({d}) Pick$\&$Place
 \endminipage\hfill
@@
   \includegraphics[width=\linewidth]{figures/chapter3/slide.png}
-  ({e}) FetchSlide \hspace{3em}       
+  ({e}) Slide     
 \endminipage
@@
 ({a}) is the Empty Room environment, in which a yellow circle indicates the position of the agent and a red star represents a target position.
-({b}--{e}) are the Fetch robotic environments.
+({b}--{e}) are the tasks in the Fetch robot environment.
 The red spot represents a target position.}
@@
 
-\subsection{Setup}
+\subsection{Environments}
 \textbf{{Empty Room (grid-world) environment}}:
@@
 The start position of the agent is at the left upper corner of the room, and the target position is randomly selected  within the room.
-When the agent chooses an action that would lead it to fall outside the grid area, the agent stays at the current position.
+When the agent chooses an action that would fall outside the grid area, the agent will stay at the current position.
 The length of each episode is 32.
-The desired goal, $g$, is a two-dimensional grid coordinate which represents the target position.
-The achieved goal, $g^{ac}_{t}$, is also a two-dimensional coordinate which represents the current position of the agent at time step $t$, and finally, the observation is a two-dimensional coordinate which represents the current position of the agent.
-The agent has five actions: left, right, up, down and stay; the agent executes a random action with probability 0.2.
-The agent can get $+1$ as a reward only when $g^{ac}_{t}=g$, otherwise, it gets a reward of $0$.
+The desired goal, $g$, is a two-dimensional grid coordinate, which represents the target position.
+The achieved goal, $g^{ac}_{t}$, is also a two-dimensional coordinate, which represents the current position of the agent, and finally, the observation is a two-dimensional coordinate which represents the current position of the agent.
+The agent has five actions: left, right, up, down, and stay; the agent executes a random action with a probability of 0.2.
+The agent can get $+1$ as a reward only when $g^{ac}_{t+1}=g$, otherwise, it gets a reward of $0$.
 
@@
 
-\textbf{{Fetch robotic (continuous) environments}%.
-}~\cite{plappert2018multi}:~The Fetch robotic environments are physically plausible simulations based on the real Fetch robot.
-The purpose of these environments is to provide a platform to tackle problems which are close to practical challenging robot manipulation tasks.
-Fetch is a 7-DoF robot arm with a two finger gripper.
-The Fetch environments include four~tasks: \emph{FetchReach}, \emph{FetchPush}, \emph{FetchPickAndPlace} and \emph{FetchSlide}.
+\textbf{{Fetch robot (continuous) environment}%.
+}~\cite{plappert2018multi}:~The Fetch robot environment is a physically plausible simulation based on the real Fetch robot.
+This environment aims to provide a platform to tackle problems that are close to practical challenging robot manipulation tasks.
+Fetch is a 7-DoF robot arm with a two-finger gripper.
+The Fetch environment include four~tasks: Reach, Push, Pick$\&$Place and Slide.
 For all Fetch tasks, the length of each episode is 50.
-The~desired goal, $g$, is a three-dimensional coordinate which represents the target position.
-If~a task has an object, the achieved goal $g^{ac}_{t}$ is a three-dimensional coordinate represents the position of the object.
-Otherwise, $g^{ac}_{t}$ is a three-dimensional coordinate represents the position of the gripper.
-Observations include the following information: position, velocity and state of the gripper.
-If a task has an object, the position, velocity and rotation information of the object is included.
-Therefore, the~observation of FetchReach is a 10-dimensional vector.~The observation of other tasks is a 25-dimensional vector.~The~action is a four-dimensional vector.
-The first three dimensions represent the relative position that the gripper needs to move in the next step.
+The~desired goal, $g$, is a three-dimensional coordinate that represents the target position.
+If~a task has an object, the achieved goal $g^{ac}_{t}$ is a three-dimensional coordinate that represents the position of the object.
+Otherwise, $g^{ac}_{t}$ is a three-dimensional coordinate that represents the position of the gripper.
+Observations include the following information: position, velocity, and state of the gripper.
+If a task has an object, the position, velocity, and rotation information of the object is included.
+Therefore, the~observation of the Reach task is a 10-dimensional vector.~The observation of other tasks is a 25-dimensional vector.~The~action is a four-dimensional vector.
+The first three dimensions represent the relative position the gripper needs to move in the next step.
 The last dimension indicates the distance between the fingers of the gripper.
-The reward function can be written as $r_{t}=-\mathds{1}\left(\left\|g^{ac}_{t} - g\right\|>\epsilon\right)$, where $\epsilon =0.05$.
+The reward function can be written as $r_{t}=-\mathds{1}\left(\left\|g^{ac}_{t+1} - g\right\|>\epsilon\right)$, where $\epsilon =0.05$.
 
-In the Fetch environments, for FetchReach, FetchPush and FetchPickAndPlace tasks, the agent is trained using 16 CPU cores.
+In the Fetch environment, for the Reach, Push, and Pick$\&$Place tasks, the agent is trained using 16 CPU cores.
 In each epoch, 50 episodes are collected for training.
-The FetchSlide task is more complex, so 32 CPU cores are used.
+The Slide task is more complex than others, so 32 CPU cores are used.
 In each epoch, 100 episodes are collected for training.
-The~{Message Passing Interface (MPI) framework is used to perform}  synchronization when updating the network.
+The~{message passing interface (MPI) framework is used to perform} synchronization when updating the network.
 After each epoch, the agent is evaluated for 10 episodes by each MPI worker.
 Finally,~the~success rate of each MPI worker is averaged.
-During training, actions are sampled from {multivariate normal} distributions.
-In the evaluation phase, the mean vector of the distribution is used as an action.
+During training, the actions are sampled from {multivariate normal} distributions.
+In the evaluation phase, the mean value of the Gaussian distribution is used as an action.
 
-The proposed method, termed PPO+ESIL, is compared with different baselines on different environments.
-All experiments are plotted based on five runs with different seeds.
+The proposed method, termed PPO+ESIL, is compared with different baselines in the different environments.
+All results are plotted based on five runs with different seeds.
 The solid line is the median value.
-The upper bound is the {75th} percentile and the lower bound is the {25th} percentile.
+The upper bound is the {75th} percentile, and the lower bound is the {25th} percentile.
 
-\subsection{Network Structure and Hyperparameters}
+\subsection{Experimental Settings}
 \textbf{Network structure}: Both the actor network and the critic network have three hidden layers with 256~neurons.
-ReLu is selected as the activation function for the hidden layers.
+ReLU is selected as the activation function for the hidden layers.
 In the grid-world environment, the actor network builds a categorical distribution.
-In the Fetch environment, the actor network builds normal {distributions by producing mean vectors and the standard deviations of the independent variables}.
+In the Fetch environment, the actor network builds Gaussian {distribution by producing mean values and the standard deviations}.
 
@@
 The discount factor $\gamma$ is 0.98.
-Adam is chosen as an optimiser with $\epsilon=0.00001$.
+Adam is chosen as an optimiser.
 For each epoch, the~actor network and critic network are updated 10 times.
 The clip ratio of the PPO algorithm is 0.2.
-For the grid-world environment, it trains networks for 100 epochs with batch size equals 160.
+For the grid-world environment, networks are trained for 100 epochs with a batch size equal to 160.
 Each~epoch consists of 100 episodes.
-For the Fetch environments, in FetchReach task, it trains networks for 100 epochs and other tasks for 1000 epochs with batch size equals to 125.
-For FetchReach, FetchPush and FetchPickAndPlace tasks, each epoch consists of 50 episodes.
-For FetchSlide task, each epoch consists of 100 episodes.
-{In designing the experiments, the number of episodes within an epoch is a balance between being able to train, the length of time required to run experiments and the maximum number of time steps that would be required to a achieve a goal.
-All environments have a fixed maximum number of time-steps , but this maximum differs depending on the problem or environment.
+For the Fetch environment, in the Reach task, networks are trained for 100 epochs and other tasks for 1000 epochs with a batch size equals to 125.
+For the Reach, Push, and Pick$\&$Place tasks, each epoch consists of 50 episodes.
+For the Slide task, each epoch consists of 100 episodes.
+In designing the experiments, the number of episodes within an epoch is a balance between being able to train, the length of time required to run experiments and the maximum number of timesteps that would be required to achieve a goal.
+All environments have a fixed maximum number of timesteps, but this maximum differs depending on the problem or environment.
 This means that the number of state--action pairs can differ between two environments that have the same number of episodes and the same number of epochs.
-We arrange the episodes to try to compensate for the number of state--action pairs collected during training to make experiments easier to compare.} {The models are trained on a machine with an Intel i7-5960X CPU and 64GB RAM.}
+We arrange the episodes to try to compensate for the number of state--action pairs collected during training to make experiments easier to compare.
+The models are trained on a machine with an Intel i7-5960X CPU and 64GB RAM.
 
-%\section{Results and Discussions}
-\subsection{Grid-World Environments}
-To understand the basic properties of the proposed method, the toy Empty Room environment is used to evaluate ESIL.
-%We compare the proposed method, named as PPO+HSL, with other baselines on the grid-world environment.
-The following baselines are considered:
-\begin{itemize}
-    \item PPO: vanilla PPO~\cite{schulman2017proximal} for discrete action spaces;
-    \item PPO+SIL/PPO+SIL+HER: Self-imitation learning (SIL) is used with PPO to solve hard exploration environments by imitating past good experiences~\cite{oh2018self}.
-In order to solve sparse rewards tasks, hindsight experience replay (HER) is applied to sampled transitions;
-    \item DQN+HER: Hindsight experience replay (HER), designed for sparse reward problems, is~combined with a deep Q-learning network (DQN)~\cite{andrychowicz2017hindsight}; this is an off policy algorithm;
-    \item Hindsight Policy Gradients (HPG): the vanilla implementation of HPG that is only suitable for discrete action spaces~\cite{rauber2018hindsight}.
-\end{itemize}
-More specifically, {PPO+ESIL is compared with above baseline methods} in Figure~\ref{fig:toy_results}a.
-This shows that PPO+ESIL converges faster than the other four baselines, and PPO+SIL converges faster than vanilla PPO, because PPO+SIL reuses past good experiences to help exploration and training.
-Hindsight~Policy Gradient (HPG) is slower than the others because goal sampling is not efficient and also unstable.
-\begin{figure}[H]
+\begin{figure}[h!]
 \minipage{0.5\textwidth}
@@
   \includegraphics[width=\linewidth]{figures/chapter3/empty_room_samples.pdf}
-  ({c}) Variation of adaptive weight coefficient $\beta$
+  ({c}) Variation of adaptive weight $\beta$
 \endminipage
@@
 
-Further, {the performance of the trajectory selection module is evaluated} in Figure~\ref{fig:toy_results}b.
-This shows that the selection strategy helps improve the performance.
+%\section{Results and Discussions}
+\subsection{Results on Grid-World Environment}
+To understand the basic properties of the proposed method, the toy Empty Room environment is used to evaluate ESIL.
+%We compare the proposed method, named as PPO+HSL, with other baselines on the grid-world environment.
+The following baselines are considered:
+\begin{itemize}
+    \item PPO: vanilla PPO~\cite{schulman2017proximal} for discrete action spaces;
+    \item PPO+SIL/PPO+SIL+HER: Self-imitation learning (SIL) is used with PPO to solve hard exploration environments by imitating past good experiences~\cite{oh2018self}.
+In order to solve sparse rewards tasks, hindsight experience replay (HER) is applied to sampled transitions;
+    \item DQN+HER: Hindsight experience replay (HER), designed for sparse reward problems, is~combined with a deep Q-learning network (DQN)~\cite{andrychowicz2017hindsight}; this is an off-policy algorithm;
+    \item Hindsight Policy Gradients (HPG): the vanilla implementation of HPG that is only suitable for discrete action spaces~\cite{rauber2018hindsight}.
+\end{itemize}
+More specifically, {PPO+ESIL is compared with the above baseline methods} in Figure~\ref{fig:toy_results}a.
+This shows that PPO+ESIL converges faster than the other four baselines, and PPO+SIL converges faster than vanilla PPO, because PPO+SIL reuses past good experiences to help exploration and training.
+Hindsight policy gradient (HPG) is slower than the others because goal sampling is not efficient and is unstable.
+
+The performance of the trajectory selection module is evaluated in Figure~\ref{fig:toy_results}b.
+It shows that the selection strategy helps improve the performance.
 Hindsight experiences are not always perfect; the trajectory selection module filters some undesirable, modified experiences.
 Through adopting this selection strategy, the chance of agents learning from poor trajectories is reduced.
-The adaptive weight coefficient $\beta$ is also investigated in these experiments.
+The adaptive weight coefficient, $\beta$, is also investigated in these experiments.
 {In Figure~\ref{fig:toy_results}c, it can be seen that at the initial stages of training, $\beta$ is high}.
-{This is because at this stage, the agent very seldom achieves the original goals}.
+{This is because, at this stage, the agent very seldom achieves the original goals}.
 The hindsight experiences can yield higher returns than the original experiences.
-{Therefore,~a~large proportion of hindsight experiences are selected to conduct self-imitation learning, helping the agent learn a policy for moving through the room}.
-In the later stages of training, the agent can achieve success frequently, and the hindsight experiences might be redundant (e.g., $R(s_{t}, g)\geq R(s_{t}, g^{\prime})$).
-In~this case, undesired hindsight experiences are removed by using the trajectory selection module and $L_{PPO}$ leads the training.
-However, when the trajectory selection module is not employed, all hindsight experiences are used through the entire training process which includes the redundant hindsight experiences.
+{Therefore,~a~large proportion of hindsight experiences is selected to conduct self-imitation learning, helping the agent learn a policy for moving through the room}.
+In the later stages of training, the agent can achieve success frequently, and the hindsight experiences might be redundant (e.g., $G(s_{t}, g)\geq G(s_{t}, g^{\prime})$).
+In~this case, undesired hindsight experiences are removed by using the trajectory selection module and $\mathcal{L}_{PPO}$ leads the training.
+However, when the trajectory selection module is not employed, all hindsight experiences are used throughout the training, including the redundant hindsight experiences.
 This leads to overfitting and makes training unstable.
-Thus, the $L_{ESIL}$ can provide the agent with a better initial policy, and the adaptive weight coefficient $\beta$ can balance the contributions of $L_{PPO}$ and $L_{ESIL}$ properly during training.
+Thus, $\mathcal{L}_{ESIL}$ can provide the agent with a better initial policy, and the adaptive weight coefficient $\beta$ can balance the contributions of $\mathcal{L}_{PPO}$ and $\mathcal{L}_{ESIL}$ properly during training.
 
-\begin{figure}[t!]
-\centering
-\minipage{0.5\textwidth}
-  \centering
-  \includegraphics[width=\linewidth]{figures/chapter3/reach_baseline.pdf}
-  ({a}) FetchReach
-\endminipage
-\minipage{0.5\textwidth}%
-  \centering
-  \includegraphics[width=\linewidth]{figures/chapter3/push_baseline.pdf}
-  ({b}) FetchPush
-\endminipage\hfill
-\minipage{0.5\textwidth}%
-  \centering
-  \includegraphics[width=\linewidth]{figures/chapter3/pick_baseline.pdf}
-  ({c}) FetchPickAndPlace
-\endminipage
-\minipage{0.5\textwidth}%
-  \centering
-  \includegraphics[width=\linewidth]{figures/chapter3/slide_baseline.pdf}
-  ({d}) FetchSlide
-\endminipage\hfill
-\caption{Results of comparison between ESIL and on-policy baselines on all Fetch environments.}
-\label{fig:baseline_compare}
-\end{figure}
 Finally, {the combination of PPO+ESIL is also compared with DQN+HER,} which is an off-policy RL algorithm, in Figure~\ref{fig:toy_results}d.
-This shows that DQN+HER works a little better than ESIL at the {start of training}.
+It shows that DQN+HER works a little better than ESIL at the {start of training}.
 However, the proposed method achieves similar results to DQN+HER later in training.
 
-\subsection{Continuous Environments}
+\subsection{Results on Continuous Control Environment}
+% continuous control
 {Continuous control problems are generally more challenging for reinforcement learning.
-In the experiments of this section, the aim is to investigate how useful the proposed method is for several hard exploration OpenAI Gym Fetch tasks.
-These environments are commonly used to assess the performance of RL methods for continuous control.} The following baselines are considered:
+In the experiments of this section, the aim is to investigate how useful the proposed method is for several hard exploration Fetch tasks.
+These robot manipulation tasks are commonly used to assess the performance of RL methods for continuous control.} The following baselines are considered:
 \begin{itemize}
@@
 \end{itemize}
-\begin{figure}[H]
+
+\begin{figure}[h!]
+\centering
+\minipage{0.5\textwidth}
+  \centering
+  \includegraphics[width=\linewidth]{figures/chapter3/reach_baseline.pdf}
+  ({a}) Reach
+\endminipage
+\minipage{0.5\textwidth}%
+  \centering
+  \includegraphics[width=\linewidth]{figures/chapter3/push_baseline.pdf}
+  ({b}) Push
+\endminipage\hfill
+\minipage{0.5\textwidth}%
+  \centering
+  \includegraphics[width=\linewidth]{figures/chapter3/pick_baseline.pdf}
+  ({c}) Pick$\&$Place
+\endminipage
+\minipage{0.5\textwidth}%
+  \centering
+  \includegraphics[width=\linewidth]{figures/chapter3/slide_baseline.pdf}
+  ({d}) Slide
+\endminipage\hfill
+\caption{Results of comparison between ESIL and on-policy baselines in all Fetch tasks.
+\td{From the results, ESIL is able to resolve all of the tasks, while other baselines can only solve the simple Reach task.}}
+\label{fig:baseline_compare}
+\end{figure}
+
+\subsubsection{Comparison to On-Policy Baselines}
+{Figure~\ref{fig:baseline_compare}, PPO+ESIL} achieves reasonable results in all Fetch tasks.
+{In contrast}, PPO, PPO+SIL, and PPO+SIL+HER do not work in all tasks, except for the Reach task.
+In comparison with the other selected tasks from the Fetch environment, Reach task is relatively simple, because there is no object to be manipulated.
+For other tasks, it is quite difficult for the agent to achieve sufficient positive rewards during exploration, because of their rare occurrence.
+Although PPO+SIL utilises past good experiences to help exploration, it is still faced with the difficulty that past experiences do not easily achieve positive rewards.
+From the experiments ({see Figure~\ref{fig:baseline_compare}}), PPO+SIL (no hindsight) converges much more slowly than using PPO only.
+Attempting to use only the original trajectories for self-imitation learning leads to unsatisfactory performance.
+For PPO+SIL+HER (with no episodic update), the sampled transitions are modified into hindsight experiences, achieving better performance in the Reach and Slide tasks.
+However, this transition-based method still cannot solve the other two manipulation tasks.
+In contrast, the proposed PPO+ESIL, utilising episodic hindsight experiences from failed trajectories, can quickly achieve positive rewards at the start of~training.
+
+\begin{figure}[h!]
 \centering
@@
   \includegraphics[width=\linewidth]{figures/chapter3/reach_hs.pdf}
-  ({a}) FetchReach
+  ({a}) Reach
 \endminipage
@@
   \includegraphics[width=\linewidth]{figures/chapter3/push_hs.pdf}
-  ({b}) FetchPush
+  ({b}) Push
 \endminipage\hfill
@@
   \includegraphics[width=\linewidth]{figures/chapter3/pick_hs.pdf}
-  ({c}) FetchPickAndPlace
+  ({c}) Pick$\&$Place
 \endminipage
@@
   \includegraphics[width=\linewidth]{figures/chapter3/slide_hs.pdf}
-  ({d}) FetchSlide
+  ({d}) Slide
 \endminipage\hfill
-\caption{Results of ablation studies with or without using trajectory selection module on all Fetch environments.}
+\caption{Results of ablation studies with and without using the trajectory selection module in all Fetch tasks.
+\td{In three challenging tasks: Push, Pick$\&$Place and Slide, the trajectory selection module can stabilise the training and prevent performance degradation in two of three cases (Push and Pick$\&$Place).}}
 \label{fig:hs_compare}
@@
 
-\subsubsection{Comparison to On-Policy Baselines}
-{Figure~\ref{fig:baseline_compare}, PPO+ESIL} achieves reasonable results on all Fetch environments.
-{In contrast}, PPO, PPO+SIL and PPO+SIL+HER do not work on all tasks, with the exception of FetchReach.
-In comparison with the other selected tasks from the Fetch environments, FetchReach is relatively simple, because there is no object to be manipulated.
-For other tasks, it is quite difficult for the agent to achieve sufficient positive rewards during exploration, because of their rare occurrence.
-Although PPO+SIL utilises past good experiences to help exploration, it is still faced with the difficulty that past experiences do not easily achieve positive rewards.
-From the experiments ({see Figure~\ref{fig:baseline_compare}}), PPO+SIL (no hindsight) converges much more slowly than using PPO only.
-Attempting to use only the original trajectories for self-imitation learning leads to unsatisfactory performance.
-For PPO+SIL+HER (with no episodic update), the sampled transitions are modified into hindsight experiences, achieving better performance in the FetchReach and FetchSlide tasks.
-However, this transition-based method still cannot solve the other two manipulation tasks.
-In contrast, the proposed PPO+ESIL, through utilizing episodic hindsight experiences from failed trajectories, can achieve positive rewards quickly at the start of~training.
-
-\subsubsection{Ablation Study of Trajectory Selection Module}
-In order to investigate the effect of trajectory selection, {ablation studies are performed} to validate the selection strategy of our approach.
-{Figure~\ref{fig:hs_compare}}, when the trajectory selection module is not used, the~ performance of the agent increases at first, and then starts to decrease.
-This suggests that the agent starts to converge to a sub-optimal location.
-However, {Figure~\ref{fig:hs_compare}d}, for the FetchSlide task, the agent converges faster without the trajectory selection module, and has better performance.
-{This is likely to be because FetchSlide is the most difficult of the Fetch environments.
-During training, the agent is very unlikely to achieve positive rewards.
-Figure~\ref{fig:her_samples_compare} also indicates that the value of $\beta$ in FetchSlide is higher than values in other environments, which means the majority of hindsight experiences have higher returns than original experiences.
-Thus, using {\em more} hindsight experiences (without filtering) accelerates training at this stage.} Nonetheless, the trajectory selection module prevents the agent overfitting the hindsight experience in the other three tasks.
-Figure~\ref{fig:her_samples_compare}, shows the adaptive weight coefficient $\beta$ on all Fetch environments.
-When the trajectory selection module is used, the value of $\beta$ decreases with the increase in training epochs.
-This implies that the agent can achieve a greater proportion of the original goals in the latter stages of training, and fewer hindsight experiences are required for  self-imitation learning.
-\begin{figure}[H]
+\begin{figure}[h!]
 \centering
@@
   \includegraphics[width=\linewidth]{figures/chapter3/reach_samples.pdf}
-  ({a}) FetchReach
+  ({a}) Reach
 \endminipage
@@
   \includegraphics[width=\linewidth]{figures/chapter3/push_samples.pdf}
-  ({b}) FetchPush
+  ({b}) Push
 \endminipage\hfill
@@
   \includegraphics[width=\linewidth]{figures/chapter3/pick_samples.pdf}
-  ({c}) FetchPickAndPlace
+  ({c}) Pick$\&$Place
 \endminipage
@@
   \includegraphics[width=\linewidth]{figures/chapter3/slide_samples.pdf}
-  ({d}) FetchSlide
+  ({d}) Slide
 \endminipage\hfill
-\caption{Variation in adaptive weight coefficient $\beta$ through the training on all Fetch environments.}
+\caption{Variation in adaptive weight coefficient $\beta$ through the training in all Fetch tasks.
+\td{In the early stage of training, the weight of the ESIL term is higher, which indicates more samples need to be modified to facilitate the exploration.
+As training progresses, the weight decreases, which means that the agent can achieve more positive samples and less modified samples are required.}}
 \label{fig:her_samples_compare}
 \end{figure}
+
+\subsubsection{Ablation Study of Trajectory Selection Module}
+To investigate the effect of trajectory selection, {ablation studies are performed} to validate the selection strategy of our approach.
+{Figure~\ref{fig:hs_compare}}, when the trajectory selection module is not used, the~ performance of the agent increases at first, and then starts to decrease.
+This suggests that the agent starts to converge to a sub-optimal location.
+However, {Figure~\ref{fig:hs_compare}d}, for the Slide task, the agent converges faster without the trajectory selection module and has better performance.
+{This is likely to be because the Slide task is the most difficult of the Fetch environment.
+During training, the agent is unlikely to achieve positive rewards.
+Figure~\ref{fig:her_samples_compare} also indicates that the value of $\beta$ in the Slide task is higher than values in other tasks, which means the majority of hindsight experiences have higher returns than original experiences.
+Thus, using {\em more} hindsight experiences (without filtering) accelerates training at this stage.} Nonetheless, the trajectory selection module prevents the agent from overfitting the hindsight experience in the other three tasks.
+Figure~\ref{fig:her_samples_compare}, shows the adaptive weight coefficient $\beta$ in all Fetch tasks.
+When the trajectory selection module is used, the value of $\beta$ decreases with the increase in training epochs.
+This implies that the agent can achieve a greater proportion of the original goals in the latter stages of training, and fewer hindsight experiences are required for self-imitation learning.
+\begin{figure}[h!]
+\minipage{0.5\textwidth}
+  \centering
+  \includegraphics[width=\linewidth]{figures/chapter3/reach_her.pdf}
+  ({a}) Reach
+\endminipage
+\minipage{0.5\textwidth}%
+  \centering
+  \includegraphics[width=\linewidth]{figures/chapter3/push_her.pdf}
+  ({b}) Push
+\endminipage\hfill
+\minipage{0.5\textwidth}%
+  \centering
+  \includegraphics[width=\linewidth]{figures/chapter3/pick_her.pdf}
+  ({c}) Pick$\&$Place
+\endminipage
+\minipage{0.5\textwidth}%
+  \centering
+  \includegraphics[width=\linewidth]{figures/chapter3/slide_her.pdf}
+  ({d}) Slide
+\endminipage\hfill
+\caption{Results of the comparison between PPO+ESIL and DDPG+HER in all fetch tasks.
+\td{From the results, PPO+ESIL has worse sample efficiency than the off-policy baseline.
+However, PPO+ESIL achieves similar or even better performance in the end of training.}}
+\label{fig:her_compare}
+\end{figure}
+
 \subsubsection{Comparison to Off-Policy Baselines}
@@
 A replay buffer is also employed to store  samples collected in the past.
-This approach has better sample efficiency than on-policy algorithms such as PPO.} Even so, {Figure~\ref{fig:her_compare}c} shows that PPO+ESIL still outperforms DDPG+HER in the FetchPickAndPlace task and the success rate is close to 1.
+This approach has better sample efficiency than on-policy algorithms such as PPO.} Even so, {Figure~\ref{fig:her_compare}c} shows that PPO+ESIL still outperforms DDPG+HER in the Pick$\&$Place task and the success rate is close to 1.
 {This suggests that PPO+ESIL approximates the characteristics of on-policy algorithms, which have low sample efficiency, but are able to obtain a comparable performance to off-policy algorithms in continuous control tasks~\cite{schulman2017proximal}.}
-\begin{figure}[H]
-\minipage{0.5\textwidth}
-  \centering
-  \includegraphics[width=\linewidth]{figures/chapter3/reach_her.pdf}
-  ({a}) FetchReach
-\endminipage
-\minipage{0.5\textwidth}%
-  \centering
-  \includegraphics[width=\linewidth]{figures/chapter3/push_her.pdf}
-  ({b}) FetchPush
-\endminipage\hfill
-\minipage{0.5\textwidth}%
-  \centering
-  \includegraphics[width=\linewidth]{figures/chapter3/pick_her.pdf}
-  ({c}) FetchPickAndPlace
-\endminipage
-\minipage{0.5\textwidth}%
-  \centering
-  \includegraphics[width=\linewidth]{figures/chapter3/slide_her.pdf}
-  ({d}) FetchSlide
-\endminipage\hfill
-\caption{Results of comparison between PPO+ESIL and DDPG+HER on all Fetch environments.}
-\label{fig:her_compare}
-\end{figure}
 
 \subsection{Overall Performance}
-{Table~\ref{tb:last}} shows the average success rate of the last 10 epochs during training of baseline methods and PPO+ESIL.
+{Table~\ref{tb:ch3_last}} shows the average success rate of the last 10 epochs during training of baseline methods and PPO+ESIL.
 The proposed ESIL achieves the best performance in four out of five tasks.
-However~PPO and PPO+SIL  only obtain reasonable results for the Empty Room and FetchReach tasks.
-With the assistance of HER, PPO+SIL+HER obtains a better performance in the FetchSlide task.
-{For the off-policy methods of DDPG+HER, all five tasks are achieved, but a better performance is  obtained than PPO+ESIL only in the FetchPush task}.
+However,~PPO and PPO+SIL only obtain reasonable results in the Empty Room and Reach task.
+With the assistance of HER, PPO+SIL+HER obtains a better performance in the Slide task.
+{For the off-policy method of DDPG+HER, all five tasks are achieved, but it only obtain a better performance than PPO+ESIL in the Push task}.
 
-\begin{table}[H]
-  \caption{Average success rate $\pm$ standard error in the last 10 epochs over five random seeds on all~environments {(\textbf{bold} indicates the best result among all methods).}}
-  \label{tb:last}
+\begin{table}[h]
   \centering
@@
     \toprule
-           & {Empty Room} & {Reach} & {Push} & {Pick} & {Slide}  \\
+           & {Empty Room} & {Reach} & {Push} & {Pick$\&$Place} & {Slide}  \\
     \midrule
@@
   \end{tabular}}
+   \caption{Average success rate $\pm$ standard error in the last 10 epochs over five random seeds in all~environments {(\textbf{bold} indicates the best result among all methods).}}
+  \label{tb:ch3_last}
 \end{table}
@@
 
-\section{Conclusions}
+\section{Discussion}
 \label{sec:conclusion}
-This paper proposed a novel method for self-imitation learning (SIL), in which an on-policy RL algorithm uses episodic modified past trajectories, i.e., hindsight experiences, to update policies.
+This chapter proposed a novel method for self-imitation learning (SIL), in which an on-policy RL algorithm uses episodic modified trajectories, i.e., hindsight experiences, to update policies.
 Compared with standard self-imitation learning, episodic self-imitation learning (ESIL) has a better performance in continuous control tasks where rewards are sparse.
-As far as we know, it is also the first time that hindsight experiences have been combined with state-of-the-art on-policy RL algorithms, such as PPO, to solve relatively hard exploration environments in continuous action spaces.
+As far as we know, it is also the first time that hindsight experiences have been combined with state-of-the-art on-policy RL algorithms, such as PPO, to solve relatively hard exploration tasks with continuous action spaces.
 
@@
 
-Future work includes extending {the proposed} method to support hierarchical reinforcement learning (HRL) algorithms for more complex manipulation control tasks, such as in-hand manipulation.
-Episodic self-imitation learning (ESIL) can also be applied to simultaneously learn sub-goal policies.
+% Future work includes extending {the proposed} method to support hierarchical reinforcement learning (HRL) algorithms for more complex manipulation control tasks, such as in-hand manipulation.
+Episodic self-imitation learning (ESIL) can also be applied to learn sub-goal policies simultaneously.

### 2022-04-23 chapters/chapter4.tex
update

@@
-\chapter{Diversity-based Trajectory and Goal Selection with Hindsight Experience Replay}
+\chapter{Efficient Sampling Strategy with Hindsight}
+\label{ch:dtgsh}
 \section{Introduction}
-Deep reinforcement learning (DRL)~\cite{arulkumaran2017deep}, in which neural networks are used as function approximators for reinforcement learning (RL), has been shown to be capable of solving complex control problems in several environments, including board games~\cite{schrittwieser2020mastering,silver2017mastering}, video games~\cite{berner2019dota,mnih2015human,vinyals2019grandmaster}, simulated and real robotic manipulation~\cite{andrychowicz2020learning,gu2017deep,levine2016end} and simulated autonomous driving~\cite{kiran2021deep}.
+% Deep reinforcement learning (DRL)~\cite{arulkumaran2017deep}, in which neural networks are used as function approximators for reinforcement learning (RL), has been shown to be capable of solving complex control problems in several environments, including board games~\cite{schrittwieser2020mastering,silver2017mastering}, video games~\cite{berner2019dota,mnih2015human,vinyals2019grandmaster}, simulated and real robotic manipulation~\cite{andrychowicz2020learning,gu2017deep,levine2016end} and simulated autonomous driving~\cite{kiran2021deep}.
 
-However, learning from a \textit{sparse} reward signal, where the only reward is provided upon the completion of a task, still remains difficult.
+% However, learning from a \textit{sparse} reward signal, where the only reward is provided upon the completion of a task, still remains difficult.
 An agent may rarely or never encounter positive examples from which to learn in a sparse-reward environment.
@@
 
-Fortunately, a common scenario is goal-oriented RL, where the RL agent is tasked with solving different goals within the same environment~\cite{kaelbling1993learning,schaul2015universal}.
+% Fortunately, a common scenario is goal-oriented RL, where the RL agent is tasked with solving different goals within the same environment~\cite{kaelbling1993learning,schaul2015universal}.
 Even if each task has a sparse reward, the agent ideally \textit{generalises} across goals, making the learning process easier.
 For example, in a robotic manipulation task, the goal during a single episode would be to achieve a specific position of a target object.
+In the last chapter, we have successfully combined hindsight goal relabelling with a state-of-the-art on-policy RL algorithm and introduced a new goal-oriented RL method -- episodic self-imitation learning (ESIL).
+Although the experiments show the effectiveness of the proposed method, ESIL still has worse sample efficiency than the vanilla hindsight experience replay (i.e., DDPG+HER).
+This is because DDPG is an off-policy RL algorithm, and it employs a replay buffer to store transitions during training, which enables the usage of past experiences.
+In this chapter, we focus on finding the shortcoming of HER and improving it to elevate learning efficiency and performance.
 
-Hindsight experience replay (HER)~\cite{NIPS2017_453fadbd} was proposed to improve the learning efficiency of goal-oriented RL agents in sparse reward settings: when past experience is replayed to train the agent, the desired goal is replaced (in ``hindsight'') with the achieved goal, generating many positive experiences.
-In the above example, the desired target position would be overwritten with the achieved target position, with the achieved reward also being overwritten correspondingly.
+As a recap, hindsight experience replay (HER)~\cite{andrychowicz2017hindsight} is proposed to improve the learning efficiency of goal-oriented RL agents in sparse reward settings: when past experiences are replayed to train the agent, the desired goal is replaced (in ``hindsight'') with the achieved goal, generating many positive experiences.
+For example, in a robot manipulation task, the goal during a single episode would be to achieve a specific position of a target object.
+After applying HER, the desired target position would be overwritten with the achieved target position, with the achieved reward also being modified correspondingly.
 
-We note that HER, whilst it enabled solutions to previously unsolved tasks, can be somewhat inefficient in its use of uniformly sampling transitions during training.
+We note that HER, whilst it enables solutions to previously unsolved tasks, can be somewhat inefficient in its use of uniformly sampling transitions during training.
 In the same way that prioritised experience replay~\cite{schaul2016prioritized} has significantly improved over the standard experience replay in RL, several approaches have improved upon HER by using data-dependent sampling~\cite{fang2019curriculum,zhao2018energy}.
@@
 Curriculum-guided HER (CHER)~\cite{fang2019curriculum} samples trajectories uniformly, and then samples transitions based on a mixture of proximity to the desired goal and the diversity of the samples; CHER adapts the weighting of these factors over time.
-In this work, we introduce diversity-based trajectory and goal selection with HER (DTGSH; See Fig.~\ref{fig:illustration}), which samples trajectories based on the diversity of the goals achieved within the trajectory, and then samples transitions based on the diversity of the set of samples.
-In this paper, DTGSH is evaluated on five challenging robotic manipulation tasks.
+In this chapter, we introduce diversity-based trajectory and goal selection with HER (DTGSH; See Figure~\ref{fig:illustration_c4}), which samples trajectories based on the diversity of the goals achieved within the trajectory, and then samples transitions based on the diversity of the set of samples.
+In the experiments, DTGSH is evaluated on five challenging robot manipulation tasks.
 From extensive experiments, our proposed method converges faster and reaches higher rewards than prior work, without requiring domain knowledge~\cite{zhao2018energy} or tuning a curriculum~\cite{fang2019curriculum}, and is based on a single concept---determinantal point processes (DPPs)~\cite{kulesza2012determinantal}.
@@
 During training, $m$ episodes are sampled according to their diversity-based priority, and then $k$ diverse, hindsight-relabelled transitions are sampled using a $k$-DPP~\cite{kulesza2011k}.}
-    \label{fig:illustration}
+    \label{fig:illustration_c4}
 \end{figure}
 
+\section{Related Work}
+The proposed work is built on HER~\cite{andrychowicz2017hindsight} as a way to effectively augment goal-oriented transitions from a replay buffer: to address the problem of sparse rewards, transitions from unsuccessful trajectories are turned into successful ones.
+HER uses an episodic replay buffer, with uniform sampling over trajectories, and uniform sampling over transitions.
+However, these samples may be redundant, and many may contribute little to the successful training of the agent.
+
+In the literature, some efforts have been made to increase the efficiency of HER by prioritising more valuable episodes/transitions.
+Motivated by the work-energy principle in physics, HEBP~\cite{zhao2018energy} assigns higher probability to trajectories in which the target object has higher energy; once the episodes are sampled, the transitions are then sampled uniformly.
+However, HEBP requires knowing the semantics of the goal space in order to calculate the probability, which is proportional to the sum of the target's potential, kinetic and rotational energies.
+
+CHER~\cite{fang2019curriculum} dynamically controls the sampling of transitions during training based on a mixture of goal proximity and diversity.
+Firstly, $m$ episodes are uniformly sampled from the episodic replay buffer, and then a minibatch of $k < m$ is sampled according to the current state of the curriculum.
+The curriculum initially biases sampling to achieved goals that are close to the desired goal (requiring a distance function), and later biases sampling towards diverse goals, using a $k$-nearest neighbour graph and a submodular function to more efficiently sample a diverse subset of goals (using the same distance function).
+
+Other work has expanded HER in orthogonal directions.
+Hindsight policy gradient \cite{rauber2018hindsight} and episodic self-imitation learning~\cite{dai2020episodic} apply HER to improve the efficiency of goal-oriented on-policy algorithms.
+Dynamic HER~\cite{fang2018dher} and competitive ER~\cite{liu2018competitive} expand HER to the dynamic goal and multi-agent settings, respectively.
+
+The use of DPPs in RL has been more limited, with applications towards modelling value functions of sets of agents in multiagent RL~\cite{osogami2019determinantal,pmlr-v119-yang20i}, and most closely related to us, finding diverse policies~\cite{parker2020effective}.
+
 \section{Methodology}
-We now formally describe the two main components of our method, DTGSH: 1) a diversity-based trajectory selection module to sample valuable trajectories for the further goal selection; 2) a diversity-based goal selection module to select transitions with diverse goal states from the previously selected trajectories.
+We now formally describe the two main components of our method, DTGSH: 1) a diversity-based trajectory selection module to sample valuable trajectories for the further goal selection, and 2) a diversity-based goal selection module to select transitions with diverse goal states from the previously selected trajectories.
 Together, these select informative transitions from a large area of the goal space, improving the agent's ability to learn and generalise.
@@
 \subsection{Diversity-based Trajectory Selection}
-We propose a diversity-based prioritization method to select valuable trajectories for efficient training.
+We propose a diversity-based prioritisation method to select valuable trajectories for efficient training.
 Related to HEBP's prioritisation of high-energy trajectories~\cite{zhao2018energy}, we hypothesise that trajectories that achieve diverse goal states $g^{ac}_{t}$ are more valuable for training; however, unlike HEBP, we do not require knowledge of the goal space semantics.
 
-In a robotic manipulation task, the agent needs to move a target object from its initial position, $g^{ac}_{1}$, to the target position, $g$.
+In a robot manipulation task, the agent needs to move a target object from its initial position, $g^{ac}_{1}$, to the target position, $g$.
 If the agent never moves the object, despite hindsight relabelling it will not be learning information that would directly help in task completion.
@@
 \begin{equation}
-    d_{\tau_{j}} = \text{det}(L_{\tau_{j}}), 
+    d_{\tau_{j}} = \text{det}(\mathbf{L}_{\tau_{j}}), 
 \label{eq:partial_div}
 \end{equation}
-where $L_{\tau_{j}}$ is the kernel matrix of partial trajectory $\tau_{j}$:
+where $\mathbf{L}_{\tau_{j}}$ is the kernel matrix of partial trajectory $\tau_{j}$:
 \begin{equation}
-    L_{\tau_{j}} = M^{T}M,
+    L_{\tau_{j}} = \mathbf{M}^{T}\mathbf{M},
 \end{equation}
-and $M=[\hat{g}^{ac}_{n}, \hat{g}^{ac}_{n+1}, \cdots, \hat{g}^{ac}_{n+b-1}]$, where each $\hat{g}^{ac}$ is the $\ell_2$-normalised version of the achieved goal $g^{ac}$~\cite{kulesza2011k}.
+and $\mathbf{M}=[\hat{g}^{ac}_{n}, \hat{g}^{ac}_{n+1}, \cdots, \hat{g}^{ac}_{n+b-1}]$, where each $\hat{g}^{ac}$ is the $\ell_2$-normalised version of the achieved goal $g^{ac}$~\cite{kulesza2011k}.
 Finally, the diversity $d_\Tau$ of trajectory $\Tau$ is the sum of the diversity of its $N_p$ constituent partial trajectories:
@@
 \subsection{Diversity-based Goal Selection}
-In prior work~\cite{NIPS2017_453fadbd,zhao2018energy}, after selecting the trajectories from the replay buffer, one transition from each selected trajectory is sampled uniformly to construct a minibatch for training.
+In prior work~\cite{andrychowicz2017hindsight,zhao2018energy}, after selecting the trajectories from the replay buffer, one transition from each selected trajectory is sampled uniformly to construct a minibatch for training.
 However, the modified goals $g^{\prime}$ in the minibatch might be similar, resulting in redundant information.
@@
 \begin{equation}
-    \mathcal{P}_{L}^{k}(\mathbf{Y}=Y) = \frac{\text{det}(L_{Y})}{\sum_{|Y^{\prime}|=k} \text{det}(L_{Y^{\prime}})}.
+    \mathcal{P}_{L}^{k}(\mathbf{Y}=Y) = \frac{\text{det}(\mathbf{L}_{Y})}{\sum_{|Y^{\prime}|=k} \text{det}(\mathbf{L}_{Y^{\prime}})}.
 \end{equation}
@@
 Finally, a $k$-DPP is used to sample a diverse set of transitions based on the relabelled goals $g'$ (which, in this context, we denote as ``candidate goals'').
-Fig.~\ref{fig:goal_sampling} gives an example of uniform vs.
+Figure~\ref{fig:goal_sampling} gives an example of uniform vs.
 $k$-DPP sampling, demonstrating the increased coverage of the latter.
-Fig.~\ref{fig:kde} provides corresponding estimated density plots; note that the density of the $k$-DPP samples is actually more uniform over the \textit{support} of the candidate goal distribution.
+Figure~\ref{fig:kde} provides corresponding estimated density plots; note that the density of the $k$-DPP samples is actually more uniform over the \textit{support} of the candidate goal distribution.
 \begin{figure}[h]
@@
  \REQUIRE set of $m$ candidate goal states $\mathcal{G}:=\{g_{i}\}_{i=1:m}$, minibatch size $k$
- \STATE $J \leftarrow\varnothing$, $M \leftarrow  [g_{1}, g_{2}, \cdots, g_{m}]$
- \STATE Calculate the DPP kernel matrix $L_M$
- \STATE $\{\boldsymbol{v}_{n}, \lambda_{n}\} \leftarrow \text{EigenDecomposition}(L_M)$
+ \STATE $J \leftarrow\varnothing$, $\mathbf{M} \leftarrow  [g_{1}, g_{2}, \cdots, g_{m}]$
+ \STATE Calculate the DPP kernel matrix $\mathbf{L}_\mathbf{M}$
+ \STATE $\{\boldsymbol{v}_{n}, \lambda_{n}\} \leftarrow \text{EigenDecomposition}(\mathbf{L}_\mathbf{M})$
  %\STATE $e_{k}(\lambda_{1}, \cdots, \lambda_{m}): = \sum\limits_{|J'|=k}\prod_{n\in J'}\lambda_{n}$ \hfill{$\triangleright$ elementary symmetric polynomial: $e^{m}_{k}$}
@@
 
-Alg.~\ref{alg:goal_sample} shows the details of the goal selection subroutine, and Alg.~\ref{alg:main_alg} gives the overall algorithm for our method, DTGSH.
+Algorithm~\ref{alg:goal_sample} shows the details of the goal selection subroutine, and Algorithm~\ref{alg:main_alg} gives the overall algorithm for our method, DTGSH.
 
 \section{Experiments}
-We evaluate our proposed method, and compare it with current state-of-the-art HER-based algorithms~\cite{NIPS2017_453fadbd,fang2019curriculum,zhao2018energy} on challenging robotic manipulation tasks~\cite{plappert2018multi}, pictured in Fig.~\ref{fig:env}.
+We evaluate our proposed method, and compare it with current state-of-the-art HER-based algorithms~\cite{andrychowicz2017hindsight,fang2019curriculum,zhao2018energy} on challenging robot manipulation tasks~\cite{plappert2018multi}, pictured in Figure~\ref{fig:env_c4}.
 Furthermore, we perform ablation studies on our diversity-based trajectory and goal selection modules.
@@
   \end{subfigure}\hfill
-  \caption{Robotic manipulation environments. (a-b) use the Fetch robot, and (c-e) use the Shadow Dexterous Hand.} 
-  \label{fig:env}
+  \caption{Robot manipulation environments. (a-b) use the Fetch robot, and (c-e) use the Shadow Dexterous Hand.} 
+  \label{fig:env_c4}
 \end{figure}
@@
 \subsection{Environments}
-The robotic manipulation environments used for training and evaluation include five different tasks.
-Two tasks use the 7-DoF Fetch robotic arm with two-fingers parallel gripper: Push, and Pick\&Place, which both require the agent to move a cube to the target position.
+The robot manipulation environments used for training and evaluation includes five different tasks.
+Two tasks use the 7-DoF Fetch robot arm with two-fingers parallel gripper: Push, and Pick\&Place, which both require the agent to move a cube to the target position.
 The remaining three tasks use a 24-DoF Shadow Dexterous Hand to manipulate an egg, a block and a pen, respectively.
-The sparse reward function is given by Eq.~\eqref{eq:sparse_reward}.
+The sparse reward function is given by Equation~\eqref{eq:sparse_reward}.
 
-In the Fetch environments, the state $s_{t}$ contains the position and velocity of the joints, and the position and rotation of the cube.
-Each action $a_{t}$ is a 4-dimensional vector, with three dimensions specifying the relative position of the gripper, and the final dimension specifying the state of the gripper (i.e., open or closed).
+In the Fetch environment, the state $s_{t}$ contains the position and velocity of the joints, and the position and rotation of the cube.
+Each action, $a_{t}$, is a 4-dimensional vector, with three dimensions specifying the relative position of the gripper, and the final dimension specifying the state of the gripper.
 The desired goal $g$ is the target position, and the achieved goal $g^{ac}_{t}$ is the position of the cube.
 Each episode is of length $T = 50$.
+%(i.e., open or closed)
 
-In the Shadow Dexterous Hand environments, the state $s_{t}$ contains the position and velocity of the joints.
+In the Shadow Dexterous Hand environment, the state $s_{t}$ contains the position and velocity of the joints.
 Each action $a_{t}$ is a 20-dimensional vector which specifies the absolute position of 20 non-coupled joints in the hand.
-The desired goal $g$ and achieved goal $g^{ac}_t$ specify the rotation of the object for the block and pen tasks, and the position + rotation of the object for the egg task.
+The desired goal, $g$, and achieved goal, $g^{ac}_t$, specify the rotation of the object for the block and pen tasks, and the position + rotation of the object for the egg task.
 Each episode is of length $T = 200$.
 
-\subsection{Training Settings}
+\subsection{Experimental Settings}
 We base our training setup on CHER~\cite{fang2019curriculum}.
 We train all agents on minibatches of size $k = 64$ for 50 epochs using MPI for parallelisation over 16 CPU cores; each epoch consists of 1600 ($16 \times 100$) episodes, with evaluation over 160 ($16 \times 10$) episodes at the end of each epoch.
-Remaining hyperparameters for the baselines are taken from the original work~\cite{NIPS2017_453fadbd,fang2019curriculum,zhao2018energy}.
+Remaining hyperparameters for the baselines are taken from the original respective works~\cite{andrychowicz2017hindsight,fang2019curriculum,zhao2018energy}.
 Our method, DTGSH, uses partial trajectories of length $b = 2$ and $m = 100$ as the number of candidate goals.
 
-\subsection{Benchmark Results}
-We compare DTGSH to DDPG~\cite{lillicrapContinuous}, DDPG+HER~\cite{NIPS2017_453fadbd}, DDPG+HEBP~\cite{zhao2018energy} and DDPG+CHER~\cite{fang2019curriculum}.
+\subsection{Results on Benchmark Environments}
+We compare DTGSH to DDPG~\cite{lillicrapContinuous}, DDPG+HER~\cite{andrychowicz2017hindsight}, DDPG+HEBP~\cite{zhao2018energy} and DDPG+\\CHER~\cite{fang2019curriculum}.
 Evaluation results are given based on repeated runs with 5 different seeds; we plot the median success rate with upper and lower bounds given by the \nth{75} and \nth{25} percentiles, respectively.
@@
 \begin{figure}
-  \begin{subfigure}[t]{0.33\textwidth}
+\centering
+  \begin{subfigure}[t]{0.49\textwidth}
     \includegraphics[width=\textwidth]{figures/chapter4/FetchPush-v1.pdf}
@@
   \end{subfigure}\hfill
-  \begin{subfigure}[t]{0.33\textwidth}
+  \begin{subfigure}[t]{0.49\textwidth}
     \includegraphics[width=\textwidth]{figures/chapter4/FetchPickAndPlace-v1.pdf}
@@
   \end{subfigure}\hfill
-  \begin{subfigure}[t]{0.33\textwidth}
+  \begin{subfigure}[t]{0.49\textwidth}
     \includegraphics[width=\textwidth]{figures/chapter4/HandManipulateEggFull-v0.pdf}
@@
   \end{subfigure}\hfill
-  \begin{subfigure}[t]{0.33\textwidth}
+  \begin{subfigure}[t]{0.49\textwidth}
     \includegraphics[width=\textwidth]{figures/chapter4/HandManipulateBlockRotateXYZ-v0.pdf}
@@
   \end{subfigure}\hfill
-  \begin{subfigure}[t]{0.33\textwidth}
+  \begin{subfigure}[t]{0.49\textwidth}
     \includegraphics[width=\textwidth]{figures/chapter4/HandManipulatePenRotate-v0.pdf}
@@
   \end{subfigure}\hfill
-  \begin{subfigure}[t]{0.33\textwidth}
-    \includegraphics[width=\textwidth]{figures/chapter4/blank.png}
-  \end{subfigure}\hfill
-  \caption{Success rate of DTGSH and baseline approaches.} 
-  \label{fig:baseline_compare}
+%   \begin{subfigure}[t]{0.49\textwidth}
+%     \includegraphics[width=\textwidth]{figures/chapter4/blank.png}
+%   \end{subfigure}\hfill
+  \caption{Success rate of DTGSH and baseline approaches. \td{DDPG+DTGSH outpeforms other baselines in all tasks, and without using HER, DDGP cannot even complete the simple Reach task.}} 
+  \label{fig:baseline_compare_c4}
 \end{figure}
 
-Fig.~\ref{fig:baseline_compare} and Tab.~\ref{tab:benchmark} show the performance of DDPG+DTGSH and baseline approaches on all five tasks.
+Figure~\ref{fig:baseline_compare_c4} and Table~\ref{tab:benchmark} show the performance of DDPG+DTGSH and baseline approaches on all five tasks.
 In the Fetch tasks, DDPG+DTGSH and DDPG+HEBP both learn significantly faster than the other methods, while in the Shadow Dexterous Hand tasks DDPG+DTGSH learns the fastest and achieves higher success rates than all other methods.
@@
 %\vspace{-15mm}
-\subsection{Ablation Studies}
-In this section, we perform the following experiments to investigate the effectiveness of each component in DTGSH: 1) diversity-based trajectory selection with HER (DTSH) and diversity-based goal selection with HER (DGSH) are evaluated independently to assess the contribution of each stage; 2) the performance using different partial trajectory lengths $b$; 3) the performance of using different candidate goal set sizes $m$.
-\begin{figure}[h]
-  \begin{subfigure}[t]{0.33\textwidth}
+\subsection{Results on Ablation Studies}
+In this section, we perform the following experiments to investigate the effectiveness of each component in DTGSH: 1) diversity-based trajectory selection with HER (DTSH) and diversity-based goal selection with HER (DGSH) are evaluated independently to assess the contribution of each stage, 2) the performance using different partial trajectory lengths $b$, and 3) the performance of using different candidate goal set sizes $m$.
+\begin{figure}[h!]
+  \centering
+  \begin{subfigure}[t]{0.49\textwidth}
     \includegraphics[width=\textwidth]{figures/chapter4/FetchPush-v1_ab1.pdf}
@@
   \end{subfigure}\hfill
-  \begin{subfigure}[t]{0.33\textwidth}
+  \begin{subfigure}[t]{0.49\textwidth}
     \includegraphics[width=\textwidth]{figures/chapter4/FetchPickAndPlace-v1_ab1.pdf}
@@
   \end{subfigure}\hfill
-  \begin{subfigure}[t]{0.33\textwidth}
+  \begin{subfigure}[t]{0.49\textwidth}
     \includegraphics[width=\textwidth]{figures/chapter4/HandManipulateEggFull-v0_ab1.pdf}
@@
   \end{subfigure}\hfill
-  \begin{subfigure}[t]{0.33\textwidth}
+  \begin{subfigure}[t]{0.49\textwidth}
     \includegraphics[width=\textwidth]{figures/chapter4/HandManipulateBlockRotateXYZ-v0_ab1.pdf}
@@
   \end{subfigure}\hfill
-  \begin{subfigure}[t]{0.33\textwidth}
+  \begin{subfigure}[t]{0.49\textwidth}
     \includegraphics[width=\textwidth]{figures/chapter4/HandManipulatePenRotate-v0_ab1.pdf}
@@
   \end{subfigure}\hfill
-  \begin{subfigure}[t]{0.33\textwidth}
-    \includegraphics[width=\textwidth]{figures/chapter4/blank.png}
-  \end{subfigure}\hfill
-  \caption{Success rate of HER, DTGSH, and ablations DTSH and DGSH.} 
+%   \begin{subfigure}[t]{0.33\textwidth}
+%     \includegraphics[width=\textwidth]{figures/chapter4/blank.png}
+%   \end{subfigure}\hfill
+  \caption{Success rate of HER, DTGSH, and ablations DTSH and DGSH.
+\td{From the results, DTSH outperforms DDPG+HER in all tasks, which verify the efficiency that selecting trajectories with diverse achieved goals for training.
+DGSH also has better performance than DDPG+HER in three tasks.
+By combining these two modules together, DTGSH achieves the best performance.} }
   \label{fig:ablation1}
@@
 
-Fig.~\ref{fig:ablation1} shows the performance of using DTSH and DGSH independently.
-DDPG+DTSH outperforms DDPG+HER substantially in all tasks, which supports the view that sampling trajectories with diverse achieved goals can substantially improve performance.
-Furthermore, unlike DDPG+HEBP, DTSH does not require knowing the structure of the goal space in order to calculate the energy of the target object; DDPG+DGSH achieves better performance than DDPG+HER in three environments, and is only worse in one environment.
-DGSH performs better in environments where it is easier to solve the task (e.g., Fetch tasks), and hence the trajectories selected are more likely to contain useful transitions.
+Figure~\ref{fig:ablation1} shows the performance of using DTSH and DGSH independently.
+DDPG+\\DTSH outperforms DDPG+HER substantially in all tasks, which supports the view that sampling trajectories with diverse achieved goals can substantially improve performance.
+Furthermore, unlike DDPG+HEBP, DTSH does not require knowing the structure of the goal space in order to calculate the energy of the target object; DDPG+DGSH achieves better performance than DDPG+HER in three tasks, and is only worse in one task.
+DGSH performs better in the environment where it is easier to solve the task (e.g., Fetch tasks), and hence the trajectories selected are more likely to contain useful transitions.
 However, DTGSH, which is the combination of both modules, performs the best overall.
 
-\begin{figure}[t]
-  \begin{subfigure}[t]{0.33\textwidth}
+\begin{figure}[h!]
+\centering
+  \begin{subfigure}[t]{0.49\textwidth}
     \includegraphics[width=\textwidth]{figures/chapter4/FetchPush-v1_ab2.pdf}
@@
   \end{subfigure}\hfill
-  \begin{subfigure}[t]{0.33\textwidth}
+  \begin{subfigure}[t]{0.49\textwidth}
     \includegraphics[width=\textwidth]{figures/chapter4/FetchPickAndPlace-v1_ab2.pdf}
@@
   \end{subfigure}\hfill
-  \begin{subfigure}[t]{0.33\textwidth}
+  \begin{subfigure}[t]{0.49\textwidth}
     \includegraphics[width=\textwidth]{figures/chapter4/HandManipulateEggFull-v0_ab2.pdf}
@@
   \end{subfigure}\hfill
-  \begin{subfigure}[t]{0.33\textwidth}
+  \begin{subfigure}[t]{0.49\textwidth}
     \includegraphics[width=\textwidth]{figures/chapter4/HandManipulateBlockRotateXYZ-v0_ab2.pdf}
@@
   \end{subfigure}\hfill
-  \begin{subfigure}[t]{0.33\textwidth}
+  \begin{subfigure}[t]{0.49\textwidth}
     \includegraphics[width=\textwidth]{figures/chapter4/HandManipulatePenRotate-v0_ab2.pdf}
@@
   \end{subfigure}\hfill
-  \begin{subfigure}[t]{0.33\textwidth}
-    \includegraphics[width=\textwidth]{figures/chapter4/blank.png}
-  \end{subfigure}\hfill
-  \caption{Success rate of DTGSH with different partial trajectory lengths $b$ and different candidate goal set sizes $m$.} 
+%   \begin{subfigure}[t]{0.33\textwidth}
+%     \includegraphics[width=\textwidth]{figures/chapter4/blank.png}
+%   \end{subfigure}\hfill
+  \caption{Success rate of DTGSH with different partial trajectory lengths $b$ and different candidate goal set sizes $m$.
+\td{From the results, when $b \gg 2$, the performance degrades in all tasks.
+When $m \gg 100$, the performance degrades in the Shadow Dexterous Hand environment.}}
   \label{fig:ablation2}
@@
 
-Fig.~\ref{fig:ablation2} shows the performance of DDPG+DTGSH with different partial trajectory lengths $b$ and different candidate goal set sizes $m$.
+Figure~\ref{fig:ablation2} shows the performance of DDPG+DTGSH with different partial trajectory lengths $b$ and different candidate goal set sizes $m$.
 In this work, we use $b = 2$ and $m = 100$ as the defaults.
-Performance degrades with $b \gg 2$, indicating that pairwise diversity is best for learning in our method.
-$m \gg 100$ does not affect performance in the Fetch environments, but degrades performance in the Shadow Dexterous Hand environments.
+Performance degrades with $b \gg 2$, indicating that pairwise diversity is the best for learning in our method.
+$m \gg 100$ does not affect performance in the Fetch environment, but degrades performance in the Shadow Dexterous Hand environment.
 
@@
 DTGSH requires an additional calculation of the diversity score of $\mathcal{O}(N_{p}b^3)$ at the end of every training episode, and sampling of $\mathcal{O}(mk^2)$ for each minibatch.
-\begin{table}[h]
+\begin{table}[h!]
     \centering
-    \resizebox{\textwidth}{!}{
+    \resizebox{0.85\textwidth}{!}{
     \begin{tabular}{c|c|c|c|c}
@@
 \section{Discussion}
-In this paper, we introduced diversity-based trajectory and goal selection with hindsight experience replay (DTGSH) to improve the learning efficiency of goal-orientated RL agents in the sparse reward setting.
-Our method can be divided into two stages: 1) valuable trajectories are selected according to diversity-based priority, as modelled by determinantal point processes (DPPs)~\cite{kulesza2012determinantal}; 2) $k$-DPPs~\cite{kulesza2011k} are leveraged to sample transitions with diverse goal states from previously selected trajectories for training.
-Our experiments empirically show that DTGSH achieves faster learning and higher final performance in five challenging robotic manipulation tasks, compared to previous state-of-the-art approaches~\cite{NIPS2017_453fadbd,fang2019curriculum,zhao2018energy}.
+In this chapter, we introduced diversity-based trajectory and goal selection with hindsight experience replay (DTGSH) to improve the learning efficiency of goal-orientated RL agents in sparse reward settings.
+Our method can be divided into two stages: 1) valuable trajectories are selected according to diversity-based priority, as modelled by determinantal point processes (DPPs)~\cite{kulesza2012determinantal}, and 2) $k$-DPPs~\cite{kulesza2011k} are leveraged to sample transitions with diverse goal states from previously selected trajectories for training.
+Our experiments empirically show that DTGSH achieves faster learning and higher final performance in five challenging robot manipulation tasks, compared to previous state-of-the-art approaches~\cite{andrychowicz2017hindsight,fang2019curriculum,zhao2018energy}.
 Furthermore, unlike prior extensions of hindsight experience replay, DTGSH does not require semantic knowledge of the goal space~\cite{zhao2018energy}, and does not require tuning a curriculum~\cite{fang2019curriculum}.

### 2022-04-23 chapters/chapter5.tex
update

\newcommand{\ourmethod}{\text{DAIM}}
\newcommand{\policypi}{\mathbf{\pi} }
\newcommand{\expection}{\mathbb{E} }
\newcommand{\tables}{{Table }}
\newcommand{\eq}{{Equation }}
\newcommand{\alg}{{Algorithm }}
\newcommand{\starcraft}{{StarCraft II}}

\chapter{Diversity-Augmented Intrinsic Motivation}
\label{ch:daim}
\section{Introduction}
The previous chapter uses the determinantal point processes (DPPs) framework to measure the diversity of achieved goals in a trajectory and proposes a sampling strategy that uses diversity as a priority to sample the trajectories for further processing, in which diversity can also quantify the contribution of each trajectory. In this chapter, we keep exploring the potential application of determinantal point processes (DPPs) in deep reinforcement learning (DRL).

%Deep reinforcement learning (RL) allows agents to learn a wide range of reward-driven skills, including playing games~\cite{mnih2015human,silver2016mastering,xu2020neurips}, controlling robots~\cite{dai2020episodic,fang2018dher,gu2017deep,levine2016end}, and making decisions in other environments~\cite{fang2017learning,vinyals2017starcraft}. 
Many important applied problems in DRL can be accurately modelled as learning from extrinsic rewards, where a function (approximately) relates features associated with a state/action pair and the expectation of the resulting rewards. Furthermore, with appropriate features, it is often the case that linear or nonlinear functions can accurately approximate the reward function. However, ill-designed reward functions usually lead to unexpected behaviors~\cite{riedmiller2018learning} and limited exploration. 
Potential-based reward shaping methods~\cite{ng1999policy} offers a general answer to what degree of reward function modifications do not affect the optimal policy. Other works have tried to design auxiliary rewards to derive policies with desired properties. For example, Jaderberg \textit{et al.}~\cite{jaderberg2016reinforcement} calculate a pseudo-reward from unsupervised auxiliary tasks, and use this reward to refine its internal representation. In count-based methods~\cite{bellemare2016unifying, ostrovski2017count, tang2017exploration}, a pseudo-count based reward is used to encourage the agent to explore novel states in the environment.

In environments or tasks for which the reward sparsity is higher, undirected, or even deceptive, the application of RL techniques is significantly more difficult, since the agent might have only occasional opportunities to improve its policy, such as when the final goal is reached. In such scenarios, agents can be endowed with mechanisms of {\it intrinsic motivation} that encourage exploration in environments and learn useful policies even without explicit supervision. Early attempts to incorporate intrinsic motivation use self-supervised prediction errors as intrinsic rewards to help the agent explore~\cite{pathak2017curiosity}. More recently, Zheng \textit{et al.}~\cite{zheng2018learning} design an augmented agent with additive intrinsic rewards. When the intrinsic value of the action is added to the task-dependent (hereafter extrinsic) rewards, the performance of policy gradient based learning methods can be improved. Another line of work uses auxiliary control tasks, such as pixel control or feature control, which can significantly speed up learning of the main task. The rationale is that learning the auxiliary tasks gives the agent features that are useful for manipulating the environment. However, these intrinsic methods are either difficult to explain or only specific to some tasks.

Encouraging effective exploration of novel states for agents without any reward signal is still a challenging problem.
Making progress on this problem requires specifying a learning objective that ensures the agent explores large parts of the state space. In this chapter, we show how a simple objective based on a diversity measure can enable RL agents to autonomously explore states. This module is useful for a number of environments, as confirmed by a selection of MuJoCo locomotion tasks and video games.

We hypothesise that in order to acquire an agent that is useful and efficient, we must train the agent so that it can attain coverage over the set of possible states. However, the common exploration strategies ({e.g.}, random sampling, $\epsilon$-greedy) can re-visit previously explored states of the agent, limiting the effectiveness of state-space coverage.

A key idea in our work is to use a measure of discrepancy between states as an intrinsic motivation. \td{Further, states that are distinguishable are not necessarily diverse in a semantically meaningful way. To combat this problem, we want the intrinsic motivation of the agent that is not only distinguishable but also is as diverse as possible.}

To model the diversity of the agent's states, we adopt the framework of determinantal point processes (DPPs)~\cite{kulesza2012determinantal} to formulate the intrinsic reward as a dynamic measure of diversity. Regarding the sequence of states as a DPP allows a measure of the diversity of subset selection, widely used in data summarization for selecting a diversified subset \cite{gong2014diverse}. We build an intrinsic reward for diversity with parameterised DPPs and propose a new general method for learning policies with intrinsic rewards. The proposed method is referred to as diversity-augmented intrinsic motivation ($\ourmethod$). Through a series of experiments, we show that the agents trained with the proposed diversity-based intrinsic rewards learn effectively and attain higher rewards than well-known baselines, especially in continuous control environments.

The contributions of this chapter can be summarized as follows: 1) The DPPs framework is proposed as a measure for the diversity of adjacent states, and this diversity is used as the intrinsic reward, 2) A bi-level optimisation~\cite{zheng2018learning} is borrowed to combine with the proposed diversity-based measure, so that the intrinsic reward will increase the extrinsic reward while increasing diversity, and 3) $\ourmethod$ is evaluated in the MuJoCo locomotion and arcade learning environments. Empirical results demonstrate that $\ourmethod$ can outperform baseline methods in most of the standard tasks and outperform baselines in 12 out of 15 tasks with delayed rewards in the MuJoCo locomotion environment. In addition, $\ourmethod$ also improves the rate of learning in the initial stage of training in the arcade learning environment.

\section{Related Work}
Intrinsic motivation is an approach to explore novel states which may lead to higher rewards in the future. Pathak \textit{et al.}~\cite{pathak2017curiosity} propose an intrinsic curiosity module (ICM), suggesting the prediction error between an actual encoded frame and the predicted encoded frame as the intrinsic reward. \td{A higher prediction error indicates the agent has moved into a novel state, and this intrinsic reward will encourage the agent to keep exploring such novel states.} In order to find an appropriate representation of pixels that could be invariant to insignificant changes in the environment, an inverse dynamic model is introduced to extract features. This requires training a network, which uses consecutive frames as input to predict the action that would transition the previous frame to the subsequent. The insight of the inverse dynamic model is that through predicting environment dynamics, features are learned to preserve the key information that predicts the action. Burda \textit{et al.}~\cite{burda2018largescale} conduct a large-scale study of curiosity-driven learning. Their findings suggest that an agent can learn to play a significant number of Atari games, even without receiving any extrinsic rewards from the environment.

However, environments with stochastic dynamics may not be suitable for the application of purely exploratory approaches. Variational Information Maximising Exploration (VIME)~\cite{NIPS2016_abd81528} formulates the intrinsic reward based on the information gain according to the agent’s belief of environment dynamics, and it shows significant improvements in the continuous control tasks with the sparse reward setting.

Zheng \textit{et al.}~\cite{zheng2018learning} introduce learning intrinsic reward for policy gradient (LIRPG), which enables a learnable intrinsic reward function to be used for training an RL agent. \td{In addition, LIRPG adopts a bi-level optimisation scheme that when the policy network is updated to maximise the accumulated intrinsic and extrinsic rewards, the intrinsic reward function is also updated to maximise the accumulated extrinsic rewards achieved by the policy network. This optimisation scheme guarantees the learned reward function can generate meaningful intrinsic rewards to facilitate the agent to reach the ultimate goal (i.e., maximising the accumulated extrinsic rewards). In this work, we also leverage this bi-level optimisation to ensure the agent can explore novel states that deliver more accumulated extrinsic rewards.}

\section{Methodology}
% start
\begin{figure}[t]
    \centering
    \includegraphics[width=\linewidth]{figures/chapter5/illustration.pdf}
    \caption{Illustration of $\ourmethod$. It shows the intrinsic rewards of an entire episode of SuperMarioBros. The red points annotate the key frames in the game. In frames \{17, 26, 32, 41\}, the agent keeps moving forward, and the intrinsic rewards are \td{relatively higher}. In the frames \{303, 327, 354, 379\}, the agent is stuck at the bottom of the flagpole. For these latter four frames, the diversity between adjacent states is very small, and the intrinsic rewards are close to zero.}
    \label{fig:illustration}
\end{figure}

In the proposed method, let $r_{t}^{\text{ex}}$ denote the extrinsic reward received from the environment and $r^{\text{in}}_{t}$ denote the intrinsic reward parameterised by $\eta$. The agent can receive a mixed reward represented in the following form:
\begin{equation}
    r^{\text{mix}}_t = \lambda^{\text{ex}} r^{\text{ex}}(s_{t}, a_t) + \lambda^{\text{in}} r^{\text{in}}(s_t, s_{t+1};\eta),
\end{equation}
where $\lambda^{\text{ex}}$ and $\lambda^{\text{in}}$ are the weight coefficients of the extrinsic reward $r^{\text{ex}}_{t}$ and the intrinsic reward $r^{\text{in}}_{t}$ at timestep $t$. The agent is updated to maximise the diversity-augmented return $G^{\text{mix}}(s_t,a_t)$, where $G^{\text{mix}}(s_t,a_t) =\sum_{k=0}^{T-1}\gamma^{k}r^{\text{mix}}_{t+k}$. Notice that the ultimate goal of the agent is still maximising the extrinsic return $G^{\text{ex}}(s_{t}, a_{t})$.

In the rest of this section, we first introduce the primary objective function and the optimisation algorithm that can ensure the maximisation of the extrinsic return while increasing the diversity-augmented returns. Subsequently, the intrinsic reward formulated under the DPPs framework is discussed.

\subsection{Primary Objective Function}
To make sure that $\ourmethod$ still increases the extrinsic reward, we adopt a bi-level optimisation framework. The objective function is defined as follows:
\begin{align}
  \label{eq:obj}
  \max_{\eta}\quad & J^{\text{ex}}(\eta;\theta^{\prime}),\\ 
  \text{subject to}\quad & \theta^{\prime} =\mathop{\arg\max}_{\theta}\ J^{\text{mix}}(\theta, \eta).
  \nonumber
\end{align}
\td{Here $J^{\text{mix}} = \mathbb{E}_{\pi_{\theta}}\bigr[G^{\text{mix}}(s_t,a_t) \bigr]$, and $\theta$ is the parameters of the policy network.
The optimisation problem of \eqref{eq:obj} can be viewed as a bi-level optimisation, which is widely used in the fields of machine learning, such as meta-learning~\cite{andrychowicz2016learning,nichol2018first,santoro2016meta,xu2018meta,zheng2018learning}, signal processing~\cite{kunapuli2008classification,flamary2014learning}, and hyperparameters optimisation~\cite{franceschi2018bilevel, shaban2019truncated}. In Equation~\eqref{eq:obj}, $J^{\text{ex}}$ and $J^{\text{mix}}$ are upper-level and inner-level objective functions, respectively. The primary goal of this bi-level objective function is to maximise extrinsic returns (upper-level objective function) $J^{\text{ex}}$ with respect to the parameters of intrinsic reward $\eta$, and the updated parameters of the policy network $\theta^{\prime}$ can be achieved via maximising the inner-level objective function.}

% The upper-level optimisation is treated as the leader, while the lower-level optimisation is the follower. The upper agent reshapes the reward to make the low-level agent's learning easier. The low-level agent observes the commands of the upper-level agent and works towards meeting the commands.

The policy parameters $\theta$ and the intrinsic parameters $\eta$ can be updated alternately.
Firstly, the update of the policy network $\pi_{\theta}$ with a learning rate $\alpha$ can be written as:
\begin{equation}
    \theta^{\prime}=\theta + \alpha\nabla_{\theta} \log \pi(a_t|s_t;\theta)G^{\text{mix}}(s_t, a_t).
\label{eq:mix_cost}
\end{equation}
Secondly, we update $\eta$ towards maximising the expected extrinsic returns $G^{\text{ex}}(s_t, a_t)$ of the agent. 
To build the connection between $J^{\text{ex}}$ and $\eta$, the updated parameters of the policy network $\theta^{\prime}$ are used to sample a new trajectory and solve the following problem:
\begin{equation}
    \max_{\eta} J^{\text{ex}} = \mathbb{E}_{\pi_{\theta^{\prime}|\eta}}[G^{\text{ex}}(s_t, a_t)].
\label{eq:cost_ex}
\end{equation}
Then, the gradient $\nabla_{\eta} J^{\text{ex}}$ can be calculated using the chain rule:
\begin{equation}
    \nabla_{\eta} J^{\text{ex}} = \nabla_{\theta^{\prime}}J^{\text{ex}}\nabla_{\eta}\theta^{\prime}, 
  \label{eq:chain_rule}
\end{equation}
with 
\begin{align}
    \nabla_{\eta}\theta^{\prime} &=  \nabla_{\eta}\alpha G^{\text{mix}}(s_t, a_t) \nabla_{\theta} \log \pi(a_t|s_t;\theta) \nonumber \\
    &= \alpha\lambda^{\text{in}}\sum_{k=0}^{\infty}\gamma^{k}\nabla_{\eta}r^{\text{in}}_{\eta,t+k}\nabla_{\theta} \log \pi(a_t|s_t;\theta). 
\end{align}

Here, instead of collecting another set of samples, importance sampling is employed to increase the sample efficiency of the algorithm:
\begin{equation}
    \nabla_{\theta^{\prime}}J^{\text{ex}} =  \nabla_{\theta^{\prime}}\Bigg (\frac{\pi(a_{t}|s_{t};\theta^{\prime})}{\pi(a_{t}|s_{t};\theta)}\Bigg ) G^{\text{ex}}(s_{t}, a_{t}).
\label{eq:importance_sampling}
\end{equation}
% The principle of the \eq \eqref{eq:chain_rule} is to formulate the effect of the change of $\eta$ on influencing $J^{\text{ex}}$ through its influence on the updated policy parameter $\theta'$.
%This is a commonly adopted technique in meta-gradient learning~\cite{andrychowicz2016learning,nichol2018first,santoro2016meta,xu2018meta,zheng2018learning}. 
Through \td{this bi-level optimisation}, the features that are learned ensure that the resulting diversity also increases the extrinsic returns. \td{Furthermore, the value network is updated via a mean squared error loss $\mathcal{L}_{value} = \mathbb{E}[(V(s_{t};\psi) - G_{t}^{\text{mix}})^2]$.} Algorithm~\ref{algo:vi} gives the pseudo code for our method -- diversity-augmented intrinsic motivation ($\ourmethod$). The code is publicly available at: \href{https://github.com/TianhongDai/daim-rl}{https://github.com/TianhongDai/daim-rl}.

\begin{algorithm}[t]
	\caption{\label{alg:all} Diversity-Augmented Intrinsic Motivation.
	}
	\begin{algorithmic}[1]
		\REQUIRE a policy network $\pi(a_{t}|s_{t};\theta)$, an intrinsic reward module $\phi(s_{t},s_{t+1};\eta)$, a value network $V(s_{t};\psi)$, learning rate $\alpha$, intrinsic module learning rate $\beta$, transition buffer $\mathcal{B}$, number of iterations $N$, length of each rollout $T$, weight coefficient of extrinsic reward $\lambda^{\text{ex}}$, weight coefficient of intrinsic reward $\lambda^{\text{in}}$
		%\STATE \textbf{Init}: initialize the policy network $\pi(a_{t}|s_{t};\theta)$ and intrinsic network $\phi(s_{t};\eta)$
		\FOR{iteration = 1, 2, 3, ..., $N$}
		\STATE Let $\mathcal{B} \leftarrow \varnothing$
		\FOR{$t$ = 1, 2, 3, ..., $T$}
        \STATE Sample an action $a_{t}$ using policy network $\pi(a_{t}|s_{t};\theta)$
        \STATE Execute the action $a_{t}$ and get the next state $s_{t+1}$ and the extrinsic reward $r^{\text{ex}}_{t}$
		\STATE Calculate the kernel matrix: $\textbf{L}_{y}=\mathbf{M}^{T}\mathbf{M}$ using $\phi(s_{t},s_{t+1};\eta)$
		\STATE Compute the intrinsic reward: $r_{t}^{\text{in}} = \det\left(\mathbf{L}_{y} \odot \left(\diag{(\mathbf{L}_{y}})^{T}\diag({\mathbf{L}_{y}})\right)^{-\frac{1}{2}}\right)$
		\STATE Store the transition $(s_{t}, r^{\text{in}}_{t}, r^{\text{ex}}_{t}, s_{t+1}, a_{t})$ in $\mathcal{B}$
		\ENDFOR
		\STATE Compute the diversity-augmented return $G^{\text{mix}}(s, a)$ using the mixed reward: $r^{\text{mix}}=\lambda^{\text{ex}}r^{\text{ex}} + \lambda^{\text{in}}r^{\text{in}}$
		\STATE Compute the extrinsic return $G^{\text{ex}}(s, a)$ using the extrinsic reward: $r^{\text{ex}}$
		\STATE Update $\theta$ according to Eq.\eqref{eq:mix_cost} with learning rate $\alpha$
		\STATE\td{Update $\psi$ according to $\mathcal{L}_{value}=\mathbb{E}[(V(s_{t};\psi)-G^{\text{mix}}_{t})^2]$ with learning rate $\alpha$}
		\STATE Update $\eta$ according to Eq.\eqref{eq:chain_rule} and Eq.\eqref{eq:importance_sampling} with learning rate $\beta$
        \ENDFOR
	\end{algorithmic}
	\label{algo:vi}
\end{algorithm}

\subsection{Parameterised Diversity as Intrinsic Reward}
We are now ready to parameterise the intrinsic reward with parameter $\eta$ in this section. The intrinsic reward at timestep $t$ is defined as:
\begin{equation}
    r^{\text{in}}(s_t, s_{t+1}; \eta) = \det\left(\mathbf{L}_{{y}=\{s_{t},s_{t+1}\}}\right),
\label{eq:intrinsic}
\end{equation}
where $\mathbf{L}_{y}$ is the kernel matrix of the subset $y=\{s_{t}, s_{t+1}\}$ with:
\begin{equation}
    \mathbf{L}_{y} = \mathbf{M}^{T}\mathbf{M},
\label{eq:kernel_matrix}
\end{equation}
and $\mathbf{M} = [\phi(s_{t};\eta),\phi(s_{t+1};\eta)]$ and $\phi(\cdot;\eta)$ is the feature vector parameterised by the intrinsic parameter $\eta$.
$r_{t}^{\text{in}}$ measures the diversity of the adjacent state sequence ${y}=\{s_{t},s_{t+1}\}$. Intuitively, it represents the diversity that an action, $a_t$, will bring to the trajectory. In order to scale the intrinsic reward, $r^{\text{in}}_{t}$, in a reasonable range ({e.g.}, [0, 1]), the kernel matrix $\mathbf{L}_{y}$ can be normalised by:
\begin{equation}
    \mathbf{L}_{y}^{\prime} = \mathbf{L}_{y} \odot \left(\diag{(\mathbf{L}_{y}})^{T}\diag({\mathbf{L}_{y}})\right)^{-\frac{1}{2}},
\end{equation}
where $\odot$ is the element-wise multiplication and $\diag(\cdot)$ outputs the diagonal elements of $\mathbf{L}_{y}$ as a vector.

Figure~\ref{fig:illustration} illustrates the proposed intrinsic reward. When the agent is moving ahead, the intrinsic rewards are \td{relatively higher}. In contrast, when the agent is stuck at the bottom of the flagpole, the intrinsic rewards are close to zero.

\section{Experiments}
In this section, we evaluate $\ourmethod$ with different selections of baseline policy gradient architecture, namely, proximal policy optimisation (PPO) and advantage actor critic (A2C). We choose MuJoCo locomotion and the arcade learning environments (Atari games and SuperMarioBro) as the benchmark environments (see Figure~\ref{fig:ch5_envs}). 
\begin{figure}[h]
    \begin{subfigure}{\textwidth}
        \centering
        \includegraphics[width=0.9\textwidth]{figures/chapter5/mujoco_env.pdf}
        \caption{MuJoCo Locomotion Environment.}
        \label{fig:mujoco_env}
    \end{subfigure}
    \begin{subfigure}{\textwidth}
        \centering
        \includegraphics[width=0.9\textwidth]{figures/chapter5/atari_env.pdf}
        \caption{Arcade Learning Environment.}
        \label{fig:ale_env}
    \end{subfigure}
    \caption{Visualisation of (a) MuJoCo Locomotion Environment: HalfCheetah, Hopper, Walker2d, Ant, and Humanoid, and (b) Arcade Learning Environment: Alien, BeamRider, Breakout, MsPacman, UpNDown, and SuperMarioBros.}
    \label{fig:ch5_envs}
\end{figure}

\subsection{Environments}
Firstly, we evaluate our agent in the MuJoCo locomotion environment~\cite{duan2016benchmarking} and use PPO as the base RL algorithm. We use two kinds of environment settings for evaluation: the standard setting and the delayed reward setting~\cite{zheng2018learning}. Specifically, {for the delayed reward setting}, the accumulated reward is only given every 10, 20 or 40 steps. This setting is more difficult for agent learning. The five tasks selected in the experiment are HalfCheetah, Hopper, Walker2d, Ant and Humanoid (see Figure~\ref{fig:mujoco_env}).

Subsequently, to investigate the performance of $\ourmethod$ with high dimensional inputs, we also evaluate our approach on the Atari games and the SuperMarioBros~\cite{gym-super-mario-bros} using A2C as our base RL algorithm. For the Atari games, Alien, BeamRider, Breakout, MsPacman and UpNDown are selected as our test bench (see Figure~\ref{fig:ale_env}).

\subsection{Experimental Settings}
In MuJoCo tasks, baseline methods are vanilla PPO~\cite{schulman2017proximal}, LIRPG~\cite{zheng2018learning} and VIME~\cite{NIPS2016_abd81528}. The intrinsic module $\phi_{\eta}(\cdot)$ has 3 hidden layers with 64 neurons with ReLU activation functions. The output of the intrinsic module is a 64 dimensional embedding feature. Adam~\cite{kingma2014adam} is chosen as the optimiser with learning rate $\alpha=0.0001$. $\lambda^{\text{in}}$ and $\lambda^{\text{ex}}$ are both set to 1 as default. The hyperparameters of PPO part are kept the same as the baseline setting: timesteps per rollout are 2048, clip ratio is 0.2, entropy coefficient is 0, optimisation epochs per iteration is 10, batch size is 64, learning rate is 0.0003, discount factor $\gamma$ is 0.99, and GAE parameter $\lambda$ is 0.95. Training for each agent lasts for 1 million frames. {The average time of one complete training run of DAIM using one random seed on the MuJoCo tasks is 1.46 hours.}

In the Atari games and SuperMarioBros, Vanilla A2C and LIRPG are selected as baseline methods. For $\ourmethod$, the same CNN architecture as A2C is used within the intrinsic module, with a 256-dimensional feature vector as output. RMSprop is selected as the optimiser with learning rate $\alpha=0.0007$. $\lambda^{\text{in}}$ and $\lambda^{\text{ex}}$ are set to 0.01 and 1 respectively. The hyperparameters of the A2C part are kept the same as the baseline setting: number of workers is 16, timesteps per rollout of each worker are 5, entropy coefficient is 0.01, learning rate is 0.0007, discount factor $\gamma$ is 0.99, and GAE parameter $\lambda$ is 0.95. Training on the Atari games and SuperMarioBros lasts for 50 million and 5 million frames respectively. One complete training run of DAIM using one random seed takes 23.96 hours and 2.83 hours on the Atari games and the SuperMarioBros, respectively.

\subsection{Results on MuJoCo Locomotion Environment}
The comparisons between DAIM and baselines are presented in Figure~\ref{fig:mujoco_results} and the related quantitative results are in Table~\ref{tab:standard} and Table~\ref{tab:delay}. In the standard setting, $\ourmethod$ achieves higher rewards than PPO and VIME in 4 out of 5 tasks, and outperforms LIRPG on all tasks. We notice that in the standard setting without delay, LIRPG has similar or lower performance than PPO. This might be because in the standard setting -- which uses dense rewards -- the agent can receive an extrinsic reward at each timestep and LIRPG does not encourage any exploration explicitly. Thus, the learned intrinsic rewards {of LIRPG} provide only minor contributions.

In the delayed reward setting, $\ourmethod$ achieves the best performance in 12 cases out of 15 delayed reward tasks. This indicates that, through the bi-level update, $\ourmethod$ still maximises the accumulated rewards, so the performance does not drop. On the other hand, it also suggests that the diversity-based intrinsic rewards help training in delayed-reward environments. Due to only receiving delayed rewards, the performance of PPO drops with the increase of the delay period. LIRPG and VIME can achieve better results than PPO, because they can provide the intrinsic reward to the agent at each step. The result shows that $\ourmethod$ adopts DPPs as a diversity measure between states, which encourages the agent to explore in the environment even with only delayed rewards.

In addition, we also investigate the performance of $\ourmethod$ under different reward weights, $\lambda^{\text{ex}}=\{0.01, 0\}$. When $\lambda^{\text{ex}}=0$, the policy is updated towards maximising the intrinsic reward only. From Figure~\ref{fig:mujoco_results}, either $\lambda^{\text{ex}}=0$ or $\lambda^{\text{ex}}=0.01$, the results achieved are comparable to $\lambda^{\text{ex}}=1$ in both the standard and delayed reward settings. More specifically, the weight coefficients $\lambda^{\text{ex}}$ and $\lambda^{\text{in}}$ determine which reward plays the major role during training.  When $\lambda^{\text{ex}}=1$, the agent can outperform other baselines in most tasks, especially in the standard and the delayed reward setting with a short delay (Delay = 10). In these settings, the extrinsic rewards can provide feedback frequently which is helpful to the training. But in the tasks with longer delayed feedback ({e.g.}, Delay=20 and Delay=40), the extrinsic rewards only rarely provide useful feedback, and the intrinsic rewards -- which can provide denser feedback -- lend valuable guidance to the agent. Thus, reducing the weighting of the extrinsic rewards can sometimes improve the performance of the agent, though this depends on the dynamics of the particular environment (see Figure~\ref{fig:mujoco_results}).

This also demonstrates that the bi-level update has led to meaningful intrinsic {rewards} that can criticize the policy.
\begin{figure}[t!]
    \centering
    \includegraphics[width=1.0\linewidth]{figures/chapter5/mujoco_result_54_seaborn.pdf}
    \caption{Comparison between $\ourmethod$ and baseline approaches (PPO, LIRPG, and VIME) in the MuJoCo locomotion environment with standard and delayed reward settings. The $x$-axis represents the number of timesteps in training. The $y$-axis represents the average extrinsic rewards over the last 100 training episodes. All of the experiments were run using 5 different seeds. The solid lines and shaded areas in the plots represent the mean and the standard errors over the performances of the different seeds.}
    \label{fig:mujoco_results}
\end{figure}

\begin{table}[h!]
    \centering
    \resizebox{0.9\textwidth}{!}{
    \begin{tabular}{c|ccccc}
    \toprule
     \multirow{2}{*}{Methods} & \multicolumn{5}{c}{Standard} \\
     \cmidrule(lr){2-6}
         & HalfCheetah & Hopper & Walker2d & Ant & Humanoid \\
    \midrule
    PPO & 2039$\pm$354 & \underline{2112$\pm$208} & 2145$\pm$514 & {512$\pm$53} & 566$\pm$28\\
    LIRPG & 1694$\pm$347 & 2004$\pm$125 & 1997$\pm$322 & -25$\pm$3 & 562$\pm$14 \\
    VIME & 2204$\pm$545 & \textbf{2126$\pm$238} & 2323$\pm$562 & \underline{711$\pm$157} & 530$\pm$29 \\
    \midrule
    DAIM & \textbf{2895$\pm$405} & {2103$\pm$129} & \underline{2766$\pm$487} & \textbf{1192$\pm$157} & \textbf{670$\pm$23} \\
    DAIM ($\lambda^{ex}=0$) & 1780$\pm$405 & 2078$\pm$377 & \textbf{2879$\pm$580} & -164$\pm$35& 516$\pm$19\\
    DAIM ($\lambda^{ex}=0.01$) & \underline{2363$\pm$158} & 1290$\pm$442 & 2285$\pm$504 & -84$\pm$12 & \underline{593$\pm$39}\\
    \bottomrule
    \end{tabular}
    }
    \caption{Quantitative results comparison between DAIM and other baseline methods (PPO, LIRPG, and VIME) in the MuJoCo locomotion environment with the standard reward setting. In the table, it shows final average extrinsic rewards $\pm$ standard error across 5 different seeds of last 100 episodes in the training. The best and the second best results are (\textbf{bold}) and (\underline{underline}), respectively.}
    \label{tab:standard}
\end{table}

\begin{table}[h!]
    \centering
    \resizebox{0.9\textwidth}{!}{
    \begin{tabular}{c|ccccc}
    \toprule
     \multirow{2}{*}{Methods} & \multicolumn{5}{c}{ Delay = 10} \\
     \cmidrule(lr){2-6}
      & HalfCheetah & Hopper & Walker2d & Ant & Humanoid \\
    \midrule
    PPO & 1674$\pm$297& 1457$\pm$396 & 1404$\pm$258 & {207$\pm$26} & 503$\pm$9 \\
    LIRPG & 1756$\pm$273 & \underline{2275$\pm$192} & 1279$\pm$215 &-27$\pm$5 & \underline{555$\pm$31} \\
    VIME & \underline{2498$\pm$638} & 992$\pm$232 & 1186$\pm$487 & \underline{249$\pm$182} & 518$\pm$24 \\
    \midrule
    DAIM & \textbf{3114$\pm$72} & \textbf{2288$\pm$181} & \textbf{2824$\pm$248} & \textbf{663$\pm$67} & \textbf{612$\pm$45} \\
    DAIM ($\lambda^{ex}=0$) &1680$\pm$367&1420$\pm$371 &2582$\pm$116 &-97$\pm$30 & 530$\pm$20 \\
    DAIM ($\lambda^{ex}=0.01$) & {1961$\pm$450} & 1495$\pm$368& \underline{2730$\pm$307}& -51$\pm$15 & 535$\pm$24\\
    \midrule
    \midrule
    \multirow{2}{*}{Methods} & \multicolumn{5}{c}{ Delay = 20} \\
     \cmidrule(lr){2-6}
      & HalfCheetah & Hopper & Walker2d & Ant & Humanoid \\
    \midrule
    PPO & 1274$\pm$148 & 1672$\pm$358 & 519$\pm$49 & {45$\pm$11} & 458$\pm$16 \\
    LIRPG & 1790$\pm$317& \underline{1748$\pm$127} & \underline{1835$\pm$313} &-50$\pm$8 & \textbf{568$\pm$20} \\
    VIME & 1948$\pm$308 & 936$\pm$354 & 1368$\pm$379 & \underline{142$\pm$86} & 477$\pm$5 \\
    \midrule
    DAIM & \textbf{2397$\pm$288} & \textbf{1810$\pm$236} & 1792$\pm$483 & \textbf{249$\pm$25}& \underline{545$\pm$24} \\
    DAIM ($\lambda^{ex}=0$) & 1736$\pm$244 & 1734$\pm$505 &1809$\pm$357 & -172$\pm$32 & 539$\pm$17 \\
    DAIM ($\lambda^{ex}=0.01$) & \underline{2350$\pm$413} & 1586$\pm$353 & \textbf{2209$\pm$406} & -40$\pm$4 & 527$\pm$14 \\
    \midrule
    \midrule
    \multirow{2}{*}{Methods} & \multicolumn{5}{c}{ Delay = 40} \\
     \cmidrule(lr){2-6}
      & HalfCheetah & Hopper & Walker2d & Ant & Humanoid \\
    \midrule
    PPO & 931$\pm$113 & \textbf{1590$\pm$300} & 305 $\pm$13 & {-50$\pm$9} & 449$\pm$7 \\
    LIRPG & 941$\pm$130 & \underline{1456$\pm$359}& 1292$\pm$169 & -51$\pm$13 & \textbf{502$\pm$6}\\
    VIME  & 1053$\pm$121 & 1178$\pm$173 & 304$\pm$14 & \underline{-26$\pm$11} & 425$\pm$16 \\
    \midrule
    DAIM & 1201$\pm$126& 879$\pm$166 & 476$\pm$38 & \textbf{-13$\pm$6} & 467$\pm$13\\
    DAIM ($\lambda^{ex}=0$) & \textbf{2034$\pm$283} & 1153$\pm$248 & \textbf{2324$\pm$475} & -208$\pm$24 & 473$\pm$6\\
    DAIM ($\lambda^{ex}=0.01$) & \underline{1597$\pm$410} & 1052$\pm$337 & \underline{1768$\pm$518}& -117$\pm$19 & \underline{488$\pm$7}\\
    \bottomrule
    \end{tabular}}
    \caption{Quantitative results comparison between DAIM and other baseline methods (PPO, LIRPG, and VIME) in the MuJoCo locomotion environment with the delayed reward setting. In the table, it shows final average extrinsic rewards $\pm$ standard error across 5 different seeds of last 100 episodes in the training. The best and the second best results are (\textbf{bold}) and (\underline{underline}), respectively.}
    \label{tab:delay}
\end{table}

% \begin{figure}[t!]
% % \vspace{-1.cm}
%     \centering
%     \includegraphics[width=\linewidth]{figures/chapter5/walker2d_result.pdf}
%     \caption{The performance of different feature embedding methods in the Walker2d task with standard and delayed reward settings. DAIM achieves higher accumulated extrinsic rewards.}
%     \label{fig:walker2d_results}
% \end{figure}
\begin{figure}[h!]
\centering
  \begin{subfigure}[t]{0.49\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter5/embedding/delay1.pdf}
    \caption{Standard}
  \end{subfigure}\hfill
  \begin{subfigure}[t]{0.49\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter5/embedding/delay10.pdf}
    \caption{Delay=10}
  \end{subfigure}\hfill
  \begin{subfigure}[t]{0.49\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter5/embedding/delay20.pdf}
    \caption{Delay=20}
  \end{subfigure}\hfill
  \begin{subfigure}[t]{0.49\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter5/embedding/delay40.pdf}
    \caption{Delay=40}
  \end{subfigure}\hfill
  \caption{The performance of different feature embedding methods in the Walker2d task with standard and delayed reward settings. DAIM achieves higher accumulated extrinsic rewards.} 
  \label{fig:walker2d_results}
\end{figure}
% \begin{figure}[t!]
%     \centering
%     \includegraphics[width=\linewidth]{figures/chapter5/walker2d_r_in_result.pdf}
%     \caption{Using {\em only} intrinsic rewards ($\lambda^{\text{ex}}=0$), we evaluate the performance (accumulated {\it extrinsic} rewards on vertical axes) of different feature embedding methods in all five MuJoCo tasks with the standard reward setting.}
%     \label{fig:walker2d_results_r_in}
% \end{figure}
\begin{figure}[h!]
\centering
  \begin{subfigure}[t]{0.49\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter5/r_in_only/halfcheetah.pdf}
    \caption{HalfCheetah}
  \end{subfigure}\hfill
  \begin{subfigure}[t]{0.49\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter5/r_in_only/hopper.pdf}
    \caption{Hopper}
  \end{subfigure}\hfill
  \begin{subfigure}[t]{0.49\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter5/r_in_only/walker2d.pdf}
    \caption{Walker2d}
  \end{subfigure}\hfill
  \begin{subfigure}[t]{0.49\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter5/r_in_only/ant.pdf}
    \caption{Ant}
  \end{subfigure}\hfill
  \begin{subfigure}[t]{0.49\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter5/r_in_only/humanoid.pdf}
    \caption{Humanoid}
  \end{subfigure}\hfill
  \caption{Using {\em only} intrinsic rewards ($\lambda^{\text{ex}}=0$), we evaluate the performance (accumulated {\it extrinsic} rewards on vertical axes) of different feature embedding methods in all five MuJoCo tasks with the standard reward setting.} 
  \label{fig:walker2d_results_r_in}
\end{figure}

\subsubsection{Feature Embedding Methods Study}
We also investigate the efficiency of the other two feature learning methods as our baselines which are summarized briefly as below.

\noindent\textbf{Inverse Dynamics Features} (\text{Inv-Dyn}) Pathak \textit{et al.}~\cite{pathak2017curiosity} propose the inverse dynamics model in the Intrinsic Curiosity Module (ICM). The inverse dynamics model uses the current state $s_{t}$ and the next state $s_{t+1}$ to predict the action $a_{t}$. The motivation of this method is that the learned features should only depends on the current action of the agent: theoretically, it would not be affected by insignificant changes in the environment. The predicted action $a_{t}$ is estimated by:
\begin{equation}
    \widehat{a_{t}} = g(\phi(s_{t};\eta), \phi(s_{t+1};\eta);\theta_{I}),
\end{equation}
where, $g$ is the inverse dynamics model and $\phi(\cdot;\eta)$ is the learned embedding feature space. The parameters of the embedding network, $\eta$, and the parameters of the inverse dynamic model $\theta_{I}$ are optimised through minimising the error of the predicted action $\widehat{a_{t}}$ and the actual action $a_{t}$:
\begin{equation}
    \min_{\eta, \theta_{I}}\mathcal{L}(\widehat{a_{t}}, a_{t}),
\end{equation}
In the MuJoCo tasks, the action space is continuous and the output of the inverse dynamics model $g$ is the mean $\mu$ and standard deviation $\sigma$ of the Gaussian distribution over the action space. The loss function $\mathcal{L}$ is used to minimise the discrepancy between the predicted action $\widehat{a_{t}}$ and the actual action $a_{t}$ through maximising the log-likelihood estimation of $\theta_{I}$ under the Gaussian distribution.

\noindent\textbf{Raw Observation Features} (\text{Raw-Obs}) We also adopt a simple way to calculate the intrinsic reward as a baseline. The raw observations are used directly, where $\phi(s_{t})=s_{t}$. However, this method is subject to observation noises and cannot scale to high-dimensional inputs.

In Figure~\ref{fig:walker2d_results}, we compare the performance of all feature embedding methods in the Walker2d task with and without delayed rewards. $\ourmethod$ achieves the best results in the standard and delayed reward settings, whereas the other two feature embedding methods have similar performance. It must be emphasised that during the training of the feature embedding module of $\ourmethod$, the {\em intrinsic} network is updated to maximise the {\em extrinsic} returns. In this way, the learned features are likely to help the agent explore in an appropriate direction, which not only increases the diversity between the states but also leads to higher extrinsic returns. For \text{Inv-Dyn} and \text{Raw-Obs}, the dissimilarity between the adjacent states is only used as an intrinsic reward but not attempting to maximise extrinsic returns. In this case, the agent will maximise the diversity regardless of its effect on overall performance in terms of the extrinsic returns.

We also train the agents with only intrinsic rewards, and results are shown in Figure~\ref{fig:walker2d_results_r_in}. $\ourmethod$ outperforms \text{Inv-Dyn} and \text{Raw-Obs}, which shows that higher diversity does not necessarily lead to better performance. In the standard reward setting, $\ourmethod$ yields higher rewards than the other two embedding feature methods in 4 out of 5 tasks. In the Ant task, the agent trained with inverse dynamic features and raw observation features yields decreasing performance. This suggests that the intrinsic reward is not, on its own, sufficient to train the agent. However, the features learned by using bi-level optimisation help to maximise the extrinsic returns, showing better performance than the other two methods.

% \begin{figure}[t!]
%     \centering
%     \includegraphics[width=\linewidth]{figures/chapter5/mujoco_kernel_ablation.pdf}
%     \caption{The performance of using different size $k$ of consecutive states for calculating intrinsic rewards in the Walker2d task with standard and delayed reward settings.  The $y$-axis represents the average extrinsic rewards over the last 100 training episodes.}
%     \label{fig:multi_frames}
% \end{figure}

\begin{figure}[h!]
\centering
  \begin{subfigure}[t]{0.49\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter5/multi_frames/delay1.pdf}
    \caption{Standard}
  \end{subfigure}\hfill
  \begin{subfigure}[t]{0.49\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter5/multi_frames/delay10.pdf}
    \caption{Delay=10}
  \end{subfigure}\hfill
  \begin{subfigure}[t]{0.49\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter5/multi_frames/delay20.pdf}
    \caption{Delay=20}
  \end{subfigure}\hfill
  \begin{subfigure}[t]{0.49\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter5/multi_frames/delay40.pdf}
    \caption{Delay=40}
  \end{subfigure}\hfill
  \caption{The performance of using different size $k$ of consecutive states for calculating intrinsic rewards in the Walker2d task with standard and delayed reward settings.  The $y$-axis represents the average extrinsic rewards over the last 100 training episodes.} 
  \label{fig:multi_frames}
\end{figure}

\subsubsection{Ablation Study: Length of State Sequence}
In this section, we further investigate the performance of $\ourmethod$ using different lengths of consecutive states to calculate the intrinsic reward during training. In this case, Equation~\eqref{eq:intrinsic} can be rewritten as:
\begin{equation}
    r^{\text{in}}(s_{t:t+k-1}; \eta) = \det\left(L_{\bm{y}=\{s_{t:t+k-1}\}}\right),
\label{eq:intrinsic_multi_frame}
\end{equation}
where, $k$ is the size of consecutive states for calculating the intrinsic reward.

In Figure~\ref{fig:multi_frames}, we report on the performance of using different sizes of $k=\{2, 4, 8\}$ in calculating intrinsic rewards.  Specifically, $k=2$ represents the default setting of $\ourmethod$ which uses adjacent states for computing intrinsic rewards. From the plot, in the standard reward setting, $\ourmethod$ with different sizes $k$ achieves similar performance. In the delayed reward setting, when $k=8$, performance is worse, indicating that larger spans of consecutive states may not benefit the training. When $k=2$ or $k=4$, DAIM achieves comparable results. For Delay = 40, both $k=2$ and $k=4$ provide improvements over baselines PPO and VIME (see Figure~\ref{fig:mujoco_results}), with $k=2$ conveying the best advantage. \td{This may be due to the fact that two adjacent states can convey information about the direction and speed of the movement.} Therefore, $k=2$ was selected as the default setting for our experiments.

% \begin{figure}[t!]
%     \centering
%     \includegraphics[width=\linewidth]{figures/chapter5/atari_mario_result.pdf}
%     \caption{Comparison between $\ourmethod$ and baseline approaches (A2C and LIRPG) in the Atari and SuperMarioBros games. The $x$-axis represents the number of timesteps in the training. The $y$-axis represents the average extrinsic rewards over the last 100  training episodes. All experiments were run using 5 different seeds. The solid lines and shaded areas in the plots represent the mean values and the standard errors, respectively, over the different seeds.}
%     \label{fig:atari_result}
% \end{figure}
\begin{figure}[h!]
\centering
  \begin{subfigure}[t]{0.49\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter5/atari_exp/alien.pdf}
    \caption{Alien}
  \end{subfigure}\hfill
  \begin{subfigure}[t]{0.49\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter5/atari_exp/beamrider.pdf}
    \caption{BeamRider}
  \end{subfigure}\hfill
  \begin{subfigure}[t]{0.49\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter5/atari_exp/breakout.pdf}
    \caption{Breakout}
  \end{subfigure}\hfill
  \begin{subfigure}[t]{0.49\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter5/atari_exp/mspacman.pdf}
    \caption{MsPacman}
  \end{subfigure}\hfill
  \begin{subfigure}[t]{0.49\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter5/atari_exp/upndown.pdf}
    \caption{UpNDown}
  \end{subfigure}\hfill
  \begin{subfigure}[t]{0.49\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter5/atari_exp/SuperMarioBros.pdf}
    \caption{SuperMarioBros}
  \end{subfigure}\hfill
  \caption{Comparison between $\ourmethod$ and baseline approaches (A2C and LIRPG) in the Atari and SuperMarioBros games. The $x$-axis represents the number of timesteps in the training. The $y$-axis represents the average extrinsic rewards over the last 100  training episodes. All experiments were run using 5 different seeds. The solid lines and shaded areas in the plots represent the mean values and the standard errors, respectively, over the different seeds.} 
  \label{fig:atari_result}
\end{figure}

% another figure
\subsection{Results on Arcade Learning Environment}
%In both Atari games and SuperMarioBros, the hyper-parameters are set to be the same to the default A2C algorithm.

%In this experiment, Vanilla A2C and LIRPG are selected as baseline methods. For $\ourmethod$, the same CNN architecture as A2C is used upon the intrinsic module, with a 256-dimensional feature vector as output. RMSprop is selected as the optimizer with learning rate $\alpha=0.0007$. $\lambda^{\text{in}}$ and $\lambda^{\text{ex}}$ are set to 0.01 and 1 respectively. The hyper-parameters of the A2C part are kept the same to the baseline setting: number of workers is 16, timesteps per rollout of each worker are 5, entropy coefficient is 0.01, learning rate is 0.0007, discount factor $\gamma$ is 0.99, and GAE parameter $\lambda$ is 0.95. Training on the Atari games and SuperMarioBros lasts for 50 million and 5 million frames respectively. One complete training run of DAIM using one random seed takes 23.96 hours and 2.83 hours on the Atari tasks and SuperMarioBros, respectively.
Figure~\ref{fig:atari_result} and Table~\ref{tab:atari} show the performance of $\ourmethod$ against baselines. In the experiments, $\ourmethod$ has faster convergence in 5 out of 6 games (Alien, BeamRider, Breakout, MsPacman and SuperMarioBros) and achieves comparable or better results than baseline methods towards the end. In the experiments of BeamRider, Breakout, MsPacman and SuperMarioBros, $\ourmethod$ demonstrates significant improvement of learning speed in the initial stages of training. In the SuperMarioBros, $\ourmethod$ displays better performance compared to the two baseline techniques.

The explanation for the observations above is that the intrinsic rewards based on diversity can be helpful in the exploration stage of training. Thus, with the assistance of the intrinsic reward, $\ourmethod$ helps the agent explore novel states at the start of training, and achieves better or comparable performance at the end of training.

\begin{table}[h]
    \centering
\resizebox{\textwidth}{!}{
    \begin{tabular}{c|cccccc}
    \toprule
        Methods & Alien & BeamRider & Breakout & MsPacman & UpNDown & SuperMarioBro \\
        \midrule
         A2C & 1698$\pm$129 & 8380$\pm$205& \textbf{567$\pm$10} & 2739$\pm$221 & \textbf{147069$\pm$3923} & 175$\pm$6\\
         LIRPG & \textbf{2429$\pm$379} & \textbf{9183$\pm$205} & \underline{566$\pm$8} & \underline{2405$\pm$77} & 75088$\pm$9299 & \underline{182$\pm$10}\\
         \midrule
         DAIM & \underline{2132$\pm$341} & \underline{9001$\pm$131} & 544$\pm$23 & \textbf{3874$\pm$542} & \underline{98683$\pm$28430} & \textbf{198$\pm$4}\\
         \bottomrule
    \end{tabular}
    }
    \caption{Quantitative results comparison between DAIM and other baseline methods (A2C and LIRPG) in the Atari and SuperMarioBro games. In the table, it shows final average extrinsic rewards $\pm$ standard error across 5 different seeds of last 100 episodes in the training. The best and the second best results are (\textbf{bold}) and (\underline{underline}), respectively.}
    \label{tab:atari}
\end{table}

\section{Discussion}
In this chapter, the DPPs framework is employed to measure the diversity of state sequences; it is shown this can be used to create an intrinsic reward that encourages state-space exploration in on-policy learning; we term this variant diversity-augmented intrinsic motivation (DAIM).

The update of the proposed intrinsic reward module is formulated as a bi-level optimisation problem. During training, as the agent tries to maximise intrinsic returns, the bi-level optimisation guarantees that extrinsic returns are also maximised.

Based on the extensive experiments, DAIM achieves improvements in continuous control in delayed-reward problems when compared to well-known on-policy training approaches of similar complexity, such as PPO~\cite{schulman2017proximal}, LIRPG~\cite{zheng2018learning} and VIME~\cite{NIPS2016_abd81528}.

DAIM also accelerates the convergence of training agents on ALE compared to the baseline training approaches. In the work of this paper, DAIM is combined with PPO and A2C; in principle, DAIM could also be applied to other on-policy RL algorithms, such as TRPO~\cite{schulman2015trust} or SVPG~\cite{liu2017stein}.

The wider implications of these experiments should be assessed against off-policy training approaches, e.g., the DQN~\cite{mnih2015human} for discrete control or DDPG~\cite{lillicrap2015continuous} for continuous control, which are considered more sample efficient than on-policy approaches. The advantage of on-policy RL algorithms is generally that they are more stable in training and are more robust to hyperparameter settings.

In the future, we will investigate how to extend the DAIM approach to off-policy sampling and address more challenging tasks, such as Montezuma's Revenge.
%with even sparser reward structure

### 2022-04-23 chapters/chapter_tmp.tex
update

\chapter{Interpretability of DRL}
\label{ch:drl_interp}
\section{Introduction}
In previous chapters, we have discussed several methods to improve sample efficiency in deep reinforcement learning (DRL), especially in complex robot manipulation tasks with continuous control. However, all of the above experiments are carried out in a simulated environment. Training becomes much more problematic when applying DRL to real-world applications. For example, in a robot manipulation task, which requires an agent to pick up a cube and place it in a target position, computer vision techniques, such as object detection, pose estimation, or semantic segmentation, are used to estimate the position and orientation of the cube. Then, at each timestep, the current information of the cube and the proprioceptive data of the robot arm are provided to the RL agent to select an action.

To reduce \td{real-world training times}, a number of end-to-end approaches~\cite{gu2017deep,levine2016end,zhu2017target} have been proposed in recent years, which enable the agent to infer the policy according to raw observations (e.g., images) directly. However, such an end-to-end training setup would require the agent to collect samples for a long time in a real-world environment to update the policy network, which may increase the risk of robot/hardware failure and thus elevate the cost of training significantly. An emerging solution is to have the agent trained in a simulated environment before transferring the policy to the real world. In this period, the agent also needs to overcome the ``reality gap"~\cite{jakobi1995noise} between the simulated environment and the real-world environment, and domain randomisation (DR)~\cite{andrychowicz2018learning,james2017transferring,peng2018sim,sadeghi2017cad2rl,tobin2017domain} is a widely used approach for this purpose. When using DR in training, the elements of the simulated environment are varied to some extent, from the visual appearance to the environment dynamics (see Figure~\ref{fig:dr_example}), in order to improve the policy's robustness to various changes in the environment.
\begin{figure}[h!]
  \centering
  \includegraphics[width=\linewidth]{figures/chapter6/domain_random.png}
  \caption{Examples of domain randomisation (visual appearances) in the simulated Fetch (top) and Jaco (bottom) robots environments.}
  \label{fig:dr_example}
\end{figure}

However, such an end-to-end training paradigm, which could involve inferring actions based on image inputs, makes the theoretical analysis of both the environment and the agents that act within it practically infeasible. Specifically, deep neural networks (NNs) trained with DRL form a direct ``black box" mapping from observation to action, presenting the challenge of understanding learned control strategies. If we would like to deploy such policies---particularly on robots in the real world---we would also want to be able to \emph{explain} them~\cite{arrieta2020explainable}. An interpretation of the inference of a model can be used not only as a way of explaining the behaviour of the model, but can also be used to characterise other properties, such as safety, fairness, and reliability ~\cite{doshi2017towards}. Although there are many existing approaches to explain NNs~\cite{guidotti2018survey}, each of these methods is subjective to varying degrees.
\begin{figure}[h!]
  \centering
  \includegraphics[width=\linewidth]{figures/chapter6/strategies.png}
  \caption{The range of strategies an agent may learn to use when trained to reach a red target, which we split into three components. Firstly, the agent must visually localise the target---which can be accomplished by detecting red, spherical objects, or simply anything with a significant red component. Secondly, the agent can use vision and/or proprioception to guide its arm to the target. Finally, the agent may accomplish the task with varying levels of robustness to changes in the rest of the visual scene. Using a range of analyses, we show that agents trained to accomplish the same task learn different subsets of these strategies, depending on their training conditions.}
  \label{fig:strategies}
\end{figure}

In this chapter, we aim to understand \td{how DRL agents make decision in the environment} and understand how agents trained with DR behave differently from agents trained without. To this end, we train DRL agents under a range of different settings and use relative differences to better interpret their policies. In contrast to most prior work examining DRL policies~\cite{atrey2020exploratory,greydanus2018visualizing, olson2021counterfactual,puri2020explain,rupprecht2020finding,such2018atari,zahavy2016graying}, we not only provide explanations for how policies react to parts of states/individual states, but use our array of experiments to characterise their \emph{global} strategies (see Figure~\ref{fig:strategies}). \td{Through these experiments and analysis, we are able to address following questions:
\begin{itemize}
    \item What is the role of DR on layer learning?
    \item To what extent is visual inputs rather than proprioceptive inputs used by the agent?
    \item How robust are agents to unseen scenarios?
    \item How redundant is the network in its representation? 
\end{itemize}
}
%For instance, we examine the relative importance of different input modalities, and whether agents need their memory to complete tasks.

In particular, we train different robots with vision-based inputs to perform the task of reaching for a (red) target, and then use different techniques to understand what features of the environment the policies react to. Under certain training settings, the agent may localise the target by simply looking for anything red, and it is only under more challenging training conditions that the agent utilises both colour and shape. Clearly, the latter strategy better represents what we want the agent to learn, while the former is merely a ``shortcut'' \cite{geirhos2020shortcut}.

To be noticed, the structure in this chapter is organised in a different way from the previous ones. In Section~\ref{ch6:sec_interp}, we first introduce the interpretability techniques used in this work. In Section~\ref{ch6:exp}, we detail our customised simulated environments, the network structure of DRL agents, training details, and the settings of test scenarios. Subsequently, in Section~\ref{ch6:results}, we conduct an exhaustive analysis on the trained models using a broad suite of interpretability techniques. Finally, in Section~\ref{ch6:discussion}, we discuss our findings and provide recommendations for applying interpretability techniques to DRL agents.

\section{Interpretability Techniques}
Neural network (NN) has achieved great success in various fields, benefiting from its powerful feature representation capabilities. However, compared with hand-crafted feature descriptors, NN lacks interpretability, and it treats the training as black boxes. Therefore, it is important that we are able to explain and understand what the NN has learned before deploying it in real-world applications, where safety and reliability are crucial. In this work, we focus on understanding the NN in a DRL training paradigm. To do so, various interpretability methods have been considered, which include saliency maps, unit ablations, layer re-initialisation and recurrent ablations. 
\label{ch6:sec_interp}
\subsection{Saliency Maps}
\begin{figure}[h!]
  \centering
  \begin{subfigure}{0.24\columnwidth}
    \includegraphics[width=\linewidth]{figures/chapter6/occ_jaco_baseline}
    \subcaption{Occ. (Gaussian)}
  \end{subfigure}
  \begin{subfigure}{0.24\columnwidth}
    \includegraphics[width=\linewidth]{figures/chapter6/average_map_jaco_large.png}
    \subcaption{Avg. $\mathbf{x}^{base}$}
  \end{subfigure}
  \begin{subfigure}{0.24\columnwidth}
    \includegraphics[width=\linewidth]{figures/chapter6/occ_jaco_average_map.png}
    \subcaption{Occ. (avg. $\mathbf{x}^{base}$)}
  \end{subfigure}
  \begin{subfigure}{0.24\columnwidth}
    \includegraphics[width=\linewidth]{figures/chapter6/average_map_mask.png}
    \subcaption{Avg. $\mathbf{x}^{base}$ mask}
  \end{subfigure}
  \caption{Occlusion-based saliency map method (a) using Gaussian mask (c) our improved method using an average image (b, d). Occlusion-based method using Gaussian blur masks (a) shows extra artifacts. Our improved saliency map method highlights important regions in the result by using average baseline image created from plentiful trajectories (c), (d) shows the usage of the average baseline image.}
  \label{fig:saliency_baseline}
\end{figure}

Saliency maps are widely used to understand and explain the output of NNs. Similarly to previous work on understanding DRL agents~\cite{greydanus2018visualizing}, we also use an occlusion-based approach for interpreting. The occlusion-based method~\cite{zeiler2014visualizing} generally adopts a gray mask to scan over the entire input image, and then measures the variation of the output in each masked region to identify which areas have significant influences on the output of the NN. The variation of the performance is also called saliency value, and it can be expressed as:
\begin{align}
S_{m, n} = \lVert F(\mathbf{x}) - F(\mathbf{x}_{m, n}^{occ}) \rVert_2,
\end{align}
where $(m, n)$ is the location on the image, $F(\mathbf{x})$ is the output of the original input, and $F(\mathbf{x}_{m, n}^{occ})$ is the output of the input with an occlusion mask at location $(m, n)$. In the work of~\cite{greydanus2018visualizing}, which aims to understand DRL agents in Atari games, the gray mask is replaced with a Gaussian mask. Because the agent only receives gray scale images in the training and testing. Thus, the gray mask can be recognised as a part of the gray object, and the Gaussian blur mask can introduce spatial uncertainty to find informative regions.

However, the agent receives an RGB image as the input in our task. As a result, the occlusion-based method using Gaussian blur masks cannot highlight any important regions (see Figure~\ref{fig:saliency_baseline}a), and thus fails to provide any valuable information to understand the agent. Atrey \textit{et al.}~\cite{atrey2020exploratory} identify this issue of applying saliency methods to DRL agents as modifying observations in a manner that is incongruent with the true environment's state and generative process, and proposed intervening directly in the environment to employ counterfactual methods. However, in general, this level of control over the environment is not easy. In this work, we are more interested in task-specific objects in the environment, such as the robot arm and the red ball. To avoid interference from the occlusion-based masks on unimportant areas, we first average the frames of approximately 1,000 trajectories to get a baseline image (see Figure~\ref{fig:saliency_baseline} (b)). Then, we use this baseline image to generate the mask in the corresponding locations (see Figure~\ref{fig:saliency_baseline}d). Finally, we can get an improved saliency map, which highlights the informative regions in the input image  (see Figure~\ref{fig:saliency_baseline}c).
%Inspired by the work which use reference images for the calculation of interpreta

Another recent study introduced a more robust DRL-specific saliency method~\cite{puri2020explain}, but it is only applicable to agents which learn a state-action value function over discrete action spaces.
%\subsection{Statistical and Structural Weight Characterisations}
\subsection{Unit Ablations}
Unit ablations evaluate the importance of a single filter in the convolutional layer by masking it to zero and examining the variations in the performance. In general, an important filter can lead to a significant performance drop. While more sophisticated unit ablations have been applied in DRL~\cite{meyes2020you}, {this has only been in the context of a control task with a low-dimensional symbolic state space}.
% Layer initialisation
\subsection{Layer Re-initialisation}
Layer re-initialisation~\cite{zhang2019all} is an extension of unit ablations, which is proposed to test the re-initialisation robustness of the trained neural network. The primary idea is to replace the parameters of a single fully trained layer with the parameters in the previous training epochs, and observe the effect on the performance of the network. In this work, the parameters of each layer can be expressed as: $\{\theta_{1}, \theta_{2}, ..., \theta_{L}\}$, where $L$ is the number of layers of the network, and $\theta_{l}^{t}$ represents the parameters of the $l^\text{th}$ layer in the training epoch $t\in [1,T]$. For example, when training is finished, to evaluate the robustness of a single layer $l$ at training epoch $t$, we can re-initialise the parameters in the layer by: $\theta_{l}^{T} \leftarrow \theta_{l}^{t}$. Furthermore, we can also measure the changes of parameters in the $\ell_{\infty}$ and $\ell_{2}$-norm. Layer re-initialisation allows us to identify the critical layers in the network structure, and these layers are greatly affected by the parameters re-initialisation. Furthermore, with a similar capacity of network, we can also investigate the task complexity by using re-initialisation robustness.

\subsection{Recurrent Ablation}
Recurrent ablation is used to evaluate if the recurrent layer (e.g., long short-term memory (LSTM)) is utilised in a network structure. To do so, we set the hidden state to a constant value. If the performance of the agent is decreased to some extent, it means the recurrent layer helps to perform the task, otherwise, the recurrent layer is redundant in this structure. While it is possible to investigate the hidden states of DRL agents over time, visualising and inspecting a high-dimensional time series can be difficult, even for domain experts~\cite{jaunet2020drlviz}.

\section{Experiments}
\label{ch6:exp}
\subsection{Environments}
In the experiment, we create two simulated environments with different robot arms (see Figure~\ref{fig:robot_setups}): 1) Fetch robot arm and 2) Kinova Jaco robot arm, to accomplish a target reaching task, where the agent needs to control the robot arm's gripper to reach a specific object (e.g., the red ball) in each episode by using visual RGB inputs with or without an extra proprioceptive input (e.g., joint position, joint velocity, and joint angles). Furthermore, these two robot arms have different appearances and control schemes, which provide us with enriched experimental conditions.
These robot environments are commonly used within the DRL literature~\cite{andrychowicz2017hindsight, gu2017deep, rusu2017sim} and, therefore, adopting these in our experiments enables both comparisons to prior work and applicability to the DRL field.

\begin{figure}[h!]
  \centering
  \begin{subfigure}{0.24\columnwidth}
    \includegraphics[width=\linewidth]{figures/chapter6/fetch_vol_resize.png}
    \subcaption{Fetch enviroment}
    \label{fig:fetch_volumn}
  \end{subfigure}
  \begin{subfigure}{0.24\columnwidth}
    \includegraphics[width=\linewidth]{figures/chapter6/fetch_view.png}
    \subcaption{Fetch visual input}
    \label{fig:fetch_view}
  \end{subfigure}
  \begin{subfigure}{0.24\columnwidth}
    \includegraphics[width=\linewidth]{figures/chapter6/jaco_vol_resize.png}
    \subcaption{Jaco environment}
    \label{fig:jaco_volumn}
  \end{subfigure}
  \begin{subfigure}{0.24\columnwidth}
    \includegraphics[width=\linewidth]{figures/chapter6/jaco_top.png}
    \subcaption{Jaco visual input}
    \label{fig:jaco_view}
  \end{subfigure}
  \caption{(a) Fetch environment, (b) an example of Fetch visual input, (c) Jaco environment, and (d) an example of Jaco visual input. \td{The blue cuboids indicate the sampling space of the target objects.}}
  \label{fig:robot_setups}
\end{figure}

The Fetch robot environment includes a 7 degrees-of-freedom (DOF) robot arm, and it is modified based on the model provided in OpenAI Gym~\cite{plappert2018multi}. In our Fetch environment, we revise the original proprioceptive input, where we delete the velocity and position of the gripper and the position of the target. It now only contains the position and velocity of 7 joints of the robot arm. In order to acquire visual inputs, we also add an extra RGB camera in front of the robot arm (see Figure~\ref{fig:fetch_view}). The control scheme remains the same, which is the relative position of the gripper in x, y and z axis. The target object is placed on the table. During training, the position of the target object is randomly sampled in the 2D space (see blue cuboid in Figure~\ref{fig:fetch_volumn}) of the table surface. During testing, 80 fixed uniformly distributed positions are used to evaluate the performance of the agent.

The Jaco robot environment includes a 6 DoF robot arm (the gripper is disabled). The proprioceptive input contains the position and velocity of 6 joints. An RGB camera is placed on the top of the robot arm to provide visual inputs (see Figure~\ref{fig:jaco_view}). The action space is the 6 joint velocities of the robot arm. During training, the position of the target object is randomly sampled in a 3D space (see blue cuboid in Figure~\ref{fig:jaco_volumn}). During testing, 250 uniformly distributed fixed positions are used to evaluate the performance of the agent.

The maximum step of each episode is up to 100 in training, which is used to encourage the agent to have enough exploration in the environment. The maximum step of each episode is limited to 20 in testing, which is used to avoid random actions to reach the target. The reward function is similar to Equation~\eqref{eq:sparse_reward}, the agent will achieve a +1 reward if the distance between the gripper and the target is smaller than a threshold $\epsilon$ (in this chapter, $\epsilon=$10cm), otherwise, it only gets 0. More details of these two environments are provided in Table~\ref{tbl:exp_setup}.

\begin{table}[h!]
  \centering
  \begin{tabular}{c|cc}
    \toprule
    Setting & Fetch & Jaco \\
    \midrule
    Active (Total) DoF & 7 (8) & 6 (9) \\
    Target Range & $21 \times 31 \text{cm}^2$ & $40 \times 40 \times 40 \text{cm}^3$ \\
    Num. Test Targets & 80 & 250 \\
    Vision Input & $3 \times 64 \times 64$ & $3 \times 64 \times 64$ \\
    Proprioceptive Inputs & 30 & 18 \\
    Control Type & Position & Velocity \\
    Num. Actions & 3 & 6 \\
    Action Discretisation & 5 & 5 \\
    Control Frequency & 6.67Hz & 6.67Hz \\
    \bottomrule
  \end{tabular}
  \caption{The details of the experimental setup of Fetch and Jaco environments.}
  \label{tbl:exp_setup}
\end{table}

\subsection{Training Configurations}
\begin{figure}[h!]
  \centering
  \includegraphics[width=0.85\linewidth]{figures/chapter6/axes.png}
  \caption{The 8 different experimental conditions: 2 robots (Fetch vs. Jaco), 2 input modalities (vision vs. vision + proprioception), 2 visual conditions (DR vs. no DR).}
  \label{fig:axes}
\end{figure}
We setup 8 different configurations to cover different possible training strategies and provide a comprehensive comparison. There include different combinations over three primitives: 1) using Fetch robot arm or Jaco robot arm, 2) using visual input only or visual input + proprioceptive input, and 3) training with or without domain randomisation (DR) (see Figure~\ref{fig:axes}).

{The two robot arms have different morphologies and control schemes. In addition,  we also set different target spaces (2D and 3D) to control the difficulties of the tasks. Thus, we can obtain some indication on whether the results and conclusions are generalised.}

For real-world deployment, one would want to use as many sensors as it is feasible to use. However, we are interested in how the absence of specific input modalities can affect learning, as this changes how the agent is able to extract information from the environment. For example, while one might expect robots equipped with proprioceptive sensors to use these alone for pose estimation, agents can also learn to utilise vision, if extracting a pose from images is not too difficult.

Finally, analogously to how \emph{data augmentation}~\cite{lecun1998gradient,shorten2019survey} can be used to improve the generalisation of models in supervised learning settings, \emph{domain randomisation} (Figure \ref{fig:dr_example}) is a common training paradigm in ``sim2real'' approaches for learning real-world robots control~\cite{andrychowicz2018learning,james2017transferring,peng2018sim,sadeghi2017cad2rl,tobin2017domain}. In DR, various properties of the simulation are varied, altering anything from the positions or dynamical properties of objects to their visual appearance, resulting in an expansion of the training domain. We therefore test how the agents trained with or without DR differ. Supporting previous results, we show that agents trained with DR are more robust to out-of-distribution (OoD) perturbations. In our experiments, we use a standard visual DR setup. At each timestep, we randomise all of the components in the environment by changing/adding different textures, colour gradients, RGB colours, and noise, except for the target object. More details about the DR textures are described in Table~\ref{tbl:random_summary} and visualised in Figure~\ref{fig:random_vis}.

%\subsection{Domain Randomisation}
\begin{table}[h!]
  \centering
  \resizebox{\linewidth}{!}{
  \begin{tabular}{c|c|c}
    \toprule
    No. & Option & Description \\
    \midrule
   1 & checkerboard & Render checkerboard pattern with two random RGB colours\\
   2 & gradient & Render gradient in vertical/horizontal direction with two random RGB colours\\
   3 & flat\_rgb & Render single random RGB colour\\
   4 & noise & Render artificial noise with two random RGB colours\\
    \bottomrule
  \end{tabular}}
   \caption{DR textures used during training. At every environment timestep, for each component (e.g., robot arm, skybox, table; the Jaco robot has 13 components, and the Fetch robot has 19 components), the appearance of the component is rendered using a sequential sampling process. In the first step of the sampling process, one of each of the 4 possible choices is made with the same probability, then the subsequent appearance of each component is determined by a second draw, described in the rightmost column of the table. Each RGB channel is drawn uniformly from $\{0 \ldots 255\}$. The details of each option are described in Table~\ref{tbl:texture_expression}.}
\label{tbl:random_summary}
\end{table}

\begin{table}[h!]
  \centering
  \resizebox{\linewidth}{!}{
  \begin{tabular}{c|c|c}
    \toprule
    No. & Option & Expression \\
    \midrule
   1 & checkerboard & texture\_mat[x, y] = rgb1 \textbf{if} (x + y) $\%$ 2 == 0 \textbf{else} rgb2\\
   2 & gradient & texture\_mat[x, y] = (1 - p) $\cdot$ rgb1 + p $\cdot$ rgb2, vertical: p = y / (height - 1), horizontal: p = x / (width - 1)\\
   3 & flat\_rgb & texture\_mat[x, y] = rgb1\\
   4 & noise & texture\_mat[x, y] = rgb1 \textbf{if} p $>$ 0.9 \textbf{else} rgb2, p $\sim$ U(0, 1)\\
    \bottomrule
  \end{tabular}
  }
\caption{Python pseudocode describing the DR texture generators. rgb1 and rgb2 are random RGB colours; the value of each RGB channel is drawn uniformly from $\{0 \ldots 255\}$.}
\label{tbl:texture_expression}
\end{table}

\begin{figure}[h!]
  \centering
  \begin{subfigure}{0.24\columnwidth}
    \includegraphics[width=\linewidth]{figures/chapter6/rand_checker.png}
    \subcaption{checkerboard}
  \end{subfigure}
  \begin{subfigure}{0.24\columnwidth}
    \includegraphics[width=\linewidth]{figures/chapter6/rand_gradient.png}
    \subcaption{gradient}
  \end{subfigure}
  \begin{subfigure}{0.24\columnwidth}
    \includegraphics[width=\linewidth]{figures/chapter6/rand_rgb.png}
    \subcaption{flat\_rgb}
  \end{subfigure}
  \begin{subfigure}{0.24\columnwidth}
    \includegraphics[width=\linewidth]{figures/chapter6/rand_noise.png}
    \subcaption{noise}
  \end{subfigure}
  \caption{Visualisation of different DR textures used during training: (a) checkerboard, (b) gradient, (c) flat\_rgb, (d) noise.}
  \label{fig:random_vis}
\end{figure}

\subsection{Network Structure and Training Details}
\begin{figure}[h!]
  \centering
  \includegraphics[width=\linewidth]{figures/chapter6/network-arch-td.pdf}
  \caption{The overview of the network architecture used in the experiments.}
  \label{fig:network}
\end{figure}
The network architecture follows the same actor-critic style used in the work of Rusu \textit{et al.}~\cite{rusu2017sim} (see Figure~\ref{fig:network}). The visual input is first passed through 2 convolutional layers, and then the features of the last convolutional layer are flattened into 1152-dimensional vector. Subsequently, if proprioceptive input is provided, it will be concatenated with the previous flatten vector and then sent into a fully connected layer. This is followed by a long short-term memory (LSTM) layer. Finally, the output of the LSTM layer will be sent to two independent fully connected layers to predict the policy $\pi(a_{t}|s_{t};\theta)$ and state value $V(s_{t};\theta)$, respectively.

In the experiment, proximal policy optimisation (PPO) is used as the base RL algorithm, and 4 MPI workers are used to parallelise the training. During training, each MPI worker uses 8 threads to collect experience for 128 timesteps per epoch. PPO update per epoch is 4 with a batch size equals to 256. Adam is selected as the optimiser, and learning rate is 0.00025. Clip ratio $\epsilon$ is 0.1. Discount factor $\gamma$ is 0.99. A complete training for one model lasts 5000 epochs, and it needs to run on a GTX-1080Ti GPU for one day. For each training configuration, we train the models with 5 different seeds. The overall setup is detailed in Algorithm~\ref{alg:overall_algo}. The code and pretrained models used in this paper are available at: \url{https://github.com/TianhongDai/domain-rand-interp}.

\begin{algorithm}[h!]
\caption{PPO + DR training.}
\label{alg:overall_algo}
\begin{algorithmic}
\REQUIRE environment $E$, actor network $\pi(a|s;\theta)$, critic network $V_{\pi}(s;\theta)$, number of training epochs $N$, number of timesteps to collect data $T$, number of PPO updates $M$, on-policy transition memory $\mathcal{B}$
\FOR{epoch $= 1, 2, \ldots, N$}
  \STATE $\mathcal{B} \gets \varnothing$
  \FOR{$t = 1, 2, \ldots, T$}
    \IF{$E$ has terminated}
      \STATE Reset $E$ and receive DR \indent\indent state $s_0$
    \ENDIF
    \STATE Sample an action $a_t \sim \pi(a|s_t;\theta)$
    \STATE Execute $a_{t}$ in $E$ and receive reward $r_t$, DR next \indent\indent state $s_{t+1}$
    \STATE Store the transition $\left(s_t, a_t, r_t, s_{t+1} \right)$ in $\mathcal{B}$
  \ENDFOR
  \FOR{PPO update $= 1, 2, \cdots, M$}
    \STATE Update $\pi(a|s;\theta)$ using gradient ascent on Eq.\eqref{eq:standard_ppo}  with data from $\mathcal{B}$
    \STATE Update $V_{\pi}(s;\theta)$ using gradient descent on Eq.\eqref{eq:mse} \indent\indent with data from $\mathcal{B}$
  \ENDFOR
\ENDFOR
\end{algorithmic}
\end{algorithm}

\section{Results and Analysis}
\label{ch6:results}
\begin{figure}[h!]
  \centering
  %\hspace{0.025\textwidth}
  \begin{subfigure}{0.49\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/training_curves/jaco_prop.pdf}
    \caption{Trained with Proprioceptive inputs}
  \end{subfigure}\hfill
  \begin{subfigure}{0.49\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/training_curves/jaco_noprop.pdf}
    \caption{Trained without proprioceptive inputs}
  \end{subfigure}
  %\hspace{0.025\textwidth}
  \caption{The learning curves of Jaco agents trained with DR and with (a) or without (b) proprioceptive inputs. (a) the agent have similar performance in both standard and DR environments. (b) the agent has lower performance in standard environment. The solid line is the median value and the shadow area is the 95$\%$ confidence interval.}
  \label{fig:domain_shift}
\end{figure}

Before formally testing and analysing the models, they are \td{fully trained} with 8 training configurations and the same hyperparameters to achieve promising performance. However, in the Jaco environment, we also find that the model trained with visual + proprioceptive inputs and DR has a performance degradation when using standard visual inputs (no DR; see Figure~\ref{fig:domain_shift}). This is unexpected when the trained policy is transferred from complex visual scenarios to a simple visual scenario, and this observation demonstrates that this agent is overfitting to DR visual inputs. Since each model has different performance in the standard environment, it is more meaningful to compare the performance variation of individual models under different test conditions.

\subsection{Test Conditions}
\begin{figure*}[h!]
  \centering
  \begin{subfigure}{0.32\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/test_observations/standard}
    \caption{Standard}
  \end{subfigure}
  \begin{subfigure}{0.32\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/test_observations/colour_b}
    \caption{Colour}
  \end{subfigure}
  \begin{subfigure}{0.32\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/test_observations/shape}
    \caption{Shape}
  \end{subfigure}
  \begin{subfigure}{0.32\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/test_observations/illumination}
    \caption{Illumination (Lvl.)}
  \end{subfigure}
  \begin{subfigure}{0.32\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/test_observations/illumination_dir}
    \caption{Illumination (Dir.)}
  \end{subfigure}
  \begin{subfigure}{0.32\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/test_observations/noise}
    \caption{Noise}
  \end{subfigure}
  \begin{subfigure}{0.32\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/test_observations/reflection}
    \caption{Reflection}
  \end{subfigure}
  \begin{subfigure}{0.32\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/test_observations/translation}
    \caption{Translation}
  \end{subfigure}
  \begin{subfigure}{0.32\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/test_observations/invisible}
    \caption{Invisibility}
  \end{subfigure}
  \caption{The visualisation of different test conditions in the Fetch environment.}
  \label{fig:tests}
\end{figure*}
We develop 8 different test conditions to evaluate the generalisation ability of the trained agents. Figure~\ref{fig:tests} shows an overview of 8 test conditions in the Fetch environment. The details of each test condition are listed below:
\begin{enumerate}
\item[a)] \textbf{Standard}: This is the default setting of the environment, which is used to evaluate the baseline performance of the agent for further comparison.

\item[b)] \textbf{Colour}: This condition introduces a distractor object, which has the same shape and size as the target object but has a different colour. It aims to evaluate the robustness of the agent when there exists an extra object with different colours. We use five different colours: yellow, blue, green, brown and purple. Some of these colours have a red component (yellow, brown and purple), while others do not (blue and green).

\item[c)] \textbf{Shape}: This condition introduces a red distractor object, which has a different shape to the target object. It aims to evaluate the robustness of the agent when there exists an extra object with different shapes. We use four different shapes: cube, ellipsoid, rectangle and diamond.

\item[d)] \textbf{Illumination (Lvl.)}: This condition changes the diffuse intensity of the source of light. We use five illumination levels: 0.4 to 0.0 for the Fetch environment, and 0.9 to 0.1 for the Jaco environment.

\item[e)] \textbf{Illumination (Dir.)}: This condition changes the location of the source of light. We use five different locations for both Fetch and Jaco environments.

\item[f)] \textbf{Noise}: This condition adds Gaussian noise $\mathcal{N}(0, 0.25)$ to the visual inputs. This noise is not the same as the one generated in DR.

\item[g)] \textbf{Reflection}: This condition makes the table of the Fetch environment and the floor of the Jaco environment reflective, and it can produces mirror-like reflections in these surfaces.

\item[h)] \textbf{Translation}: This condition shifts the RGB camera's location. We use 5 locations: -20 to 20cm in the y direction for the Fetch environment, and -20 to 20cm in the x direction for the Jaco environment.

\item[i)] \textbf{Invisibility}: This condition is an extreme case, where the robot arm becomes transparent, and it is only used to investigate the importance of the visual inputs to localise the robot arm.
\end{enumerate}

\subsection{Quantitative Results on Testing Conditions}
We first discuss how local visual changes affect the performance of these trained agents. Table~\ref{tbl:tests} demonstrates the performance of agents with different training configurations under different test conditions. For both the Fetch and Jaco environments, the agents trained with DR perform slightly worse than those trained without DR in the standard test condition. However, agents trained with DR show strong robustness to both kinds of distractors. In Table~\ref{tbl:diff_colours}, we also provide the performance of the agents to different colours and different shapes of the distractors.

In the Fetch environment, the trained agents are more robust to the colour distractors. However, the shape distractors decrease the performance of the agents trained without DR, especially for non-DR agents trained with proprioceptive inputs, which drops to a 0.48 success rate. This implies that the non-DR agent trained with proprioceptive inputs localises the target object primarily by colour information and partly with shape information. The non-DR agent trained with visual inputs only merely drops 0.17 success rate. \td{This is because the agent trained with only visual inputs requires better representation/feature learning to localise the position of the robot arm (i.e., self-localisation), and these learned features can help the agent to distinguish between different shapes and colours.}

In the Jaco environment, both non-DR agents show performance degradation when there exists a distractor with red components, and both DR agents have significant performance drops when the shape of the distractor is a diamond, which has a similar appearance to a sphere in a low-resolution viewed input. It also indicates that the non-DR agents prefer to localise the target object via red colour.

In addition, we also test the effect of the distractors on the agent in different positions. In Table~\ref{tbl:vary_distractor}, it only has a few influences on performance.

\begin{table}[H]
%\begin{sidewaystable}[h!]
  %\label{tbl:tests}
  \resizebox{\linewidth}{!}{
  \begin{tabular}{c|cc|c|cc|cc|ccc|c}
    \toprule
    Robot & DR & Prop. & Standard & Colours & Shapes & Illumination (Lvl.) & Illumination (Dir.) & Noise & Reflection & Translation & Invisibility\\
    \midrule
    Fetch & {\xmark}   & {\xmark}   & 1.000$\pm 0.000$ & 0.993$\pm 0.004$ & 0.827$\pm 0.056$ & 0.694$\pm 0.049$ & 0.539$\pm 0.069$ & 0.980$\pm 0.006$ & 0.447$\pm 0.039$ & 0.270$\pm 0.079$ & 0.000$\pm 0.000$\\
    Fetch & {\xmark}   & {\cmark} & 1.000$\pm 0.000$ & 0.937$\pm 0.027$ & 0.524$\pm 0.087$ & 0.593$\pm 0.064$ & 0.491$\pm 0.066$ & 0.988$\pm 0.004$ & 0.570$\pm 0.078$ & 0.231$\pm 0.078$ & 0.000$\pm 0.000$\\
    Fetch & {\cmark} & {\xmark}   & 0.983$\pm 0.004$ & 0.974$\pm 0.004$ & 0.939$\pm 0.018$ & 0.957$\pm 0.008$ & 0.740$\pm 0.058$ & 0.985$\pm 0.007$ & 0.972$\pm 0.011$ & 0.480$\pm 0.064$ & 0.000$\pm 0.000$\\
    Fetch & {\cmark} & {\cmark} & 0.997$\pm 0.002$ & 0.997$\pm 0.001$ & 0.978$\pm 0.009$ & 0.992$\pm 0.002$ & 0.883$\pm 0.038$ & 0.970$\pm 0.008$ & 0.985$\pm 0.005$ & 0.536$\pm 0.063$ & 0.023$\pm 0.015$\\
    \hline
    Jaco  & {\xmark}   & {\xmark}   & 0.995$\pm 0.003$ & 0.772$\pm 0.058$ & 0.367$\pm 0.069$ & 0.966$\pm 0.012$ & 0.838$\pm 0.035$ & 0.635$\pm 0.028$ & 0.734$\pm 0.032$ & 0.556$\pm 0.063$ & 0.000$\pm 0.000$\\
    Jaco  & {\xmark} & {\cmark} & 0.995$\pm 0.001$ & 0.808$\pm 0.044$ & 0.382$\pm 0.064$ & 0.876$\pm 0.032$ & 0.911$\pm 0.014$ & 0.478$\pm 0.059$ & 0.618$\pm 0.061$ & 0.688$\pm 0.048$ & 0.001$\pm 0.001$\\
    Jaco  & {\cmark} & {\xmark}   & 0.650$\pm 0.056$ & 0.652$\pm 0.022$ & 0.593$\pm 0.031$ & 0.643$\pm 0.027$ & 0.549$\pm 0.032$ & 0.575$\pm 0.040$ & 0.429$\pm 0.060$ & 0.310$\pm 0.041$ & 0.007$\pm 0.002$\\
    Jaco  & {\cmark} & {\cmark} & 0.991$\pm 0.004$ & 0.989$\pm 0.002$ & 0.843$\pm 0.061$ & 0.816$\pm 0.043$ & 0.882$\pm 0.031$ & 0.896$\pm 0.007$ & 0.946$\pm 0.006$ & 0.658$\pm 0.054$ & 0.916$\pm 0.022$\\
    \bottomrule
  \end{tabular}}
  \caption{Test performance of all models with local visual changes (distractors), global visual changes, and invisibility (visual self-localisation test). Colour and shape results are averaged over 9 different distractor locations, as well as all colours and shapes, respectively. Checkmarks and crosses indicate enabling/disabling DR and proprioceptive inputs (Prop.), respectively. Mean $\pm$ standard error are calculated over all models (seeds) and test target locations.}
  \label{tbl:tests}
%\end{sidewaystable}
\end{table}

\begin{table}[h!]
%\begin{sidewaystable}[h!]
  \resizebox{\linewidth}{!}{
  \begin{tabular}{c|cc|ccccc|cccc}
    \toprule
    Robot & DR & Prop. & Yellow & Blue & Green & Brown & Purple & Cube & Ellipsoid & Rectangle & Diamond\\
    \midrule
    Fetch & {\xmark}   & {\xmark}   & 0.993$\pm 0.007$ & 1.000$\pm 0.000$ & 1.000$\pm 0.000$ & 0.978$\pm 0.017$ & 0.995$\pm 0.004$ & 0.775$\pm 0.085$ & 0.993$\pm 0.007$ & 0.960$\pm 0.036$ & 0.580$\pm 0.142$\\
    Fetch & {\xmark}   & {\cmark} & 0.875$\pm 0.088$ & 1.000$\pm 0.000$ & 1.000$\pm 0.000$ & 0.825$\pm 0.069$ & 0.983$\pm 0.006$ & 0.243$\pm 0.064$ & 0.980$\pm 0.006$ & 0.783$\pm 0.079$ & 0.090$\pm 0.048$\\
    Fetch & {\cmark} & {\xmark}   & 0.970$\pm 0.011$ & 0.980$\pm 0.006$ & 0.975$\pm 0.006$ & 0.973$\pm 0.010$ & 0.973$\pm 0.011$ & 0.913$\pm 0.042$ & 0.975$\pm 0.009$ & 0.965$\pm 0.013$ & 0.905$\pm 0.048$\\
    Fetch &{\cmark} & {\cmark} & 0.995$\pm 0.003$ & 0.998$\pm 0.002$ & 1.000$\pm 0.000$ & 0.998$\pm 0.002$ & 0.995$\pm 0.003$ & 0.963$\pm 0.020$ & 0.998$\pm 0.002$ & 0.995$\pm 0.003$ & 0.958$\pm 0.022$\\
    \hline
    Jaco  & {\xmark}   & {\xmark}   & 0.281$\pm 0.067$ & 0.871$\pm 0.036$ & 0.994$\pm 0.003$ & 0.730$\pm 0.095$ & 0.984$\pm 0.008$ & 0.274$\pm 0.077$ & 0.407$\pm 0.098$ & 0.759$\pm 0.066$ & 0.027$\pm 0.006$\\
    Jaco  & {\xmark}   & {\cmark} & 0.451$\pm 0.040$ & 0.941$\pm 0.032$ & 0.993$\pm 0.002$ & 0.706$\pm 0.064$ & 0.950$\pm 0.028$ & 0.258$\pm 0.044$ & 0.440$\pm 0.058$ & 0.784$\pm 0.044$ & 0.045$\pm 0.009$\\
    Jaco  & {\cmark} & {\xmark}   & 0.640$\pm 0.046$ & 0.655$\pm 0.055$ & 0.651$\pm 0.055$ & 0.668$\pm 0.046$ & 0.648$\pm 0.046$ & 0.636$\pm 0.040$ & 0.642$\pm 0.035$ & 0.666$\pm 0.047$ & 0.426$\pm 0.054$\\
    Jaco  & {\cmark} & {\cmark} & 0.987$\pm 0.005$ & 0.990$\pm 0.003$ & 0.987$\pm 0.003$ & 0.990$\pm 0.003$ & 0.990$\pm 0.003$ & 0.970$\pm 0.017$ & 0.970$\pm 0.011$ & 0.990$\pm 0.003$ & 0.441$\pm 0.128$\\
    \bottomrule
  \end{tabular}}
  \caption{Test performance of all models with local visual changes (distractors). Checkmarks and crosses indicate enabling/disabling DR and proprioceptive inputs (Prop.), respectively. Mean $\pm$ standard error are calculated over all models (seeds), test target locations, and 9 distractor locations.}
  \label{tbl:diff_colours}
%\end{sidewaystable}
\end{table}

\begin{table}[H]
  \centering
  \begin{tabular}{c|cc|cc}
    \toprule
    Robot & DR & Prop. & Colours & Shapes\\
    \midrule
    Fetch & {\xmark}   & {\xmark}   & 0.993$\pm 0.004$ & 0.567$\pm 0.047$\\
    Fetch & {\xmark}   & {\cmark} & 0.937$\pm 0.027$ & 0.507$\pm 0.047$\\
    Fetch & {\cmark} & {\xmark}   & 0.974$\pm 0.004$ & 0.639$\pm 0.051$\\
    Fetch & {\cmark} & {\cmark} & 0.997$\pm 0.001$ & 0.622$\pm 0.050$\\
    \hline
    Jaco  & {\xmark} & {\xmark}   & 0.772$\pm 0.058$ & 0.567$\pm 0.047$\\
    Jaco  & {\xmark} & {\cmark} & 0.808$\pm 0.044$ & 0.507$\pm 0.047$\\
    Jaco  & {\cmark} & {\xmark}   & 0.652$\pm 0.022$ & 0.639$\pm 0.051$\\
    Jaco  & {\cmark} & {\cmark} & 0.989$\pm 0.002$ & 0.622$\pm 0.050$\\
    \bottomrule
  \end{tabular}
  \caption{Test performance of a single model with distractors locations varying over 9 different on the ground plane (Jaco) and table (Fetch). Checkmarks and crosses indicate enabling/disabling DR and proprioceptive inputs (Prop.), respectively. Mean $\pm$ standard error are calculated for the best model (seed), over all test target locations and all distractor locations.}
  \label{tbl:vary_distractor}
\end{table}

Next, we discuss how global visual changes affect the performance of these trained agents. From Table~\ref{tbl:tests}, in this case, the DR agents show better robustness but still sacrifice a lot of performance in many test conditions.

{In the Illumination (Lvl.) condition, the performance of all trained agents decreases as the light intensity reduces ({see Table~\ref{tbl:diff_levels}}). In the Fetch environment, the DR agents are more robust to variations of the illumination intensity. In the Jaco environment, however, the DR agent trained with visual inputs only is more robust in this case. This is probably because the agent trained with visual inputs must learn better features for self-localisation, thus the learned features are more robust to the change of the environment.}

In the Illumination (Dir.) condition, the performance of all agents decreases as the direction of the light source changes (see Table~\ref{tbl:diff_dirs}). Similarly, in the Fetch environment, the DR agents are more robust. In the Jaco environment, compared to training with DR, proprioceptive inputs would be more helpful for robustness.

\begin{table}[H]
  \resizebox{0.9\linewidth}{!}{
  \begin{tabular}{c|cc|ccccc}
    \toprule
    Robot & DR & Prop. & 0.0/0.1 & 0.1/0.3 & 0.2/0.5 & 0.3/0.7 & 0.4/0.9 \\
    \midrule
    Fetch & {\xmark}   & {\xmark}   & 0.468$\pm 0.670$ & 0.513$\pm 0.081$ & 0.643$\pm 0.076$ & 0.860$\pm 0.047$ & 0.988$\pm 0.009$ \\
    Fetch & {\xmark}   & {\cmark} & 0.325$\pm 0.115$ & 0.383$\pm 0.112$ & 0.508$\pm 0.110$ & 0.775$\pm 0.069$ & 0.973$\pm 0.013$ \\
    Fetch & {\cmark} & {\xmark}   & 0.893$\pm 0.013$ & 0.948$\pm 0.018$ & 0.975$\pm 0.004$ & 0.985$\pm 0.004$ & 0.983$\pm 0.004$ \\
    Fetch & {\cmark} & {\cmark} & 0.983$\pm 0.006$ & 0.990$\pm 0.002$ & 0.995$\pm 0.003$ & 0.995$\pm 0.003$ & 0.995$\pm 0.003$ \\
    \hline
    Jaco  & {\xmark} & {\xmark}   & 0.874$\pm 0.034$ & 0.972$\pm 0.008$ & 0.990$\pm 0.001$ & 0.996$\pm 0.001$ & 0.996$\pm 0.001$ \\
    Jaco  & {\xmark} & {\cmark} & 0.587$\pm 0.043$ & 0.846$\pm 0.022$ & 0.966$\pm 0.007$ & 0.989$\pm 0.002$ & 0.994$\pm 0.001$ \\
    Jaco  & {\cmark} & {\xmark}   & 0.473$\pm 0.049$ & 0.626$\pm 0.042$ & 0.707$\pm 0.039$ & 0.711$\pm 0.046$ & 0.698$\pm 0.041$ \\
    Jaco  & {\cmark} & {\cmark} & 0.442$\pm 0.018$ & 0.713$\pm 0.013$ & 0.946$\pm 0.012$ & 0.986$\pm 0.002$ & 0.994$\pm 0.002$ \\
    \bottomrule
  \end{tabular}}
  \caption{Test performance of all models with different global illumination intensities; illumination is specificied as Fetch/Jaco. Checkmarks and crosses indicate enabling/disabling DR and proprioceptive inputs (Prop.), respectively. Mean $\pm$ standard error are calculated over all models (seeds) and test target locations.}
  \centering
  \label{tbl:diff_levels}
\end{table}

\begin{table}[H]
  \resizebox{0.9\linewidth}{!}{
  \begin{tabular}{c|cc|ccccc}
    \toprule
    Robot & DR & Prop. & 0.0,0.0,-1.0/ & -1.0,0.0,0.0/ & -1.0,0.0,-1.0/ & 0.0,-1.0,0.0/ & 0.0,1.0,1.0/   \\
          &    &       & 0.0,0.0,-1.3  & -1.3,0.0,0.0  & -1.3,0.0,-1.3  & 0.0,-1.3,-1.3 & -1.3,-1.3,-1.3 \\
    \midrule
    Fetch & {\xmark}   & {\xmark}   & 1.000$\pm 0.000$ & 0.020$\pm 0.004$ & 0.593$\pm 0.095$ & 0.518$\pm 0.091$ & 0.563$\pm 0.066$ \\
    Fetch & {\xmark}   & {\cmark} & 1.000$\pm 0.000$ & 0.068$\pm 0.041$ & 0.493$\pm 0.074$ & 0.453$\pm 0.079$ & 0.440$\pm 0.066$ \\
    Fetch & {\cmark} & {\xmark}   & 0.983$\pm 0.004$ & 0.338$\pm 0.117$ & 0.850$\pm 0.087$ & 0.763$\pm 0.097$ & 0.765$\pm 0.075$ \\
    Fetch & {\cmark} & {\cmark} & 0.998$\pm 0.002$ & 0.540$\pm 0.076$ & 0.948$\pm 0.025$ & 0.985$\pm 0.009$ & 0.945$\pm 0.021$ \\
    \hline
    Jaco  & {\xmark}   & {\xmark}   & 0.995$\pm 0.003$ & 0.818$\pm 0.022$ & 0.650$\pm 0.121$ & 0.805$\pm 0.044$ & 0.922$\pm 0.015$ \\
    Jaco  & {\xmark}   & {\cmark} & 0.995$\pm 0.001$ & 0.852$\pm 0.019$ & 0.930$\pm 0.029$ & 0.855$\pm 0.023$ & 0.922$\pm 0.011$ \\
    Jaco  & {\cmark} & {\xmark}   & 0.650$\pm 0.056$ & 0.608$\pm 0.044$ & 0.455$\pm 0.070$ & 0.506$\pm 0.068$ & 0.529$\pm 0.074$ \\
    Jaco  & {\cmark} & {\cmark} & 0.991$\pm 0.004$ & 0.586$\pm 0.017$ & 0.932$\pm 0.016$ & 0.968$\pm 0.010$ & 0.934$\pm 0.016$ \\
    \bottomrule
  \end{tabular}}
  \caption{Test performance of all models with different main illumination directions; direction is specified as Fetch/Jaco. Checkmarks and crosses indicate enabling/disabling DR and proprioceptive inputs (Prop.), respectively. Mean $\pm$ standard error are calculated over all models (seeds) and test target locations.}
  \centering
  \label{tbl:diff_dirs}
\end{table}

For the case of additive Gaussian noise, the agents trained in the Fetch environment are all robust to the noise. However, in the Jaco environment, the non-DR agents have approximately 30$\%$ performance drop, while the DR agents only have 10$\%$.

In the reflection condition, the agents trained with DR are more robust. However, in the Fetch environment, it decreases the performance of the non-DR agents by 50$\%$. While in the Jaco environment, there is not a significant influence on the performance of the non-DR agents. One possible explanation is the difference in the sizes of the robot arm. The Jaco robot arm is relatively small in the simulation, so its reflection has little change on the visual inputs.

In the translation condition, the performance of all agents reduces heavily (see Table~\ref{tbl:diff_trans}). In the Fetch environment, just as before, DR agents exhibit better robustness. In the Jaco environment, non-DR and DR agents have similar performance, which reflects that these agents are able to learn the extent of translation invariance for the policies. This could be a benefit for the target reaching task in a 3D space, which leads to a more generalisable spatial representation.

\begin{table}[H]
  \resizebox{\linewidth}{!}{
  \begin{tabular}{c|cc|ccccc}
    \toprule
    Robot & DR & Prop. & -20cm & -10cm & 0cm & 10cm & 20cm \\
    \midrule
    Fetch & {\xmark}   & {\xmark}   & 0.013$\pm 0.004$ & 0.168$\pm 0.101$ & 1.000$\pm 0.000$ & 0.163$\pm 0.091$ & 0.008$\pm 0.004$ \\
    Fetch & {\xmark}   & {\cmark} & 0.043$\pm 0.027$ & 0.075$\pm 0.039$ & 1.000$\pm 0.000$ & 0.035$\pm 0.010$ & 0.000$\pm 0.000$ \\
    Fetch & {\cmark} & {\xmark}   & 0.268$\pm 0.042$ & 0.625$\pm 0.056$ & 0.983$\pm 0.004$ & 0.433$\pm 0.056$ & 0.093$\pm 0.040$ \\
    Fetch & {\cmark} & {\cmark} & 0.358$\pm 0.051$ & 0.685$\pm 0.033$ & 0.998$\pm 0.002$ & 0.488$\pm 0.094$ & 0.153$\pm 0.055$ \\
    \hline
    Jaco  & {\xmark}   & {\xmark}   & 0.107$\pm 0.030$ & 0.570$\pm 0.065$ & 0.995$\pm 0.003$ & 0.714$\pm 0.022$ & 0.394$\pm 0.055$ \\
    Jaco  & {\xmark}   & {\cmark} & 0.464$\pm 0.062$ & 0.826$\pm 0.026$ & 0.995$\pm 0.001$ & 0.757$\pm 0.022$ & 0.399$\pm 0.040$ \\
    Jaco  & {\cmark} & {\xmark}   & 0.175$\pm 0.025$ & 0.265$\pm 0.049$ & 0.650$\pm 0.056$ & 0.321$\pm 0.030$ & 0.141$\pm 0.034$ \\
    Jaco  & {\cmark} & {\cmark} & 0.338$\pm 0.034$ & 0.853$\pm 0.014$ & 0.991$\pm 0.004$ & 0.751$\pm 0.021$ & 0.356$\pm 0.029$ \\
    \bottomrule
  \end{tabular}}
  \caption{Test performance of all models with different levels of camera translation. Checkmarks and crosses indicate enabling/disabling DR and proprioceptive inputs (Prop.), respectively. Mean $\pm$ standard error are calculated over all models (seeds) and test target locations.}
  \label{tbl:diff_trans}
\end{table}

Finally, in the case of the arm being invisible condition, the performance of almost all agents is reduced to zero, except for the Jaco DR agent trained with proprioceptive inputs. For those failure cases, the agents need to infer the position of the robot arm visually. For the Jaco DR agent trained with proprioceptive inputs, the position of the robot arm is estimated by using only proprioceptive inputs.

From these results, we can conclude that, with the existing visual DR setting that randomises textures and colours, the agents can be made more robust to local visual changes, but less robust to global visual changes, such as the change of illumination intensity or the translation of the RGB camera. It is reasonable, because we do not include these variations in the DR configuration. Through finding such failure scenarios for the agent, it can also help us improve the configuration of DR.

% Saliency Map
\subsection{Saliency Maps}
We first employ the saliency map technique to infer which part of the visual input will affect the output of the policy network. Then, our proposed occlusion-based saliency map is used to investigate the trained agents in standard condition, colour change condition (yellow), and shape change condition (cube), respectively.

Figure~\ref{fig:saliency_fetch_distractor} demonstrates the saliency maps of the trained agents in the Fetch environment. From the results, DR agents in Figure~\ref{fig:saliency_fetch_distractor}j-l can self-localise via proprioceptive inputs. The others need to use visual inputs for self-localisation, where the saliency maps display highlighted regions in the robot arm. In Figure~\ref{fig:saliency_fetch_distractor}d-f, although the proprioceptive inputs have been provided to non-DR agents, these agents still use the visual inputs for self-localisation, which means not all inputs will be utilised efficiently. In addition, the saliency maps of the non-DR agents highlight more regions on the distractors (see Figure~\ref{fig:saliency_fetch_distractor}a-f), while the saliency maps of the DR agents do not (see Figure~\ref{fig:saliency_fetch_distractor}g-k), except for the case in Figure~\ref{fig:saliency_fetch_distractor}l.

Figure~\ref{fig:saliency_jaco_distractor} shows the saliency maps of the trained agents in the Jaco environment, which are more in line with our expectations. When the proprioceptive inputs are not available, the saliency maps have more highlighted regions on the robot arm for self-localisation. Conversely, the saliency maps only focus on the target object. Furthermore, from Figure~\ref{fig:saliency_jaco_distractor}b,c,e,f, the non-DR agents also pay more attention to the distractors. A close observation reveals that there exist weak highlighted regions around distractors in Figure~\ref{fig:saliency_jaco_distractor}e,f and robot arm in  Figure~\ref{fig:saliency_jaco_distractor}g-i, respectively.

While these saliency maps can provide some intuitive explanations, they do not completely replace the quantitative results in Table~\ref{tbl:diff_colours}. For example, the Fetch DR agent trained with proprioception inputs pays attention to the distractors. However, it has less performance drop than the Fetch DR agent trained with visual inputs only, whose saliency map only has highlighted regions on the target ball. Therefore, we cannot rely solely on saliency maps to interpret and analyse trained agents.

\begin{figure}[h!]
  \centering
  \begin{subfigure}{0.24\columnwidth}
    \includegraphics[width=\linewidth]{figures/chapter6/distractor_saliency_fetch_pro_off/standard_visual_std}
    \subcaption{Standard (no DR, no proprioception)}
  \end{subfigure}
  \begin{subfigure}{0.24\columnwidth}
    \includegraphics[width=\linewidth]{figures/chapter6/distractor_saliency_fetch_pro_off/color_visual_std}
    \subcaption{Colour (no DR, no proprioception)}
  \end{subfigure}
  \begin{subfigure}{0.24\columnwidth}
    \includegraphics[width=\linewidth]{figures/chapter6/distractor_saliency_fetch_pro_off/shape_visual_std}
    \subcaption{Shape (no DR, no proprioception)}
  \end{subfigure}
  \begin{subfigure}{0.24\columnwidth}
    \includegraphics[width=\linewidth]{figures/chapter6/distractor_saliency_fetch_pro_on/standard_sensor_std}
    \subcaption{Standard (no DR, proprioception)}
  \end{subfigure}

  \begin{subfigure}{0.24\columnwidth}
    \includegraphics[width=\linewidth]{figures/chapter6/distractor_saliency_fetch_pro_on/color_sensor_std}
    \subcaption{Colour (no DR, proprioception)}
  \end{subfigure}
  \begin{subfigure}{0.24\columnwidth}
    \includegraphics[width=\linewidth]{figures/chapter6/distractor_saliency_fetch_pro_on/shape_sensor_std}
    \subcaption{Shape (no DR, proprioception)}
  \end{subfigure}

  \begin{subfigure}{0.24\columnwidth}
    \includegraphics[width=\linewidth]{figures/chapter6/distractor_saliency_fetch_pro_off/standard_visual_random}
    \subcaption{Standard (DR, no proprioception)}
  \end{subfigure}
  \begin{subfigure}{0.24\columnwidth}
    \includegraphics[width=\linewidth]{figures/chapter6/distractor_saliency_fetch_pro_off/color_visual_random}
    \subcaption{Colour (DR, no proprioception)}
  \end{subfigure}
  \begin{subfigure}{0.24\columnwidth}
    \includegraphics[width=\linewidth]{figures/chapter6/distractor_saliency_fetch_pro_off/shape_visual_random}
    \subcaption{Shape (DR, no proprioception)}
  \end{subfigure}
  \begin{subfigure}{0.24\columnwidth}
    \includegraphics[width=\linewidth]{figures/chapter6/distractor_saliency_fetch_pro_on/standard_sensor_random}
    \subcaption{Standard (DR, proprioception)}
  \end{subfigure}

  \begin{subfigure}{0.24\columnwidth}
    \includegraphics[width=\linewidth]{figures/chapter6/distractor_saliency_fetch_pro_on/color_sensor_random}
    \subcaption{Colour (DR, proprioception)}
  \end{subfigure}
  \begin{subfigure}{0.24\columnwidth}
    \includegraphics[width=\linewidth]{figures/chapter6/distractor_saliency_fetch_pro_on/shape_sensor_random}
    \subcaption{Shape (DR, proprioception)}
  \end{subfigure}
  \caption{Occlusion-based saliency maps of Fetch agents trained with (g-l) or without (a-f) DR and with (d-f, j-l) or without proprioceptive inputs (a-c, g-i) in three different test conditions (standard, colour and shape). The best Fetch model is used for each training configuration.}
  \label{fig:saliency_fetch_distractor}
\end{figure}

\begin{figure}[h!]
  \centering
  \begin{subfigure}{0.24\columnwidth}
    \includegraphics[width=\linewidth]{figures/chapter6/distractor_saliency_jaco_pro_off/standard_visual_std}
    \subcaption{Standard (no DR, no proprioception)}
  \end{subfigure}
  \begin{subfigure}{0.24\columnwidth}
    \includegraphics[width=\linewidth]{figures/chapter6/distractor_saliency_jaco_pro_off/color_visual_std}
    \subcaption{Colour (no DR, no proprioception)}
  \end{subfigure}
  \begin{subfigure}{0.24\columnwidth}
    \includegraphics[width=\linewidth]{figures/chapter6/distractor_saliency_jaco_pro_off/shape_visual_std}
    \subcaption{Shape (no DR, no proprioception)}
  \end{subfigure}
  \begin{subfigure}{0.24\columnwidth}
    \includegraphics[width=\linewidth]{figures/chapter6/distractor_saliency_jaco_pro_on/standard_sensor_std}
    \subcaption{Standard (no DR, proprioception)}
  \end{subfigure}

  \begin{subfigure}{0.24\columnwidth}
    \includegraphics[width=\linewidth]{figures/chapter6/distractor_saliency_jaco_pro_on/color_sensor_std}
    \subcaption{Colour (no DR, proprioception)}
  \end{subfigure}
  \begin{subfigure}{0.24\columnwidth}
    \includegraphics[width=\linewidth]{figures/chapter6/distractor_saliency_jaco_pro_on/shape_sensor_std}
    \subcaption{Shape (no DR, proprioception)}
  \end{subfigure}

  \begin{subfigure}{0.24\columnwidth}
    \includegraphics[width=\linewidth]{figures/chapter6/distractor_saliency_jaco_pro_off/standard_visual_random}
    \subcaption{Standard (DR, no proprioception)}
  \end{subfigure}
  \begin{subfigure}{0.24\columnwidth}
    \includegraphics[width=\linewidth]{figures/chapter6/distractor_saliency_jaco_pro_off/color_visual_random}
    \subcaption{Colour (DR, no proprioception)}
  \end{subfigure}
  \begin{subfigure}{0.24\columnwidth}
    \includegraphics[width=\linewidth]{figures/chapter6/distractor_saliency_jaco_pro_off/shape_visual_random}
    \subcaption{Shape (DR, no proprioception)}
  \end{subfigure}
  \begin{subfigure}{0.24\columnwidth}
    \includegraphics[width=\linewidth]{figures/chapter6/distractor_saliency_jaco_pro_on/standard_sensor_random}
    \subcaption{Standard (DR, proprioception)}
  \end{subfigure}

  \begin{subfigure}{0.24\columnwidth}
    \includegraphics[width=\linewidth]{figures/chapter6/distractor_saliency_jaco_pro_on/color_sensor_random}
    \subcaption{Colour (DR, proprioception)}
  \end{subfigure}
  \begin{subfigure}{0.24\columnwidth}
    \includegraphics[width=\linewidth]{figures/chapter6/distractor_saliency_jaco_pro_on/shape_sensor_random}
    \subcaption{Shape (DR, proprioception)}
  \end{subfigure}
  \caption{Occlusion-based saliency maps of Jaco models trained with (g-l) or without (a-f) DR and with (d-f, j-l) or without proprioceptive inputs (a-c, g-i) in three different test conditions (standard, colour and shape). The best Jaco model is used for each training configuration.}
  \label{fig:saliency_jaco_distractor}
\end{figure}

%\subsection{Statistical and Structural Weight Characterisations}
\subsection{Unit Ablations}

\begin{figure}[h!]
  \centering
  \begin{subfigure}{0.49\linewidth}
    \includegraphics[width=\textwidth]{figures/chapter6/unitablation/conv1_ablations}
    \caption{Standard test condition (Layer 1)}
  \end{subfigure}
  \begin{subfigure}{0.49\linewidth}
    \includegraphics[width=\textwidth]{figures/chapter6/unitablation/conv1_ablations_noisy}
    \caption{Noise test condition (Layer 1)}
  \end{subfigure}

  \begin{subfigure}{0.49\linewidth}
    \includegraphics[width=\textwidth]{figures/chapter6/unitablation/conv2_ablations}
    \caption{Standard test condition (Layer 2)}
  \end{subfigure}
  \begin{subfigure}{0.49\linewidth}
    \includegraphics[width=\textwidth]{figures/chapter6/unitablation/conv2_ablations_noisy}
    \caption{Noise test condition (Layer 2)}
  \end{subfigure}
  \caption{The performance of unit ablation tests of the trained Fetch and Jaco agents under 8 different training configurations in standard and visual noise conditions. Each point represents one unit in layer 1 (a, b) or layer 2 (c, d). The vertical bars indicate baseline performance in the test environment. The training settings correspond to the Fetch (F) and Jaco (J) robots, whether additional proprioceptive inputs are available (Prop), and if DR is activated. The best model is used for each training configuration. To be noticed that the Jaco agent trained with DR but without proprioceptive inputs has a lower baseline performance on the standard visuals than the other models (Table \ref{tbl:tests}).}
  \label{fig:unit_ablations}
\end{figure}
We use unit ablations to provide white box and quantitative analysis. We first select the best agent from each of 8 training configurations across five different seeds as our test models. Subsequently, for each test model, we mask one of the output channels to zero in the first or second convolutional layers. Then, the modified model is evaluated on both standard and noise conditions (80 fixed positions for Fetch agents and 250 fixed positions for Jaco agents). Finally, we repeat the above process until all output channels have been tested. The results of the unit ablation tests are illustrated in Figure~\ref{fig:unit_ablations}

From Figure~\ref{fig:unit_ablations}, we observe that the impact of unit ablation on the Fetch agents is small. However, the effect on the Jaco agents is relatively large, as the performance of the Jaco agents varies significantly. This is probably because of the increased complexity of the Jaco task (e.g., control scheme) and which requires more powerful channels to extract relevant information to train the agents.

% We further find that the unit ablations can improve the performance in the noise condition, even for the DR agents. \td{This may}

Furthermore, in the first convolutional layer (see Figure~\ref{fig:unit_ablations}a,b), there exist several extremely critical channels/units that can substantially affect the performance of the agent. However, there are no such channels in the second convolutional layer (see Figure~\ref{fig:unit_ablations}c,d).

\subsection{Layer Re-initialisation}
Next, we move to the layer re-initialisation test, where we evaluate models' robustness to the layer re-initialisation and the change of the parameters in $\ell_{\infty}$ and $\ell_{2}$ norm. The related plots of Fetch and Jaco agents are in Figure~\ref{fig:fetch_robustness} and Figure~\ref{fig:jaco_robustness}, respectively. Generally, after a period of training, the LSTM layer and policy layer are more robust to be re-initialised with the initial weights. For the DR agents, they show less robustness to re-initialisation from the early to middle stages of training, which indicates DR agents require longer periods to learn useful representations from tasks.

Another interesting finding is that when the proprioceptive inputs are available, the first fully connected layer (fc1) of the agents needs more extended periods for training. This is probably because the fc1 layer has to learn a sophisticated combination between the visual inputs and proprioceptive inputs. This is especially the case for the Jaco DR agent, which can use only proprioceptive inputs for self-localisation: it requires an even longer time to train the fc1 layer than other agents (see Figure~\ref{fig:jaco_robustness}j).

Finally, in almost all cases, the LSTM layer is robust to be re-initialised with the initial weights. Therefore, we suppose that the LSTM layer is redundant in the network structure for these tasks. Further experiments to verify this assumption are provided in the following subsection.

\begin{figure}[h!]
  \centering
  \begin{subfigure}{0.32\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/robustness/fetch/visual_std/error}
    \caption{Robustness (no DR, no proprioception)}
  \end{subfigure}
  \begin{subfigure}{0.32\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/robustness/fetch/visual_std/inf_dist}
    \caption{$\ell_\infty$-norm (no DR, no proprioception)}
  \end{subfigure}
  \begin{subfigure}{0.32\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/robustness/fetch/visual_std/l2_dist}
    \caption{$\ell_2$-norm (no DR, no proprioception)}
  \end{subfigure}
  \begin{subfigure}{0.32\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/robustness/fetch/sensor_std/error}
    \caption{Robustness (no DR, proprioception)}
  \end{subfigure}
  \begin{subfigure}{0.32\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/robustness/fetch/sensor_std/inf_dist}
    \caption{$\ell_\infty$-norm (no DR, proprioception)}
  \end{subfigure}
  \begin{subfigure}{0.32\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/robustness/fetch/sensor_std/l2_dist}
    \caption{$\ell_2$-norm (no DR, proprioception)}
  \end{subfigure}
  \begin{subfigure}{0.32\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/robustness/fetch/visual_random/error}
    \caption{Robustness (DR, no proprioception)}
  \end{subfigure}
  \begin{subfigure}{0.32\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/robustness/fetch/visual_random/inf_dist}
    \caption{$\ell_\infty$-norm (DR, no proprioception)}
  \end{subfigure}
  \begin{subfigure}{0.32\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/robustness/fetch/visual_random/l2_dist}
    \caption{$\ell_2$-norm (DR, no proprioception)}
  \end{subfigure}
  \begin{subfigure}{0.32\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/robustness/fetch/sensor_random/error}
    \caption{Robustness (DR, proprioception)}
  \end{subfigure}
  \begin{subfigure}{0.32\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/robustness/fetch/sensor_random/inf_dist}
    \caption{$\ell_\infty$-norm (DR, proprioception)}
  \end{subfigure}
  \begin{subfigure}{0.32\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/robustness/fetch/sensor_random/l2_dist}
    \caption{$\ell_2$-norm (DR, proprioception)}
  \end{subfigure}
  \caption{Performance of re-initialisation robustness (0 for 100$\%$ success rate, and 1 for 0$\%$ success rate), and change of parameters in $\ell_\infty$- and $\ell_2$-norm of Fetch agents. The best Fetch model is selected for each training configuration.}
  \label{fig:fetch_robustness}
\end{figure}
%\vspace{-15mm}
\begin{figure}[h!]
  \centering
  \begin{subfigure}{0.32\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/robustness/jaco/visual_std/error}
    \caption{Robustness (no DR, no proprioception)}
  \end{subfigure}
  \begin{subfigure}{0.32\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/robustness/jaco/visual_std/inf_dist}
    \caption{$\ell_\infty$-norm (no DR, no proprioception)}
  \end{subfigure}
  \begin{subfigure}{0.32\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/robustness/jaco/visual_std/l2_dist}
    \caption{$\ell_2$-norm (no DR, no proprioception)}
  \end{subfigure}
  \begin{subfigure}{0.32\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/robustness/jaco/sensor_std/error}
    \caption{Robustness (no DR, proprioception)}
  \end{subfigure}
  \begin{subfigure}{0.32\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/robustness/jaco/sensor_std/inf_dist}
    \caption{$\ell_\infty$-norm (no DR, proprioception)}
  \end{subfigure}
  \begin{subfigure}{0.32\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/robustness/jaco/sensor_std/l2_dist}
    \caption{$\ell_2$-norm (no DR, proprioception)}
  \end{subfigure}
  \begin{subfigure}{0.32\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/robustness/jaco/visual_random/error}
    \caption{Robustness (DR, no proprioception)}
  \end{subfigure}
  \begin{subfigure}{0.32\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/robustness/jaco/visual_random/inf_dist}
    \caption{$\ell_\infty$-norm (DR, no proprioception)}
  \end{subfigure}
  \begin{subfigure}{0.32\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/robustness/jaco/visual_random/l2_dist}
    \caption{$\ell_2$-norm (DR, no proprioception)}
  \end{subfigure}
  \begin{subfigure}{0.32\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/robustness/jaco/sensor_random/error}
    \caption{Robustness (DR, proprioception)}
  \end{subfigure}
  \begin{subfigure}{0.32\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/robustness/jaco/sensor_random/inf_dist}
    \caption{$\ell_\infty$-norm (DR, proprioception)}
  \end{subfigure}
  \begin{subfigure}{0.32\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/robustness/jaco/sensor_random/l2_dist}
    \caption{$\ell_2$-norm (DR, proprioception)}
  \end{subfigure}
  \caption{Performance of re-initialisation robustness (0 for 100$\%$ success rate, and 1 for 0$\%$ success rate), and change of parameters in $\ell_\infty$- and $\ell_2$-norm of Jaco agents. The best Fetch model is selected for each training configuration.}
  %Note that the final failure rate of the best Jaco agent trained with DR and without proprioception on the standard environment is around 20\%. The re-initialisation robustness plot for this condition (g) indicates that all layers are necessary and that training continues to improve performance in the epochs depicted and beyond.
  \label{fig:jaco_robustness}
\end{figure}

\subsection{Recurrent Ablation}
We then evaluate the effectiveness of the LSTM layer for the training. In this case, the hidden states and cell states are set to constant values for the evaluation. From Table~\ref{tbl:hidden_ablation}, the LSTM layer with constant values only has a minor influence on non-DR agents, while it has a significant influence on the DR agents. This suggests that the LSTM layer is functional when DR is activated during training, otherwise it is unnecessary to learn a mapping from observations to policy.

\begin{table}[h!]
  \centering
  \begin{tabular}{c|cc|cc}
    \toprule
    Robot & DR & Prop. & Standard & Constant Hidden\\
    \midrule
    Fetch & {\xmark} & {\xmark} & 1.000$\pm 0.000$ & 0.988$\pm 0.016$\\
    Fetch & {\xmark} & {\cmark} & 1.000$\pm 0.000$ & 0.990$\pm 0.012$\\
    Fetch & {\cmark} & {\xmark} & 0.983$\pm 0.004$ & 0.658$\pm 0.106$\\
    Fetch & {\cmark} & {\cmark} & 0.997$\pm 0.002$ & 0.838$\pm 0.051$\\
    \hline
    Jaco  & {\xmark} & {\xmark} & 0.995$\pm 0.003$ & 0.919$\pm 0.032$\\
    Jaco  & {\xmark} & {\cmark} & 0.995$\pm 0.001$ & 0.943$\pm 0.022$\\
    Jaco  & {\cmark} & {\xmark} & 0.650$\pm 0.056$ & 0.422$\pm 0.083$\\
    Jaco  & {\cmark} & {\cmark} & 0.991$\pm 0.004$ & 0.746$\pm 0.060$\\
    \bottomrule
  \end{tabular}
  \caption{Performance of recurrent ablations of all models with standard operation and constant (empirical average) hidden states. Checkmarks and crosses indicate activate/deactivate DR and proprioceptive inputs (Prop.), respectively. Mean $\pm$ standard error are calculated over all models (seeds) and test target locations.}
  \label{tbl:hidden_ablation}
\end{table}
% discussion
\section{Discussion}
This chapter aims to analyse and understand what strategies and representations the agents have learned in a simulated robot manipulation task, especially with domain randomisation (DR). In this experiment, we construct 8 different training configurations, which cover different morphologies and control schemes of the robot arm (Fetch and Jaco robot arms), different modalities of inputs (visual inputs or visual + proprioceptive inputs), and using or not using DR. We also setup 8 different test conditions to evaluate the agents' robustness to local and global visual changes and the ability of visual self-localisation. Finally, various interpretability techniques are adopted to provide both quantitative and qualitative analysis of the agents.

From the results, the agent trained with DR shows better generalisation ability to local visual changes, such as distractors with different colours and shapes. While DR agents also present a degree of robustness to global visual changes, such as illumination (direction and intensity), RGB camera translation, additive noise, and reflective surfaces, they still show significant performance degradation in several scenarios. The above failure cases can remind us to consider more factors in the DR design, where we only focus on visual DR in this chapter (e.g., texture, colours and noise). From saliency maps, the non-DR agents show more highlighted regions over the distractors, and DR agents focus more attention on the task related objects (e.g., robot arm and target ball).

\td{In unit ablations, due to the task complexity and the control scheme of the Jaco environment, the Jaco agents require more channels for learning the representations to solve the task.} In layer re-initialisation, DR leads to distinguishable changes in $\ell_{\infty}$-norm of parameters in the convolutional layers and therefore acquires more robust features. It also shows that, when the proprioceptive inputs are available, the fc1 layer of the Jaco DR agent requires longer periods for training,  because it needs to learn a better representation from proprioceptive inputs for self-localisation. This trend is also in line with the results in the test of visual self-localisation, where this agent can still perform the task without the rendered robot arm. In this case, the agent only utilises proprioceptive inputs to infer the location of the robot arm. In the test of recurrent ablations, the LSTM layer is only necessary when DR is activated.

Finally, some recommendations are also provided to those who intend to investigate DRL agents.
\begin{itemize}
\item[1)] Do not assume that results generalise. While the Jaco agent trained with DR and without proprioception obeys some trends, it remains an outlier in many other aspects.
\item[2)] Do not expect clear results. While some methods, such as the saliency map, uncovered clear trends, others, such as unit ablations, did not. This does not mean that unit ablations are generally useless---one can imagine that if several units were highly important without DR and no units are individually important with DR, we would have found a significant trend using this method.
\item[3)] If possible, use a range of interpretability methods. Uncertainty about results from one method may be resolved by results from another method.
\item[4)] No single interpretability method is more useful than another---each type of method reveals different, complementary pieces of information.
\item[5)] Finally, use interpretability methods before making claims about the strategies used by agents. While one may assume that an agent that performs well at a task on average is using an intelligent strategy, in all likelihood, it may be using heuristics that fail to generalise \cite{ruderman2019uncovering}.
\end{itemize}

\label{ch6:discussion}

### 2022-04-23 conclusion/conclusion.tex
update

\chapter{Conclusion}

\label{ch:conclusions}

\section{Summary of Thesis Achievements}
In this thesis, we have proposed several approaches to improve the sample efficiency of deep reinforcement learning (DRL) in tasks with sparse and delayed rewards. Moreover, we provide a comprehensive quantitative and qualitative analysis on the agents trained with and without domain randomisation (DR) to visualise and explain why DR can improve the robustness of the agents against unseen scenarios.

Firstly, in \textbf{Chapter}~\ref{ch:esil} and \textbf{Chapter}~\ref{ch:dtgsh}, our methods aim to improve the learning efficiency of goal-oriented RL in continuous control tasks with sparse rewards. In \textbf{Chapter}~\ref{ch:esil}, to overcome the difficulty that the agent rarely achieves positive rewards from the environment, hindsight goal relabelling is leveraged to modify failed trajectories into successful ones. Then, these modified trajectories are used to accelerate the training/exploration in the initial stages of training via a self-imitation learning scheme. In addition, a trajectory selection module and an adaptive weight are proposed to remove redundant hindsight experiences and maintain the balance between the exploration and exploiting. In \textbf{Chapter}~\ref{ch:dtgsh}, we argue that when using hindsight goal relabelling, not all agent's experiences contribute equally to the training, and the naive uniform sampling may lead to inefficient learning. In order to achieve valuable experience, we first sample trajectories according to the diversity of the goal states as modelled by determinantal point processes (DPPs). Then, transitions with diverse goal states are selected from the trajectories by using $k$-DPPs. Extensive experiments in these two chapters demonstrate the effectiveness of our methods and show that our methods achieve state-of-the-art results on several simulated robot manipulation tasks.

Secondly, in \textbf{Chapter}~\ref{ch:daim}, diversity is again used via a mechanism of intrinsic motivation, a diversity-augmented intrinsic motivation is proposed to encourage the agent keep exploring novel states in the environment. This intrinsic reward is formulated based on the diversity of adjacent states, measured under the framework of DPPs. Furthermore, a bi-level optimisation is used to guarantee that when increasing the intrinsic reward to encourage the agent to explore novel states, the extrinsic returns are still maximised. Experimental results show that this method outperforms other baselines on the MuJoCo continuous control tasks with standard and delayed reward settings, and it accelerates exploration in the initial stage of training in the arcade learning environments.

Finally, in \textbf{Chapter}~\ref{ch:drl_interp}, we investigate the agents trained with and without DR to interpret why the agents trained with DR are more robust to unseen scenarios. In this work, we built two simulated robot environments that provide both visual and proprioceptive inputs. The agents are trained to manipulate the robot arm to reach a specific target with different training conditions (different types of inputs and with or without DR). In order to analyse the trained agents, various previously unseen test conditions are used to evaluate the robustness of the trained agents, and both quantitative and qualitative analysis are applied to provide sufficient insights into the behaviour of the trained agents. We further propose an improved occlusion-based saliency map method to provide better attention maps of the agents at each timestep. \td{From results, we show that the DR agents have better generalisation ability to local visual changes, and also exhibit a degree of robustness to global visual changes. The saliency maps of the DR agents focus more on the target object, while non-DR agents show attention to the distractors. In the layer re-initialisation, DR agents also show significant changes in the parameters, which leads to more robust features. The recurrent ablations also reflect the LSTM is only necessary when DR is activated.} In the end, some recommendations for applying interpretability methods to DRL agents are also suggested.

%\section{Applications}

%Applications.
\section{Future Work and Potential Applications}
The topic of my thesis focuses on improving sample efficiency in DRL and interpretability in DRL. Thus, this section can be split into two branches: 1) discussion of limitations and potential extensions to the proposed methods. 2) potential applications to solve scientific problems.

\subsection{Limitations and Extensions}
In \textbf{Chapter}~\ref{ch:esil}, the proposed episodic self-imitation learning (ESIL) allows on-policy DRL algorithms (e.g., PPO) and hindsight experience replay (HER) to be combined and can be used to solve continuous control tasks with the sparse rewards. However, this still require more samples for training than off-policy DRL algorithms (e.g., DDPG+HER). One possible future study is to improve the sample efficiency with ESIL. Another future work could incorporate self-imitation learning mechanisms with DDPG+HER. Nair \textit{et al.}~\cite{nair2018overcoming} have already done relevant work, where they collect expert demonstrations for imitation learning. In stead of manually collecting demonstrations, \td{a potential improvement is to utilise} past good trajectories for self-imitation learning in the off-policy DRL paradigm.

In \textbf{Chapter}~\ref{ch:dtgsh}, the diversity-based sampling strategy is only applicable to environments that contain manipulable objects, such as picking and placing a cube at a target position. In the future, we can extend this method to more general goal-oriented tasks, such as navigating an agent to arrive at a target location in a maze~\cite{florensa2018automatic}.

In \textbf{Chapter}~\ref{ch:daim}, the proposed diversity-augmented intrinsic motivation (DAIM) achieves limited performance in the arcade learning environment (ALE), and it is also unable to address hard exploration tasks, such as Montezuma's Revenge. Therefore, \td{tackling such challenging hard exploration tasks is a potential direction for improvement.}

In \textbf{Chapter}~\ref{ch:drl_interp}, the current work focuses solely on the analysis of visual domain randomisation. In the future, we will also analysis the trained agents with dynamic DR (e.g., changing the frictional coefficient.). Furthermore, instead of using a simple reach task as the test environment, we will further develop more complicated tasks for analysing, such as a stacking cube task or a pick and place task, which are more in line with real-world applications.

In addition, although there exist plentiful methods to improve sample efficiency in DRL. However, sample efficiency in DRL is still a challenging problem. Therefore, I will further discuss some interesting and novel directions for my future research work.

Firstly, rather than using HER to address goal-oriented RL problems, C-learning~\cite{eysenbach2020c} provides a new paradigm to solve goal-oriented RL by estimating the probability density over future states. In practice, estimating the probability density function is very difficult, and C-learning estimates this density function indirectly via training a binary classifier to predict if the current state--action pair can generate a specific future state (i.e. the goal state) in the current episode, and the state--action pair has a higher probability of generating the goal state will be assigned with a higher return for training. Extensive experiments also show that C-learning outperforms other baselines in various goal-oriented tasks. Therefore, C-learning proposes a new way to solve goal-oriented RL, and it deserves further investigation.

Another way to improve sample efficiency in DRL is to reuse prior knowledge from previously unseen tasks. This kind of method can differ from imitation learning methods~\cite{bojarski2016end,xu2017end,torabi2018behavioral} which require collecting demonstrations from the same task. Some related methods like meta-learning based RL~\cite{duan2016rl,wang2017learning,finn2017model} and meta-imitation learning~\cite{duan2017one,finn2017one,huang2019continuous} try to leverage prior knowledge from unseen tasks to accelerate the learning of new tasks. Another interesting method called prior accelerated reinforcement (PARROT)~\cite{singh2020parrot}, first pretrains a flow-based generative model that maps the latent variables, which are conditioned on the current state, to an action by using expert demonstrations from the previous unseen tasks. Subsequently, this pretrained generative model can be used as a behaviour prior to facilitate the exploration in the new tasks. Thus, how to efficiently reuse prior knowledge/skills is an interesting research topic in the future.

\subsection{Potential Applications}
\begin{figure}[h]
  \begin{subfigure}[t]{0.24\textwidth}
    \includegraphics[width=\textwidth]{figures/conclusion/real_img.jpg}
    \caption{Microscopy Image}
    \label{subfig:real_img}
  \end{subfigure}\hfill
  \begin{subfigure}[t]{0.24\textwidth}
    \includegraphics[width=\textwidth]{figures/conclusion/synthetic_image_resize.png}
    \caption{Synthetic Image A}
    \label{subfig:synthetic}
  \end{subfigure}\hfill
  \begin{subfigure}[t]{0.24\textwidth}
    \includegraphics[width=\textwidth]{figures/conclusion/CapGAN.png}
    \caption{Synthetic Image B}
    \label{subfig:synthetic_b}
  \end{subfigure}\hfill
  \begin{subfigure}[t]{0.24\textwidth}
    \includegraphics[width=\textwidth]{figures/conclusion/3d_resize.png}
    \caption{3D Synthetic Image}
    \label{subfig:3d_synthetic}
  \end{subfigure}\hfill
  \caption{Examples of different types of axon images. (a) is a real microscopy image, (b) is the basic 2D synthetic axon image, (c) is the 2D synthetic axon image generated by CapsPix2Pix~\cite{bass2018image}, and (d) is the 3D synthetic axon image.} 
  \label{fig:axon_env}
\end{figure}
The recent successful achievement of machine learning for science is AlphaFold~\cite{jumper2021highly}, which leverages state-of-the-art deep learning based techniques to predict the 3D structure of a protein based solely on its genetic sequence, and the predicted structures are far more accurate than the outputs of any previous approaches. Predicting protein structures is important to understand the functionality of proteins in any biological system as well as in the diagnosis and treatment of especially genetic diseases (e.g., Alzheimer). Thus, it is meaningful if DRL can be applied to address some real-world/scientific problems. One potential application of DRL is to track the elongated structure in biomedical/microscopy images, such as axons and blood vessels. Automatically tracking elongated structures is a challenging problem in the field of biomedical imaging. \td{In contrast to segmentation, tracking also allows the establishment of correspondences and state estimation to be performed.} Some existing algorithms, such as Kalman filtering, particle filtering, or semi-heuristic connectivity algorithms, require tuning parameters manually for different datasets. To alleviate the need for hand-engineering trackers for different biomedical image datasets, we
first formulate the task of tracing paths along the centrelines of elongated structures as a RL problem, and then explore the use of DRL to train deep convolutional neural networks (CNNs) to learn tracking policies. Up to now, we have developed a simulated environment that can generate synthetic axon images with customised settings, such as the number of axons, the number of branches, and the type of noise (see Figure~\ref{subfig:synthetic}). To reduce the reality gap between the real microscopy image (see Figure~\ref{subfig:real_img}) and the synthetic image, we have proposed CapsPix2Pix~\cite{bass2018image} to generate more realistic synthetic images (see Figure~\ref{subfig:synthetic_b}). We have proposed two DRL trackers~\cite{dai2019deep,balaram2019maximum} which are able to solve 2D axon tracking and achieve comparable results with state-of-the-art axon tracking method -- Vaa3D~\cite{peng2010v3d}. However, the above DRL trackers \td{do not yet} scale to the 3D axon tracking task (see Figure~\ref{subfig:3d_synthetic}), because they lack effective exploration in the environment. In the future, we will apply different techniques (e.g., intrinsic motivation) to overcome the exploration problems in the 3D tracking task.
%Thus, the intrinsic motivation method can be combined with the DRL tracker, which encourages the agent to explore novel states to find a promising policy and solve the problem.

%Preliminary Results and Future Work.

### 2022-04-23 dedication/copyright.tex
update

\cleardoublepage

\begin{mycopyright}
The copyright of this thesis rests with the author. Unless otherwise indicated,
its contents are licensed under a Creative Commons Attribution-Non
Commercial 4.0 International Licence (CC BY-NC).
Under this licence, you may copy and redistribute the material in any medium
or format. You may also create and distribute modified versions of the work.
This is on the condition that: you credit the author and do not use it, or any
derivative works, for a commercial purpose.
When reusing or sharing this work, ensure you make the licence terms clear to
others by naming the licence and linking to the licence text. Where a work has
been adapted, you should indicate that the work has been changed and
describe those changes.
Please seek permission from the copyright holder for uses of this work that are
not included in this licence or permitted under UK Copyright Law.

\hfill Tianhong Dai
\end{mycopyright}

### 2022-04-23 dedication/dedication.tex
update

--- 2021-12-27 dedication/dedication.tex
@@
 I hereby declare that the content of this thesis is the product of my own research work.
-The main parts of this thesis have been published in the following list of my publications.
+The main parts of this thesis have been published in related conferences and journals.
 I made the major contribution in design, implementation and experiments for these work under the guidance of Prof.
-Anil Anthony Bharath and some others which have been acknowledged appropriately in text.
-The information, codes, and former ideas guided from other sources have been cited in a list of references is in the bibliography.
+Anil Anthony Bharath and some others which have been acknowledged appropriately.\\
+
+\hfill Tianhong Dai
 \end{dedication}
 % List the publications used in the thesis
-\textbf{Publication List}:
-\begin{enumerate}
-	\item \underline{Tianhong Dai}, Magda Dubois, Kai Arulkumaran, Jonathan Campbell, Cher Bass, Benjamin Billot, Fatmatulzehra Uslu, Vincenzo de Paola, Claudia Clopath, and Anil Anthony Bharath, ``Deep Reinforcement Learning for Subpixel Neural Tracking", International Conference on Medical Imaging with Deep Learning, 2019 (Spotlight).
-	\item \underline{Tianhong Dai}, Hengyan Liu, Anil Anthony Bharath, ``Episodic Self-Imitation Learning with Hindsight", Electronics, Special Issue on Deep Reinforcement Learning: Methods and Applications, 2020.
-	\item \underline{Tianhong Dai}, Hengyan Liu, Kai Arulkumaran, Guangyu Ren, and Anil Anthony Bharath, ``Diversity‐based Trajectory and Goal Selection with Hindsight Experience Replay", Pacific Rim International Conference on Artificial Intelligence, 2021.
-	\item \underline{Tianhong Dai}, Yali Du, Meng Fang, and Anil Anthony Bharath, ``Diversity-Augmented Intrinsic Motivation for Deep Reinforcement Learning", Neurocomputing, 2021.
-	\item \underline{Tianhong Dai}, Kai Arulkumaran, Tamara Gerbert, Samyakh Tukra, Feryal Behbahani, and Anil Anthony Bharath, ``Analysing Deep Reinforcement Learning Agents Trained with Domain Randomisation", Neurocomputing, 2022.
-\end{enumerate}
+% \textbf{Publication List}:
+% \begin{enumerate}
+% 	\item \underline{Tianhong Dai}, Magda Dubois, Kai Arulkumaran, Jonathan Campbell, Cher Bass, Benjamin Billot, Fatmatulzehra Uslu, Vincenzo de Paola, Claudia Clopath, and Anil Anthony Bharath, ``Deep Reinforcement Learning for Subpixel Neural Tracking", International Conference on Medical Imaging with Deep Learning, 2019 (Spotlight).
+% 	\item \underline{Tianhong Dai}, Hengyan Liu, Anil Anthony Bharath, ``Episodic Self-Imitation Learning with Hindsight", Electronics, Special Issue on Deep Reinforcement Learning: Methods and Applications, 2020.
+% 	\item \underline{Tianhong Dai}, Hengyan Liu, Kai Arulkumaran, Guangyu Ren, and Anil Anthony Bharath, ``Diversity‐based Trajectory and Goal Selection with Hindsight Experience Replay", Pacific Rim International Conference on Artificial Intelligence, 2021.
+% 	\item \underline{Tianhong Dai}, Yali Du, Meng Fang, and Anil Anthony Bharath, ``Diversity-Augmented Intrinsic Motivation for Deep Reinforcement Learning", Neurocomputing, 2021.
+% 	\item \underline{Tianhong Dai}, Kai Arulkumaran, Tamara Gerbert, Samyakh Tukra, Feryal Behbahani, and Anil Anthony Bharath, ``Analysing Deep Reinforcement Learning Agents Trained with Domain Randomisation", Neurocomputing, 2022.
+% \end{enumerate}
 
-% List extra publications
-In addition to the above papers, these following publications that I authored/co-authored during my PhD are out of the main scope of this thesis, and are briefly mentioned in the background or future directions.
-\begin{enumerate}
-	\item Shafa Balaram, Kai Arulkumaran, \underline{Tianhong Dai}, and Anil Anthony Bharath, ``A Maximum Entropy Deep Reinforcement Learning Neural Tracker", International Workshop on Machine Learning in Medical Imaging, 2019.
-	\item Yali Du, Lei Han, Meng Fang, \underline{Tianhong Dai}, Ji Liu, and Dacheng Tao, ``LIIR: Learning Individual Intrinsic Reward in Multi-Agent Reinforcement Learning", Advances in Neural Information Processing Systems, 2019.
-	\item \underline{Tianhong Dai}, Wei Li, Xilei Cao, Jianzhuang Liu, Xu Jia, Ales Leonardis, Youliang Yan, and Shanxin Yuan, ``Wavelet-Based Network For High Dynamic Range Imaging", Submitted to IEEE Transactions on Imaging Processing, 2021.
-\end{enumerate}
+% % List extra publications
+% In addition to the above papers, these following publications that I authored/co-authored during my PhD are out of the main scope of this thesis, and are briefly mentioned in the background or future directions.
+% \begin{enumerate}
+% 	\item Cher Bass, \underline{Tianhong Dai}, Benjamin Billot, Kai Arulkumaran, Antonia Creswell, Claudia Clopath, Vincenzo de Paola, and Anil Anthony Bharath, ``Image Synthesis with a Convolutional Capsule Generative Adversarial Network", International Conference on Medical Imaging with Deep Learning, 2019 (Oral).
+% 	\item Shafa Balaram, Kai Arulkumaran, \underline{Tianhong Dai}, and Anil Anthony Bharath, ``A Maximum Entropy Deep Reinforcement Learning Neural Tracker", International Workshop on Machine Learning in Medical Imaging, 2019.
+% 	\item Yali Du, Lei Han, Meng Fang, \underline{Tianhong Dai}, Ji Liu, and Dacheng Tao, ``LIIR: Learning Individual Intrinsic Reward in Multi-Agent Reinforcement Learning", Advances in Neural Information Processing Systems, 2019.
+% \end{enumerate}

### 2022-04-23 introduction/introduction.tex
update

\chapter{Introduction}

% TODO List:
% Focus on two parts: sample efficiency (sparse reward setting and general exploration strategy), interperbility.

% chang the title of subsection 1.2 -> general exploration strategies in DRL (-> intrinsic motivation... -> we propose a method based on intrinisc).

% Finally, we discuss the interperbility stuff.

\label{chapter1}
%\section{Reinforcement Learning}
\textit{Reinforcement Learning} (RL) is a machine learning method that aims to solve sequential decision-making problems with a trial-and-error scheme~\cite{sutton2018reinforcement}. In RL, an agent needs to interact with the environment to get a reward signal, thus adjusting its behaviour (policy) to maximise the accumulated rewards. In contrast to the other two machine learning paradigms: \textit{supervised learning} and \textit{unsupervised learning}, RL needs to find the optimal behaviour using the experiences of an agent instead of labelled training data, and the ultimate goal of RL is to maximise the accumulated rewards of an agent from the environment, rather than finding a specific hidden structure of data.

% reinforcement learning histroy
In recent years, with the success of deep learning (DL) and rapid iteration of computer hardware (e.g., Graphic Processing Unit (GPU)), deep neural networks have significantly improved the state-of-the-art in computer vision, such as for object classification~\cite{krizhevsky2012imagenet} and detection~\cite{ren2015faster}. \td{Incorporating deep neural networks with RL is a common approach}, which is capable of scaling to previously unsolved problems with high-dimensional inputs (e.g., images), and this learning paradigm is known as deep reinforcement learning (DRL). \td{Some other methods~\cite{guo2016k} reduce the dimension of inputs by adopting clustering algorithms.}

Deep reinforcement learning has demonstrated its impressive performance in controlling the agent in complex environments. For example, in 2015, DeepMind proposed the deep Q-network (DQN)~\cite{mnih2015human} which enabled the agent to outperform previous RL algorithms on Atari 2600 games and achieve comparable results to professional human players by only receiving the image of a game screen as the inputs. In 2016, AlphaGo~\cite{silver2016mastering} defeated the world-famous Go player Mr. Lee Sedol. In 2019, DRL also successfully addressed one of the most challenging real-time strategy (RTS) games - StarCraft II~\cite{vinyals2019grandmaster} and multiplayer online battle arena (MOBA) games - DOTA2~\cite{berner2019dota} which exhibited the potential of using DRL to tackle scientific and real-world problems. In addition to the milestone achievements above, DRL is also utilised as a solution in some other scenarios, such as dexterous robot manipulation~\cite{andrychowicz2020learning}, autonomous driving~\cite{kiran2021deep}, and urban traffic signal control~\cite{wu2020multi}. However, some existing challenges may become obstacles to impede people from using RL in real-world applications.

One crucial challenge is sample/data efficiency. Sample efficiency indicates the number of experiences/samples that an agent needs to collect in an environment to reach a certain performance level. Like DL, DRL also requires large amounts of interaction data to train the agent, which can lead to unacceptable costs in real-world applications. For example, even for training agents in a video game, such as StarCraft II, \td{16 Tensor Processing Units (TPUs) over 14 days were required} to complete the training (approximately 200 years of \td{human rate} StarCraft II play)~\cite{vinyals2019grandmaster}. Therefore, improving the sample efficiency of DRL algorithms is an important research direction for the community.

In this thesis, we mainly focus on overcoming some problems that lead to sample inefficiency in DRL. The proposed methods and investigation are based on the following three aspects: 1) solving DRL with a sparse reward setting, 2) improving general exploration strategies in DRL, and 3) generalisation and interpretability of DRL.

\section{DRL with Sparse Reward}
In RL, the goal or skills that the agent is expected to learn can be formalised into a specific reward function~\cite{sutton2018reinforcement}. The purpose of reward shaping is to design a task-specific reward function towards known optimal behaviours, and this usually incorporates domain knowledge and hand-crafted design. For example, in a locomotion task, a two-legged robot is required to move forward, and the reward function can be designed as~\cite{lee2018composing}:
\begin{align}
\mathcal{R}(s_{t}, a_{t}) = & \lambda_{\text{vel}}\cdot |v_{\text{x}} - v_{\text{target}}| + \lambda_{\text{alive}} - \lambda_{\text{height}}\cdot |1.1 - \min (1.1, \Delta h)| \\
        & + \lambda_{\text{angle}}\cdot \cos (\theta) \lambda_{\text{foot}}(v_{\text{left\_foot}} - v_{\text{right\_foot}}), \nonumber
\label{eq:designed_reward_function}
\end{align}
where $v_{x}$, $v_{\text{left\_foot}}$ and $v_{\text{right\_foot}}$ are the forward velocity, right foot angular velocity, left foot angular velocity. $\Delta h$ is the height between the foot and the torso. $\theta$ is the angle of the torso. $\lambda_{*}$ are the hyperparameters for the task. Therefore, even for simple moving forward behaviour, a complicated reward function needs to be designed, and this reward function cannot be generalised across different tasks and morphology of the robot. Furthermore, ill-designed reward functions usually lead to unexpected behaviors~\cite{riedmiller2018learning} and limit exploration.

Sparse reward functions can alleviate the complexity of design, where the non-negative reward is only given when the agent reaches a specific goal. However, sparse reward functions also mean that the agent rarely achieves a non-negative reward, and this affects the efficiency of training~\cite{andrychowicz2017hindsight}. In Chapter~\ref{ch:esil} and Chapter~\ref{ch:dtgsh}, we explore two approaches to address goal-oriented RL within a sparse reward setting, based on hindsight experience replay (HER)~\cite{andrychowicz2017hindsight}. Details about the goal-oriented RL and sparse reward function are described in Chapter~\ref{ch:background}.

\section{Exploration Strategies in DRL}
The trade-off between \textit{exploration} and \textit{exploitation} is a challenging topic in DRL. In order to achieve sufficient rewards from the environment, an agent needs to exploit previously learned actions that can lead to higher rewards; however, it also needs to explore in the environment to acquire better action selection in the future. Without enough exploration, the agent might be stuck at local-minima or even fail to solve tasks.

The classic exploration strategies adopted in DRL include: $\epsilon$-greedy~\cite{mnih2015human,wang2016dueling}, entropy regularisation term~\cite{mnih2016asynchronous,schulman2017proximal,schulman2015trust}, and noise-based exploration~\cite{lillicrap2015continuous,plappert2018parameter,fortunato2018noisy}. For $\epsilon$-greedy, the agent selects optimal actions with probability of $1-\epsilon$ and selects random actions with probability of $\epsilon$; for entropy regularisation, to encourage the agent to select diverse actions in the environment, an additional entropy term $\mathcal{H}(\pi(a_{t}|s_{t}))$ is used to increase the uncertainty of the policy $\pi$; for noise-based exploration, the noise (e.g., Gaussian noise) is added to the action space or parameter space to improve the exploration. However, these classic exploration strategies are relatively simple and can only provide limited contributions when the complexity of the environment is increased.

In this thesis, we discuss a more efficient exploration strategy called \textit{intrinsic motivation}, which can achieve better performance than previous classic exploration strategies in complex environments. From a cognitive psychological perspective, \textit{extrinsic motivation} denotes a behaviour driven by some specific incentive mechanism, while \textit{intrinsic motivation} refers to a behaviour which is self-driven~\cite{singh2005intrinsically}. More specifically, without the guidance of task-specific reward signals, intrinsic motivation can still keep organisms exploring and emerging broad behaviours or skills based on curiosity. Although such emerged skills may not be able to solve one specific task directly, they can be used as low-level skills to address various high-level tasks and thereby improve the learning efficiency.

This mechanism of curiosity can also be extended to RL. In RL, the agent typically learns the skills to tackle a specific task through a designed reward function, and this learning process is driven by \textit{extrinsic motivation}, and the corresponding reward function is also called \textit{extrinsic reward}. However, in the task where the extrinsic reward is difficult to acquire (e.g., sparse reward or delayed reward settings), the agent cannot achieve a sufficient supervised signal for learning. In this case, \textit{intrinsic motivation} can encourage the agent to explore novel states in the environment and learn valuable skills to improve learning efficiency and reach the ultimate goal. In Chapter~\ref{ch:daim}, we propose a diversity-based intrinsic motivation to help the agent explore an environment with delayed rewards.

\section{Generalisation and Interpretability}
Generate a policy that can be generalised to unseen environments is also essential when deploying RL models in real-world applications~\cite{kirk2021survey}. In the DRL community, instead of using a physical robot, a robot simulator is often selected to collect samples during training. The reason is that collecting samples in the simulation environment is faster and cheaper, and it can also avoid unexpected movements of the physical robot during the exploration stage. However, the distribution of the observations in the simulation environment is different from that of the real world. Thus, it is challenging to transfer the policy trained in the simulation environment to the real world.

Interpretability is also another challenge in DRL. Although DRL has achieved impressive performance in many tasks. However, they are often treated as black boxes during inference. When applying DRL models in real-world tasks where security and reliability are important (e.g., autonomous driving and robot manipulation), it is necessary to understand how DRL models make decisions, in order to prevent potential risks during deployment. In Chapter~\ref{ch:drl_interp}, we train agents with and without one of the sim-to-real technologies -- domain randomisation (DR)~\cite{andrychowicz2018learning,james2017transferring,peng2018sim,sadeghi2017cad2rl,tobin2017domain}, and evaluate their robustness to various previously unseen scenarios. Furthermore, different interpretability methods are adopted to analyse and understand how these trained agents work during inference.

%s still restricted to be applied to real-world problems, because RL requires large amount of samples for training, and .
\section{Contributions and Outline}
The contributions and outline of this thesis are organized as follows:
% make a list to summarize the contribution
\begin{enumerate}
    \item In \textbf{Chapter}~\ref{ch:background}, we provide an introduction to reinforcement learning, deep neural networks, optimisation methods, and deep reinforcement learning algorithms that are used in the following chapters.

    \item In \textbf{Chapter}~\ref{ch:esil}, we propose the episodic self-imitation learning (ESIL) algorithm, in order to improve the sample efficiency of RL in continuous control tasks within the sparse reward setting. The proposed method utilises episodic modified trajectories to update the policy. In addition, an auxiliary trajectory selection module and an adaptive weight are leveraged to remove undesired trajectories and balance the contribution between the RL loss term and self-imitation learning loss term. Extensive experiments show that ESIL achieves better performance than other baselines in four robot simulation tasks.

    \item In \textbf{Chapter}~\ref{ch:dtgsh}, we propose a diversity-based sampling strategy to improve the learning efficiency of hindsight experience replay (HER)~\cite{andrychowicz2017hindsight}. The diversity-based sampling strategy can be divided into two stages in this work. During training, the trajectories that are more valuable for training will be sampled in the first stage. In the second stage, one transition will be sampled from each of the selected trajectories to form candidate transitions. Then, a minibatch size of transitions with diverse goal states will be selected for the final training by using $k$-DPP. From experiments, our method outperforms other baselines in all benchmark tasks.

    \item In \textbf{Chapter}~\ref{ch:daim}, we design a diversity-based intrinsic reward to encourage the agent explore novel states in the tasks with sparse or delayed reward settings. The key idea in our work is to use determinantal point processes (DPPs) to model the discrepancy between states as an intrinsic reward. Furthermore, A bi-level optimisation method is utilised to combine with the proposed diversity-based measure, so that the intrinsic reward will increase the extrinsic reward while increasing diversity. From extensive experiments, our method achieves state-of-the-art performance in the MuJoCo locomotion environment within standard and delayed reward settings, and it also improves the learning efficiency at the initial stage of training in the arcade learning environment (ALE).

    \item In \textbf{Chapter}~\ref{ch:drl_interp}, we investigate the trained DRL agents with various types of interpretability methods. Large amounts of quantitative and qualitative results are used to analyse the agent trained with and without domain randomisation (DR) to understand the effect of DR.

    % Discussion and Conclusion
    \item In \textbf{Chapter}~\ref{ch:conclusions}, we briefly summarise the contributions of this thesis, and introduce future works and potential applications.
\end{enumerate}
% \section{Statement of Originality}
% Statement here.
\section{Publications}
This thesis includes the content and figures of the listed publications ($^{*}$ denotes equal contributions).
\begin{enumerate}
	\item \underline{\textbf{Tianhong Dai}}, Hengyan Liu, Anil Anthony Bharath, ``Episodic Self-Imitation Learning with Hindsight", Electronics, Special Issue on Deep Reinforcement Learning: Methods and Applications, vol.9, pp.1742, 2020. [\textbf{Chapter}~\ref{ch:esil}]
	\item \underline{\textbf{Tianhong Dai}}, Hengyan Liu, Kai Arulkumaran, Guangyu Ren, and Anil Anthony Bharath, ``Diversity‐based Trajectory and Goal Selection with Hindsight Experience Replay", Pacific Rim International Conference on Artificial Intelligence, pp.32--45, 2021. [\textbf{Chapter}~\ref{ch:dtgsh}]
	\item \underline{\textbf{Tianhong Dai}}, Yali Du, Meng Fang, and Anil Anthony Bharath, ``Diversity-Augmented Intrinsic Motivation for Deep Reinforcement Learning", Neurocomputing, vol.468, pp.396--406, 2021. [\textbf{Chapter}~\ref{ch:daim}]
	\item \underline{\textbf{Tianhong Dai}$^{*}$}, Kai Arulkumaran$^{*}$, Tamara Gerbert, Samyakh Tukra, Feryal Behbahani, and Anil Anthony Bharath, ``Analysing Deep Reinforcement Learning Agents Trained with Domain Randomisation", Neurocomputing, vol.493, pp.143--165, 2022. [\textbf{Chapter}~\ref{ch:drl_interp}]
	\item \underline{\textbf{Tianhong Dai}}, Magda Dubois, Kai Arulkumaran, Jonathan Campbell, Cher Bass, Benjamin Billot, Fatmatulzehra Uslu, Vincenzo de Paola, Claudia Clopath, and Anil Anthony Bharath, ``Deep Reinforcement Learning for Subpixel Neural Tracking", International Conference on Medical Imaging with Deep Learning, pp.130--150, 2019 (Spotlight). [\textbf{Chapter}~\ref{ch:conclusions}]
\end{enumerate}

% List extra publications
\td{In addition to the above publications, the following publications that I authored/co-authored during my Ph.D. study also contribute to deep reinforcement learning and its related applications.}
\begin{enumerate}
	\item Cher Bass, \underline{\textbf{Tianhong Dai}}, Benjamin Billot, Kai Arulkumaran, Antonia Creswell, Claudia Clopath, Vincenzo de Paola, and Anil Anthony Bharath, ``Image Synthesis with a Convolutional Capsule Generative Adversarial Network", International Conference on Medical Imaging with Deep Learning, pp.39--62, 2019 (Oral).
	\item Shafa Balaram, Kai Arulkumaran, \underline{\textbf{Tianhong Dai}}, and Anil Anthony Bharath, ``A Maximum Entropy Deep Reinforcement Learning Neural Tracker", International Workshop on Machine Learning in Medical Imaging, pp.400--408, 2019.
	\item Yali Du, Lei Han, Meng Fang, \underline{\textbf{Tianhong Dai}}, Ji Liu, and Dacheng Tao, ``LIIR: Learning Individual Intrinsic Reward in Multi-Agent Reinforcement Learning", Advances in Neural Information Processing Systems, vol.32, pp.4403--4414, 2019.
\end{enumerate}

### 2022-08-21 response.tex
update the final version

\documentclass[11pt]{article}
% Language setting
% Replace `english' with e.g. `spanish' to change the document language
\usepackage[english]{babel}

% Set page size and margins
% Replace `letterpaper' with`a4paper' for UK/EU standard size
\usepackage[letterpaper,top=2cm,bottom=2cm,left=3cm,right=3cm,marginparwidth=1.75cm]{geometry}
\usepackage{setspace}
% Useful packages
\usepackage{amsmath}
\usepackage{mathrsfs}
\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{bm}
%\usepackage{hyperref}
 \usepackage{float}
\usepackage{graphicx}
\newcommand{\td}[1]{\textcolor{red}{#1}}
\usepackage[colorlinks=true, allcolors=blue]{hyperref}
\renewcommand{\thefigure}{\Alph{figure}}
\title{Response Letter For Ph.D. Thesis Revision}
\author{Candidate: Tianhong Dai (CID: 01085460) \\ External Examiner: Prof. Jonathan Hare \\ Internal Examiner: Prof. Aldo Faisal \\ Supervisor: Prof. Anil Anthony Bharath}
\date{}
\begin{document}
%\doublespacing
\onehalfspacing
\maketitle

\noindent I thank the examiners for their constructive feedback. In light of examiners' comments and suggestions, I have revised my thesis manuscript. The modified text in the manuscript is indicated with a \td{red} color. Detailed responses to the comments are listed below.

\section*{General Modification}
\begin{enumerate}
\item[1)] \textbf{Modification:} According to the suggestions in the viva examination, I have moved Chapter 6 in the original manuscript to Chapter 3 in the revised manuscript. Furthermore, I have also changed the order in the summarization of each chapter in Chapter 1 and Chapter 7 to keep the consistency. The structure of the revised manuscript is listed as follows and the \textcolor{blue}{blue} numbers in brackets represent the original chapter numbers:
\begin{itemize}
    \item Chapter 1 - Introduction
    \item Chapter 2 - Background Theory
    \item Chapter 3 (\textcolor{blue}{6}) - Analysing DRL Agents Trained with Domain Randomisation
    \item Chapter 4 (\textcolor{blue}{3}) - Episodic Self-Imitation Learning with Hindsight
    \item Chapter 5 (\textcolor{blue}{4}) - Efficient Sampling Strategy with Hindsight
    \item Chapter 6 (\textcolor{blue}{5}) - Diversity-Augmented Intrinsic Motivation
    \item Chapter 7 - Conclusion
\end{itemize}

% \item[2)] \textbf{Comment:} Revise the manuscript according to Dr. Hare's notes.\\ \\
% \textbf{Response}: I have corrected the typos and texts according to Dr. Hare's notes.

\end{enumerate}
% new page
\newpage

\section*{Response to Chapter 1}
\begin{enumerate}
\item[1)] \textbf{Comment:} List of Tables and Figures: please define short titles ([ ] in latex) to tidy up the structure and give a neat label to the figure.\\ \\
\textbf{Response:} I have added short titles for all tables and figures in the latest manuscript. Now the titles in the List of Tables and Figures are much more concise.

\item[2)] \textbf{Comment:} Chapter 6 but also 1.3 set yourself apart from the IAI and XAI debate where interpretability in ML has specific connotations, while here you actually mean analysing the policy. \\ \\
\textbf{Response:} I have changed the title of Chapter 6 (Chapter 3 in the revised manuscript) to ``Analysing DRL Agents trained with Domain Randomisation". Furthermore, I also have changed the main idea in Section 1.3, which separates my thesis from IAI and XAI.

\end{enumerate}

% start a new page for review 2
\newpage
\section*{Response to Chapter 2}
\begin{enumerate}
\item[1)] \textbf{Comment:} Figure 2.1 which is the highlighted neuron in relation to Eq 2.1 please make more clear to make it a bit more general or in vector form.\\ \\
%\textbf{Response:} Some results have been removed from the abstract.
\textbf{Response:} I have highlighted the neurons in Figure 2.1, and I make Equation 2.1 in a more general form.
%\\ \\
\item[2)] \textbf{Comment:} 2.1.2. First sentence grammar wrong (perceptrons $<->$ its). Please fix. 2nd sentence is generally true also for conv nets. The reason why Conv nets are more compact or have fewer parameters has not been explained. Make this explicit why there are fewer parameters. Please revise, why not give orders of scaling.
Make more precise the fact that the architectures are not necessarily translation-invariant, etc. \\ \\
\textbf{Response:} In Section 2.1.2, I have fixed the grammar in the first sentence. I have also added an explanation of why CNNs can have fewer parameters and clarified that the architectures of CNNs are not necessarily translation-invariant as follows:

``
When the MLP is used to classify an image, each node in the hidden layer (the next layer of the input layer) is connected to all pixels of the input image. Thus, there will be a large number of parameters in the MLP when the size of the input image is large. For example, when the size of the input image is $256\times256$, the first hidden layer has $256\times256\times N$ parameters to learn, and $N$ is the number of neurons in this layer. Therefore, it is essential to find an approach to substantially reduce parameters. Convolutional neural networks (CNNs) [29] alleviate this problem by employing the parameter sharing strategy and local connectivity. Instead of connecting each node to all nodes in the previous layer, CNNs use a set of learnable filters with relatively small size (e.g. $9\times9$) to slide across the entire image to produce the features. In this case, the first convolution layer only has $9\times9\times M$ parameters to learn, and $M$ is the number of filters, therefore CNNs greatly reduce the number of parameters. Benefiting from parameter sharing, CNNs might have an implicit strong inductive bias that reflects translation invariance in some desired mapping from physical world to feature space.
"

\item[3)] \textbf{Comment:} 2.4.1 Clarify the benefit of the target network, why not also highlight its introduction in the nature paper.\\ \\
\textbf{Response:} In Section 2.4.1, I have added more clarification about the benefit of using the target network according to the original nature paper as follows:

``In DQN, the target network is proposed to stabilise the training. This is because in the standard online Q-learning update, the update that leads to an increase in the current Q-value, $Q(s_{t}, a_{t})$, can also increase the Q-value of the next state and all actions, $Q(s_{t+1}, a)$. Hence, the estimated target Q-value will also increase and may cause divergence or oscillations of the policy. Therefore, using the target network with the old parameters can introduce a delay between the updated Q-value and the estimated target Q-value, and thus prevent policy divergence [5]."

\item[4)] \textbf{Comment:} Uncovered statements on TRPO not sufficiently flexible but either an argument or a situation should be given to back it up. Why is PPO more accessible? Explain.
\\ \\
\textbf{Response:} In Section 2.4.3, I have added the argument that leads to the inflexible implementation of TRPO -- conjugate gradient, and also add a citation, which is a lecture note from John Schulman, who is the author of TRPO to support this claim. Furthermore, I also have added the explanation that why PPO is more accessible. This is because PPO only needs to maximise a single objective function and does not require additional algorithms (e.g., conjugate gradient and line search).
\end{enumerate}

\newpage
\section*{Response to Chapter 3 \textcolor{blue}{(Original Chapter 6)}}
\begin{enumerate}
\item[1)] \textbf{Comment:} Review the title if it actually does what it says on the tin. Chapter 6 is called “Interpretability of Deep RL” How is this intepretability in the sense of IAI? Provide in the introduction chapter and here a motivation for what you actually are doing, what other people mean to do in the space of XAI/*I*AI and your take on it. Framing this chapter is key to make it part of a consistent body of work.
\\ \\
\textbf{Response:} I have changed the title of this chapter to ``Analysing DRL Agents Trained with Domain Randomisation" to reduce ambiguity between XAI/IAI.

\item[2)] \textbf{Comment:} Also why not move the Chapter 6 to after Chapter 2 (as the story flows better” you do interpretability of features learned by PPO, then you go further and improve on PPO with different ways of improving it in the “old” chapter 3-5).\\ \\
\textbf{Response:} I have moved the Chapter 6 (original manuscript) to Chapter 3 (revised manuscript). Then, I also have changed the order in the summarization of each chapter in Chapter 1 and Chapter 7 to keep the consistency. Furthermore, I have modified the first paragraph in Section 3.1 to discuss why using PPO as the DRL algorithm and make the story flow better as follows:

``In this thesis, we first provide a comprehensive analysis on the trained DRL agents, which aims to understand features and policies that the agents learned in the environment. Specifically, proximal policy optimisation (PPO) [17] is selected as the DRL algorithm to train the model, because it is widely used in a variety of scenarios [8, 52, 7], we believe this analysis can provide more insight to the community. In the following chapters, we will further improve the sample efficiency of the PPO algorithm with different ways in the tasks with sparse or delayed rewards."

\item[3)] \textbf{Comment:} 6.1b Here make clear you are ignoring your agents but focus on PPO. maybe say why? e.g., everyone uses PPO and that is why I analysed PPO.\\ \\
\textbf{Response:} I have added the motivation that we select PPO as our DRL algorithm for the analysis in the first paragraph of Section 3.1 (more details can be found in the above response).

\item[4)] \textbf{Comment:} What does the LSTM actually learn? Briefly summarise the finding. Is it surprising?\\ \\
\textbf{Response:} I have added the corresponding finding about LSTM layers in Section 3.4.6 as follows:

``One possible explanation is that when DR is not activated during training, the textures and colours of the background in the environment remain unchanged, the agent only needs to learn the policy that keeps moving towards the target object in each frame. Therefore, in this case, the sequence information is not that crucial, and the LSTM layer does not play an important role. When DR is activated during training, the textures and colours of the background in the environment change at every frame. When the textures or colours of the background are similar to the target object, the agent needs to use the sequence information to remember the location of the target object in the previous frames to distinguish the target from irrelevant information and complete the task. Therefore, when DR is activated during training, the LSTM layer needs to capture the location of the target object in previous frames."

\item[5)] \textbf{Comment:} 6.4.2 Make it clear that the numbers in the second paragraph are relative changes and not absolute and relate these to the tables.\\ \\
\textbf{Response:} I have clarified that the numbers in the second paragraph are relative changes in Section 3.4.2 as follows:

``...which reduces the success rate by 0.48 (the relative changes in its performance in the standard test environment, which is 1.00$\pm$0.00; see Table 3.4)."

``The agent trained with visual inputs and without DR merely drops 0.17 success rate (the relative changes in its performance in the standard test environment, which is 1.00$\pm$0.00; see Table 3.4)."

\item[6)] \textbf{Comment:} Table 6.4 and 6.5 are not well readable, drop accuracy after the decimal point to make the fonts bigger.\\ \\
\textbf{Response:} I have adjusted the font size and precision of Table 3.4 and Table 3.5  to make them more readable.

\end{enumerate}

\newpage
\section*{Response to Chapter 4 \textcolor{blue}{(Original Chapter 3)}}
\begin{enumerate}
\item[1)] \textbf{Comment:} 3.1 Appendix A figure in slides please in thesis and explain there. Work it into the HER, SIL, ESIL explanations.\\ \\
\textbf{Response:} I have added the illustration in Figure 4.2 and added the corresponding explanation in Section 4.1 as follows:

``We further provide an illustration (Figure 4.2) to distinguish between the different types of buffer used in each algorithm in this chapter. Figure 4.2a shows the structure of an \textit{episodic buffer}, which is used in the proposed ESIL. In the episodic buffer, each entry stores a complete episode (i.e., trajectory) with the same length $T$. In each update iteration, all episodes in this buffer will be used for training. After each update iteration, the stored episodes in this buffer will be removed. The episodic buffer only stores the episodes that are collected by the current policy of the agent. Figure 4.2b shows the structure of a standard \textit{replay buffer}, which is used by SIL and SIL+HER in this chapter. Instead of storing a complete episode in each entry, the replay buffer stores a transition in each entry. In each update iteration, batch-sized transitions are sampled for training. After each update iteration, the buffer will not be cleaned. The replay buffer will remove the oldest transition when it is full and a new transition needs to be stored. Figure 4.2c shows the structure of an \textit{episodic replay buffer}, which is used by DDPG+HER in this chapter. Similar to episodic buffer, the episodic replay buffer stores a complete episode with the same length in each entry. In each update iteration, batch-sized episodes will be uniformly sampled. Then, one transition will be uniformly sampled from each selected episode to form batch-sized transitions for training. After each update iteration, the buffer will not be cleaned. The episodic replay buffer will remove the oldest episode when it is full and a new episode needs to be stored."

\item[2)] \textbf{Comment:} Figure 3.2 please write an expanded explanation in the main text working us through what the figures show (relate it to the other content in the chapter).\\ \\
\textbf{Response:} I have added an expanded explanation in Section 4.3.2 as follows:

``An original trajectory collected by an agent can be divided into two categories: 1) the agent never reaches the goal and 2) the agent has reached its original goal before. Figure 4.3A shows the case where the agent never reaches the goal. In this case, the original experiences never contain a non-negative reward. However, the hindsight experience will have at least one transition that has a non-negative reward (e.g., the final step). According to the rules for calculating the returns, all hindsight experiences (solid orange trajectory) have higher returns than the original experiences (solid blue trajectory). Therefore, in this case, all hindsight experiences can be selected to benefit the training. Figure 4.3B demonstrates the case that the agent has achieved its original goal in the third timestep and thus received a non-negative reward. However, for the hindsight experience, it only achieves a non-negative reward in the end of the trajectory. Therefore, in the first three steps, the original experiences have higher returns than the hindsight experiences (dashed orange trajectory), which means that the original experiences have better quality than the hindsight experiences, and these hindsight experiences are ignored from the training. For the remainder of the episode of Figure 4.3B, the hindsight experiences (solid orange trajectory) have higher returns than the original ones, and thus these hindsight experiences are selected to accelerate the training."

\item[3)] \textbf{Comment:} Linking $\rightarrow$ In section 2.3.2. Clarify the notion of state and goal in the context of the illustrations and more generally. Is a goal a state?
\\ \\
\textbf{Response:} I have added the clarification of the notion of state and goal in Section 2.3.2 as follows:

``To be clarified, the desired goal $g$ should be part of the state in RL. However, in order to highlight the specificity of different goal positions in goal-oriented RL, we split the desired goal out of the state. During training or evaluation, the desired goal and state are still concatenated together as the input of the neural network."

\item[4)] \textbf{Comment:} In section 3.4.1. Empty room paragraph, target position is randomly selected *at* each episode, needs to be clarified. Add citation if you want to say this is standard.
\\ \\
\textbf{Response:} I have added the clarification that the target position in Empty Room is randomly selected at the beginning in Section 4.4.1. I also add the related citation to explain this is a standard solution.
\end{enumerate}

\newpage
\section*{Response to Chapter 5 \textcolor{blue}{(Original Chapter 4)}}
\begin{enumerate}
\item[1)] \textbf{Comment:} Why is this not one chapter with Chapter 3:
Chapter 3 changes loss function, Chapter 4 changes sampling.\\ \\
\textbf{Response:} I have discussed the reason that why use two independent chapters to introduce two algorithms that related to hindsight experience replay in Section 5.1.

\item[2)] \textbf{Comment:} 4.2 in preview to 4.3 add small table that lists in each row a sampling strategy (method), paper citation, and description, to structure ”the field”. Perhaps add briefly a shortcoming for each method. Perhaps use this table to “derive” why your method was the logical conclusion to overcome the shortcomings of those methods.\\ \\
\textbf{Response:} I have added Table 5.1 in Section 5.2 to list the sampling strategies for each method and discussed the limitations of these methods and how our approach can overcome these shortcomings.

\item[3)] \textbf{Comment:} Figure 4.5 and 4.6 and content - why not training models longer, e.g. 4.5.d seems not to have converged. If you think about these figures as evidence of your methods being better, you need to qualify that (4.5.a is clearly better, 4.5.c not so clear). This needs to be discussed in the discussion in terms of the computational limitations of chapter 4, and in results clearly stated that for some cases this is inclusive as you did not have enough computer time. Make clear that you ran these models for months, so it does not appear you were lazy.\\ \\
\textbf{Response:} I have added the discussion about the limitations of the experiments, which are caused by the lack of sufficient computing resources in Section 5.5 as follows:

``There are still several limitations in some of the experiments, where the models do not have enough training time to converge in Figure 4.5c and Figure 4.5d. There are two main reasons for this, on the one hand we use the same experimental setup as CHER [69] for fair comparison and on the other hand we do not have enough computing resources to complete training with more epochs. In this chapter, it took us more than one month to finish these experiments. In the future, once we have enough computing resources, we will re-evaluate the method for longer training epochs."

\item[4)] \textbf{Comment:} 4.4.5 This subsubsection is just 2 sentences long this is not appropriate, it should be completed with reasoning of what the results in the table mean and where your method stands.\\ \\
\textbf{Response:} Followed by the comment 5 in this chapter, I have moved and added more discussion about the time complexity of baseline methods and DTGSH in Section 5.5 as follows:

``Furthermore, Table 5.3 gives example training times of all of the HER-based algorithms. From the table, HER requires the minimum time, because it only uses the naive uniform sampling strategy for both trajectory and transition. HEBP needs a bit more training time than HER, because it requires additional computation for the energy of each trajectory. CHER displays the longest training time, because it requires an inefficient greedy algorithm [69] for the sampling according to the proximity and diversity of candidate goals. DTGSH is more efficient than CHER, but slower than the other two methods. It needs an additional calculation of the diversity score of $\mathcal{O}(N_{p}b^3)$ at the end of every training episode, and sampling of $\mathcal{O}(mk^2)$ for each minibatch. Although DTGSH has better sample efficiency than other baselines, it needs a longer training period. In the future, we will improve the method towards achieving the balance between performance and time complexity."

\item[5)] \textbf{Comment:} 4.5 needs more of a discussion (1 paragraph is a bit short and short on references to the field). You need to relate to the table mentioned above (this could be even moved here) you should relate to other methods. You should also discuss the limitations of your computer simulation results.\\ \\
\textbf{Response:} In Section 5.5, I have expanded more discussions, moved related discussion about time complexity here, and discussed the limitations and future work about DTGSH.
\end{enumerate}

\newpage
\section*{Response to Chapter 6 \textcolor{blue}{(Original Chapter 5)}}
\begin{enumerate}
\item[1)] \textbf{Comment:} Chapter 3 what is the strategic difference to chapter 5?
Chapter 3 increases the chance of seeing a (sparse) reward and Chapter 5 adds an endogenously defined reward term to help with the reward sparsity problem. This logic should be framed and phrased at the start of chapter 5 to make clear what changes. This difference should be also highlighted in 7.1 when relating to the chapter\\ \\
\textbf{Response:} I have added the changes between previous chapters and the current chapter at the beginning of Section 6.1. I have also highlighted the difference between different proposed methods in the summary of thesis achievements in Section 7.1.

\item[2)] \textbf{Comment:} If so, why not combine both methods - if you were planning to do it, suggest it as an extension of your work in the discussion of chapter 5 (sec 5.5) and also thus in 7.2\\ \\
\textbf{Response:} I have added the combination between ESIL and intrinsic motivation as the future work in both Section 6.5 and Section 7.2.1.

\item[3)] \textbf{Comment:} 5.4 (p.103) extend the end of the top paragraph to give the good explanation. in fact you drew a nice graphic in the viva, you can use it to illustrate your point.\\ \\
\textbf{Response:} I have added more explanation in Section 6.4 ({p.128}) with an illustration (see Figure 6.7) as follows:

``For example, in Figure 6.7, the agent selects an action in $s_{t}$, which causes it to fall into a trap in $s_{t+1}$. When the raw observation feature is used as the feature embedding method, the selected action leads to high diversity between the adjacent states, however the agent fails in this episode. Therefore, simply increasing diversity may not lead to better performance and it is necessary to learn the features that can better quantify the diversity between adjacent states to improve the task-specific performance (i.e., extrinsic returns) of the agent."

\item[4)] \textbf{Comment:} 5.5 you state that complexities are similar but you have not discussed these, a table and explanatory words in the main text would help here a lot. Also link the logic of the paragraphs together, it is not good style to have one sentence paragraphs.\\ \\
\textbf{Response:} In Section 6.5, I have added the training time of DAIM and other baselines in Table 6.4 to support the state that DAIM has similar complexity to other baseline methods. Furthermore, I have linked the paragraphs together in this section to improve the style.
\end{enumerate}

\newpage
\section*{Response to Chapter 7}
\begin{enumerate}
\item[1)] \textbf{Comment:} 7.1 summarise first how your achievements relate to your thesis title - a few sentences suffices.\\ \\
\textbf{Response:} I have re-organised the first paragraph of Section 7.1 to demonstrate how my achievements relate to my thesis title as follows:

``In this thesis, in order to improve the sample efficiency of deep reinforcement learning (DRL) in the environments with sparse or delayed rewards, three novel DRL methods are proposed. In these methods, we propose 1) a novel self-imitation learning loss term that allows the agent to overcome hard exploration problems by using hindsight experiences, 2) a novel sampling strategy that increases the chance to sample valuable transitions for the training and 3) a diversity-based intrinsic reward that can provide auxiliary dense feedback to the agent for helping the exploration in the environments with delayed or sparse rewards. From extensive experiments, these methods achieve better performance than other baseline approaches when using the same number of samples for training, which demonstrates that our methods indeed improve the sample efficiency of DRL algorithms. Moreover, we provide a comprehensive quantitative and qualitative analysis on the agents trained with and without domain randomisation (DR) to visualise and explain why DR can improve the robustness of the agents against unseen scenarios."

\item[2)] \textbf{Comment:} Chapter 7.2 you tie it up by saying you would like to tie the interpretability with your new methods together and compare to PPO, as perhaps the features have changed.\\ \\
\textbf{Response:} I have added this work in Section 7.2.1 as the future work of Chapter 3 as follows:

``Another future work is to apply the related interpretability methods to our proposed methods in other chapters and then compare the results with the PPO algorithm, because the learned feature may have changed and thus some interesting findings may be obtained to explain why these methods can improve the performance of vanilla PPO algorithm."
\end{enumerate}

\end{document}

### 2022-08-21 abstract/abstract.tex
update the final version

@@
 \begin{abstract}
-Deep reinforcement learning (DRL) has made great progress in dealing with complex control problems in \td{various testing scenarios}, such as playing video games, playing board games, and dexterous robotic manipulation; \td{and real-world applications, such as controlling plasmas for nuclear fusion.} However, DRL requires large amounts of interactions with an environment to find an optimal policy to solve the task, limiting its application in real-world problems.
+Deep reinforcement learning (DRL) has made great progress in dealing with complex control problems in various test scenarios, such as playing video games, playing board games, and dexterous robotic manipulation, with the promise of critical real-world applications, such as controlling plasmas for nuclear fusion.
+However, DRL requires large amounts of interactions with an environment to find an optimal policy to solve the task, limiting its application in real-world problems.
 In this thesis, we focus on two aspects to improve sample efficiency in DRL: 1) solving sparse reward tasks and 2) improving general exploration strategies.
 
-First, we propose two methods to overcome exploration difficulties and improve learning efficiency in goal-oriented RL with the sparse reward setting, where an agent can rarely achieve positive feedback.
-In the first method, to provide sufficient positive samples for training an agent, \td{hindsight goal relabelling is used to replace goals in original samples with intermediate goals, and these augmented positive samples are leveraged to accelerate the training via a self-imitation learning paradigm.} An additional selection module is also designed to remove undesirable modified samples and stabilise training.
+First, we analyse the trained agents with and without domain randomisation (DR), a technique that can reduce the reality gap between a simulator and real-world scenarios.
+Through evaluating their robustness to previous unseen environments and applying both qualitative and quantitative interpretability methods, we provide the insight into the behaviour of trained agents.
+Finally, some suggestions are also given to researchers who intend to adopt interpretability methods to analyse DRL agents.
+
+Second, we propose two methods to overcome exploration difficulties and improve learning efficiency in goal-oriented RL with the sparse reward setting, where an agent can rarely achieve positive feedback.
+In the first method, to provide sufficient positive samples for training an agent, hindsight goal relabelling is used to replace goals in original samples with intermediate goals, and these augmented positive samples are leveraged to accelerate the training via a self-imitation learning paradigm.
+An additional selection module is also designed to remove undesirable modified samples and stabilise training.
 In the second method, to alleviate the inefficiency of hindsight experience replay (HER) caused by its uniform sampling strategy, a diversity-based sampling method is employed to select valuable and diverse experiences for efficient training.
 
-Second, a diversity-augmented intrinsic motivation is introduced to encourage the agent to explore novel states in an environment with sparse or delayed rewards.
+Furthermore, diversity-augmented intrinsic motivation is introduced to encourage the agent to explore novel states in an environment with sparse or delayed rewards.
 During training, the diversity of adjacent state sequences is measured under the framework of determinantal point processes (DPPs) and this measurement is used as an auxiliary reward to facilitate the exploration of the agent, thus improving the final performance.
 
-Furthermore, we analyse the trained agents with and without domain randomisation (DR), a technique that can reduce the reality gap between a simulator and real-world scenarios.
-Through evaluating their robustness to previous unseen environments and applying both qualitative and quantitative interpretability methods, we provide the insight into the behaviour of the trained agents.
-Finally, some suggestions are also given to researchers who intend to adopt interpretability methods to analyse DRL agents.
-
 \end{abstract}

### 2022-08-21 acknowledgements/acknowledgements.tex
update the final version

@@
  \item Kai Arulkumaran: Kai is my closest colleague/collaborator in the BICI group.
-He has helped me a lot in my research.
-In addition, he also gives me many suggestions in daily life.
+He has been a constant and continuing collaborator.
+In addition, he also has provided me many suggestions in daily life.
 He is also my ``leader" in the field of reinforcement learning.
@@
  \vspace*{3mm}
- \item My parents: From the day I was born, my parents have provided me with unfailing love and support. Without them, I wouldn't be here today.
+ \item Meng Fang: Dr. Meng Fang was my mentor during my internship at Tencent AI Lab and he helped me a lot in the research of deep reinforcement learning.
+ \item Shanxin Yuan: Dr. Shanxin Yuan was my mentor during my internship at Huawei Noah's Ark Lab and he helped me a lot in the research of low-level computer vision.
  \vspace*{3mm}
- \item My wife: From the day we meet, she continually inspires me to chase my dreams and gives me much love and care.
-During our two years together in the UK, she has been there to guide me through difficult times, to help me with my worries and to allow me to focus on my work.
-She makes my life so full of hope and joy that no words can express my gratitude and love to her.
- % 从我们认识的那天起，她就给予我充足的关怀以及爱。我们一起在英国的这两年间，她每天为我准备美味的食物以及提供舒适的生活环境，让我能够专注于我自己的工作。也是她让我的生活充满了希望与快乐，再多的文字也表达不出我对她的感激与爱。
+ \item My Parents: From the day I was born, my parents have provided me with unfailing love and support. Without them, I wouldn't be here today.
+ \vspace*{3mm}
+ \item My Wife: From the day we meet, she continually inspires me to pursue my dreams and gives me much love and care.
+During our two years together in the UK, she has been there to guide me through difficult times, to help me with my worries, and to allow me to focus on my work.
+She makes my life so full of hope and joy that no words can express my gratitude and love for her.
 \end{itemize}

### 2022-08-21 background/background.tex
update the final version

@@
 \label{ch:background}
-In this chapter, the background covered in this thesis is presented according to the following structure: in Section~\ref{ch3:dnn} and Section~\ref{ch3:bp}, we first briefly discuss deep neural network (DNN), back-propagation and optimisation methods.
+In this chapter, the background covered in this thesis is presented according to the following structure: in Section~\ref{ch3:dnn} and Section~\ref{ch3:bp}, we first briefly discuss deep neural networks (DNNs), back-propagation and optimisation methods.
 Then, in Section~\ref{ch3:rl}, some background to reinforcement learning (RL) is introduced.
@@
     \centering
-    \includegraphics[width=0.7\textwidth]{figures/background/mlp_background.pdf}
-    \caption{A simple structure of a multi-layer perceptron (MLP).
+    \includegraphics[width=0.9\textwidth]{figures/background/NN.pdf}
+    \caption[A simple Structure of an MLP.]{A simple structure of a multi-layer perceptron (MLP).
 It consists of an input layer, two hidden layers and an output layer.
-The highlighted lines (\textcolor{red}{red}) show the connection between a single neuron and its previous layer.}
+The highlighted lines (red) show the connection between a single neuron and its previous layer.}
     \label{fig:nn}
@@
 A feed-forward neural network (FFNN) or multi-layer perceptron (MLP) is a standard structure of ANN, and it consists of several neuron layers, which include an input layer, one or several hidden layers and an output layer (see Figure~\ref{fig:nn}).
-In each layer, there are multiple neurons, and each neuron is connected to those in the previous layer with corresponding weights $w_{i}$ and a bias $b$.
-For example, in Figure~\ref{fig:nn}, the transmission of the signal between the highlighted neuron and the previous layer can be expressed as:
+In each layer, there are multiple neurons.
+Consider a single neuron connected to those in the previous layer with weights $w_{i}$ and a bias $b$.
+For example, in Figure~\ref{fig:nn}, the transmission of the signal between the highlighted neuron and the previous layer can be expressed in a general form as:
+% \begin{equation}
+%     f = \sigma (w_{0}x_{0} + w_{1}x_{1} + w_{2}x_{2} + b)
+% \label{eq:nn_compute}
+% \end{equation}
 \begin{equation}
-    f = \sigma (w_{0}x_{0} + w_{1}x_{1} + w_{2}x_{2} + b)
+    %f = \sigma (w_{0}x_{0} + w_{1}x_{1} + w_{2}x_{2} + b)
+    f = \sigma (b + \sum_{i=0}^{N} w_{i}x_{i})
 \label{eq:nn_compute}
 \end{equation}
-here, $\sigma(\cdot)$ represents an activation function, such as a rectifier linear unit (ReLU)~\cite{nair2010rectified}, sigmoid$(\cdot)$ and tanh$(\cdot)$, which are used to add non-linearity to the network, because most real-world problems are nonlinear.
-The hidden layers in a neural network are used to learn various non-linear combinations of the input data, and the learned non-linear combinations are also called features.
+where, $N$ is the number of input neurons from the previous layer and $N$ is 3 in the above example.
+$\sigma(\cdot)$ represents an activation function, such as a rectifier linear unit (ReLU)~\cite{nair2010rectified}, sigmoid$(\cdot)$ and tanh$(\cdot)$, which are used to add non-linearity to the network, because most real-world problems are nonlinear.
+The hidden layers in a neural network are used to learn various non-linear combinations of the input data, and the learned non-linear combinations are sometimes called features.
 These features are the key factors to decide the prediction of the network.
@@
 Then, in each hidden layer, the output of each neuron is computed according to Equation~\eqref{eq:nn_compute}, and this computation proceeds from the first hidden layer to the last hidden layer.
-Finally, in the output layer, the output of the network can be the probability of each class in a classification task, the predicted value in a regression task, or the policy/predicted state value in a DRL task.
+Finally, in the output layer, the output of the network can be the preferences of each class in a classification task, the predicted value in a regression task, or the policy/predicted state value in a DRL task.
 
@@
 % CNN
+%As the input size increases, the number of parameters (weight and bias) also increases.
 \subsection{Convolutional Neural Network}
-Multi-layer perceptrons (MLPs) are somewhat inefficient in dealing with computer vision tasks because of its fully connected structure.
-As the input size increases, the number of parameters (weight and bias) also increases.
-Convolutional neural network (CNN)~\cite{lecun1998gradient} alleviates this problem by using learnable kernels/filters, where each filter traverses the entire image.
-\td{CNNs are a strong inductive bias that reflects translation invariance of the physical world.} Compared with an MLP, the parameters of these kernels are shared, less, and easy to train.
+Multi-layer perceptrons (MLPs) are somewhat inefficient in dealing with computer vision tasks because of their fully connected structures.
+When the MLP is used to classify an image, each node in the hidden layer (the next layer of the input layer) is connected to all pixels of the input image.
+Thus, there will be a large number of parameters in the MLP when the size of the input image is large.
+For example, when the size of the input image is $256\times256$, the first hidden layer has $256\times256\times N$ parameters to learn, and $N$ is the number of neurons in this layer.
+Therefore, it is essential to find an approach to substantially reduce parameters.
+Convolutional neural networks (CNNs)~\cite{lecun1998gradient} alleviate this problem by employing the parameter sharing strategy and local connectivity.
+Instead of connecting each node to all nodes in the previous layer, CNNs use a set of learnable filters with relatively small size (e.g.
+$9\times9$) to slide across the entire image to produce the features.
+In this case, the first convolution layer only has $9\times9\times M$ parameters to learn, and $M$ is the number of filters, therefore CNNs greatly reduce the number of parameters.
+Benefiting from parameter sharing, CNNs might have an implicit strong inductive bias that reflects translation invariance in some desired mapping from physical world to feature space.
+Compared with an MLP, the parameters of these kernels are shared, fewer in number, and easier to train.
 In previous works~\cite{krizhevsky2012imagenet,girshick2014rich,detone2016deep}, CNN has demonstrated its better feature representation ability than other hand-crafted methods, such as histogram of oriented gradient (HOG)~\cite{dalal2005histograms} and scale-invariant feature transform (SIFT)~\cite{lowe2004distinctive}.
@@
 %\td{[shifting and scaling in-variance makes CNN important]} 
+%\newpage
+%otherwise it may cause aliasing of the input.
 
@@
 Gradient descent~\cite{bottou2010large} is a common technique used to update the parameters of the deep neural network (DNN) to fit a model, and it relies on the gradient of an objective function/loss function with respect to the parameters of the network.
-The gradient can be calculated efficiently through using the back-propagation algorithm~\cite{kelley1960gradient}.
+The gradient can be calculated efficiently through the back-propagation algorithm~\cite{kelley1960gradient}.
 In the forward propagation of an MLP with $N$ layers, the output of each hidden layer can be formulated as:
@@
 
-Once get the gradient $\nabla_{\bm{\theta}}\mathcal{L}$, the parameters $\bm{\theta}$ can be updated in the opposite direction of the gradient with a step size $\eta$ (i.e.
+Once we get the gradient $\nabla_{\bm{\theta}}\mathcal{L}$, the parameters $\bm{\theta}$ can be updated in the opposite direction of the gradient with a step size $\eta$ (i.e.
 learning rate) via gradient descent:
@@
 
-Root mean square propagation (RMSProp) can adaptively adjust the learning rate for different parameters (i.e., per-parameter learning rate) to accelerate the convergence, the update rule can be written as:
+Root mean square propagation (RMSProp) can adaptively adjust the learning rate for different parameters (i.e., per-parameter learning rate) to accelerate the convergence, and the update rule can be written as:
 \begin{align}
@@
 \end{align}
-In Equation~\eqref{eq:squared_grad}, $\mathbf{v}$ is the moving average of the squared gradient, $\odot$ is the element-wise multiplication, and $\beta$ is the decay coefficient which indicates the ratio of history information used to compute the average value.
-In Equation~\eqref{eq:rmsprop_update}, $\epsilon$ is a small value to prevent division by zero, $\frac{1}{(\epsilon + \sqrt{\mathbf{v}})}$ is the weight to balance the learning rate for individual parameter, in doing so, the learning rate for large gradient will decrease to avoid exploding, and the learning rate for small gradient will increase to avoid vanishing.
+In Equation~\eqref{eq:squared_grad}, $\mathbf{v}$ is the moving average of the squared gradient, $\odot$ is element-wise multiplication, and $\beta$ is the decay coefficient which indicates the ratio of history information used to compute the average value.
+In Equation~\eqref{eq:rmsprop_update}, $\epsilon$ is a small value to prevent division by zero, $\frac{1}{(\epsilon + \sqrt{\mathbf{v}})}$ is the weight to balance the learning rate for individual parameter, in doing so, the learning rate for a large gradient will decrease to avoid exploding, and the learning rate for small gradient will increase to avoid vanishing.
 
-Adam is a combination of momentum~\cite{qian1999momentum} and RMSProp, and it is the most popular optimiser in DRL community. The update rule of Adam is:
+Adam is a combination of momentum~\cite{qian1999momentum} and RMSProp, and is the most popular optimiser in the DRL community. The update rule of Adam is:
 \begin{align}
@@
 Equation~\eqref{eq:momentum_adam} is the momentum term, $\mathbf{m}$ is the running average of the gradient, and $\beta_{1}$ is the decay coefficient.
-The momentum term can correct the gradient in the right direction and thus accelerate the convergence.
+The momentum term can correct the gradient in the right direction, and thus accelerate the convergence.
 The other parts are similar to RMSProp.
@@
 \end{equation}
-Here, the discount factor $\gamma$ exponentially downplays the influence of future rewards.
+Here, the discount factor $\gamma$ exponentially diminishes the influence of future rewards.
 {In a robot manipulation task, the state $s_{t}$ can be the velocity and position of each joint of a robot arm.
@@
 In this thesis, the experiments in Chapter~\ref{ch:esil} and Chapter~\ref{ch:dtgsh} follow the terminology suggested by OpenAI~\cite{plappert2018multi}, in which the possible goals are drawn from a goal space $\mathcal{G}$, and the goal being pursued does not influence the environment dynamics.
-Two types of goal are recognised.
+Two types of goals are recognised.
 One is the desired goal $g\in \mathcal{G}$, which is the target position {or state}, and may be different for different episodes.
@@
 In an episode, each transition can be represented as $(s_t|\langle g, g^{ac}_{t} \rangle, a_t, r_t, s_{t+1}|\langle g, g^{ac}_{t+1} \rangle)$, where $s_t$ indicates a state, $a_t$ indicates an action and $r_t$ indicates a~reward; $\langle \, , \rangle$ is simply used to represent the grouping of goals.
+To be clarified, the desired goal $g$ should be part of the state in RL.
+However, in order to highlight the specificity of different goal positions in goal-oriented RL, we split the desired goal out of the state.
+During training or evaluation, the desired goal and state are still concatenated together as the input of the neural network.
 
@@
     \centering
-    \includegraphics[width=0.85\textwidth]{figures/background/her_background.pdf}
-    \caption{A brief illustration of the hindsight goal relabelling in a robotic manipulation task~\cite{plappert2018multi}.
+    \includegraphics[width=0.85\textwidth]{figures/background/her_background_modified.pdf}
+    \caption[A brief illustration of Hindsight Experience Replay (HER).]{A brief illustration of the hindsight goal relabelling in a robotic manipulation task~\cite{plappert2018multi}.
 In this task, the agent needs to control the gripper to hit the puck and make it slide to a specific position, where the achieved goal is the position of the puck and the desired goal is the target position (the red spot).}
@@
 Thus, hindsight experiences~\cite{andrychowicz2017hindsight} (see Figure~\ref{fig:her_back}) are created through replacing the original desired goal $g$ with the current achieved goal $g_{t+1}^{ac}$ to augment successful samples, and then reward $r_{t}$ can be recomputed according to Equation~\eqref{eq:sparse_reward}.
-The modification of the desired goal can be denoted as $g^\prime$ and transitions from hindsight experiences can be represented as $(s_t|\langle g^\prime, g^{ac}_{t} \rangle, a_t, r_t, s_{t+1}|\langle g^\prime, g^{ac}_{t+1} \rangle)$.} {Intuitively, introducing $g^\prime$ serves a useful purpose in the early stages of training; taking, for example, a robot reaching task, the agent has no prior ``concept" of how to move its effector to a specific location in space.
-Thus, even these original failed episodes contain valuable information for ultimately learning a useful control policy for the original, desired goal $g$.}
+The modification of the desired goal can be denoted as $g^\prime$ and transitions from hindsight experiences can be represented as $(s_t|\langle g^\prime, g^{ac}_{t} \rangle, a_t, r_t, s_{t+1}|\langle g^\prime, g^{ac}_{t+1} \rangle)$.}
+
+Intuitively, introducing $g^\prime$ serves a useful purpose in the early stages of training; taking, for example, a robot reaching task, the agent has no prior ``concept" of how to move its effector to a specific location in space.
+Thus, even these original failed episodes contain valuable information for ultimately learning a useful control policy for the original, desired goal $g$.
+\newpage
 
@@
 Value-based methods aim to learn a state or state--action value function instead of an explicit policy.
-The value function can evaluate the quality of a given state or the quality of an action performed at a given state.
-Then, the agent can select an action according to the corresponding estimate.
+The value function can evaluate the quality of a given state (state value function) or the quality of an action performed at a given state (Q-value function).
+Then, the agent can select an action according to the corresponding value.
 This section will discuss how to derive a policy through the value function.
@@
 \centering
-    V_{\pi}(s_{t}) = \mathbb{E}_{\pi}[G_{t}] = \mathbb{E}_{\pi}\biggr[\sum_{k=0}^{T-1} \gamma^{k}r_{t+k}\biggr].
+    V_{\pi}(s_{t}) = \mathbb{E}_{\pi}[G_{t}|s_{t}] = \mathbb{E}_{\pi}\biggr[\sum_{k=0}^{T-1} \gamma^{k}r_{t+k} | s_{t}\biggr].
 \end{equation}
-Through the Bellman equation, the state value function can be rewritten as the combination of the immediate reward, $r_{t}$, and the discounted state value of the next state, $V_{\pi}(s_{t+1})$:
+Through the Bellman equation~\cite{sutton2018reinforcement}, the state value function can be rewritten as the combination of the immediate reward, $r_{t}$, and the discounted state value of the next state, $V_{\pi}(s_{t+1})$:
 \begin{equation}
@@
 % state-action value function
-\td{Apart from estimating the state value,} we can also predict the value of a particular state--action pair, denoted as $Q_{\pi}(s_{t},a_{t})$, and this has a similar definition to the state value function:
+In addition to the state value, we can also use state-action value function to evaluate how good it is for an agent to take an action $a_{t}$ in a state $s_{t}$ with a policy $\pi$.
+This state-action value function can be denoted as $Q_{\pi}(s_{t},a_{t})$, and has a definition similar to that of the state value function:
 \begin{equation}
 \centering
-     Q_{\pi}(s_{t},a_{t}) = \mathbb{E}_{\pi}[G_{t}] = \mathbb{E}_{\pi}\biggr[\sum_{k=0}^{T-1} \gamma^{k}r_{t+k}\biggr].
+     Q_{\pi}(s_{t},a_{t}) = \mathbb{E}_{\pi}[G_{t}|s_{t}, a_{t}] 
+     = \mathbb{E}_{\pi}\biggr[\sum_{k=0}^{T-1} \gamma^{k}r_{t+k} | s_{t}, a_{t}\biggr].
 \end{equation}
-The state--action value function $Q_{\pi}(s_{t},a_{t})$ is also called the Q-value function. Similarly, the Q-value function can also be expressed in a recursive form:
+The state--action value function $Q_{\pi}(s_{t},a_{t})$ is also called the Q-value function. Similarly, the Q-value function can also be expressed in a recursive form via the Bellman equation:
 \begin{equation}
@@
 Next, the optimal policy can be derived using these value functions.
-In RL, the agent aims to follow the policy that can achieve more accumulated rewards from the environment.
-Thus, an optimal policy $\pi_{*}$ is defined as the policy that achieves the maximum/optimal state value function, $V_{*}(s_{t})$:
+In RL, the agent aims to learn the policy that can achieve more accumulated rewards from the environment.
+Thus, an optimal policy $\pi_{*}(s_{t})$ is defined as the policy that achieves the optimal state value function, $V_{*}(s_{t})=\max_{\pi}V_{\pi}(s_{t})$.
+For a state $s_{t}$, the optimal policy $\pi_{*}$ can be derived through evaluating all possible actions at the state $s_{t}$ and selecting the action that can maximise the value function:
 \begin{equation}
-    V_{*}(s_{t}) = \max_{\pi}V_{\pi}(s_{t}).
+    \pi_{*}(s_{t}) = \argmax_{a_{t}\in\mathcal{A}} \sum_{s_{t+1}\in\mathcal{S}} p(s_{t+1}|s_{t}, a_{t})V_{*}(s_{t+1}).
 \end{equation}
-Similarly, an optimal policy is also defined as the policy that achieves the maximum/optimal Q-value function $Q_{*}(s_{t},a_{t})$:
-\begin{equation}
-    Q_{*}(s_{t},a_{t}) = \max_{\pi}Q_{\pi}(s_{t},a_{t}).
-\end{equation}
-Particularly, the optimal policy $\pi_{*}$ can be achieved through this optimal Q-value function directly:
+% \begin{equation}
+%     V_{*}(s_{t}) = \max_{\pi}V_{\pi}(s_{t}).
+% \end{equation}
+Similarly, an optimal policy is also defined as the policy that achieves the optimal Q-value function, $Q_{*}(s_{t},a_{t})=\max_{\pi}Q_{\pi}(s_{t},a_{t})$.
+% \begin{equation}
+%     Q_{*}(s_{t},a_{t}) = \max_{\pi}Q_{\pi}(s_{t},a_{t}).
+% \end{equation}
+In particular, the optimal policy $\pi_{*}(s_{t})$ can be achieved directly through this optimal Q-value function, without the need to know the transition dynamics of the environment explicitly (i.e., model-free):
 \begin{equation}
@@
 \end{equation}
-Therefore, finding the optimal Q-value function is equivalent to finding the optimal policy. The optimal Q-value function can be reorganised using the Bellman optimality equation:
-% bellman optimality equation
-\begin{equation}
-\centering
-     Q_{*}(s_{t},a_{t}) = \mathbb{E}[r_{t} + \gamma\max_{a_{t+1}} Q_{*}(s_{t+1}, a_{t+1})].
-\end{equation}
-Q-learning~\cite{watkins1992q} is a tabular and iterative update method to solve this equation.
+Therefore, finding the optimal Q-value function is equivalent to finding the optimal policy. 
+% The optimal Q-value function can be reorganised using the Bellman optimality equation:
+% % bellman optimality equation
+% \begin{equation}
+% \centering
+%      Q_{*}(s_{t},a_{t}) = \mathbb{E}[r_{t} + \gamma\max_{a^{\prime}} Q_{*}(s_{t+1}, a^{\prime})].
+% \end{equation}
+
+In order to learn the above Q-value function, Q-learning algorithm~\cite{watkins1992q}, which is a tabular and iterative update method, can be adopted.
 In Q-learning, the state--action pairs and corresponding Q-values are stored in a table.
@@
 \begin{equation}
-    Q_{\pi}(s_{t}, a_{t}) \leftarrow Q_{\pi}(s_{t}, a_{t}) + \alpha(r_{t} + \gamma\max_{a_{t+1}}Q_{*}(s_{t+1}, a_{t+1}) - Q_{\pi}(s_{t}, a_{t})),
+    Q_{\pi}(s_{t}, a_{t}) \leftarrow Q_{\pi}(s_{t}, a_{t}) + \alpha(r_{t} + \gamma\max_{a^{\prime}}Q_{*}(s_{t+1}, a^{\prime}) - Q_{\pi}(s_{t}, a_{t})),
 \end{equation}
@@
 \subsubsection{Function Approximator and Deep Q-Network}
-\td{Q-learning in its tabular form}, is impractical to scale to tasks with high-dimensional states, such as playing a video game using images as inputs.
+Q-learning in its tabular form is impractical for scaling to tasks with high-dimensional states, such as playing a video game using images as inputs.
 The Deep Q-Network (DQN)~\cite{mnih2015human} overcomes this difficulty by combining deep learning and Q-learning.
@@
 here, $\theta$ represents the parameters of a main network.
-During training, the agent collect samples from the environment with the $\epsilon$-greedy exploration strategy, and each transition ($s_{t}$, $a_{t}$, $r_{t}$, $s_{t+1}$) is stored in a replay buffer $\mathcal{D}$.
+During training, the agent collects samples from the environment with the $\epsilon$-greedy exploration strategy, and each transition ($s_{t}$, $a_{t}$, $r_{t}$, $s_{t+1}$) is stored in a replay buffer $\mathcal{D}$.
 The replay buffer is designed to reuse past experiences.
@@
 \begin{equation}
-    \min_{\theta}\mathcal{L} = \mathbb{E}_{(s_{t}, a_{t}, r_{t}, s_{t+1})\sim\mathcal{D}}[(Q(s_{t}, a_{t};\theta) - (r_{t} + \gamma\max_{a_{t+1}}Q(s_{t+1}, a_{t+1};\theta^{-})))^{2}],
+    \min_{\theta}\mathcal{L} = \mathbb{E}_{(s_{t}, a_{t}, r_{t}, s_{t+1})\sim\mathcal{D}}[(Q(s_{t}, a_{t};\theta) - (r_{t} + \gamma\max_{a^{\prime}}Q(s_{t+1}, a^{\prime};\theta^{-})))^{2}],
 \end{equation}
@@
 The target network has the same structure as the main network, and its parameters are copied from the main network at each fixed step.
-The target network is also used to stabilise the training.
+In DQN, the target network is proposed to stabilise the training.
+This is because in the standard online Q-learning update, the update that leads to an increase in the current Q-value, $Q(s_{t}, a_{t})$, can also increase the Q-value of the next state and all actions, $Q(s_{t+1}, a)$.
+Hence, the estimated target Q-value will also increase and may cause divergence or oscillations of the policy.
+Therefore, using the target network with the old parameters can introduce a delay between the updated Q-value and the estimated target Q-value, and thus prevent policy divergence~\cite{mnih2015human}.
+\newpage
 
@@
 % policy based
-Policy-based methods learn a policy function $\pi(a_{t}|s_{t};{\theta})$, parameterised by $\theta$, and it directly maps a given state $s_{t}$ to an action $a_{t}$.
+Policy-based methods learn a policy function $\pi(a_{t}|s_{t};{\theta})$, parameterised by $\theta$, which directly maps a given state $s_{t}$ to an action $a_{t}$.
 In discrete control tasks, the output of the policy function is a categorical distribution over the action space.
@@
 During training, the actions are sampled from the corresponding distribution for exploration.
-In policy-based methods, the optimal parameters $\theta$ can be found by maximising expected returns in the environment:
+In policy-based methods, the optimal parameters $\theta$ can be found by maximising the expected returns in the environment:
 \begin{align}
@@
 \end{align}
-where $d(s_{t})$ is the stationary distribution in the Markov chain under the policy $\pi_{\theta}$.
-In order to maximise the expected returns, we need to increase the probability of the actions which lead to higher returns, and the gradient of the objective function can be expressed as:
+where $d(s_{t})$ is the stationary distribution of states.
+In order to maximise the expected returns, we need to increase the probability of taking actions which lead to higher returns, and the gradient of the objective function can be expressed as:
 \begin{align}
@@
 where Equation~\eqref{eq:reinforce} is also known as REINFORCE~\cite{williams1992simple}, and the parameters are updated via gradient ascent.
-In practice, the return value $G_{t}$ is calculated using Monte Carlo methods, leads to a high variance in the estimated gradient, and slows the convergence of the policy function.
+In practise, the return value $G_{t}$ is calculated using Monte Carlo methods, leads to a high variance in the estimated gradient, and slows the convergence of the policy function.
 In order to alleviate the above shortcomings, actor-critic methods are introduced.
-
+\newpage
 \subsection{Actor-Critic Methods}
@@
 The actor is the policy function $\pi(a_{t}|s_{t};\theta)$ to generate an action for a given state.
-The critic is the value function $V(s_{t};\psi)$ or Q-value function $Q(s_{t},a_{t};\psi)$ to evaluate the quality of a state or a state--action pair and thus guides the training of the actor.
-To reduce the variance in policy-based methods, a common practice is to subtract a baseline from the estimated return $G_{t}$~\cite{sutton2018reinforcement}.
+The critic is the state value function $V(s_{t};\psi)$ or Q-value function $Q(s_{t},a_{t};\psi)$ to evaluate the quality of a state or a state--action pair and thus guides the training of the actor.
+To reduce the variance in policy-based methods, a common practise is to subtract a baseline from the estimated return $G_{t}$~\cite{sutton2018reinforcement}.
 In REINFORCE, the state value predicted by the critic $V(s_{t};\psi)$ is selected as the baseline:
@@
 \end{equation}
-where $A(s_{t}, a_{t})$ is an advantage function, and the gradient of the policy function can be rewritten as:
+where $A(s_{t}, a_{t})$ is an advantage function and the gradient of the policy function can be rewritten as:
 \begin{equation}
@@
 However, A2C does not perform well in continuous control tasks.
-In this paper, two actor-critic DRL methods -- deep deterministic policy gradient (DDPG)~\cite{lillicrap2015continuous} and proximal policy gradient (PPO)~\cite{schulman2017proximal} -- are adopted as the base algorithms to address the continuous control tasks.
+In this thesis, two actor-critic DRL methods -- deep deterministic policy gradient (DDPG)~\cite{lillicrap2015continuous} and proximal policy gradient (PPO)~\cite{schulman2017proximal} -- are adopted as the base DRL algorithms to address continuous control tasks.
+\newpage
 
@@
 For each update iteration, a minibatch sized transitions is sampled to update the actor and the critic.
-The actor aims to maximise the expected returns $\mathbb{E}[Q(s_{t}, a_{t};\psi)|_{a_{t}=\pi(s_{t};\theta)}]$ and the gradient of the objective function can be estimated using chain rule:
-
+The actor aims to maximise the expected return $\mathbb{E}[Q(s_{t}, a_{t};\psi)|_{a_{t}=\pi(s_{t};\theta)}]$ and the gradient of the objective function can be estimated using the chain rule:
 \begin{align}
@@
 \end{align}
-The update of the critic is similar to DQN which minimises the MSE between the target Q-vaule and the estimated Q-value:
+The update of the critic is similar to DQN, which minimises the MSE between the target Q-value and the estimated Q-value:
 \begin{equation}
@@
 here, $\tau\in$(0,1) is a coefficient that decides the updating speed of the target networks. 
+\newpage
 
@@
 In addition, a line search method can be used to estimate an appropriate step size, with the conjugate gradient method being used to estimate the gradient.
-Although TRPO achieves great performance in continuous control tasks, it is not flexible to be implemented and it performs poorly on tasks with high-dimensional inputs, such as playing Atari games.
+Although TRPO achieves great performance in continuous control tasks, the conjugate gradient makes it inflexible to be implemented, and it performs poorly on tasks with high-dimensional inputs, such as those requiring visual inputs~\cite{schulman2017advanced}.
 
@@
 \centering
-\max_{\theta}J = \mathbb{E}_{\theta}\biggr[\min\bigg (\frac{\pi(a_{t}|s_{t};\theta)}{\pi(a_{t}|s_{t};\theta_{old})},
-	\clip \bigg(\frac{\pi(a_{t}|s_{t};\theta)}{\pi(a_{t}|s_{t};\theta_{old})}, 1-\epsilon, 1+\epsilon \bigg)\bigg )A(s_t, a_t)\biggr],
+\max_{\theta}J = \mathbb{E}_{\theta}\biggr[\min\bigg (\frac{\pi(a_{t}|s_{t};\theta)}{\pi(a_{t}|s_{t};\theta_{old})}A(s_t, a_t),
+	\clip \bigg(\frac{\pi(a_{t}|s_{t};\theta)}{\pi(a_{t}|s_{t};\theta_{old})}, 1-\epsilon, 1+\epsilon \bigg)A(s_t, a_t)\bigg )\biggr],
 \label{eq:standard_ppo}
 \end{equation}
-where $\epsilon$ represents the clipping ratio. In practice, PPO makes the implementation more accessible and improves the performance in tasks with high-dimensional inputs.
-
+where $\epsilon$ represents the clipping ratio.
+In practise, PPO makes the implementation more accessible, because it only needs to maximise a single objective function by using gradient ascent methods and does not require additional algorithms (e.g., conjugate gradient and line search).
+Furthermore, it also improves the performance in tasks with high-dimensional inputs.
+%\newpage
 \section{Determinantal Point Processes}

### 2022-08-21 chapters/chapter3.tex
update the final version

@@
 For such settings, the application of deep reinforcement learning requires algorithms that can learn from unshaped, and usually sparse, reward signals.
+
 The complicated dynamics of robot manipulation exacerbates the difficulty posed by sparse rewards, especially for on-policy RL algorithms.
@@
 These aspects of robot control result in a situation where a naive RL agent so rarely receives a reward at the start of training that it cannot learn at all.
-A common solution in the \td{DRL} community is collecting a sufficient number of expert demonstrations, then use imitation learning to train the agent.
+A common solution in the DRL community is collecting a sufficient number of expert demonstrations, then use imitation learning to train the agent.
 However, in some scenarios, demonstrations are expensive to collect and the performance of a trained agent is restricted by their quantity.
@@
 
-There are two kinds of approaches to alleviate the problems associated with sparse rewards: imitation learning and hindsight experience replay (HER)~\cite{andrychowicz2017hindsight}.
+There are two types of approach to alleviate the problems associated with sparse rewards: imitation learning and hindsight experience replay (HER)~\cite{andrychowicz2017hindsight}.
 First, the standard approach of imitation learning is to use supervised learning algorithms and minimise a surrogate loss with respect to an oracle.
@@
 While being able to learn policies for continuous control tasks with dense or delayed rewards~\cite{oh2018self}, the present experiments suggest that SIL struggles when the rewards are sparse.
+
 Recently, {hindsight experience replay} has been proposed to solve such goal-oriented, sparse reward problems.
-The main idea of HER is that during replay, the selected transitions are sampled from state--action pairs derived from {achieved goals} that are substituted for the real goals of the task; this increases the frequency of positive rewards.
+The main idea of HER is that during replay, the selected transitions are sampled from state--action pairs derived from {achieved goals}.
+This increases the frequency of positive rewards.
 {Hindsight experience replay} is used with off-policy RL algorithms, such as DQN~\cite{mnih2015human} and DDPG~\cite{lillicrap2015continuous}, for experience replay and has several extensions~\cite{schaul2015prioritized,liu2018competitive}.
@@
 In summary, self-imitation learning with on-policy algorithms for tasks that require continuous control, and for which rewards are sparse, remains unsolved.
+
+\begin{figure}[t]
+\centering
+\includegraphics[width=0.95\textwidth]{figures/chapter3/buffer.png}
+\caption[Illustration between SIL+HER and ESIL.]{Illustration of the difference between self-imitation learning (SIL)+hindsight experience replay (HER) and episodic self-imitation learning (ESIL).}
+%% PPO+SIL+HER samples a batch of transitions from the replay buffer and modify them into hindsight experiences and conduct self-imitation learning directly.
+PPO+ESIL utilizes entire current collected episodes and convert them into hindsight episodes.
+Another trajectory selection module is used to remove undesired transitions in each hindsight episodes.
+\label{fig:esil}
+\end{figure}
 
@@
 Figure~\ref{fig:esil} shows the difference between naive SIL+HER and ESIL.
+We further provide an illustration (Figure~\ref{fig:type_buffer}) to distinguish between the different types of buffer used in each algorithm in this chapter.
+Figure~\ref{fig:type_buffer}a shows the structure of an \textit{episodic buffer}, which is used in the proposed ESIL.
+In the episodic buffer, each entry stores a complete episode (i.e., trajectory) with the same length $T$.
+In each update iteration, all episodes in this buffer will be used for training.
+After each update iteration, the stored episodes in this buffer will be removed.
+The episodic buffer only stores the episodes that are collected by the current policy of the agent.
+Figure~\ref{fig:type_buffer}b shows the structure of a standard \textit{replay buffer}, which is used by SIL and SIL+HER in this chapter.
+Instead of storing a complete episode in each entry, the replay buffer stores a transition in each entry.
+In each update iteration, batch-sized transitions are sampled for training.
+After each update iteration, the buffer will not be cleaned.
+The replay buffer will remove the oldest transition when it is full and a new transition needs to be stored.
+Figure~\ref{fig:type_buffer}c shows the structure of an \textit{episodic replay buffer}, which is used by DDPG+HER in this chapter.
+Similar to episodic buffer, the episodic replay buffer stores a complete episode with the same length in each entry.
+In each update iteration, batch-sized episodes will be uniformly sampled.
+Then, one transition will be uniformly sampled from each selected episode to form batch-sized transitions for training.
+After each update iteration, the buffer will not be cleaned.
+The episodic replay buffer will remove the oldest episode when it is full and a new episode needs to be stored.
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=0.95\textwidth]{figures/chapter3/buffer_type.pdf}
+\caption[Illustration of different types of buffer.]{Illustration of different types of buffers: (a) is the episodic buffer, which is used in ESIL; (b) is the standard replay buffer, which is used in SIL and SIL+HER; (c) is the episodic replay buffer, which is used in HER.
+Here, $N$ is the number of samples or episodes, T is the length of each episode, D is dimension of the stored elements (e.g., state $s$).}
+%% PPO+SIL+HER samples a batch of transitions from the replay buffer and modify them into hindsight experiences and conduct self-imitation learning directly.
+PPO+ESIL utilizes entire current collected episodes and convert them into hindsight episodes.
+Another trajectory selection module is used to remove undesired transitions in each hindsight episodes.
+\label{fig:type_buffer}
+\end{figure}
+
 During training by SIL+HER, a batch of transitions is sampled from the replay buffer; these are modified into ``hindsight experiences'' and used directly in self-imitation learning.
@@
 Using tasks from the OpenAI Fetch environment~\cite{plappert2018multi}, this chapter demonstrates that the proposed ESIL approach effectively trains the agent which is required to solve continuous control problems, and shows that it achieves state-of-the-art results on several tasks.
-
-\begin{figure}[t]
-\centering
-\includegraphics[width=0.95\textwidth]{figures/chapter3/buffer.png}
-\caption{Illustration of the difference between self-imitation learning (SIL)+hindsight experience replay (HER) and episodic self-imitation learning (ESIL).}
-%% PPO+SIL+HER samples a batch of transitions from the replay buffer and modify them into hindsight experiences and conduct self-imitation learning directly.
-PPO+ESIL utilizes entire current collected episodes and convert them into hindsight episodes.
-Another trajectory selection module is used to remove undesired transitions in each hindsight episodes.
-\label{fig:esil}
-\end{figure}
 
@@
 \subsection{Episodic Self-Imitation Learning}
-The present method aims to use episodic hindsight experiences to guide the exploration of the PPO algorithm.
+The method aims to use episodic hindsight experiences to guide the exploration of the PPO algorithm.
 To this end, hindsight experiences are created from current episodes.
@@
 Simply, to create hindsight experiences, {the achieved goal $g^{ac}_{T}$ in the last state $s_{T}$ is selected and considered as the modified desired goal $g^{\prime}$, i.e., $g^\prime = g^{ac}_{T}$.} Next, a new reward $r_{t}^{\prime}$ is computed under the new goal $g^\prime$.
-Then, {a new ``imagined'' episode is achieved}, and a new series of transitions $\tau_i^{\prime} = \{(s_{t}, a_{t}, r_{t}^{\prime},s_{t+1}, g^\prime, g^{ac}_{t+1})\}_{t=0:T-1}$ is collected.
+Then, {a new ``imagined'' episode is created}, and a new series of transitions $\tau_i^{\prime} = \{(s_{t}, a_{t}, r_{t}^{\prime},s_{t+1}, g^\prime, g^{ac}_{t+1})\}_{t=0:T-1}$ is collected.
 Then, an approach to self-imitation learning based on episodic hindsight experiences is proposed, which applies the policy updates to both hindsight and in-environment episodes.
@@
 The proposed method only selects hindsight experiences that can achieve higher returns.
-The illustration of the trajectory selection module is in Figure~\ref{fig:hs_princple}.
+%The illustration of the trajectory selection module is in Figure~\ref{fig:hs_princple}. 
+%\td{An original trajectory can be divided into two categories: 1) the agent never reaches the goal and 2) the agent has reached its original before.
+Figure~\ref{fig:hs_princple} A shows the case that the agent never reaches the goal.
+In this case, the }
+
 For an episodic experience and its hindsight experience, the returns of the episodic experience and its hindsight experience can be calculated, respectively. 
@@
 Then, for a trajectory $\tau_{i}$, we have $\left\{G_{0}^{i}, G_{1}^{i}, G_{2}^{i}, ..., G_{T-1}^{i}\right\}$.
-
 For the hindsight experiences, similarly, the return $G_{t}'$ for each timestep $t$,~with respect to the hindsight goal $g^\prime$, can be calculated.
 Based on the modified trajectory $\tau_{i}^{\prime}$ with the same length of $\tau_{i}$, we therefore have the returns $\left\{G_{0}^{i^\prime}, G_{1}^{i^\prime}, G_{2}^{i^\prime}, ..., G_{T-1}^{i^\prime}\right\}$.
-During training, the~hindsight experiences with higher returns are used for self-imitation learning.
-The rest of the hindsight experiences will be \td{assumed to be of lower importance} and ignored.
+
+The illustration of the trajectory selection module is shown in Figure~\ref{fig:hs_princple}.
+An original trajectory collected by an agent can be divided into two categories: 1) the agent never reaches the goal and 2) the agent has reached its original goal before.
+Figure~\ref{fig:hs_princple}A shows the case where the agent never reaches the goal.
+In this case, the original experiences never contain a non-negative reward.
+However, the hindsight experience will have at least one transition that has a non-negative reward (e.g., the final step).
+According to the rules for calculating the returns, all hindsight experiences (solid orange trajectory) have higher returns than the original experiences (solid blue trajectory).
+Therefore, in this case, all hindsight experiences can be selected to benefit the training.
+Figure~\ref{fig:hs_princple}B demonstrates the case that the agent has achieved its original goal in the third timestep and thus received a non-negative reward.
+However, for the hindsight experience, it only achieves a non-negative reward in the end of the trajectory.
+Therefore, in the first three steps, the original experiences have higher returns than the hindsight experiences (dashed orange trajectory), which means that the original experiences have better quality than the hindsight experiences, and these hindsight experiences are ignored during training.
+For the remainder of the episode of Figure~\ref{fig:hs_princple}B, the hindsight experiences (solid orange trajectory) have higher returns than the original ones, and thus these hindsight experiences are selected to accelerate training.
+
+During training, the~hindsight experiences with higher returns are used for self-imitation learning. The rest of the hindsight experiences will be assumed to be of lower importance and ignored.
 Then, Equation~\eqref{eq:imitaion} can be rewritten as:
@@
   \includegraphics[width=\columnwidth]{figures/chapter3/hgs.png}
-  \caption{A simplified illustration of trajectory selection.
+  \caption[A simplified illustration of trajectory selection.]{A simplified illustration of trajectory selection.
 Blue trajectories indicate original experiences.
@@
 
-From a goal perspective, episodic self-imitation learning (ESIL) tries to explore (desired) goals to get positive returns.~It can be viewed as a form of multi-task learning, {because ESIL has two objective functions to be optimised jointly.}~It is also related to self-imitation learning (SIL)~\cite{oh2018self}.~However,~ the~difference is that SIL uses $\left(G-V(s;\theta)\right)_{+}$ on past experiences to learn to choose the action chosen in the past in a given state, rather than goals.
+From a goal perspective, episodic self-imitation learning (ESIL) tries to explore (desired) goals to get positive returns.~It can be viewed as a form of multi-task learning, {because ESIL has two objective functions to be optimised jointly.}~It is also related to self-imitation learning (SIL)~\cite{oh2018self}.
+However, the difference is that SIL uses $\left(G-V(s;\theta)\right)_{+}$ on past experiences to learn to choose the action chosen in the past in a given state, rather than goals.
 The full description of ESIL can be found in Algorithm~\ref{alg}.
@@
     \STATE Execute the action $a_{t}$ and observe a new state $s_{t+1}$
-    \STATE Store the transition $\left(s_{t}, a_{t}, r_{t}, s_{t+1}, g^{ac}_{t+1}\right)$ in $\tau$
+    \STATE Store the transition $\left(s_{t}, a_{t}, r_{t}, s_{t+1}, g, g^{ac}_{t+1}\right)$ in $\tau$
     \ENDFOR
@@
 \endminipage
-\caption{Evaluation environments.
+\caption[Evaluation environments of ESIL.]{Evaluation environments.
 ({a}) is the Empty Room environment, in which a yellow circle indicates the position of the agent and a red star represents a target position.
@@
 The goal of the agent is to reach a target position in the room.
-The start position of the agent is at the left upper corner of the room, and the target position is randomly selected  within the room.
+The start position of the agent is at the left upper corner of the room, and the target position is randomly selected within the room at the beginning of each episode~\cite{rauber2018hindsight}.
 When the agent chooses an action that would fall outside the grid area, the agent will stay at the current position.
@@
 The upper bound is the {75th} percentile, and the lower bound is the {25th} percentile.
+The code is publicly available at: \url{https://github.com/TianhongDai/esil-hindsight}.
 
@@
 \endminipage\hfill
-\caption{Results of the grid-world environment.
+\caption[Results of ESIL in the grid-world environment.]{Results of the grid-world environment.
 ({a}) Comparing the performance of PPO+ESIL between the on-policy approaches.
@@
 The hindsight experiences can yield higher returns than the original experiences.
-{Therefore,~a~large proportion of hindsight experiences is selected to conduct self-imitation learning, helping the agent learn a policy for moving through the room}.
+{Therefore,~a~large proportion of hindsight experiences are selected to conduct self-imitation learning, helping the agent learn a policy for moving through the room}.
 In the later stages of training, the agent can achieve success frequently, and the hindsight experiences might be redundant (e.g., $G(s_{t}, g)\geq G(s_{t}, g^{\prime})$).
@@
 \endminipage\hfill
-\caption{Results of comparison between ESIL and on-policy baselines in all Fetch tasks.
-\td{From the results, ESIL is able to resolve all of the tasks, while other baselines can only solve the simple Reach task.}}
+\caption[Results of ESIL in the robotic manipulation environment.]{Results of comparison between ESIL and on-policy baselines in all Fetch tasks.
+From the results, ESIL is able to resolve all of the tasks, while other baselines can only solve the simple Reach task.}
 \label{fig:baseline_compare}
@@
 {In contrast}, PPO, PPO+SIL, and PPO+SIL+HER do not work in all tasks, except for the Reach task.
-In comparison with the other selected tasks from the Fetch environment, Reach task is relatively simple, because there is no object to be manipulated.
+In comparison with the other selected tasks from the Fetch environment, the Reach task is relatively simple, because there is no object to be manipulated.
 For other tasks, it is quite difficult for the agent to achieve sufficient positive rewards during exploration, because of their rare occurrence.
 Although PPO+SIL utilises past good experiences to help exploration, it is still faced with the difficulty that past experiences do not easily achieve positive rewards.
+
 From the experiments ({see Figure~\ref{fig:baseline_compare}}), PPO+SIL (no hindsight) converges much more slowly than using PPO only.
@@
 \endminipage\hfill
-\caption{Results of ablation studies with and without using the trajectory selection module in all Fetch tasks.
-\td{In three challenging tasks: Push, Pick$\&$Place and Slide, the trajectory selection module can stabilise the training and prevent performance degradation in two of three cases (Push and Pick$\&$Place).}}
+\caption[Results of ablation studies of the trajectory selection module.]{Results of ablation studies with and without using the trajectory selection module in all Fetch tasks.
+In three challenging tasks: Push, Pick$\&$Place and Slide, the trajectory selection module can stabilise the training and prevent performance degradation in two of three cases (Push and Pick$\&$Place).}
 \label{fig:hs_compare}
@@
 \endminipage\hfill
-\caption{Variation in adaptive weight coefficient $\beta$ through the training in all Fetch tasks.
-\td{In the early stage of training, the weight of the ESIL term is higher, which indicates more samples need to be modified to facilitate the exploration.
-As training progresses, the weight decreases, which means that the agent can achieve more positive samples and less modified samples are required.}}
+\caption[Variation in adaptive weight coefficient $\beta$ during training.]{Variation in adaptive weight coefficient $\beta$ through the training in all Fetch tasks.
+In the early stage of training, the weight of the ESIL term is higher, which indicates more samples need to be modified to facilitate the exploration.
+As training progresses, the weight decreases, which means that the agent can achieve more positive samples and less modified samples are required.}
 \label{fig:her_samples_compare}
@@
 However, {Figure~\ref{fig:hs_compare}d}, for the Slide task, the agent converges faster without the trajectory selection module and has better performance.
-{This is likely to be because the Slide task is the most difficult of the Fetch environment.
+{This is likely to be because the Slide task is the most difficult of the Fetch environments.
 During training, the agent is unlikely to achieve positive rewards.
 Figure~\ref{fig:her_samples_compare} also indicates that the value of $\beta$ in the Slide task is higher than values in other tasks, which means the majority of hindsight experiences have higher returns than original experiences.
-Thus, using {\em more} hindsight experiences (without filtering) accelerates training at this stage.} Nonetheless, the trajectory selection module prevents the agent from overfitting the hindsight experience in the other three tasks.
+Thus, using {\em more} hindsight experiences (without filtering) accelerates training at this stage.}
+
+Nonetheless, the trajectory selection module prevents the agent from overfitting the hindsight experience in the other three tasks.
 Figure~\ref{fig:her_samples_compare}, shows the adaptive weight coefficient $\beta$ in all Fetch tasks.
@@
 \endminipage\hfill
-\caption{Results of the comparison between PPO+ESIL and DDPG+HER in all fetch tasks.
-\td{From the results, PPO+ESIL has worse sample efficiency than the off-policy baseline.
-However, PPO+ESIL achieves similar or even better performance in the end of training.}}
+\caption[Comparison between ESIL and DDPG+HER.]{Results of the comparison between PPO+ESIL and DDPG+HER in all fetch tasks.
+From the results, PPO+ESIL has worse sample efficiency than the off-policy baseline.
+However, PPO+ESIL achieves similar or even better performance in the end of training.}
 \label{fig:her_compare}
@@
   \end{tabular}}
-   \caption{Average success rate $\pm$ standard error in the last 10 epochs over five random seeds in all~environments {(\textbf{bold} indicates the best result among all methods).}}
+   \caption[Average success rate of ESIL and other baselines in the last 10 epochs during training.]{Average success rate $\pm$ standard error in the last 10 epochs over five random seeds in all~environments {(\textbf{bold} indicates the best result among all methods).}}
   \label{tb:ch3_last}
@@
 
+\begin{table}[h!]
+    \centering
+    \resizebox{0.85\textwidth}{!}{
+    \begin{tabular}{c|c|c|c|c|c}
+    \toprule
+           & PPO & PPO+SIL & PPO+SIL+HER & DDPG+HER & PPO+ESIL\\
+        \midrule
+        Time  & 00:57:09  & 00:58:26  & 01:05:17  & 03:18:53  & 00:59:14\\
+        \bottomrule
+    \end{tabular}
+    }
+    \vspace{0.2em}
+    \caption[Training time comparison between ESIL and other baselines.]{Training time (hours:minutes:seconds) of ESIL and baseline approaches on the Reach task for 100 epochs.}
+    \label{tab:hsl_time}
+\end{table}
+
 The experiments that we have conducted suggest that simply using self-imitation learning with the PPO algorithm, even with hindsight experience, leads to disappointing performance in continuous control Fetch tasks.
@@
 The auxiliary trajectory selection module and the adaptive weight $\beta$ help the training process to remove undesired experiences and balance the contributions to learning between the PPO term and the ESIL term automatically, and also increase the stability of training.
-
 Our experiments suggest that the selection module is useful to prevent overfitting to sub-optimal hindsight experiences, but also that it does not always lead to learning a better policy faster.
 Despite~this, selection filtering appears to support learning a useful policy in challenging environments.
+Table~\ref{tab:hsl_time} shows the training time of ESIL and other baseline methods.
+From the table, PPO+ESIL shows a training time similar to other on-policy RL baselines, which indicates that the ESIL term only increases minor computational complexity.
+However, it can substantially improve performance on the continuous control tasks for on-policy RL algorithms (e.g., PPO).
+Although DDPG+HER achieves better sample efficiency in most tasks, it requires the longest training time in these methods.
+Therefore, our method achieves the balance between computational complexity and performance.
+
 The experiments we have conducted to date have utilised relatively small networks, and it would be appropriate to extend the experiments to consider more complex observation spaces, and to actor/critic networks, which are consequently more elaborate.

### 2022-08-21 chapters/chapter4.tex
update the final version

@@
 For example, in a robotic manipulation task, the goal during a single episode would be to achieve a specific position of a target object.
-In the last chapter, we have successfully combined hindsight goal relabelling with a state-of-the-art on-policy RL algorithm and introduced a new goal-oriented RL method -- episodic self-imitation learning (ESIL).
+In the last chapter, we successfully combined hindsight goal relabelling with a state-of-the-art on-policy RL algorithm and introduced a new goal-oriented RL method -- episodic self-imitation learning (ESIL).
+In ESIL, a novel self-imitation learning loss is used to accelerate the training by using the hindsight experience.
 Although the experiments show the effectiveness of the proposed method, ESIL still has worse sample efficiency than the vanilla hindsight experience replay (i.e., DDPG+HER).
 This is because DDPG is an off-policy RL algorithm, and it employs a replay buffer to store transitions during training, which enables the usage of past experiences.
-In this chapter, we focus on finding the shortcoming of HER and improving it to elevate learning efficiency and performance.
+In this chapter, we focus on finding the shortcoming of HER and improving it to elevate learning efficiency and performance by proposing a new sampling strategy.
 
@@
     \includegraphics[width=\textwidth]{figures/chapter4/illustration_DTGSH_latest.pdf}
-    \caption{Overview of DTGSH.
+    \caption[Overview of DTGSH.]{Overview of DTGSH.
 Every time a new episode is completed, its diversity is calculated, and it is stored in the episodic replay buffer.
@@
 The use of DPPs in RL has been more limited, with applications towards modelling value functions of sets of agents in multiagent RL~\cite{osogami2019determinantal,pmlr-v119-yang20i}, and most closely related to us, finding diverse policies~\cite{parker2020effective}.
+
+% \begin{table}[h]
+%     \centering
+%     \resizebox{\textwidth}{!}{
+%     \begin{tabular}{c|c@{\hskip 0.1in}|c@{\hskip 0.1in}|c|c}
+%     \toprule
+%          & DDPG+HER~\cite{andrychowicz2017hindsight} & DDPG+HEBP~\cite{zhao2018energy} & DDPG+CHER~\cite{fang2019curriculum} & DDPG+DTGSH (Ours) \\
+%         \midrule
+%         Sampling Strategy (Trajectory) & & & & \\
+%         Sampling Strategy (Transition) & & & & \\
+%         \bottomrule
+%     \end{tabular}
+%     }
+%     \vspace{0.2em}
+%     \caption[Average success rate of DTGSH and other baselines in the final epoch during training.]{Final mean success rate $\pm$ standard deviation, with best results in \textbf{bold}.}
+%     %\vspace{-14mm}
+%     \label{tab:benchmark}
+% \end{table}
+
+\begin{table}[h]
+\resizebox{\textwidth}{!}{
+\begin{tabular}{|c|c|c|c|}
+\hline
+& DDPG+HER~\cite{andrychowicz2017hindsight} & DDPG+HEBP~\cite{zhao2018energy} & DDPG+CHER~\cite{fang2019curriculum} \\ \hline
+\begin{tabular}[c]{@{}c@{}}Sampling Strategy\\       (Trajectory)\end{tabular} & uniform  & \begin{tabular}[c]{@{}c@{}}according to the energy\\ of each trajectory\end{tabular} & uniform                                                                                                \\ \hline
+\begin{tabular}[c]{@{}c@{}}Sampling Strategy\\ (Transition)\end{tabular}       & uniform  & uniform                                                                              & \begin{tabular}[c]{@{}c@{}}according to the proximity and \\ diversity of candidate goals\end{tabular} \\ \hline
+\end{tabular}
+}
+\caption[Comparison of sampling strategies between different HER related methods.]{The comparison of sampling strategies between HER related methods.}
+\label{tab:sample_strategy}
+\end{table}
+
+We further use Table~\ref{tab:sample_strategy} to summarize the sampling strategies of baseline HER related algorithms in this chapter.
+Although HEBP~\cite{zhao2018energy} and CHER~\cite{fang2019curriculum} improve the sample efficiency of HER by proposing new sampling strategies, these methods still use uniform sampling strategy for either trajectory or transition selection, which may sample redundant trajectories and transitions, and thus limit the performance.
+In contrast to these methods, in DTGSH, the sampling strategies for both trajectory and transition are improved, which increase the chance to sample trajectories and transitions that contribute more to the training.
 
@@
 \end{equation}
-and $\mathbf{M}=[\hat{g}^{ac}_{n}, \hat{g}^{ac}_{n+1}, \cdots, \hat{g}^{ac}_{n+b-1}]$, where each $\hat{g}^{ac}$ is the $\ell_2$-normalised version of the achieved goal $g^{ac}$~\cite{kulesza2011k}.
+and $\mathbf{M}=[\hat{g}^{ac}_{n}, \hat{g}^{ac}_{n+1}, \cdots, \hat{g}^{ac}_{n+b-1}]$, where each $\hat{g}^{ac}$ is the $\ell_2$-normalised vector of the achieved goal $g^{ac}$~\cite{kulesza2011k}.
 Finally, the diversity $d_\Tau$ of trajectory $\Tau$ is the sum of the diversity of its $N_p$ constituent partial trajectories:
@@
     \end{subfigure}
-    \caption{Visualisation of $k=32$ goals selected from $m=100$ candidate goals of the Push task using either uniform sampling or $k$-DPP sampling, respectively.
+    \caption[Visualisation of goal selection between uniform sampling and $k$-DPP sampling.]{Visualisation of $k=32$ goals selected from $m=100$ candidate goals of the Push task using either uniform sampling or $k$-DPP sampling, respectively.
 The candidate goals are distributed over a 2D ($xy$) space.
@@
   \end{subfigure}\hfill
-  \caption{Robot manipulation environments. (a-b) use the Fetch robot, and (c-e) use the Shadow Dexterous Hand.} 
+  \caption[Evaluation environments of DTGSH.]{Robot manipulation environments. (a-b) use the Fetch robot, and (c-e) use the Shadow Dexterous Hand.} 
   \label{fig:env_c4}
@@
 
-\begin{figure}
+\begin{figure}[H]
 \centering
@@
 %   \end{subfigure}\hfill
-  \caption{Success rate of DTGSH and baseline approaches. \td{DDPG+DTGSH outpeforms other baselines in all tasks, and without using HER, DDGP cannot even complete the simple Reach task.}} 
+  \caption[Results of DTGSH in the robotic manipulation environment.]{Success rate of DTGSH and baseline approaches.
+DDPG+DTGSH outperforms other baselines in all tasks, and without using HER, DDGP cannot even complete the simple Reach task.}
   \label{fig:baseline_compare_c4}
@@
     \vspace{0.2em}
-    \caption{Final mean success rate $\pm$ standard deviation, with best results in \textbf{bold}.}
+    \caption[Average success rate of DTGSH and other baselines in the final epoch during training.]{Final mean success rate $\pm$ standard deviation, with best results in \textbf{bold}.}
     %\vspace{-14mm}
@@
 %   \end{subfigure}\hfill
-  \caption{Success rate of HER, DTGSH, and ablations DTSH and DGSH.
-\td{From the results, DTSH outperforms DDPG+HER in all tasks, which verify the efficiency that selecting trajectories with diverse achieved goals for training.
+  \caption[Results of ablation studies of different modules.]{Success rate of HER, DTGSH, and ablations DTSH and DGSH.
+From the results, DTSH outperforms DDPG+HER in all tasks, which verify the efficiency that selecting trajectories with diverse achieved goals for training.
 DGSH also has better performance than DDPG+HER in three tasks.
-By combining these two modules together, DTGSH achieves the best performance.} }
+By combining these two modules together, DTGSH achieves the best performance.}
   \label{fig:ablation1}
@@
 %   \end{subfigure}\hfill
-  \caption{Success rate of DTGSH with different partial trajectory lengths $b$ and different candidate goal set sizes $m$.
-\td{From the results, when $b \gg 2$, the performance degrades in all tasks.
-When $m \gg 100$, the performance degrades in the Shadow Dexterous Hand environment.}}
+  \caption[Results of ablation studies of different parameters of each module.]{Success rate of DTGSH with different partial trajectory lengths $b$ and different candidate goal set sizes $m$.
+From the results, when $b \gg 2$, the performance degrades in all tasks.
+When $m \gg 100$, the performance degrades in the Shadow Dexterous Hand environment.}
   \label{fig:ablation2}
@@
 
-\subsection{Time Complexity}
-Table~\ref{tab:time} gives example training times of all of the HER-based algorithms.
+%\subsection{Time Complexity}
+% Table~\ref{tab:time} gives example training times of all of the HER-based algorithms.
 DTGSH requires an additional calculation of the diversity score of $\mathcal{O}(N_{p}b^3)$ at the end of every training episode, and sampling of $\mathcal{O}(mk^2)$ for each minibatch.
+
+\section{Discussion}
+In this chapter, we introduced diversity-based trajectory and goal selection with hindsight experience replay (DTGSH) to improve the learning efficiency of goal-orientated RL agents in sparse reward settings.
+Our method can be divided into two stages: 1) valuable trajectories are selected according to diversity-based priority, as modelled by determinantal point processes (DPPs)~\cite{kulesza2012determinantal}, and 2) $k$-DPPs~\cite{kulesza2011k} are leveraged to sample transitions with diverse goal states from previously selected trajectories for training.
+Our experiments empirically show that DTGSH achieves faster learning and higher final performance in five challenging robot manipulation tasks, compared to previous state-of-the-art approaches~\cite{andrychowicz2017hindsight,fang2019curriculum,zhao2018energy}.
+In addition, unlike prior extensions of hindsight experience replay, DTGSH does not require semantic knowledge of the goal space~\cite{zhao2018energy}, and does not require tuning a curriculum~\cite{fang2019curriculum}.
+
 \begin{table}[h!]
@@
     \vspace{0.2em}
-    \caption{Training time (hours:minutes:seconds) of DTGSH and baseline approaches on the Push task for 50 epochs.}
+    \caption[Training time comparison between DTGSH and other baselines.]{Training time (hours:minutes:seconds) of DTGSH and baseline approaches on the Push task for 50 epochs.}
     \label{tab:time}
@@
 
-\section{Discussion}
-In this chapter, we introduced diversity-based trajectory and goal selection with hindsight experience replay (DTGSH) to improve the learning efficiency of goal-orientated RL agents in sparse reward settings.
-Our method can be divided into two stages: 1) valuable trajectories are selected according to diversity-based priority, as modelled by determinantal point processes (DPPs)~\cite{kulesza2012determinantal}, and 2) $k$-DPPs~\cite{kulesza2011k} are leveraged to sample transitions with diverse goal states from previously selected trajectories for training.
-Our experiments empirically show that DTGSH achieves faster learning and higher final performance in five challenging robot manipulation tasks, compared to previous state-of-the-art approaches~\cite{andrychowicz2017hindsight,fang2019curriculum,zhao2018energy}.
-Furthermore, unlike prior extensions of hindsight experience replay, DTGSH does not require semantic knowledge of the goal space~\cite{zhao2018energy}, and does not require tuning a curriculum~\cite{fang2019curriculum}.
+Furthermore, Table~\ref{tab:time} gives example training times of all of the HER-based algorithms.
+From the table, HER requires the minimum time, because it only uses the naive uniform sampling strategy for both trajectory and transition.
+HEBP needs a bit more training time than HER, because it requires additional computation for the energy of each trajectory.
+CHER displays the longest training time, because it requires an inefficient greedy algorithm~\cite{fang2019curriculum} for the sampling according to the proximity and diversity of candidate goals.
+DTGSH is more efficient than CHER, but slower than the other two methods.
+It needs an additional calculation of the diversity score of $\mathcal{O}(N_{p}b^3)$ at the end of every training episode, and sampling of $\mathcal{O}(mk^2)$ for each minibatch.
+Although DTGSH has better sample efficiency than other baselines, it needs a longer training period.
+In the future, we will improve the method towards achieving the balance between performance and time complexity.
+
+There are still several limitations in some of the experiments, where the models do not have sufficient training time to converge in Figure~\ref{fig:ablation1}c and Figure~\ref{fig:ablation1}d.
+There are two main reasons for this, on the one hand we use the same experimental setup as CHER~\cite{fang2019curriculum} for fair comparison and on the other hand we do not have enough computing resources to complete training with more epochs.
+In this chapter, it took us more than one month to finish these experiments.
+In the future, once we have enough computing resources, we will re-evaluate the method for longer training epochs.

### 2022-08-21 chapters/chapter5.tex
update the final version

@@
 \section{Introduction}
-The previous chapter uses the determinantal point processes (DPPs) framework to measure the diversity of achieved goals in a trajectory and proposes a sampling strategy that uses diversity as a priority to sample the trajectories for further processing, in which diversity can also quantify the contribution of each trajectory.
+In order to improve sample efficiency of DRL algorithms in the environments offering sparse rewards, Chapter~\ref{ch:esil} proposes a novel self-imitation loss function to increase the frequency that the agent achieves positive samples, and Chapter~\ref{ch:dtgsh} uses the determinantal point processes (DPPs) framework to measure the diversity of achieved goals in a trajectory and proposes a sampling strategy that uses diversity as a priority to sample the trajectories for further processing, in which diversity can also quantify the contribution of each trajectory.
 In this chapter, we keep exploring the potential application of determinantal point processes (DPPs) in deep reinforcement learning (DRL).
+In contrast to previous chapters, DPPs are used to quantify the discrepancy between adjacent states as additional endogenous rewards to accelerate the training of DRL agents in the environment with the sparse reward setting.
 
 %Deep reinforcement learning (RL) allows agents to learn a wide range of reward-driven skills, including playing games~\cite{mnih2015human,silver2016mastering,xu2020neurips}, controlling robots~\cite{dai2020episodic,fang2018dher,gu2017deep,levine2016end}, and making decisions in other environments~\cite{fang2017learning,vinyals2017starcraft}.
-Many important applied problems in DRL can be accurately modelled as learning from extrinsic rewards, where a function (approximately) relates features associated with a state/action pair and the expectation of the resulting rewards.
+
+%Many significant application problems in DRL can be formulated as a 
+% Many important applied problems in DRL can be accurately modelled as learning from extrinsic rewards, where a function (approximately) relates features associated with a state/action pair and the expectation of the resulting rewards.
 Furthermore, with appropriate features, it is often the case that linear or nonlinear functions can accurately approximate the reward function.
-However, ill-designed reward functions usually lead to unexpected behaviors~\cite{riedmiller2018learning} and limited exploration.
-Potential-based reward shaping methods~\cite{ng1999policy} offers a general answer to what degree of reward function modifications do not affect the optimal policy.
+However, hand-crafted reward functions usually lead to unexpected behaviors~\cite{riedmiller2018learning} and limited exploration.
+% Potential-based reward shaping methods~\cite{ng1999policy} offers a general answer to what degree of reward function modifications do not affect the optimal policy.
 Other works have tried to design auxiliary rewards to derive policies with desired properties.
@@
 
-In environments or tasks for which the reward sparsity is higher, undirected, or even deceptive, the application of RL techniques is significantly more difficult, since the agent might have only occasional opportunities to improve its policy, such as when the final goal is reached.
+% In environments or tasks for which the reward sparsity is higher, undirected, or even deceptive, the application of RL techniques is significantly more difficult, since the agent might have only occasional opportunities to improve its policy, such as when the final goal is reached.
 In such scenarios, agents can be endowed with mechanisms of {\it intrinsic motivation} that encourage exploration in environments and learn useful policies even without explicit supervision.
@@
 More recently, Zheng \textit{et al.}~\cite{zheng2018learning} design an augmented agent with additive intrinsic rewards.
-When the intrinsic value of the action is added to the task-dependent (hereafter extrinsic) rewards, the performance of policy gradient based learning methods can be improved.
-Another line of work uses auxiliary control tasks, such as pixel control or feature control, which can significantly speed up learning of the main task.
+When the intrinsic value of the action is added to the task-dependent (i.e., extrinsic) rewards, the performance of policy gradient based learning methods can be improved.
+%Another line of work uses auxiliary control tasks, such as pixel control or feature control, which can significantly speed up learning of the main task.
 The rationale is that learning the auxiliary tasks gives the agent features that are useful for manipulating the environment.
@@
 
-Encouraging effective exploration of novel states for agents without any reward signal is still a challenging problem.
+Encouraging the agent to explore previously unseen/novel states in the environments with sparse or delayed reward settings is still a challenge in DRL.
+In this chapter, we demonstrate how a simple generated intrinsic reward function based on a diversity measure can enable RL agents to explore novel states efficiently.
+This generated auxiliary reward is functional for a number of environments, as confirmed by a selection of MuJoCo locomotion tasks and video games.
+
+% Encouraging effective exploration of novel states for agents without any reward signal is still a challenging problem.
 Making progress on this problem requires specifying a learning objective that ensures the agent explores large parts of the state space.
-In this chapter, we show how a simple objective based on a diversity measure can enable RL agents to autonomously explore states.
-This module is useful for a number of environments, as confirmed by a selection of MuJoCo locomotion tasks and video games.
+In this chapter, we demonstrate how a simple generated intrinsic reward function based on a diversity measure can enable RL agents to explore novel states efficiently.
+This auxiliary reward is functional for a number of environments, as confirmed by a selection of MuJoCo locomotion tasks and video games.
 
-We hypothesise that in order to acquire an agent that is useful and efficient, we must train the agent so that it can attain coverage over the set of possible states.
-However, the common exploration strategies ({e.g.}, random sampling, $\epsilon$-greedy) can re-visit previously explored states of the agent, limiting the effectiveness of state-space coverage.
+%We hypothesise that in order to acquire an agent that is useful and efficient, we must train the agent so that it can attain coverage over the set of possible states.
+
+In order to obtain an agent that can solve the tasks with sparse or delayed rewards, we need the agent to explore as many novel states as possible during training.
+However, common exploration strategies ({e.g.}, random sampling, $\epsilon$-greedy) can re-visit previously explored states in the environment, which can limit the effectiveness of state-space coverage.
 
 A key idea in our work is to use a measure of discrepancy between states as an intrinsic motivation.
-\td{Further, states that are distinguishable are not necessarily diverse in a semantically meaningful way.
-To combat this problem, we want the intrinsic motivation of the agent that is not only distinguishable but also is as diverse as possible.}
-
 To model the diversity of the agent's states, we adopt the framework of determinantal point processes (DPPs)~\cite{kulesza2012determinantal} to formulate the intrinsic reward as a dynamic measure of diversity.
-Regarding the sequence of states as a DPP allows a measure of the diversity of subset selection, widely used in data summarization for selecting a diversified subset \cite{gong2014diverse}.
+Regarding the sequence of states as a DPP allows a measure of the diversity of subset selection, widely used in data summarization for selecting a diversified subset~\cite{gong2014diverse}.
 We build an intrinsic reward for diversity with parameterised DPPs and propose a new general method for learning policies with intrinsic rewards.
@@
 In addition, $\ourmethod$ also improves the rate of learning in the initial stage of training in the arcade learning environment.
+\newpage
 
@@
 Pathak \textit{et al.}~\cite{pathak2017curiosity} propose an intrinsic curiosity module (ICM), suggesting the prediction error between an actual encoded frame and the predicted encoded frame as the intrinsic reward.
-\td{A higher prediction error indicates the agent has moved into a novel state, and this intrinsic reward will encourage the agent to keep exploring such novel states.} In order to find an appropriate representation of pixels that could be invariant to insignificant changes in the environment, an inverse dynamic model is introduced to extract features.
+A higher prediction error indicates the agent has moved into a novel state, and this intrinsic reward will encourage the agent to keep exploring such novel states.
+In order to find an appropriate representation of pixels that could be invariant to insignificant changes in the environment, an inverse dynamic model is introduced to extract features.
 This requires training a network, which uses consecutive frames as input to predict the action that would transition the previous frame to the subsequent.
@@
 Zheng \textit{et al.}~\cite{zheng2018learning} introduce learning intrinsic reward for policy gradient (LIRPG), which enables a learnable intrinsic reward function to be used for training an RL agent.
-\td{In addition, LIRPG adopts a bi-level optimisation scheme that when the policy network is updated to maximise the accumulated intrinsic and extrinsic rewards, the intrinsic reward function is also updated to maximise the accumulated extrinsic rewards achieved by the policy network.
-This optimisation scheme guarantees the learned reward function can generate meaningful intrinsic rewards to facilitate the agent to reach the ultimate goal (i.e., maximising the accumulated extrinsic rewards).
-In this work, we also leverage this bi-level optimisation to ensure the agent can explore novel states that deliver more accumulated extrinsic rewards.}
+In addition, LIRPG adopts a bi-level optimisation scheme that when the policy network is updated to maximise the accumulated intrinsic and extrinsic rewards, the intrinsic reward function is also updated to maximise the accumulated extrinsic rewards achieved by the policy network.
+This optimisation scheme guarantees the learned reward function can generate meaningful intrinsic rewards to facilitate the agent reaching the ultimate goal (i.e., maximising the accumulated extrinsic rewards).
+In this work, we also leverage this bi-level optimisation to ensure the agent can explore novel states that deliver more accumulated extrinsic rewards.
 
@@
     \includegraphics[width=\linewidth]{figures/chapter5/illustration.pdf}
-    \caption{Illustration of $\ourmethod$.
+    \caption[Illustration of $\ourmethod$.]{Illustration of $\ourmethod$.
 It shows the intrinsic rewards of an entire episode of SuperMarioBros.
 The red points annotate the key frames in the game.
-In frames \{17, 26, 32, 41\}, the agent keeps moving forward, and the intrinsic rewards are \td{relatively higher}.
+In frames \{17, 26, 32, 41\}, the agent keeps moving forward, and the intrinsic rewards are relatively higher.
 In the frames \{303, 327, 354, 379\}, the agent is stuck at the bottom of the flagpole.
@@
 \end{align}
-\td{Here $J^{\text{mix}} = \mathbb{E}_{\pi_{\theta}}\bigr[G^{\text{mix}}(s_t,a_t) \bigr]$, and $\theta$ is the parameters of the policy network.
+Here $J^{\text{mix}} = \mathbb{E}_{\pi_{\theta}}\bigr[G^{\text{mix}}(s_t,a_t) \bigr]$, and $\theta$ represents the parameters of the policy network.
 The optimisation problem of \eqref{eq:obj} can be viewed as a bi-level optimisation, which is widely used in the fields of machine learning, such as meta-learning~\cite{andrychowicz2016learning,nichol2018first,santoro2016meta,xu2018meta,zheng2018learning}, signal processing~\cite{kunapuli2008classification,flamary2014learning}, and hyperparameters optimisation~\cite{franceschi2018bilevel, shaban2019truncated}.
-In Equation~\eqref{eq:obj}, $J^{\text{ex}}$ and $J^{\text{mix}}$ are upper-level and inner-level objective functions, respectively.
-The primary goal of this bi-level objective function is to maximise extrinsic returns (upper-level objective function) $J^{\text{ex}}$ with respect to the parameters of intrinsic reward $\eta$, and the updated parameters of the policy network $\theta^{\prime}$ can be achieved via maximising the inner-level objective function.}
+In Equation~\eqref{eq:obj}, $J^{\text{ex}}$ and $J^{\text{mix}}$ are upper-level and lower-level objective functions, respectively.
+The primary goal of this bi-level objective function is to maximise extrinsic returns (upper-level objective function) $J^{\text{ex}}$ with respect to the parameters of intrinsic reward $\eta$, and the updated parameters of the policy network $\theta^{\prime}$ can be achieved via maximising the lower-level objective function.
 
@@
 %This is a commonly adopted technique in meta-gradient learning~\cite{andrychowicz2016learning,nichol2018first,santoro2016meta,xu2018meta,zheng2018learning}. 
-Through \td{this bi-level optimisation}, the features that are learned ensure that the resulting diversity also increases the extrinsic returns.
-\td{Furthermore, the value network is updated via a mean squared error loss $\mathcal{L}_{value} = \mathbb{E}[(V(s_{t};\psi) - G_{t}^{\text{mix}})^2]$.} Algorithm~\ref{algo:vi} gives the pseudo code for our method -- diversity-augmented intrinsic motivation ($\ourmethod$).
-The code is publicly available at: \href{https://github.com/TianhongDai/daim-rl}{https://github.com/TianhongDai/daim-rl}.
+Through this bi-level optimisation, the features that are learned ensure that the resulting diversity also increases the extrinsic returns.
+Furthermore, the value network is updated via a mean squared error loss $\mathcal{L}_{value} = \mathbb{E}[(V(s_{t};\psi) - G_{t}^{\text{mix}})^2]$.
+Algorithm~\ref{algo:vi} gives the pseudo code for our method -- diversity-augmented intrinsic motivation ($\ourmethod$).
+The code is publicly available at: \url{https://github.com/TianhongDai/daim-rl}.
 
@@
 		\STATE Update $\theta$ according to Eq.\eqref{eq:mix_cost} with learning rate $\alpha$
-		\STATE\td{Update $\psi$ according to $\mathcal{L}_{value}=\mathbb{E}[(V(s_{t};\psi)-G^{\text{mix}}_{t})^2]$ with learning rate $\alpha$}
+		\STATE Update $\psi$ according to $\mathcal{L}_{value}=\mathbb{E}[(V(s_{t};\psi)-G^{\text{mix}}_{t})^2]$ with learning rate $\alpha$
 		\STATE Update $\eta$ according to Eq.\eqref{eq:chain_rule} and Eq.\eqref{eq:importance_sampling} with learning rate $\beta$
@@
 \end{algorithm}
-
 \subsection{Parameterised Diversity as Intrinsic Reward}
-We are now ready to parameterise the intrinsic reward with parameter $\eta$ in this section. The intrinsic reward at timestep $t$ is defined as:
+We are now able to parameterise the intrinsic reward with parameter $\eta$ in this section. The intrinsic reward at timestep $t$ is defined as:
 \begin{equation}
@@
 Figure~\ref{fig:illustration} illustrates the proposed intrinsic reward.
-When the agent is moving ahead, the intrinsic rewards are \td{relatively higher}.
-In contrast, when the agent is stuck at the bottom of the flagpole, the intrinsic rewards are close to zero.
+When the agent is moving ahead (e.g., timesteps $\in$[10, 200]), the intrinsic rewards are higher than when the agent is stuck.
+In contrast, when the agent is stuck at the bottom of the flagpole, the intrinsic rewards are close to zero (e.g., timesteps $\in$[250, 380]).
+Figure~\ref{fig:structure_overview} shows the overview of three networks used in DAIM.
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=0.7\linewidth]{figures/chapter5/daim-struct.pdf}
+    \caption[Overview of three networks in DAIM.]{Overview of three networks in DAIM: intrinsic reward module, policy network, and value network.}
+    \label{fig:structure_overview}
+\end{figure}
 
@@
     \end{subfigure}
-    \caption{Visualisation of (a) MuJoCo Locomotion Environment: HalfCheetah, Hopper, Walker2d, Ant, and Humanoid, and (b) Arcade Learning Environment: Alien, BeamRider, Breakout, MsPacman, UpNDown, and SuperMarioBros.}
+    \caption[Evaluation environments of DAIM.]{Visualisation of (a) MuJoCo Locomotion Environment: HalfCheetah, Hopper, Walker2d, Ant, and Humanoid, and (b) Arcade Learning Environment: Alien, BeamRider, Breakout, MsPacman, UpNDown, and SuperMarioBros.}
     \label{fig:ch5_envs}
@@
     \includegraphics[width=1.0\linewidth]{figures/chapter5/mujoco_result_54_seaborn.pdf}
-    \caption{Comparison between $\ourmethod$ and baseline approaches (PPO, LIRPG, and VIME) in the MuJoCo locomotion environment with standard and delayed reward settings.
+    \caption[Results of DAIM in the MuJoCo environment with standard and delayed reward settings.]{Comparison between $\ourmethod$ and baseline approaches (PPO, LIRPG, and VIME) in the MuJoCo locomotion environment with standard and delayed reward settings.
 The $x$-axis represents the number of timesteps in training.
@@
     }
-    \caption{Quantitative results comparison between DAIM and other baseline methods (PPO, LIRPG, and VIME) in the MuJoCo locomotion environment with the standard reward setting.
+    \caption[Average reward of DAIM in the MuJoCo environments with the standard reward setting in the final epoch during training.]{Quantitative results comparison between DAIM and other baseline methods (PPO, LIRPG, and VIME) in the MuJoCo locomotion environment with the standard reward setting.
 In the table, it shows final average extrinsic rewards $\pm$ standard error across 5 different seeds of last 100 episodes in the training.
@@
     \end{tabular}}
-    \caption{Quantitative results comparison between DAIM and other baseline methods (PPO, LIRPG, and VIME) in the MuJoCo locomotion environment with the delayed reward setting.
+    \caption[Average reward of DAIM in the MuJoCo environments with the delayed reward settings in the final epoch during training.]{Quantitative results comparison between DAIM and other baseline methods (PPO, LIRPG, and VIME) in the MuJoCo locomotion environment with the delayed reward setting.
 In the table, it shows final average extrinsic rewards $\pm$ standard error across 5 different seeds of last 100 episodes in the training.
@@
   \end{subfigure}\hfill
-  \caption{The performance of different feature embedding methods in the Walker2d task with standard and delayed reward settings. DAIM achieves higher accumulated extrinsic rewards.} 
+  \caption[Results of different feature embedding methods in the Walker2d task with standard and delayed reward settings.]{The performance of different feature embedding methods in the Walker2d task with standard and delayed reward settings.
+DAIM achieves higher accumulated extrinsic rewards.}
   \label{fig:walker2d_results}
@@
   \end{subfigure}\hfill
-  \caption{Using {\em only} intrinsic rewards ($\lambda^{\text{ex}}=0$), we evaluate the performance (accumulated {\it extrinsic} rewards on vertical axes) of different feature embedding methods in all five MuJoCo tasks with the standard reward setting.}
+  \caption[Results of using intrinsic reward only with different feature embedding methods in the MuJoCo environment with the standard reward setting.]{Using {\em only} intrinsic rewards ($\lambda^{\text{ex}}=0$), we evaluate the performance (accumulated {\it extrinsic} rewards on vertical axes) of different feature embedding methods in all five MuJoCo tasks with the standard reward setting.}
   \label{fig:walker2d_results_r_in}
@@
 In this case, the agent will maximise the diversity regardless of its effect on overall performance in terms of the extrinsic returns.
+For example, in Figure~\ref{fig:why_daim}, the agent selects an action in $s_{t}$, which causes it to fall into a trap in $s_{t+1}$.
+When the raw observation feature is used as the feature embedding method, the selected action leads to high diversity between the adjacent states, however the agent fails in this episode.
+Therefore, simply increasing diversity may not lead to better performance and it is necessary to learn the features that can better quantify the diversity between adjacent states to improve the task-specific performance (i.e., extrinsic returns) of the agent.
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=0.9\textwidth]{figures/chapter5/diversity_fail.pdf}
+    \caption{An example to show that higher diversity may not lead to better performance.}
+    \label{fig:why_daim}
+\end{figure}
 
@@
   \end{subfigure}\hfill
-  \caption{The performance of using different size $k$ of consecutive states for calculating intrinsic rewards in the Walker2d task with standard and delayed reward settings.
+  \caption[Results of using different length of consecutive states for calculating the intrinsic reward.]{The performance of using different size $k$ of consecutive states for calculating intrinsic rewards in the Walker2d task with standard and delayed reward settings.
 The $y$-axis represents the average extrinsic rewards over the last 100 training episodes.}
@@
 For Delay = 40, both $k=2$ and $k=4$ provide improvements over baselines PPO and VIME (see Figure~\ref{fig:mujoco_results}), with $k=2$ conveying the best advantage.
-\td{This may be due to the fact that two adjacent states can convey information about the direction and speed of the movement.} Therefore, $k=2$ was selected as the default setting for our experiments.
+This may be due to the fact that two adjacent states can convey information about the direction and speed of the movement; this is sufficient to convey some advantage in training.
+Therefore, $k=2$ was selected as the default setting for our experiments.
 
@@
   \end{subfigure}\hfill
-  \caption{Comparison between $\ourmethod$ and baseline approaches (A2C and LIRPG) in the Atari and SuperMarioBros games.
+  \caption[Results of DAIM in the arcade learning environment.]{Comparison between $\ourmethod$ and baseline approaches (A2C and LIRPG) in the Atari and SuperMarioBros games.
 The $x$-axis represents the number of timesteps in the training.
@@
     }
-    \caption{Quantitative results comparison between DAIM and other baseline methods (A2C and LIRPG) in the Atari and SuperMarioBro games.
+    \caption[Average reward of DAIM in the arcade learning environment in the final epoch during training.]{Quantitative results comparison between DAIM and other baseline methods (A2C and LIRPG) in the Atari and SuperMarioBro games.
 In the table, it shows final average extrinsic rewards $\pm$ standard error across 5 different seeds of last 100 episodes in the training.
@@
 \section{Discussion}
+\begin{table}[h!]
+    \centering
+    \resizebox{0.7\textwidth}{!}{
+    \begin{tabular}{c|c|c|c|c}
+    \toprule
+           & PPO~\cite{schulman2017proximal} & LIRPG~\cite{zheng2018learning} & VIME~\cite{NIPS2016_abd81528} & DAIM\\
+        \midrule
+        Time  & 00:28:50 & 1:23:43 & 03:18:36 & 01:26:21 \\
+        \bottomrule
+    \end{tabular}
+    }
+    \vspace{0.2em}
+    \caption[Training time comparison between DAIM and other baselines.]{Training time (hours:minutes:seconds) of DAIM and baseline approaches on the MuJoCo tasks.}
+    \label{tab:time_daim}
+\end{table}
+
 In this chapter, the DPPs framework is employed to measure the diversity of state sequences; it is shown this can be used to create an intrinsic reward that encourages state-space exploration in on-policy learning; we term this variant diversity-augmented intrinsic motivation (DAIM).
-
 The update of the proposed intrinsic reward module is formulated as a bi-level optimisation problem.
-During training, as the agent tries to maximise intrinsic returns, the bi-level optimisation guarantees that extrinsic returns are also maximised.
-
-Based on the extensive experiments, DAIM achieves improvements in continuous control in delayed-reward problems when compared to well-known on-policy training approaches of similar complexity, such as PPO~\cite{schulman2017proximal}, LIRPG~\cite{zheng2018learning} and VIME~\cite{NIPS2016_abd81528}.
-
+During training, as the agent tries to maximise mixture returns, the bi-level optimisation guarantees that extrinsic returns are also maximised.
+Based on the extensive experiments, DAIM achieves improvements in continuous control in delayed-reward problems when compared to well-known on-policy training approaches of similar complexity, such as LIRPG~\cite{zheng2018learning} and VIME~\cite{NIPS2016_abd81528}.
+Table~\ref{tab:time_daim} displays the training time of DAIM and other baselines on one MuJoCo task with a single random seed.
+From the table, DAIM requires longer training time than the vanilla PPO because of its bi-level optimisation steps.
+However, DAIM and other extension algorithms for PPO have similar training time, such as LIRPG and VIME.
 DAIM also accelerates the convergence of training agents on ALE compared to the baseline training approaches.
-In the work of this paper, DAIM is combined with PPO and A2C; in principle, DAIM could also be applied to other on-policy RL algorithms, such as TRPO~\cite{schulman2015trust} or SVPG~\cite{liu2017stein}.
+In the work of this chapter, DAIM is combined with PPO and A2C; in principle, DAIM could also be applied to other on-policy RL algorithms, such as TRPO~\cite{schulman2015trust} or SVPG~\cite{liu2017stein}.
 
@@
 In the future, we will investigate how to extend the DAIM approach to off-policy sampling and address more challenging tasks, such as Montezuma's Revenge.
+Furthermore, we can also combine the self-imitation learning loss in Chapter~\ref{ch:esil} with the diversity-based intrinsic rewards to achieve a further improvement in tasks with a sparse reward setting.
 %with even sparser reward structure

### 2022-08-21 chapters/chapter6.tex
update the final version

--- 2022-04-23 chapters/chapter_tmp.tex
@@
-\chapter{Interpretability of DRL}
+\chapter{Analysing DRL Agents Trained with Domain Randomisation}
 \label{ch:drl_interp}
 \section{Introduction}
-In previous chapters, we have discussed several methods to improve sample efficiency in deep reinforcement learning (DRL), especially in complex robot manipulation tasks with continuous control.
-However, all of the above experiments are carried out in a simulated environment.
+% In previous chapters, we have discussed several methods to improve sample efficiency in deep reinforcement learning (DRL), especially in complex robot manipulation tasks with continuous, rather than discrete, control.
+In this thesis, we first provide a comprehensive analysis on the trained DRL agents, which aims to understand features and policies that the agents learned in the environment.
+Specifically, proximal policy optimisation (PPO)~\cite{schulman2017proximal} is selected as the DRL algorithm to train the model, because it is widely used in a variety of scenarios~\cite{berner2019dota,rajeswaran2017learning,vinyals2019grandmaster}, we believe that this analysis can provide more insight to the community.
+In the following chapters, we will further improve the sample efficiency of PPO algorithm with different ways in the tasks with sparse or delayed rewards.
+Recently, most DRL experiments are performed in a simulated environment.
 Training becomes much more problematic when applying DRL to real-world applications.
@@
 
-To reduce \td{real-world training times}, a number of end-to-end approaches~\cite{gu2017deep,levine2016end,zhu2017target} have been proposed in recent years, which enable the agent to infer the policy according to raw observations (e.g., images) directly.
+%To reduce \td{real-world training times}, 
+A number of end-to-end approaches~\cite{gu2017deep,levine2016end,zhu2017target} have been proposed in recent years, which enable the agent to infer the policy according to raw observations (e.g., images) directly.
 However, such an end-to-end training setup would require the agent to collect samples for a long time in a real-world environment to update the policy network, which may increase the risk of robot/hardware failure and thus elevate the cost of training significantly.
@@
   \includegraphics[width=\linewidth]{figures/chapter6/domain_random.png}
-  \caption{Examples of domain randomisation (visual appearances) in the simulated Fetch (top) and Jaco (bottom) robots environments.}
+  \caption[Examples of visual domain randomisation in the simulated robot environments.]{Examples of domain randomisation (visual appearances) in the simulated Fetch (top) and Jaco (bottom) robots environments.}
   \label{fig:dr_example}
@@
   \includegraphics[width=\linewidth]{figures/chapter6/strategies.png}
-  \caption{The range of strategies an agent may learn to use when trained to reach a red target, which we split into three components.
+  \caption[The range of strategies an agent may learn to use when trained to reach a red target.]{The range of strategies an agent may learn to use when trained to reach a red target, which we split into three components.
 Firstly, the agent must visually localise the target---which can be accomplished by detecting red, spherical objects, or simply anything with a significant red component.
@@
 Finally, the agent may accomplish the task with varying levels of robustness to changes in the rest of the visual scene.
-Using a range of analyses, we show that agents trained to accomplish the same task learn different subsets of these strategies, depending on their training conditions.}
+Using a range of analyses, we show that agents trained to accomplish the same task learn different subsets of these strategies, depending on their training conditions.
+[This diagram was created by Dr.Kai Arulkumaran and Tamara Gerbert~\cite{dai2022analysing}.]}
   \label{fig:strategies}
@@
 
-In this chapter, we aim to understand \td{how DRL agents make decision in the environment} and understand how agents trained with DR behave differently from agents trained without.
+In this chapter, we aim to understand how DRL agents make decision in the environment and understand how agents trained with visual DR behave differently from agents trained without.
 To this end, we train DRL agents under a range of different settings and use relative differences to better interpret their policies.
 In contrast to most prior work examining DRL policies~\cite{atrey2020exploratory,greydanus2018visualizing, olson2021counterfactual,puri2020explain,rupprecht2020finding,such2018atari,zahavy2016graying}, we not only provide explanations for how policies react to parts of states/individual states, but use our array of experiments to characterise their \emph{global} strategies (see Figure~\ref{fig:strategies}).
-\td{Through these experiments and analysis, we are able to address following questions:
+Through these experiments and analysis, we are able to provide some insight into the following questions:
 \begin{itemize}
-    \item What is the role of DR on layer learning?
-    \item To what extent is visual inputs rather than proprioceptive inputs used by the agent?
     \item How robust are agents to unseen scenarios?
+    \item To what extent are visual inputs rather than proprioceptive inputs used by the agent?
     \item How redundant is the network in its representation? 
+    \item What is rate of learning in different layers under different training configurations?
 \end{itemize}
-}
 %For instance, we examine the relative importance of different input modalities, and whether agents need their memory to complete tasks.
@@
 In particular, we train different robots with vision-based inputs to perform the task of reaching for a (red) target, and then use different techniques to understand what features of the environment the policies react to.
+\textit{Unlike} many experiments reported in the literature~\cite{andrychowicz2017hindsight,fang2018dher,fang2019curriculum}, the agent does not have direct access to the coordinates of the ball as a state.
 Under certain training settings, the agent may localise the target by simply looking for anything red, and it is only under more challenging training conditions that the agent utilises both colour and shape.
-Clearly, the latter strategy better represents what we want the agent to learn, while the former is merely a ``shortcut'' \cite{geirhos2020shortcut}.
-
-To be noticed, the structure in this chapter is organised in a different way from the previous ones.
+Clearly, the latter strategy better represents what we want a real-world agent to learn, while the former is merely a ``shortcut'' \cite{geirhos2020shortcut}.
+
+The structure in this chapter is organised in a different way from the previous ones.
 In Section~\ref{ch6:sec_interp}, we first introduce the interpretability techniques used in this work.
@@
 \section{Interpretability Techniques}
-Neural network (NN) has achieved great success in various fields, benefiting from its powerful feature representation capabilities.
-However, compared with hand-crafted feature descriptors, NN lacks interpretability, and it treats the training as black boxes.
+Neural networks (NNs) have achieved great success when used with reinforcement learning.
+However, compared to hand-crafted feature descriptors, NNs lack interpretability, and the networks are often treated as black boxes.
 Therefore, it is important that we are able to explain and understand what the NN has learned before deploying it in real-world applications, where safety and reliability are crucial.
@@
   \begin{subfigure}{0.24\columnwidth}
+    \includegraphics[width=\linewidth]{figures/chapter6/average_map_mask.png}
+    \subcaption{Avg. $\mathbf{x}^{base}$ mask}
+  \end{subfigure}
+  \begin{subfigure}{0.24\columnwidth}
     \includegraphics[width=\linewidth]{figures/chapter6/occ_jaco_average_map.png}
@@
   \end{subfigure}
-  \begin{subfigure}{0.24\columnwidth}
-    \includegraphics[width=\linewidth]{figures/chapter6/average_map_mask.png}
-    \subcaption{Avg. $\mathbf{x}^{base}$ mask}
-  \end{subfigure}
-  \caption{Occlusion-based saliency map method (a) using Gaussian mask (c) our improved method using an average image (b, d).
+  \caption[Visualisation of different occlusion-based saliency map methods.]{Occlusion-based saliency map methods: (a) using Gaussian blur mask and (d) using an average image (b, c).
 Occlusion-based method using Gaussian blur masks (a) shows extra artifacts.
-Our improved saliency map method highlights important regions in the result by using average baseline image created from plentiful trajectories (c), (d) shows the usage of the average baseline image.}
+Our improved saliency map method (d) highlights important regions in the result by using average baseline image created from plentiful trajectories (b).
+(c) shows one instance step that using the average baseline image as the occlusion in the original input.
+The occlusion region is highlighted with a blue outline.}
   \label{fig:saliency_baseline}
@@
 Similarly to previous work on understanding DRL agents~\cite{greydanus2018visualizing}, we also use an occlusion-based approach for interpreting.
-The occlusion-based method~\cite{zeiler2014visualizing} generally adopts a gray mask to scan over the entire input image, and then measures the variation of the output in each masked region to identify which areas have significant influences on the output of the NN.
-The variation of the performance is also called saliency value, and it can be expressed as:
+The occlusion-based method~\cite{zeiler2014visualizing} generally adopts a gray mask to scan the entire input image and then measures the variation of the output in each masked region to identify which areas have significant influences on the output of the NN.
+The variation of the performance is also called the saliency value, and it can be expressed as:
 \begin{align}
@@
 \end{align}
-where $(m, n)$ is the location on the image, $F(\mathbf{x})$ is the output of the original input, and $F(\mathbf{x}_{m, n}^{occ})$ is the output of the input with an occlusion mask at location $(m, n)$.
-In the work of~\cite{greydanus2018visualizing}, which aims to understand DRL agents in Atari games, the gray mask is replaced with a Gaussian mask.
+where $(m, n)$ is the location on the image $\mathbf{x}$, $F(\mathbf{x})$ is the output of the original input, and $F(\mathbf{x}_{m, n}^{occ})$ is the output of the input with an occlusion mask at location $(m, n)$.
+In the work of~\cite{greydanus2018visualizing}, which aims to understand DRL agents in Atari games, the gray mask is replaced with a Gaussian blur mask.
 Because the agent only receives gray scale images in the training and testing.
@@
 
-However, the agent receives an RGB image as the input in our task.
+However, the agent receives an RGB image as an input in our task.
 As a result, the occlusion-based method using Gaussian blur masks cannot highlight any important regions (see Figure~\ref{fig:saliency_baseline}a), and thus fails to provide any valuable information to understand the agent.
@@
 In this work, we are more interested in task-specific objects in the environment, such as the robot arm and the red ball.
-To avoid interference from the occlusion-based masks on unimportant areas, we first average the frames of approximately 1,000 trajectories to get a baseline image (see Figure~\ref{fig:saliency_baseline} (b)).
-Then, we use this baseline image to generate the mask in the corresponding locations (see Figure~\ref{fig:saliency_baseline}d).
-Finally, we can get an improved saliency map, which highlights the informative regions in the input image  (see Figure~\ref{fig:saliency_baseline}c).
+To avoid interference from the occlusion-based masks on unimportant areas, we first sample 1,000 input frames with the trained policy from the environments (the policy is not updated during this process).
+Next, these 1000 frames are averaged to produce a baseline image (see Figure~\ref{fig:saliency_baseline}b).
+Then, we use this baseline image to generate the occlusion in the corresponding locations (see Figure~\ref{fig:saliency_baseline}c) in the input frame.
+Finally, we can measure the performance change and get an improved saliency map, which highlights the informative regions in the input image  (see Figure~\ref{fig:saliency_baseline}d).
 %Inspired by the work which use reference images for the calculation of interpreta
@@
 %\subsection{Statistical and Structural Weight Characterisations}
-\subsection{Unit Ablations}
-Unit ablations evaluate the importance of a single filter in the convolutional layer by masking it to zero and examining the variations in the performance.
+\subsection{Channel-wise Unit Ablations}
+Unit ablations evaluate the importance of a single unit in the convolutional layer by masking it to zero and examining the variations in the performance.
 In general, an important filter can lead to a significant performance drop.
+If a unit is set to zero and it has no effect on the ability to achieve a task, then it means one of two things: either the unit is redundant and performs no useful function in the network, or the function of the network is well-distributed amongst multiple neurons.
 While more sophisticated unit ablations have been applied in DRL~\cite{meyes2020you}, {this has only been in the context of a control task with a low-dimensional symbolic state space}.
+We further provide a Python style code snippet below to demonstrate the process of unit ablations.
+\begin{minted}[frame=lines, mathescape, linenos]{python}
+# Python code of unit ablations
+import torch
+for i in range(N_c): # N_c is the number of output channels
+    net.load_state_dict(torch.load(ckpt_path))
+    net.conv1.weight[i, :, :, :] = 0
+    net.conv1.bias[i] = 0
+    success_rate = eval(net)
+\end{minted}
 % Layer initialisation
@@
 The primary idea is to replace the parameters of a single fully trained layer with the parameters in the previous training epochs, and observe the effect on the performance of the network.
-In this work, the parameters of each layer can be expressed as: $\{\theta_{1}, \theta_{2}, ..., \theta_{L}\}$, where $L$ is the number of layers of the network, and $\theta_{l}^{t}$ represents the parameters of the $l^\text{th}$ layer in the training epoch $t\in [1,T]$.
+In this work, the parameters of each layer can be expressed as: $\{\theta_{1}, \theta_{2}, ..., \theta_{L}\}$, where $L$ is the number of layers in the network, and $\theta_{l}^{t}$ represents the parameters of the $l^\text{th}$ layer in the training epoch $t\in [1,T]$.
 For example, when training is finished, to evaluate the robustness of a single layer $l$ at training epoch $t$, we can re-initialise the parameters in the layer by: $\theta_{l}^{T} \leftarrow \theta_{l}^{t}$.
@@
 Recurrent ablation is used to evaluate if the recurrent layer (e.g., long short-term memory (LSTM)) is utilised in a network structure.
-To do so, we set the hidden state to a constant value.
-If the performance of the agent is decreased to some extent, it means the recurrent layer helps to perform the task, otherwise, the recurrent layer is redundant in this structure.
+To do so, we set the hidden and cell states to constant values.
+If the performance of the agent is decreased to some extent, it means that the recurrent layer helps to perform the task, otherwise, the recurrent layer is redundant in this structure.
 While it is possible to investigate the hidden states of DRL agents over time, visualising and inspecting a high-dimensional time series can be difficult, even for domain experts~\cite{jaunet2020drlviz}.
+More details and explanations are in Section~\ref{sec:detail_reucrrent_ablation}.
 
@@
   \end{subfigure}
-  \caption{(a) Fetch environment, (b) an example of Fetch visual input, (c) Jaco environment, and (d) an example of Jaco visual input.
-\td{The blue cuboids indicate the sampling space of the target objects.}}
+  \caption[Examples of Fetch and Jaco robot manipulation environments.]{(a) Fetch environment, (b) an example of Fetch visual input, (c) Jaco environment, and (d) an example of Jaco visual input.
+The blue cuboids indicate the sampling space of the target objects.}
   \label{fig:robot_setups}
@@
 In order to acquire visual inputs, we also add an extra RGB camera in front of the robot arm (see Figure~\ref{fig:fetch_view}).
-The control scheme remains the same, which is the relative position of the gripper in x, y and z axis.
+The control scheme remains the same, which is the relative position of the gripper in $x$, $y$ and $z$ axis.
 The target object is placed on the table.
-During training, the position of the target object is randomly sampled in the 2D space (see blue cuboid in Figure~\ref{fig:fetch_volumn}) of the table surface.
+During training, the position of the target object is randomly sampled in the 2D space (see the blue cuboid in Figure~\ref{fig:fetch_volumn}) of the table surface.
 During testing, 80 fixed uniformly distributed positions are used to evaluate the performance of the agent.
@@
 The action space is the 6 joint velocities of the robot arm.
-During training, the position of the target object is randomly sampled in a 3D space (see blue cuboid in Figure~\ref{fig:jaco_volumn}).
+During training, the position of the target object is randomly sampled in a 3D space (see the blue cuboid in Figure~\ref{fig:jaco_volumn}).
 During testing, 250 uniformly distributed fixed positions are used to evaluate the performance of the agent.
@@
   \includegraphics[width=0.85\linewidth]{figures/chapter6/axes.png}
-  \caption{The 8 different experimental conditions: 2 robots (Fetch vs. Jaco), 2 input modalities (vision vs. vision + proprioception), 2 visual conditions (DR vs. no DR).}
+  \caption[Visualisation of 8 different experimental conditions.]{The 8 different experimental conditions: 2 robots (Fetch vs.
+Jaco), 2 input modalities (vision vs.
+vision + proprioception), 2 visual conditions (DR vs.
+no DR).
+[This diagram was created by Dr.Kai Arulkumaran and Tamara Gerbert~\cite{dai2022analysing}.]}
   \label{fig:axes}
@@
 We setup 8 different configurations to cover different possible training strategies and provide a comprehensive comparison.
-There include different combinations over three primitives: 1) using Fetch robot arm or Jaco robot arm, 2) using visual input only or visual input + proprioceptive input, and 3) training with or without domain randomisation (DR) (see Figure~\ref{fig:axes}).
+There are different combinations over three primitives: 1) using Fetch robot arm or Jaco robot arm, 2) using visual input only or visual input + proprioceptive input, and 3) training with or without domain randomisation (DR) (see Figure~\ref{fig:axes}).
 
@@
 
-For real-world deployment, one would want to use as many sensors as it is feasible to use.
-However, we are interested in how the absence of specific input modalities can affect learning, as this changes how the agent is able to extract information from the environment.
+For a real-world deployment, one would want to use as many sensors as feasible.
+However, we are interested in how the absence of specific input modalities can affect learning, as this changes the way the agent is able to extract information from the environment.
 For example, while one might expect robots equipped with proprioceptive sensors to use these alone for pose estimation, agents can also learn to utilise vision, if extracting a pose from images is not too difficult.
@@
 We therefore test how the agents trained with or without DR differ.
-Supporting previous results, we show that agents trained with DR are more robust to out-of-distribution (OoD) perturbations.
+In support of previous results, we show that agents trained with DR are more robust to out-of-distribution (OoD) perturbations.
 In our experiments, we use a standard visual DR setup.
@@
   \end{tabular}}
-   \caption{DR textures used during training.
+   \caption[DR textures used during training.]{DR textures used during training.
 At every environment timestep, for each component (e.g., robot arm, skybox, table; the Jaco robot has 13 components, and the Fetch robot has 19 components), the appearance of the component is rendered using a sequential sampling process.
@@
   }
-\caption{Python pseudocode describing the DR texture generators. rgb1 and rgb2 are random RGB colours; the value of each RGB channel is drawn uniformly from $\{0 \ldots 255\}$.}
+\caption[Python pseudocode describing the DR texture generators.]{Python pseudocode describing the DR texture generators.
+rgb1 and rgb2 are random RGB colours; the value of each RGB channel is drawn uniformly from $\{0 \ldots 255\}$.}
 \label{tbl:texture_expression}
@@
   \end{subfigure}
-  \caption{Visualisation of different DR textures used during training: (a) checkerboard, (b) gradient, (c) flat\_rgb, (d) noise.}
+  \caption[Visualisation of different DR textures used during training.]{Visualisation of different DR textures used during training: (a) checkerboard, (b) gradient, (c) flat\_rgb, (d) noise.}
   \label{fig:random_vis}
@@
   %\hspace{0.025\textwidth}
-  \caption{The learning curves of Jaco agents trained with DR and with (a) or without (b) proprioceptive inputs.
+  \caption[The learning curves of Jaco agents trained with DR and with or without proprioceptive inputs.]{The learning curves of Jaco agents trained with DR and with (a) or without (b) proprioceptive inputs.
 (a) the agent have similar performance in both standard and DR environments.
@@
 
-Before formally testing and analysing the models, they are \td{fully trained} with 8 training configurations and the same hyperparameters to achieve promising performance.
+Before formally testing and analysing the models, they are fully trained with 8 training configurations and the same hyperparameters to achieve promising performance.
 However, in the Jaco environment, we also find that the model trained with visual + proprioceptive inputs and DR has a performance degradation when using standard visual inputs (no DR; see Figure~\ref{fig:domain_shift}).
@@
   \end{subfigure}
-  \caption{The visualisation of different test conditions in the Fetch environment.}
+  \caption[The visualisation of different test conditions in the Fetch environment.]{The visualisation of different test conditions in the Fetch environment.
+The agents are trained only in the standard condition.
+There are no extra distractors introduced during training.}
   \label{fig:tests}
@@
 
-\item[g)] \textbf{Reflection}: This condition makes the table of the Fetch environment and the floor of the Jaco environment reflective, and it can produces mirror-like reflections in these surfaces.
+\item[g)] \textbf{Reflection}: This condition makes the table of the Fetch environment and the floor of the Jaco environment reflective, and it can produce mirror-like reflections in these surfaces.
 
 \item[h)] \textbf{Translation}: This condition shifts the RGB camera's location.
-We use 5 locations: -20 to 20cm in the y direction for the Fetch environment, and -20 to 20cm in the x direction for the Jaco environment.
+We use 5 locations: -20 to 20cm in the $y$ direction for the Fetch environment, and -20 to 20cm in the $x$ direction for the Jaco environment.
 
@@
 
-In the Fetch environment, the trained agents are more robust to the colour distractors.
-However, the shape distractors decrease the performance of the agents trained without DR, especially for non-DR agents trained with proprioceptive inputs, which drops to a 0.48 success rate.
-This implies that the non-DR agent trained with proprioceptive inputs localises the target object primarily by colour information and partly with shape information.
-The non-DR agent trained with visual inputs only merely drops 0.17 success rate.
-\td{This is because the agent trained with only visual inputs requires better representation/feature learning to localise the position of the robot arm (i.e., self-localisation), and these learned features can help the agent to distinguish between different shapes and colours.}
-
-In the Jaco environment, both non-DR agents show performance degradation when there exists a distractor with red components, and both DR agents have significant performance drops when the shape of the distractor is a diamond, which has a similar appearance to a sphere in a low-resolution viewed input.
-It also indicates that the non-DR agents prefer to localise the target object via red colour.
-
-In addition, we also test the effect of the distractors on the agent in different positions. In Table~\ref{tbl:vary_distractor}, it only has a few influences on performance.
-
-\begin{table}[H]
+\begin{table}[h]
 %\begin{sidewaystable}[h!]
@@
     \toprule
-    Robot & DR & Prop. & Standard & Colours & Shapes & Illumination (Lvl.) & Illumination (Dir.) & Noise & Reflection & Translation & Invisibility\\
+    Robot & DR & Prop. & Standard & Colours & Shapes & Illum (Lvl.) & Illum (Dir.) & Noise & Reflection & Translation & Invisibility\\
     \midrule
-    Fetch & {\xmark}   & {\xmark}   & 1.000$\pm 0.000$ & 0.993$\pm 0.004$ & 0.827$\pm 0.056$ & 0.694$\pm 0.049$ & 0.539$\pm 0.069$ & 0.980$\pm 0.006$ & 0.447$\pm 0.039$ & 0.270$\pm 0.079$ & 0.000$\pm 0.000$\\
-    Fetch & {\xmark}   & {\cmark} & 1.000$\pm 0.000$ & 0.937$\pm 0.027$ & 0.524$\pm 0.087$ & 0.593$\pm 0.064$ & 0.491$\pm 0.066$ & 0.988$\pm 0.004$ & 0.570$\pm 0.078$ & 0.231$\pm 0.078$ & 0.000$\pm 0.000$\\
-    Fetch & {\cmark} & {\xmark}   & 0.983$\pm 0.004$ & 0.974$\pm 0.004$ & 0.939$\pm 0.018$ & 0.957$\pm 0.008$ & 0.740$\pm 0.058$ & 0.985$\pm 0.007$ & 0.972$\pm 0.011$ & 0.480$\pm 0.064$ & 0.000$\pm 0.000$\\
-    Fetch & {\cmark} & {\cmark} & 0.997$\pm 0.002$ & 0.997$\pm 0.001$ & 0.978$\pm 0.009$ & 0.992$\pm 0.002$ & 0.883$\pm 0.038$ & 0.970$\pm 0.008$ & 0.985$\pm 0.005$ & 0.536$\pm 0.063$ & 0.023$\pm 0.015$\\
+    Fetch & {\xmark}   & {\xmark}   & 1.00$\pm 0.00$ & 0.99$\pm 0.00$ & 0.83$\pm 0.06$ & 0.69$\pm 0.05$ & 0.54$\pm 0.07$ & 0.98$\pm 0.00$ & 0.45$\pm 0.04$ & 0.27$\pm 0.08$ & 0.00$\pm 0.00$\\
+    Fetch & {\xmark}   & {\cmark} & 1.00$\pm 0.00$ & 0.94$\pm 0.08$ & 0.52$\pm 0.09$ & 0.59$\pm 0.06$ & 0.49$\pm 0.07$ & 0.99$\pm 0.00$ & 0.57$\pm 0.08$ & 0.23$\pm 0.08$ & 0.00$\pm 0.00$\\
+    Fetch & {\cmark} & {\xmark}   & 0.98$\pm 0.00$ & 0.97$\pm 0.00$ & 0.94$\pm 0.02$ & 0.96$\pm 0.01$ & 0.74$\pm 0.06$ & 0.99$\pm 0.01$ & 0.97$\pm 0.01$ & 0.48$\pm 0.06$ & 0.00$\pm 0.00$\\
+    Fetch & {\cmark} & {\cmark} & 1.00$\pm 0.00$ & 1.00$\pm 0.00$ & 0.98$\pm 0.01$ & 0.99$\pm 0.00$ & 0.88$\pm 0.04$ & 0.97$\pm 0.01$ & 0.99$\pm 0.01$ & 0.54$\pm 0.06$ & 0.02$\pm 0.02$\\
     \hline
-    Jaco  & {\xmark}   & {\xmark}   & 0.995$\pm 0.003$ & 0.772$\pm 0.058$ & 0.367$\pm 0.069$ & 0.966$\pm 0.012$ & 0.838$\pm 0.035$ & 0.635$\pm 0.028$ & 0.734$\pm 0.032$ & 0.556$\pm 0.063$ & 0.000$\pm 0.000$\\
-    Jaco  & {\xmark} & {\cmark} & 0.995$\pm 0.001$ & 0.808$\pm 0.044$ & 0.382$\pm 0.064$ & 0.876$\pm 0.032$ & 0.911$\pm 0.014$ & 0.478$\pm 0.059$ & 0.618$\pm 0.061$ & 0.688$\pm 0.048$ & 0.001$\pm 0.001$\\
-    Jaco  & {\cmark} & {\xmark}   & 0.650$\pm 0.056$ & 0.652$\pm 0.022$ & 0.593$\pm 0.031$ & 0.643$\pm 0.027$ & 0.549$\pm 0.032$ & 0.575$\pm 0.040$ & 0.429$\pm 0.060$ & 0.310$\pm 0.041$ & 0.007$\pm 0.002$\\
-    Jaco  & {\cmark} & {\cmark} & 0.991$\pm 0.004$ & 0.989$\pm 0.002$ & 0.843$\pm 0.061$ & 0.816$\pm 0.043$ & 0.882$\pm 0.031$ & 0.896$\pm 0.007$ & 0.946$\pm 0.006$ & 0.658$\pm 0.054$ & 0.916$\pm 0.022$\\
+    Jaco  & {\xmark}   & {\xmark}   & 1.00$\pm 0.00$ & 0.77$\pm 0.06$ & 0.37$\pm 0.07$ & 0.97$\pm 0.01$ & 0.84$\pm 0.04$ & 0.64$\pm 0.03$ & 0.73$\pm 0.03$ & 0.56$\pm 0.06$ & 0.00$\pm 0.00$\\
+    Jaco  & {\xmark} & {\cmark} & 1.00$\pm 0.00$ & 0.81$\pm 0.04$ & 0.38$\pm 0.06$ & 0.88$\pm 0.03$ & 0.91$\pm 0.01$ & 0.48$\pm 0.06$ & 0.62$\pm 0.06$ & 0.69$\pm 0.05$ & 0.00$\pm 0.00$\\
+    Jaco  & {\cmark} & {\xmark}   & 0.65$\pm 0.06$ & 0.65$\pm 0.02$ & 0.59$\pm 0.03$ & 0.64$\pm 0.03$ & 0.55$\pm 0.03$ & 0.58$\pm 0.04$ & 0.43$\pm 0.06$ & 0.31$\pm 0.04$ & 0.01$\pm 0.00$\\
+    Jaco  & {\cmark} & {\cmark} & 0.99$\pm 0.00$ & 0.99$\pm 0.00$ & 0.84$\pm 0.06$ & 0.82$\pm 0.04$ & 0.88$\pm 0.03$ & 0.90$\pm 0.01$ & 0.95$\pm 0.01$ & 0.66$\pm 0.05$ & 0.92$\pm 0.02$\\
     \bottomrule
   \end{tabular}}
-  \caption{Test performance of all models with local visual changes (distractors), global visual changes, and invisibility (visual self-localisation test).
+  \caption[Test performance of all models under different test conditions.]{Test performance of all models with local visual changes (distractors), global visual changes, and invisibility (visual pose estimation test).
 Colour and shape results are averaged over 9 different distractor locations, as well as all colours and shapes, respectively.
@@
 \end{table}
+
+In the Fetch environment, the trained agents are more robust to the colour distractors.
+However, the shape distractors decrease the performance of the agents trained without DR, especially for agents trained with proprioceptive inputs and without DR, which reduces the success rate by 0.48 (the relative changes in its performance in the standard test environment, which is 1.00$\pm$0.00; see Table~\ref{tbl:tests}).
+This implies that the agent trained with proprioceptive inputs and without DR localises the target object primarily by colour information and partly with shape information.
+The agent trained with visual inputs and without DR merely drops 0.17 success rate (the relative changes in its performance in the standard test environment, which is 1.00$\pm$0.00; see Table~\ref{tbl:tests}).
+This is because the agent trained with only visual inputs requires better representation/feature learning to estimate the pose of the robot arm (i.e., self-localisation), and these learned features can help the agent to distinguish between different shapes and colours.
+
+In the Jaco environment, both agents trained without DR show performance degradation when there exists a distractor with red components, and both agents trained with DR have significant performance drops when the shape of the distractor is a diamond, which has a similar appearance to a sphere in a low-resolution viewed input.
+It also indicates that the agents trained without DR prefer to localise the target object via red colour.
+
+In addition, we also test the effect of the distractors on the agent in different positions. In Table~\ref{tbl:vary_distractor}, it only has a few influences on performance.
+
+% \begin{table}[H]
+% %\begin{sidewaystable}[h!]
+%   %\label{tbl:tests}
+%   \resizebox{\linewidth}{!}{
+%   \begin{tabular}{c|cc|c|cc|cc|ccc|c}
+%     \toprule
+%     Robot & DR & Prop. & Standard & Colours & Shapes & Illumination (Lvl.) & Illumination (Dir.) & Noise & Reflection & Translation & Invisibility\\
+%     \midrule
+%     Fetch & {\xmark}   & {\xmark}   & 1.000$\pm 0.000$ & 0.993$\pm 0.004$ & 0.827$\pm 0.056$ & 0.694$\pm 0.049$ & 0.539$\pm 0.069$ & 0.980$\pm 0.006$ & 0.447$\pm 0.039$ & 0.270$\pm 0.079$ & 0.000$\pm 0.000$\\
+%     Fetch & {\xmark}   & {\cmark} & 1.000$\pm 0.000$ & 0.937$\pm 0.027$ & 0.524$\pm 0.087$ & 0.593$\pm 0.064$ & 0.491$\pm 0.066$ & 0.988$\pm 0.004$ & 0.570$\pm 0.078$ & 0.231$\pm 0.078$ & 0.000$\pm 0.000$\\
+%     Fetch & {\cmark} & {\xmark}   & 0.983$\pm 0.004$ & 0.974$\pm 0.004$ & 0.939$\pm 0.018$ & 0.957$\pm 0.008$ & 0.740$\pm 0.058$ & 0.985$\pm 0.007$ & 0.972$\pm 0.011$ & 0.480$\pm 0.064$ & 0.000$\pm 0.000$\\
+%     Fetch & {\cmark} & {\cmark} & 0.997$\pm 0.002$ & 0.997$\pm 0.001$ & 0.978$\pm 0.009$ & 0.992$\pm 0.002$ & 0.883$\pm 0.038$ & 0.970$\pm 0.008$ & 0.985$\pm 0.005$ & 0.536$\pm 0.063$ & 0.023$\pm 0.015$\\
+%     \hline
+%     Jaco  & {\xmark}   & {\xmark}   & 0.995$\pm 0.003$ & 0.772$\pm 0.058$ & 0.367$\pm 0.069$ & 0.966$\pm 0.012$ & 0.838$\pm 0.035$ & 0.635$\pm 0.028$ & 0.734$\pm 0.032$ & 0.556$\pm 0.063$ & 0.000$\pm 0.000$\\
+%     Jaco  & {\xmark} & {\cmark} & 0.995$\pm 0.001$ & 0.808$\pm 0.044$ & 0.382$\pm 0.064$ & 0.876$\pm 0.032$ & 0.911$\pm 0.014$ & 0.478$\pm 0.059$ & 0.618$\pm 0.061$ & 0.688$\pm 0.048$ & 0.001$\pm 0.001$\\
+%     Jaco  & {\cmark} & {\xmark}   & 0.650$\pm 0.056$ & 0.652$\pm 0.022$ & 0.593$\pm 0.031$ & 0.643$\pm 0.027$ & 0.549$\pm 0.032$ & 0.575$\pm 0.040$ & 0.429$\pm 0.060$ & 0.310$\pm 0.041$ & 0.007$\pm 0.002$\\
+%     Jaco  & {\cmark} & {\cmark} & 0.991$\pm 0.004$ & 0.989$\pm 0.002$ & 0.843$\pm 0.061$ & 0.816$\pm 0.043$ & 0.882$\pm 0.031$ & 0.896$\pm 0.007$ & 0.946$\pm 0.006$ & 0.658$\pm 0.054$ & 0.916$\pm 0.022$\\
+%     \bottomrule
+%   \end{tabular}}
+%   \caption[Test performance of all models under different test conditions.]{Test performance of all models with local visual changes (distractors), global visual changes, and invisibility (visual pose estimation test).
+Colour and shape results are averaged over 9 different distractor locations, as well as all colours and shapes, respectively.
+Checkmarks and crosses indicate enabling/disabling DR and proprioceptive inputs (Prop.), respectively.
+Mean $\pm$ standard error are calculated over all models (seeds) and test target locations.}
+%   \label{tbl:tests}
+% %\end{sidewaystable}
+% \end{table}
+
+% \begin{table}[h!]
+% %\begin{sidewaystable}[h!]
+%   \resizebox{\linewidth}{!}{
+%   \begin{tabular}{c|cc|ccccc|cccc}
+%     \toprule
+%     Robot & DR & Prop. & Yellow & Blue & Green & Brown & Purple & Cube & Ellipsoid & Rectangle & Diamond\\
+%     \midrule
+%     Fetch & {\xmark}   & {\xmark}   & 0.993$\pm 0.007$ & 1.000$\pm 0.000$ & 1.000$\pm 0.000$ & 0.978$\pm 0.017$ & 0.995$\pm 0.004$ & 0.775$\pm 0.085$ & 0.993$\pm 0.007$ & 0.960$\pm 0.036$ & 0.580$\pm 0.142$\\
+%     Fetch & {\xmark}   & {\cmark} & 0.875$\pm 0.088$ & 1.000$\pm 0.000$ & 1.000$\pm 0.000$ & 0.825$\pm 0.069$ & 0.983$\pm 0.006$ & 0.243$\pm 0.064$ & 0.980$\pm 0.006$ & 0.783$\pm 0.079$ & 0.090$\pm 0.048$\\
+%     Fetch & {\cmark} & {\xmark}   & 0.970$\pm 0.011$ & 0.980$\pm 0.006$ & 0.975$\pm 0.006$ & 0.973$\pm 0.010$ & 0.973$\pm 0.011$ & 0.913$\pm 0.042$ & 0.975$\pm 0.009$ & 0.965$\pm 0.013$ & 0.905$\pm 0.048$\\
+%     Fetch &{\cmark} & {\cmark} & 0.995$\pm 0.003$ & 0.998$\pm 0.002$ & 1.000$\pm 0.000$ & 0.998$\pm 0.002$ & 0.995$\pm 0.003$ & 0.963$\pm 0.020$ & 0.998$\pm 0.002$ & 0.995$\pm 0.003$ & 0.958$\pm 0.022$\\
+%     \hline
+%     Jaco  & {\xmark}   & {\xmark}   & 0.281$\pm 0.067$ & 0.871$\pm 0.036$ & 0.994$\pm 0.003$ & 0.730$\pm 0.095$ & 0.984$\pm 0.008$ & 0.274$\pm 0.077$ & 0.407$\pm 0.098$ & 0.759$\pm 0.066$ & 0.027$\pm 0.006$\\
+%     Jaco  & {\xmark}   & {\cmark} & 0.451$\pm 0.040$ & 0.941$\pm 0.032$ & 0.993$\pm 0.002$ & 0.706$\pm 0.064$ & 0.950$\pm 0.028$ & 0.258$\pm 0.044$ & 0.440$\pm 0.058$ & 0.784$\pm 0.044$ & 0.045$\pm 0.009$\\
+%     Jaco  & {\cmark} & {\xmark}   & 0.640$\pm 0.046$ & 0.655$\pm 0.055$ & 0.651$\pm 0.055$ & 0.668$\pm 0.046$ & 0.648$\pm 0.046$ & 0.636$\pm 0.040$ & 0.642$\pm 0.035$ & 0.666$\pm 0.047$ & 0.426$\pm 0.054$\\
+%     Jaco  & {\cmark} & {\cmark} & 0.987$\pm 0.005$ & 0.990$\pm 0.003$ & 0.987$\pm 0.003$ & 0.990$\pm 0.003$ & 0.990$\pm 0.003$ & 0.970$\pm 0.017$ & 0.970$\pm 0.011$ & 0.990$\pm 0.003$ & 0.441$\pm 0.128$\\
+%     \bottomrule
+%   \end{tabular}}
+%   \caption[Test performance of all models with local visual changes.]{Test performance of all models with local visual changes (distractors).
+Checkmarks and crosses indicate enabling/disabling DR and proprioceptive inputs (Prop.), respectively.
+Mean $\pm$ standard error are calculated over all models (seeds), test target locations, and 9 distractor locations.}
+%   \label{tbl:diff_colours}
+% %\end{sidewaystable}
+% \end{table}
 
@@
     \midrule
-    Fetch & {\xmark}   & {\xmark}   & 0.993$\pm 0.007$ & 1.000$\pm 0.000$ & 1.000$\pm 0.000$ & 0.978$\pm 0.017$ & 0.995$\pm 0.004$ & 0.775$\pm 0.085$ & 0.993$\pm 0.007$ & 0.960$\pm 0.036$ & 0.580$\pm 0.142$\\
-    Fetch & {\xmark}   & {\cmark} & 0.875$\pm 0.088$ & 1.000$\pm 0.000$ & 1.000$\pm 0.000$ & 0.825$\pm 0.069$ & 0.983$\pm 0.006$ & 0.243$\pm 0.064$ & 0.980$\pm 0.006$ & 0.783$\pm 0.079$ & 0.090$\pm 0.048$\\
-    Fetch & {\cmark} & {\xmark}   & 0.970$\pm 0.011$ & 0.980$\pm 0.006$ & 0.975$\pm 0.006$ & 0.973$\pm 0.010$ & 0.973$\pm 0.011$ & 0.913$\pm 0.042$ & 0.975$\pm 0.009$ & 0.965$\pm 0.013$ & 0.905$\pm 0.048$\\
-    Fetch &{\cmark} & {\cmark} & 0.995$\pm 0.003$ & 0.998$\pm 0.002$ & 1.000$\pm 0.000$ & 0.998$\pm 0.002$ & 0.995$\pm 0.003$ & 0.963$\pm 0.020$ & 0.998$\pm 0.002$ & 0.995$\pm 0.003$ & 0.958$\pm 0.022$\\
+    Fetch & {\xmark}   & {\xmark}   & 0.99$\pm 0.01$ & 1.00$\pm 0.00$ & 1.00$\pm 0.00$ & 0.98$\pm 0.02$ & 1.00$\pm 0.00$ & 0.78$\pm 0.09$ & 0.99$\pm 0.01$ & 0.96$\pm 0.04$ & 0.58$\pm 0.14$\\
+    Fetch & {\xmark}   & {\cmark} & 0.88$\pm 0.09$ & 1.00$\pm 0.00$ & 1.00$\pm 0.00$ & 0.83$\pm 0.07$ & 0.98$\pm 0.01$ & 0.24$\pm 0.06$ & 0.98$\pm 0.01$ & 0.78$\pm 0.08$ & 0.09$\pm 0.05$\\
+    Fetch & {\cmark} & {\xmark}   & 0.97$\pm 0.01$ & 0.98$\pm 0.01$ & 0.98$\pm 0.01$ & 0.97$\pm 0.01$ & 0.97$\pm 0.01$ & 0.91$\pm 0.04$ & 0.98$\pm 0.01$ & 0.97$\pm 0.01$ & 0.91$\pm 0.05$\\
+    Fetch &{\cmark} & {\cmark} & 1.00$\pm 0.00$ & 1.00$\pm 0.00$ & 1.00$\pm 0.00$ & 1.00$\pm 0.00$ & 1.00$\pm 0.00$ & 0.96$\pm 0.02$ & 1.00$\pm 0.00$ & 1.00$\pm 0.00$ & 0.96$\pm 0.02$\\
     \hline
-    Jaco  & {\xmark}   & {\xmark}   & 0.281$\pm 0.067$ & 0.871$\pm 0.036$ & 0.994$\pm 0.003$ & 0.730$\pm 0.095$ & 0.984$\pm 0.008$ & 0.274$\pm 0.077$ & 0.407$\pm 0.098$ & 0.759$\pm 0.066$ & 0.027$\pm 0.006$\\
-    Jaco  & {\xmark}   & {\cmark} & 0.451$\pm 0.040$ & 0.941$\pm 0.032$ & 0.993$\pm 0.002$ & 0.706$\pm 0.064$ & 0.950$\pm 0.028$ & 0.258$\pm 0.044$ & 0.440$\pm 0.058$ & 0.784$\pm 0.044$ & 0.045$\pm 0.009$\\
-    Jaco  & {\cmark} & {\xmark}   & 0.640$\pm 0.046$ & 0.655$\pm 0.055$ & 0.651$\pm 0.055$ & 0.668$\pm 0.046$ & 0.648$\pm 0.046$ & 0.636$\pm 0.040$ & 0.642$\pm 0.035$ & 0.666$\pm 0.047$ & 0.426$\pm 0.054$\\
-    Jaco  & {\cmark} & {\cmark} & 0.987$\pm 0.005$ & 0.990$\pm 0.003$ & 0.987$\pm 0.003$ & 0.990$\pm 0.003$ & 0.990$\pm 0.003$ & 0.970$\pm 0.017$ & 0.970$\pm 0.011$ & 0.990$\pm 0.003$ & 0.441$\pm 0.128$\\
+    Jaco  & {\xmark}   & {\xmark} & 0.28$\pm 0.07$ & 0.87$\pm 0.04$ & 0.99$\pm 0.00$ & 0.73$\pm 0.10$ & 0.98$\pm 0.01$ & 0.27$\pm 0.08$ & 0.41$\pm 0.10$ & 0.76$\pm 0.07$ & 0.03$\pm 0.01$\\
+    Jaco  & {\xmark}   & {\cmark} & 0.45$\pm 0.04$ & 0.94$\pm 0.03$ & 0.99$\pm 0.00$ & 0.71$\pm 0.06$ & 0.95$\pm 0.03$ & 0.26$\pm 0.04$ & 0.44$\pm 0.06$ & 0.78$\pm 0.04$ & 0.05$\pm 0.01$\\
+    Jaco  & {\cmark} & {\xmark}   & 0.64$\pm 0.05$ & 0.66$\pm 0.06$ & 0.65$\pm 0.06$ & 0.67$\pm 0.05$ & 0.65$\pm 0.05$ & 0.64$\pm 0.04$ & 0.64$\pm 0.04$ & 0.67$\pm 0.05$ & 0.43$\pm 0.05$\\
+    Jaco  & {\cmark} & {\cmark} & 0.99$\pm 0.01$ & 0.99$\pm 0.00$ & 0.99$\pm 0.00$ & 0.99$\pm 0.00$ & 0.99$\pm 0.00$ & 0.97$\pm 0.02$ & 0.97$\pm 0.01$ & 0.99$\pm 0.00$ & 0.44$\pm 0.13$\\
     \bottomrule
   \end{tabular}}
-  \caption{Test performance of all models with local visual changes (distractors).
+  \caption[Test performance of all models with local visual changes.]{Test performance of all models with local visual changes (distractors).
 Checkmarks and crosses indicate enabling/disabling DR and proprioceptive inputs (Prop.), respectively.
@@
     \hline
-    Jaco  & {\xmark} & {\xmark}   & 0.772$\pm 0.058$ & 0.567$\pm 0.047$\\
-    Jaco  & {\xmark} & {\cmark} & 0.808$\pm 0.044$ & 0.507$\pm 0.047$\\
-    Jaco  & {\cmark} & {\xmark}   & 0.652$\pm 0.022$ & 0.639$\pm 0.051$\\
-    Jaco  & {\cmark} & {\cmark} & 0.989$\pm 0.002$ & 0.622$\pm 0.050$\\
+    Jaco  & {\xmark} & {\xmark}   & 0.772$\pm 0.058$ & 0.880$\pm 0.007$\\
+    Jaco  & {\xmark} & {\cmark} & 0.808$\pm 0.044$ & 0.897$\pm 0.007$\\
+    Jaco  & {\cmark} & {\xmark}   & 0.652$\pm 0.022$ & 0.759$\pm 0.004$\\
+    Jaco  & {\cmark} & {\cmark} & 0.989$\pm 0.002$ & 0.930$\pm 0.004$\\
     \bottomrule
   \end{tabular}
-  \caption{Test performance of a single model with distractors locations varying over 9 different on the ground plane (Jaco) and table (Fetch).
+  \caption[Test performance of a single model with the distractor presented in 9 different locations.]{Test performance of a single model with distractors locations varying over 9 different on the ground plane (Jaco) and table (Fetch).
 Checkmarks and crosses indicate enabling/disabling DR and proprioceptive inputs (Prop.), respectively.
@@
 Next, we discuss how global visual changes affect the performance of these trained agents.
-From Table~\ref{tbl:tests}, in this case, the DR agents show better robustness but still sacrifice a lot of performance in many test conditions.
-
-{In the Illumination (Lvl.) condition, the performance of all trained agents decreases as the light intensity reduces ({see Table~\ref{tbl:diff_levels}}).
-In the Fetch environment, the DR agents are more robust to variations of the illumination intensity.
-In the Jaco environment, however, the DR agent trained with visual inputs only is more robust in this case.
-This is probably because the agent trained with visual inputs must learn better features for self-localisation, thus the learned features are more robust to the change of the environment.}
+From Table~\ref{tbl:tests}, in this case, the agents trained with DR show better robustness but still sacrifice a lot of performance in many test conditions.
+
+In the Illumination (Lvl.) condition, the performance of all trained agents decreases as the light intensity reduces ({see Table~\ref{tbl:diff_levels}}).
+In the Fetch environment, the agents trained with DR are more robust to variations of the illumination intensity.
+In the Jaco environment, however, the agents trained without proprioception are more robust in this case, compared to agents trained with.
+This is probably because, in the Jaco environment, the agents trained without proprioception require a more complex visual system for the pose estimation, whereas a simpler visual system can be disrupted by a reduction in contrast or even just a change in the pixel value of the target.
+Thus the learned features are more robust to the change of the illumination in the environment.
+%must learn better features for pose estimation, thus the learned features are more robust to the change of the illumination in the environment.}
 
 In the Illumination (Dir.) condition, the performance of all agents decreases as the direction of the light source changes (see Table~\ref{tbl:diff_dirs}).
-Similarly, in the Fetch environment, the DR agents are more robust.
-In the Jaco environment, compared to training with DR, proprioceptive inputs would be more helpful for robustness.
+Similarly, in the Fetch environment, the agents trained with DR are more robust.
+However, in the Jaco environment, compared to training with DR, providing proprioceptive inputs would be more helpful in this case.
 
 \begin{table}[H]
+\centering
   \resizebox{0.9\linewidth}{!}{
@@
   \end{tabular}}
-  \caption{Test performance of all models with different global illumination intensities; illumination is specificied as Fetch/Jaco.
+  \caption[Test performance of all models with different global illumination intensities.]{Test performance of all models with different global illumination intensities; illumination is specificied as Fetch/Jaco.
 Checkmarks and crosses indicate enabling/disabling DR and proprioceptive inputs (Prop.), respectively.
@@
 \begin{table}[H]
+ \centering
   \resizebox{0.9\linewidth}{!}{
@@
   \end{tabular}}
-  \caption{Test performance of all models with different main illumination directions; direction is specified as Fetch/Jaco.
+  \caption[Test performance of all models with different main illumination directions.]{Test performance of all models with different main illumination directions; direction is specified as Fetch/Jaco.
 Checkmarks and crosses indicate enabling/disabling DR and proprioceptive inputs (Prop.), respectively.
@@
 For the case of additive Gaussian noise, the agents trained in the Fetch environment are all robust to the noise.
-However, in the Jaco environment, the non-DR agents have approximately 30$\%$ performance drop, while the DR agents only have 10$\%$.
-
-In the reflection condition, the agents trained with DR are more robust.
-However, in the Fetch environment, it decreases the performance of the non-DR agents by 50$\%$.
-While in the Jaco environment, there is not a significant influence on the performance of the non-DR agents.
-One possible explanation is the difference in the sizes of the robot arm.
-The Jaco robot arm is relatively small in the simulation, so its reflection has little change on the visual inputs.
-
-In the translation condition, the performance of all agents reduces heavily (see Table~\ref{tbl:diff_trans}).
-In the Fetch environment, just as before, DR agents exhibit better robustness.
-In the Jaco environment, non-DR and DR agents have similar performance, which reflects that these agents are able to learn the extent of translation invariance for the policies.
+However, in the Jaco environment, the agents trained without DR have approximately 30$\%$ performance drop, while the agents with DR only have 10$\%$ performance drop.
+Although both environments share the same training configurations, either the difficulty of the task or visual layout can make the Fetch agents be more robust to noise than Jaco agents.
+
+In the reflection condition, the agents trained with DR are generally robust.
+In the Fetch environment, the performance of the agents trained without DR is decreased by 50$\%$.
+In the Jaco environment, it only causes a minor influence on the performance of the agents trained without DR.
+One possible explanation is the difference in the sizes of the robot arms.
+The Jaco robot arm is relatively small in the simulation, so its reflection has insignificant changes in the visual inputs.
+When the proprioceptive inputs are provided, both Jaco and Fetch agents trained with DR demonstrate better robustness to the changes.
+
+In the translation condition, the performance of all agents is greatly reduced.
+In the Fetch environment, just as before, agents trained with DR exhibit better robustness.
+In the Jaco environment, all agents have similar performance, which reflects that these agents are able to learn the extent of translation invariance for the policies.
 This could be a benefit for the target reaching task in a 3D space, which leads to a more generalisable spatial representation.
+Performance drops monotonically with deviation from the original location (see Table~\ref{tbl:diff_trans}).
+The drop is symmetric for the Fetch agents, but asymmetric for the Jaco agents -- a relative shift in the target towards the centre of the image input is better than a shift away.
 
 \begin{table}[H]
+ \centering
   \resizebox{\linewidth}{!}{
@@
   \end{tabular}}
-  \caption{Test performance of all models with different levels of camera translation.
+  \caption[Test performance of all models with different levels of camera translation.]{Test performance of all models with different levels of camera translation.
 Checkmarks and crosses indicate enabling/disabling DR and proprioceptive inputs (Prop.), respectively.
@@
 
-Finally, in the case of the arm being invisible condition, the performance of almost all agents is reduced to zero, except for the Jaco DR agent trained with proprioceptive inputs.
-For those failure cases, the agents need to infer the position of the robot arm visually.
-For the Jaco DR agent trained with proprioceptive inputs, the position of the robot arm is estimated by using only proprioceptive inputs.
+Finally, in the case of the arm being invisible, the performance of almost all agents is reduced to zero, except for the Jaco agent trained with proprioception and DR.
+For those failure cases, the agents need to infer the pose/position of the robot arm visually.
+For the Jaco agent trained with proprioception and DR, the pose/position of the robot arm is estimated by using proprioceptive inputs \textit{only}.
 
@@
 \subsection{Saliency Maps}
-We first employ the saliency map technique to infer which part of the visual input will affect the output of the policy network.
-Then, our proposed occlusion-based saliency map is used to investigate the trained agents in standard condition, colour change condition (yellow), and shape change condition (cube), respectively.
+We first employ our proposed saliency map technique (using the average baseline image) to infer which part of the visual input will determine the output of the policy network.
+Then, the generated saliency maps are used to investigate the trained agents in standard condition, colour change condition (yellow), and shape change condition (cube), respectively.
 
 Figure~\ref{fig:saliency_fetch_distractor} demonstrates the saliency maps of the trained agents in the Fetch environment.
-From the results, DR agents in Figure~\ref{fig:saliency_fetch_distractor}j-l can self-localise via proprioceptive inputs.
-The others need to use visual inputs for self-localisation, where the saliency maps display highlighted regions in the robot arm.
-In Figure~\ref{fig:saliency_fetch_distractor}d-f, although the proprioceptive inputs have been provided to non-DR agents, these agents still use the visual inputs for self-localisation, which means not all inputs will be utilised efficiently.
-In addition, the saliency maps of the non-DR agents highlight more regions on the distractors (see Figure~\ref{fig:saliency_fetch_distractor}a-f), while the saliency maps of the DR agents do not (see Figure~\ref{fig:saliency_fetch_distractor}g-k), except for the case in Figure~\ref{fig:saliency_fetch_distractor}l.
+From the results, agents trained with DR in Figure~\ref{fig:saliency_fetch_distractor}j-l can self-localise via proprioceptive inputs.
+The others need to use visual inputs for pose estimation, where the saliency maps display highlighted regions in the gripper of the robot arm.
+In Figure~\ref{fig:saliency_fetch_distractor}d-f, although the proprioceptive inputs have been provided to agents trained without DR, these agents still use the visual inputs for pose estimation, which means not all inputs will be utilised efficiently.
+In addition, the saliency maps of the agents trained without DR highlight more regions on the distractors (see Figure~\ref{fig:saliency_fetch_distractor}a-f), while the saliency maps of the agents trained with DR do not (see Figure~\ref{fig:saliency_fetch_distractor}g-k), except for the case in Figure~\ref{fig:saliency_fetch_distractor}l.
 
 Figure~\ref{fig:saliency_jaco_distractor} shows the saliency maps of the trained agents in the Jaco environment, which are more in line with our expectations.
-When the proprioceptive inputs are not available, the saliency maps have more highlighted regions on the robot arm for self-localisation.
+When the proprioceptive inputs are not available, the saliency maps have more highlighted regions on the robot arm for pose estimation.
 Conversely, the saliency maps only focus on the target object.
-Furthermore, from Figure~\ref{fig:saliency_jaco_distractor}b,c,e,f, the non-DR agents also pay more attention to the distractors.
-A close observation reveals that there exist weak highlighted regions around distractors in Figure~\ref{fig:saliency_jaco_distractor}e,f and robot arm in  Figure~\ref{fig:saliency_jaco_distractor}g-i, respectively.
+Furthermore, from Figure~\ref{fig:saliency_jaco_distractor}b,c,e,f, the agents trained without DR also pay more attention to the distractors.
+A close observation reveals that there exist weak highlighted regions around the distractors in Figure~\ref{fig:saliency_jaco_distractor}e,f and robot arm in Figure~\ref{fig:saliency_jaco_distractor}g-i, respectively.
 
 While these saliency maps can provide some intuitive explanations, they do not completely replace the quantitative results in Table~\ref{tbl:diff_colours}.
-For example, the Fetch DR agent trained with proprioception inputs pays attention to the distractors.
-However, it has less performance drop than the Fetch DR agent trained with visual inputs only, whose saliency map only has highlighted regions on the target ball.
+For example, the Fetch agent trained with proprioception inputs and DR pays attention to the distractors.
+However, it has less performance drop than the Fetch agent trained with visual inputs only and DR, whose saliency map only has highlighted regions on the target ball.
 Therefore, we cannot rely solely on saliency maps to interpret and analyse trained agents.
@@
   \end{subfigure}
-  \caption{Occlusion-based saliency maps of Fetch agents trained with (g-l) or without (a-f) DR and with (d-f, j-l) or without proprioceptive inputs (a-c, g-i) in three different test conditions (standard, colour and shape).
+  \caption[Occlusion-based saliency maps of Fetch agents.]{Occlusion-based saliency maps of Fetch agents trained with (g-l) or without (a-f) DR and with (d-f, j-l) or without proprioceptive inputs (a-c, g-i) in three different test conditions (standard, colour and shape).
 The best Fetch model is used for each training configuration.}
@@
   \end{subfigure}
-  \caption{Occlusion-based saliency maps of Jaco models trained with (g-l) or without (a-f) DR and with (d-f, j-l) or without proprioceptive inputs (a-c, g-i) in three different test conditions (standard, colour and shape).
+  \caption[[Occlusion-based saliency maps of Jaco agents.]]{Occlusion-based saliency maps of Jaco agents trained with (g-l) or without (a-f) DR and with (d-f, j-l) or without proprioceptive inputs (a-c, g-i) in three different test conditions (standard, colour and shape).
 The best Jaco model is used for each training configuration.}
@@
 %\subsection{Statistical and Structural Weight Characterisations}
-\subsection{Unit Ablations}
+\subsection{Channel-wise Unit Ablations}
 
@@
   \end{subfigure}
-  \caption{The performance of unit ablation tests of the trained Fetch and Jaco agents under 8 different training configurations in standard and visual noise conditions.
+  \caption[Test performance of unit ablation tests of the best models under 8 training configurations in standard and noise conditions.]{The performance of unit ablation tests of the trained Fetch and Jaco agents under 8 different training configurations in standard and visual noise conditions.
 Each point represents one unit in layer 1 (a, b) or layer 2 (c, d).
@@
 Finally, we repeat the above process until all output channels have been tested.
-The results of the unit ablation tests are illustrated in Figure~\ref{fig:unit_ablations}
+The results of the unit ablation tests are illustrated in Figure~\ref{fig:unit_ablations}.
 
@@
 However, the effect on the Jaco agents is relatively large, as the performance of the Jaco agents varies significantly.
-This is probably because of the increased complexity of the Jaco task (e.g., control scheme) and which requires more powerful channels to extract relevant information to train the agents.
+This is probably because of the increased complexity of the Jaco task (e.g., control scheme), which requires more powerful channels to extract relevant information to train the agents.
 
 % We further find that the unit ablations can improve the performance in the noise condition, even for the DR agents. \td{This may}
-
 Furthermore, in the first convolutional layer (see Figure~\ref{fig:unit_ablations}a,b), there exist several extremely critical channels/units that can substantially affect the performance of the agent.
 However, there are no such channels in the second convolutional layer (see Figure~\ref{fig:unit_ablations}c,d).
+
+Next, we find that the variability in the noisy environment (see Figure~\ref{fig:unit_ablations}b,d) is greater.
+Intriguingly, ablations can improve performance beyond the baseline results, even for agents trained with DR---perhaps indicating a sensitivity to high-frequency noise.
+Although a detailed discussion is beyond the scope of this work, we note that research on corruption and adversarial robustness in supervised learning settings could provide further insights on properties such as these~\cite{gilmer2019adversarial, hendrycks2019benchmarking}.
 
@@
 Generally, after a period of training, the LSTM layer and policy layer are more robust to be re-initialised with the initial weights.
-For the DR agents, they show less robustness to re-initialisation from the early to middle stages of training, which indicates DR agents require longer periods to learn useful representations from tasks.
+For the agents trained with DR, they show less robustness to re-initialisation from the early to middle stages of training, which indicates that the agents trained with DR require longer periods to learn useful representations from tasks.
 
@@
 This is probably because the fc1 layer has to learn a sophisticated combination between the visual inputs and proprioceptive inputs.
-This is especially the case for the Jaco DR agent, which can use only proprioceptive inputs for self-localisation: it requires an even longer time to train the fc1 layer than other agents (see Figure~\ref{fig:jaco_robustness}j).
+This is especially the case for the Jaco agent trained with DR, which can use only proprioceptive inputs for pose estimation: it requires an even longer time to train the fc1 layer than other agents (see Figure~\ref{fig:jaco_robustness}j).
 
 Finally, in almost all cases, the LSTM layer is robust to be re-initialised with the initial weights.
-Therefore, we suppose that the LSTM layer is redundant in the network structure for these tasks.
+One possible explanation is, the LSTM layer learns a trivial mapping within the network.
 Further experiments to verify this assumption are provided in the following subsection.
+%Therefore, we suppose that the LSTM layer is redundant in the network structure for these tasks.
 
@@
   \end{subfigure}
-  \caption{Performance of re-initialisation robustness (0 for 100$\%$ success rate, and 1 for 0$\%$ success rate), and change of parameters in $\ell_\infty$- and $\ell_2$-norm of Fetch agents.
+  \caption[Test performance of re-initialisation robustness and change of parameters of Fetch agents.]{Performance of re-initialisation robustness (0 for 100$\%$ success rate, and 1 for 0$\%$ success rate), and change of parameters in $\ell_\infty$- and $\ell_2$-norm of Fetch agents.
 The best Fetch model is selected for each training configuration.}
@@
   \end{subfigure}
-  \caption{Performance of re-initialisation robustness (0 for 100$\%$ success rate, and 1 for 0$\%$ success rate), and change of parameters in $\ell_\infty$- and $\ell_2$-norm of Jaco agents.
+  \caption[Test performance of re-initialisation robustness and change of parameters of Jaco agents.]{Performance of re-initialisation robustness (0 for 100$\%$ success rate, and 1 for 0$\%$ success rate), and change of parameters in $\ell_\infty$- and $\ell_2$-norm of Jaco agents.
 The best Fetch model is selected for each training configuration.}
@@
 \subsection{Recurrent Ablation}
+\label{sec:detail_reucrrent_ablation}
 We then evaluate the effectiveness of the LSTM layer for the training.
-In this case, the hidden states and cell states are set to constant values for the evaluation.
-From Table~\ref{tbl:hidden_ablation}, the LSTM layer with constant values only has a minor influence on non-DR agents, while it has a significant influence on the DR agents.
-This suggests that the LSTM layer is functional when DR is activated during training, otherwise it is unnecessary to learn a mapping from observations to policy.
+In this case, the \textit{hidden} and \textit{cell} states are set to constant values for the evaluation.
+Instead of simply setting the hidden and cell states to zero -- as this may not represent the normal value in the rollouts, at each timestep -- we set the hidden and cell states to empirical average values which are calculated over the generated values of the models in testing (80 episodes for the Fetch agents and 250 episodes for the Jaco agents).
+From Table~\ref{tbl:hidden_ablation}, the LSTM layer with constant values only has a minor influence on agents trained without DR, while it has a significant influence on the agents trained with DR.
+This suggests that the LSTM layer is functional when DR is activated during training, otherwise it only learns a trivial mapping within the network.
+One possible explanation is that when DR is not activated during training, the textures and colours of the background in the environment remain unchanged, the agent only needs to learn the policy that keeps moving towards the target object in each frame.
+Therefore, in this case, the sequence information is not that crucial, and the LSTM layer does not play an important role.
+When DR is activated during training, the textures and colours of the background in the environment change in every frame.
+When the textures or colours of the background are similar to the target object, the agent needs to use the sequence information to remember the location of the target object in the previous frames to distinguish the target from irrelevant information and complete the task.
+Therefore, when DR is activated during training, the LSTM layer needs to capture the location of the target object in previous frames.
+In terms of the strategies \emph{learned} by the agents (Figure \ref{fig:strategies}), memory is sufficient for the agents trained without DR, but necessary for the agents trained with DR.
 
@@
   \end{tabular}
-  \caption{Performance of recurrent ablations of all models with standard operation and constant (empirical average) hidden states.
+  \caption[Test performance of recurrent ablations of all models.]{Performance of recurrent ablations of all models with standard operation and constant (empirical average) hidden states.
 Checkmarks and crosses indicate activate/deactivate DR and proprioceptive inputs (Prop.), respectively.
@@
 In this experiment, we construct 8 different training configurations, which cover different morphologies and control schemes of the robot arm (Fetch and Jaco robot arms), different modalities of inputs (visual inputs or visual + proprioceptive inputs), and using or not using DR.
-We also setup 8 different test conditions to evaluate the agents' robustness to local and global visual changes and the ability of visual self-localisation.
+We also setup 8 different test conditions to evaluate the agents' robustness to local and global visual changes and the ability of visual pose estimation.
 Finally, various interpretability techniques are adopted to provide both quantitative and qualitative analysis of the agents.
@@
 From the results, the agent trained with DR shows better generalisation ability to local visual changes, such as distractors with different colours and shapes.
-While DR agents also present a degree of robustness to global visual changes, such as illumination (direction and intensity), RGB camera translation, additive noise, and reflective surfaces, they still show significant performance degradation in several scenarios.
-The above failure cases can remind us to consider more factors in the DR design, where we only focus on visual DR in this chapter (e.g., texture, colours and noise).
-From saliency maps, the non-DR agents show more highlighted regions over the distractors, and DR agents focus more attention on the task related objects (e.g., robot arm and target ball).
-
-\td{In unit ablations, due to the task complexity and the control scheme of the Jaco environment, the Jaco agents require more channels for learning the representations to solve the task.} In layer re-initialisation, DR leads to distinguishable changes in $\ell_{\infty}$-norm of parameters in the convolutional layers and therefore acquires more robust features.
-It also shows that, when the proprioceptive inputs are available, the fc1 layer of the Jaco DR agent requires longer periods for training,  because it needs to learn a better representation from proprioceptive inputs for self-localisation.
-This trend is also in line with the results in the test of visual self-localisation, where this agent can still perform the task without the rendered robot arm.
+While agents trained with DR also present a degree of robustness to global visual changes, such as illumination (direction and intensity), RGB camera translation, additive noise, and reflective surfaces, they still show significant performance degradation in several scenarios.
+The above failure cases can remind us to consider more factors in the design of the algorithms/settings to perform domain randomisation, where we only focus on visual DR in this chapter (e.g., texture, colours, and noise).
+From saliency maps, the agents trained without DR show more highlighted regions over the distractors, and the agents trained with DR focus more attention on the task-related objects (e.g., robot arm and target ball).
+
+In unit ablations, due to the task complexity and the control scheme of the Jaco environment, the Jaco agents require more channels for learning the representations to solve the task (see  Figure~\ref{fig:unit_ablations}).
+In layer re-initialisation, DR leads to distinguishable changes in $\ell_{\infty}$-norm of parameters in the convolutional layers and therefore acquires more robust features.
+Figure~\ref{fig:jaco_robustness} also shows that, when the proprioceptive inputs are available, the fc1 layer of the Jaco agent trained with DR requires longer periods for training,  because it needs to learn a better representation from proprioceptive inputs for pose estimation.
+This trend is also in line with the results in the test condition of invisible robot arms, where this agent can still perform the task without the rendered robot arm.
 In this case, the agent only utilises proprioceptive inputs to infer the location of the robot arm.
@@
 
-Finally, some recommendations are also provided to those who intend to investigate DRL agents.
+Finally, some recommendations can also be provided as a result of the experiments performed in this chapter.
+%are also provided to those who intend to investigate DRL agents.
 \begin{itemize}
 \item[1)] Do not assume that results generalise. While the Jaco agent trained with DR and without proprioception obeys some trends, it remains an outlier in many other aspects.
-\item[2)] Do not expect clear results.
+\item[2)] Do not expect to obtain distinct trends in the results of an interpretability method in response to apparently clear differences between two environments.
 While some methods, such as the saliency map, uncovered clear trends, others, such as unit ablations, did not.
-This does not mean that unit ablations are generally useless---one can imagine that if several units were highly important without DR and no units are individually important with DR, we would have found a significant trend using this method.
-\item[3)] If possible, use a range of interpretability methods. Uncertainty about results from one method may be resolved by results from another method.
-\item[4)] No single interpretability method is more useful than another---each type of method reveals different, complementary pieces of information.
-\item[5)] Finally, use interpretability methods before making claims about the strategies used by agents.
+If possible, use a range of interpretability methods.
+Uncertainty about results from one method may be resolved by results from another method.
+%This does not mean that unit ablations are generally useless---one can imagine that if several units were highly important without DR and no units are individually important with DR, we would have found a significant trend using this method.
+\item[3)] Finally, use interpretability methods before making claims about the strategies used by agents.
 While one may assume that an agent that performs well at a task on average is using an intelligent strategy, in all likelihood, it may be using heuristics that fail to generalise \cite{ruderman2019uncovering}.

### 2022-08-21 chapters/old.tex
update the final version

\chapter{DRL Interpretability}
\label{ch:drl_interp}
\section{Introduction}
Deep reinforcement learning (DRL) is currently one of the most prominent subfields in AI, with applications to many domains ~\cite{arulkumaran2017deep}. One of the most enticing possibilities that DRL affords is the ability to train robots to perform complex tasks in the real world, all from raw sensory inputs. For instance, while robotics has traditionally relied on hand-crafted pipelines, each performing well-defined estimation tasks---such as ground-plane estimation, object detection, segmentation and classification, ~\cite{kragic2009vision,martinez2014taxonomy}---it is now possible to
learn visual perception and control in an ``end-to-end'' fashion~\cite{gu2017deep,levine2016end,zhu2017target}, without explicit specification and training of networks for specific sub-tasks.

A major advantage of using reinforcement learning (RL) versus the more traditional approach to robotic system design based on optimal control is that the latter requires a transition model for the task in order to solve for the optimal sequence of actions. While optimal control, when applicable, is more efficient, modelling certain classes of objects (e.g., deformable objects) can require expensive simulation steps, and often physical parameters (e.g., frictional coefficients) of real objects that are not known in detail. Instead, approaches that use RL can learn a direct mapping from observations to the optimal sequence of actions, purely through interacting with the environment. Through the powerful function approximation capabilities of neural networks (NNs), deep learning (DL) has allowed RL algorithms to scale to domains with significantly more complex input and action spaces than previously considered tractable. Such domains, which could involve inferring actions based on image inputs, make theoretical analysis of both the environment and agents that act within it practically infeasible.

In particular, NNs trained with DRL form a ``black box'' mapping from observations directly to actions, leaving us with the challenge of understanding the learned control policies. If we would like to deploy such policies---particularly on robots in the real world---we would also want to be able to \emph{explain} them~\cite{arrieta2020explainable}. An interpretation of a model's ``reasoning'' can not only be used as a way to provide an explanation of the model's behaviour, but can also be used to characterise other properties, such as safety, fairness and reliability ~\cite{doshi2017towards}. While there are now many methods available to interpret NNs~\cite{guidotti2018survey}, these methods are subjective to varying degrees. As such, we train DRL agents under a range of different settings, and use relative differences to better interpret their policies. {In contrast to most prior work examining DRL policies}~\cite{atrey2020exploratory, greydanus2018visualizing, olson2021counterfactual, puri2020explain, rupprecht2020finding, such2018atari, zahavy2016graying}, {we not only provide explanations for how policies react to parts of states/individual states, but use our array of experiments (Section \ref{sec:experiments}) to characterise their \emph{global} strategies (Figure \ref{fig:strategies}). For instance, we examine the relative importance of different input modalities, and whether agents need their memory to complete tasks (Section \ref{sec:model_analysis}).}

\begin{figure*}
  \centering
  \includegraphics[width=\linewidth]{figures/chapter6/strategies.png}
  \caption{The range of strategies an agent may learn to use when trained to reach a red target, which we split into three components. Firstly, the agent must visually localise the target---which can be accomplished by detecting red, spherical objects, or simply anything with a significant red component. Secondly, the agent can use vision and/or proprioception to guide its arm to the target. Finally, the agent may accomplish the task with varying levels of robustness to changes in the rest of the visual scene. Using a range of analyses, we show that agents trained to accomplish the same task learn different subsets of these strategies, depending on their training conditions.}
  \label{fig:strategies}
\end{figure*}

Specifically, we train different robots with vision-based inputs to perform the task of reaching for a (red) target, and then use different techniques (Subsection \ref{sec:nn_interpretability}) to understand what features of the environment the policies react to. Under certain training settings, the agent may localise the target by simply looking for anything red, and it is only under more challenging training conditions that the agent utilises both colour and shape. Clearly, the latter strategy better represents what we want the agent to learn, while the former is merely a ``shortcut'' \cite{geirhos2020shortcut}.

\hypertarget{training-conditions}{%
\subsection{Training Conditions}\label{training-conditions}}
We vary training conditions over 3 different axes: 2 different robots with different morphologies and control schemes, using vision-only vs.~vision + proprioception, and training with/without visual \emph{domain randomisation} (DR). This results in 8 different configurations (Figure \ref{fig:axes}), which we found sufficient to uncover a large compositional space of strategies (Figure
\ref{fig:strategies}).

\begin{figure}
  \centering
  \includegraphics[width=0.85\linewidth]{figures/chapter6/axes.png}
  \caption{The 8 different experimental conditions: 2 robots (Fetch vs. Jaco), 2 input modalities (vision vs. vision + proprioception), 2 visual conditions (DR vs. no DR).}
  \label{fig:axes}
\end{figure}

We chose these 3 axes for the purpose of studying \emph{representation learning} in robotics tasks. The difference in the design of the robots, as well as a 2D vs.~3D target space (Subsection \ref{sec:environments}), results in one task being more difficult---in terms of both inferring the position of the target, as well as in the control. While task complexity is not a transitive property, the representations learned do reflect the difference between these tasks: for instance, localisation of the target in the latter case requires more complex image processing (Subsection \ref{sec:activation_maximisation}).

For real-world deployment, one would want to use as many sensors as it is feasible to use. However, we are interested in how the absence of certain input modalities can affect learning, as this changes how the agent is able to extract information from the environment. While one might expect robots equipped with proprioceptive sensors to use these alone for pose estimation, agents can also learn to utilise vision if extracting pose from images is simple (Subsection \ref{sec:saliency_maps}).

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{figures/chapter6/domain_random.png}
  \caption{Examples of visual domain randomisation in our Fetch (top) and Jaco (bottom) robotics experiments.}
  \label{fig:dr_example}
\end{figure}
\setlength{\textfloatsep}{5pt}

Finally, analogously to how \emph{data augmentation}~\cite{lecun1998gradient,shorten2019survey} can be used to improve the generalisation of models in supervised learning settings, \emph{domain randomisation} (Figure \ref{fig:dr_example}) is a common training paradigm in ``sim2real'' approaches for learning real-world robotics control~\cite{andrychowicz2018learning,james2017transferring,peng2018sim,sadeghi2017cad2rl,tobin2017domain}. In DR, various properties of the simulation are varied, altering anything from the positions or dynamical properties of objects to their visual appearance, resulting in an expansion of the training domain. We therefore test how the agents trained with or without DR differ. Supporting previous results, we show that agents trained with DR are more robust to out-of-distribution (OoD) perturbations (Subsection \ref{sec:test_scenarios}). In our experiments, we use a standard visual DR setup, which is described in Table \ref{tbl:random_summary} and visualised in Figure \ref{fig:random_vis}.

\begin{figure}
  \centering
  \begin{subfigure}{0.32\columnwidth}
    \includegraphics[width=\linewidth]{figures/chapter6/rand_checker.png}
    \subcaption{checkerboard}
  \end{subfigure}
  \begin{subfigure}{0.32\columnwidth}
    \includegraphics[width=\linewidth]{figures/chapter6/rand_gradient.png}
    \subcaption{gradient}
  \end{subfigure}

  \begin{subfigure}{0.32\columnwidth}
    \includegraphics[width=\linewidth]{figures/chapter6/rand_rgb.png}
    \subcaption{flat\_rgb}
  \end{subfigure}
  \begin{subfigure}{0.32\columnwidth}
    \includegraphics[width=\linewidth]{figures/chapter6/rand_noise.png}
    \subcaption{noise}
  \end{subfigure}
  \caption{Visualisation of different DR textures used during training: (a) checkerboard, (b) gradient, (c) flat\_rgb, (d) noise.}
  \label{fig:random_vis}
\end{figure}

DRL agents trained in standard simulators (due to the sample inefficiency of DRL algorithms) do not directly transfer to the real world. However, agents trained in simulators with DR can, which makes them of particular interest to study. It is common knowledge that while training in simulation is simpler, differences between the simulated and real worlds introduces a \emph{reality gap}~\cite{jakobi1995noise}. Of the several ways to address this gap---including fine-tuning a DRL agent on the real world~\cite{rusu2017sim}, performing system identification to reduce the domain gap~\cite{chebotar2018closing}, or explicitly performing domain adaptation~\cite{tzeng2015towards}---DR is unique in that it is essentially a form of data augmentation that only affects the input data, without changing the initial model or training objective. Thus, by comparing agents trained with DR against agents trained without, we expect to uncover what \emph{strategies} they learn that would allow generalisation in the real world (as opposed to, adaptation to the real world). For example, some of the robustness of our DR agents comes through learning to perform some form of state estimation implicitly (Subsection \ref{sec:recurrent_ablation}).

\begin{table}
  \label{tbl:random_summary}
  \centering
  \resizebox{\linewidth}{!}{
  \begin{tabular}{c|c|c}
    \toprule
    No. & Option & Description \\
    \midrule
   1 & checkerboard & Render checkerboard pattern with two random RGB colours\\
   2 & gradient & Render gradient in vertical/horizontal direction with two random RGB colours\\
   3 & flat\_rgb & Render single random RGB colour\\
   4 & noise & Render artificial noise with two random RGB colours\\
    \bottomrule
  \end{tabular}}
   \caption{DR textures used during training. At every environment timestep, for each component (e.g., robotic arm, skybox, table; the Jaco robot has 13 components, and the Fetch robot has 19 components), the appearance of the component is rendered using a sequential sampling process. In the first step of the sampling process, one of each of the 4 possible choices is made with the same probability, then the subsequent appearance of each component is determined by a second draw, described in the rightmost column of the table. Each RGB channel is drawn uniformly from $\{0 \ldots 255\}$. The details of each option are described in Table~\ref{tbl:texture_expression}.}
\end{table}

\begin{table}
  \label{tbl:texture_expression}
  \centering
  \resizebox{\linewidth}{!}{
  \begin{tabular}{c|c|c}
    \toprule
    No. & Option & Expression \\
    \midrule
   1 & checkerboard & texture\_mat[x, y] = rgb1 \textbf{if} (x + y) $\%$ 2 == 0 \textbf{else} rgb2\\
   2 & gradient & texture\_mat[x, y] = (1 - p) $\cdot$ rgb1 + p $\cdot$ rgb2, vertical: p = y / (height - 1), horizontal: p = x / (width - 1)\\
   3 & flat\_rgb & texture\_mat[x, y] = rgb1\\
   4 & noise & texture\_mat[x, y] = rgb1 \textbf{if} p > 0.9 \textbf{else} rgb2, p $\sim$ U(0, 1)\\
    \bottomrule
  \end{tabular}
  }
\caption{Python pseudocode describing the DR texture generators. rgb1 and rgb2 are random RGB colours; the value of each RGB channel is drawn uniformly from $\{0 \ldots 255\}$.}
\end{table}

\hypertarget{generalisation}{%
\subsection{Generalisation}\label{generalisation}}

The study of how DRL agents generalise has received considerable
interest in the DRL community recently
\cite{cobbe2018quantifying, justesen2018procedural, packer2018assessing, witty2018measuring, zhang2018dissection, zhao2019investigating}. As RL agents are typically trained and evaluated on the same environment, generalisation requires changes to the traditional paradigm. In particular, works in this area have also focused on procedural content generation and OoD tests to evaluate generalisation. In this paper, we not only quantitatively evaluate generalisation, but also use a wide suite of interpretability methods to try and understand why our trained agents act the way they do. Furthermore, we are the first to perform an in-depth study focused on DR. One of our findings is that, depending on the training conditions, we can observe a failure of agents trained with DR to generalise to the much simpler default visuals of the simulator (Subsection \ref{sec:domain_shift}), highlighting that conditions other than DR can have a significant effect on generalisation.

\hypertarget{interpretability}{%
\subsection{Interpretability}\label{interpretability}}

While our OoD tests (Subsection \ref{sec:test_scenarios}) provide a quantitative measure by which we can probe the performance of trained agents under various conditions, they treat the trained agents as black boxes. However, with full access to the internals of the trained models and even control over the training process, we can delve even further into the models. Using common interpretability tools such as saliency maps~\cite{morch1995visualization, selvaraju2017grad, sundararajan2017axiomatic, zeiler2014visualizing}
and dimensionality reduction methods
~\cite{maaten2008visualizing, mcinnes2018umap, pearson1901liii} for
visualising NN activations~\cite{rauber2017visualizing}, we can obtain
information on why agents act the way they do. The results of these
methods work in tandem with our OoD tests, as matching performance to
the qualitative results allows us to have greater confidence in
interpreting the latter; in fact, this process allowed us to debug and
improve upon an existing saliency map method, as detailed in Subsection
\ref{sec:back_saliency_maps}.
{Similarly to prior work}\footnote{Based on interpreting
  standard DRL agents, as opposed to constructing more interpretable
  agents.} on interpretability in DRL agents~\cite{atrey2020exploratory, greydanus2018visualizing, jaunet2020drlviz, meyes2020you, olson2021counterfactual, puri2020explain, rupprecht2020finding, such2018atari, zahavy2016graying}
{we use common methods: saliency maps (Subsection \ref{sec:back_saliency_maps}), activation maximisation (Subsection \ref{sec:act_max}), unit ablations (Subsection \ref{sec:unit_abl}) and dimensionality reduction (Subsection \ref{sec:act_an}).}
In addition, we use layer re-initialisation (Subsection
\ref{sec:layer_abl}), recurrent ablations (Subsection
\ref{sec:recurrent_abl}), entanglement (Subsection \ref{sec:act_an}),
and a novel quantitative measure for spatial structure in convolutional
filters (Subsection \ref{sec:spectral}).

\hypertarget{summary-of-findings}{%
\subsection{Summary of Findings}\label{summary-of-findings}}

Under our set of experimental conditions, we show that our agents
trained with DR:

\begin{itemize}
\item require more representational learning capacity (Subsection \ref{sec:layer_ablations}),
\item are more robust to visual changes in the scene, exhibiting generalisation to unseen local/global perturbations (Subsection \ref{sec:test_scenarios}),
\item use a smaller set of more reliable visual cues when not provided proprioceptive inputs (Subsection \ref{sec:saliency_maps}),
\item more heavily rely on recurrent processing (Subsection \ref{sec:recurrent_ablation}),
\item have filters that have higher norms or greater spatial structure (Subsection \ref{sec:statistical_and_structural}), which respond to more complex spatial patterns (Subsection \ref{sec:activation_maximisation}),
\item learn more robust (Subsection \ref{sec:unit_ablations}) and \textit{entangled} \cite{frosst2019analyzing} representations (Subsection \ref{sec:activation_analysis}),
\item and can ``overfit'' to DR visuals (Subsection \ref{sec:domain_shift}).
\end{itemize}

Furthermore, we characterise what strategies (Figure
\ref{fig:strategies}) agents learn as a result of being trained under
different conditions. In our discussion (Section \ref{sec:discussion}),
we revisit what this entails for future work on sim2real methods, and
end with a set of recommendations for those interesting in applying
interpretability methods to DRL agents.

\hypertarget{overview}{%
\subsection{Overview}\label{overview}}

The rest of the paper is organised as the follows: In Section
\ref{sec:methods}, we introduce RL, and the neural network
interpretability techniques used in this work. In Section
\ref{sec:experiments}, we detail our simulated environments, DRL agent
network structure, training details, and test scenarios. In Section
\ref{sec:model_analysis}, we conduct an exhaustive analysis on the
trained models using a broad suite of interpretability techniques.
Finally, in Section \ref{sec:discussion}, we discuss our findings, and
provide recommendations for applying interpretability techniques to DRL
agents. The code used in this paper is available at:
\url{https://github.com/TianhongDai/domain-rand-interp}.
% method
\hypertarget{methods}{%
\section{Methods}\label{methods}}
\label{sec:methods}

\hypertarget{neural-network-interpretability}{%
\subsection{Neural Network
Interpretability}\label{neural-network-interpretability}}

\label{sec:nn_interpretability}

The recent success of machine learning (ML) methods has led to a renewed
interest in trying to interpret trained models. In this work, we are
primarily concerned with scientific understanding, but our
considerations are grounded in other properties necessary for eventual
real-world deployment, such as robustness to OoD inputs.

The challenge that we face is that, unlike other ML algorithms that are
considered interpretable by design (such as decision trees or nearest
neighbours \cite{freitas2014comprehensible}), standard NNs are
generally considered black boxes. However, given decades of research
into methods for interpreting NNs
\cite{craven1996extracting, morch1995visualization}, we now have a
range of techniques at our disposal \cite{guidotti2018survey}. Beyond
simply looking at test performance (a measure of interpretability in its
own right \cite{doshi2017towards}), we will focus on a variety of
techniques that will let us examine trained NNs both in the context of,
and independently of, task performance. In particular, we discuss
saliency maps (Subsection \ref{sec:back_saliency_maps}), activation
maximisation (Subsection \ref{sec:act_max}), weight visualisations
(Subsection \ref{sec:weight_vis}), statistical and structural weight
characterisations (Subsection \ref{sec:stat_struct_weight}), unit
ablations (Subsection \ref{sec:unit_abl}), layer re-initialisation
(Subsection \ref{sec:layer_abl}), recurrent ablations (Subsection
\ref{sec:recurrent_abl}) and activation analysis (Subsection
\ref{sec:act_an}). By utilising a range of techniques we hope to cover
various points along the trade-off between fidelity and interpretability
\cite{ribeiro2016should}.

\hypertarget{saliency-maps}{%
\subsubsection{Saliency Maps}\label{saliency-maps}}

\label{sec:back_saliency_maps}

Saliency maps are one of the most common techniques used for
understanding the decisions made by NNs, and in particular,
convolutional NNs (CNNs). In line with prior work on interpreting DRL
agents \cite{greydanus2018visualizing}, we use occlusion-based methods,
in which parts of the image are masked to perform a sensitivity analysis
with respect to the change in the network's outputs. The original method
introduced by Zeiler et al. \cite{zeiler2014visualizing} proposes
running a (grey, square) mask over the input and tracking how the
network's outputs change in response. Greydanus et al.
\cite{greydanus2018visualizing} applied this method to understanding
actor-critic-based DRL agents, using the resulting saliency maps to
examine strong and overfitting policies; they however noted that a grey
square may be perceived as part of a grey object, and instead used a
localised Gaussian blur to add ``spatial uncertainty''. The saliency
value \(S_{m,n}\) for each input location \((m, n)\) is the Euclidean
distance between the original output and the output given the input
\(\mathbf{x}_{m, n}^{occ}\) which has been occluded at location
\((m, n)\): \begin{align}
S_{m, n} = \lVert F(\mathbf{x}) - F(\mathbf{x}_{m, n}^{occ}) \rVert_2,
\end{align} \noindent where \(\lVert\cdot\rVert_p\) denotes the
\(\ell_p\)-norm.

However, we found that certain trained agents sometimes confused the
blurred location with the target location---a failing of the attribution
method against noise/distractors \cite{kindermans2016investigating},
and not necessarily the model itself. {Atrey et al.}
\cite{atrey2020exploratory} {identified this issue of applying saliency methods to DRL agents as modifying observations in a manner that is incongruent with the true environment's state and generative process, and proposed intervening directly in the environment in order to employ counterfactual methods. However, in general this level of control over the environment is not possible.}
Motivated by the methods that compute interpretations against reference
inputs
\cite{bach2015pixel, ribeiro2016should, shrikumar2017learning, sundararajan2017axiomatic},
we replaced the Gaussian blur with a mask\footnote{Replacing a circular
  region of 5px radius around the \((m, n)\) location.} derived from a
baseline input, which roughly represents what the model would expect to
see on average. Intuitively, this acts as a counterfactual, revealing
what would happen if the specific part of the input was not there. For
this we averaged over frames collected from our standard evaluation
protocol (see Subsection \ref{sec:environments} for details), creating
an average input to be used as an improved mask for the occlusion-based
method (Figure \ref{fig:saliency_baseline}). Unless specified otherwise,
we use our average input baseline for all occlusion-based saliency maps.
Contemporaneous work has examined the use of more advanced baselines for
gradient-based saliency map methods \cite{sturmfels2020visualizing}. {Another recent work has introduced a more robust DRL-specific saliency method}
\cite{puri2020explain}{, but it is only applicable to agents which learn a state-action value function over discrete action spaces.}

\begin{figure}
  \centering
  \begin{subfigure}{0.32\columnwidth}
    \includegraphics[width=\linewidth]{figures/chapter6/occ_jaco_baseline}
    \subcaption{Occ. (Gaussian)}
  \end{subfigure}
  \begin{subfigure}{0.32\columnwidth}
    \includegraphics[width=\linewidth]{figures/chapter6/average_map_jaco_large.png}
    \subcaption{Avg. $\mathbf{x}^{base}$}
  \end{subfigure}

  \begin{subfigure}{0.32\columnwidth}
    \includegraphics[width=\linewidth]{figures/chapter6/occ_jaco_average_map.png}
    \subcaption{Occ. (avg. $\mathbf{x}^{base}$)}
  \end{subfigure}
  \begin{subfigure}{0.32\columnwidth}
    \includegraphics[width=\linewidth]{figures/chapter6/average_map_mask.png}
    \subcaption{Avg. $\mathbf{x}^{base}$ mask}
  \end{subfigure}
  \caption{Occlusion-based saliency map method (a) improved (c) by the use of an \textit{average image} (b, d). Occlusion (occ.) with the \textit{default Gaussian blur} (a) shows additional artifacts. By creating an average image (b) from a large set of trajectories, we can form an improved baseline for occlusion (c); the usage of the average image as the occlusion (with a blue outline added for emphasis) is pictured in (d).}
  \label{fig:saliency_baseline}
\end{figure}

\hypertarget{activation-maximisation}{%
\subsubsection{Activation Maximisation}\label{activation-maximisation}}

\label{sec:act_max}

Gradients can also be used to try and visualise what maximises the
activation of a given neuron/channel. This can be formulated as an
optimisation problem, using projected gradient ascent in the input space
(where after every gradient step the input is clamped back to within
\([0, 1]\)) \cite{erhan2009visualizing}. Although this would ideally
show what a neuron/channel is selective for, unconstrained optimisation
may end up in solutions far from the training manifold
\cite{mahendran2015understanding}, and so a variety of regularisation
techniques have been suggested for making qualitatively superior
visualisations. We experimented with some of ``weak regularisers''
\cite{olah2017feature}, and found that a combination of frequency
penalisation (Gaussian blur) \cite{nguyen2015deep} and transformation
robustness (random scaling and translation/jitter)
\cite{mordvintsev2015inceptionism} worked best, although they were not
sufficient to completely rid the resulting visualisations of the high
frequency patterns caused by strided convolutions
\cite{odena2016deconvolution}. We performed the optimisation procedure
for activation maximisation for 20 iterations, applying the
regularisation transformations and taking gradient steps in the
\(\ell_2\)-norm \cite{madry2018towards} with a step size of 0.1.
Pseudocode for our method, applied to a trained network \(f\), is
detailed in Algorithm \ref{alg:act_max}.

\begin{algorithm}
\caption{Activation maximisation procedure with transformation robustness, frequency penalisation and $\ell_2$-norm gradient updates.}
\label{alg:act_max}
\begin{algorithmic}
\STATE $f' \gets \text{network } f \text{ truncated at intermediate layer}$
\STATE $i \gets \text{optimisation iterations}$
\STATE $n \gets \text{neuron/channel index}$
\STATE $\alpha \gets \text{step size}$
\STATE $x \sim U(0, 1) \text{ with dimensionality } 3 \times \text{height} \times \text{width}$
\FOR{$i$ steps}
  \STATE $x \gets RandomScale(x)$
  \STATE $x \gets RandomJitter(x)$
  \STATE $x \gets GaussianBlur(x)$
  \STATE $\mathcal{L} \gets \text{mean}(f'(x)_n)$
  \STATE $x \gets x + \alpha\frac{\nabla\mathcal{L}_x}{\Vert \nabla\mathcal{L}_x \Vert_2}$
  \STATE $x \gets \min(\max(x, 0), 1)$
\ENDFOR
%\STATE 
\RETURN $x$
\end{algorithmic}
\end{algorithm}

\hypertarget{weight-visualisations}{%
\subsubsection{Weight Visualisations}\label{weight-visualisations}}

\label{sec:weight_vis}

It is possible to visualise both convolutional filters and
fully-connected weight matrices as images. Part of the initial
excitement around DL was the observation that CNNs trained on object
recognition would learn frequency-, orientation- and colour-selective
filters \cite{krizhevsky2012imagenet}, and more broadly might reflect
the hierarchical feature extraction within the visual cortex
\cite{yamins2016using}. However, as demonstrated by Such et al.
\cite{such2018atari}, DRL agents can perform well with spatially
unstructured filters, although they did find a positive correlation
between spatial structure and performance for RL agents trained with
gradients. Consistent with these findings, we find that spatial
structure is correlated with OoD performance (Subsection
\ref{sec:test_scenarios}). To support this, we developed a novel
quantitative measure to compare filters, which we discuss below.

\hypertarget{statistical-and-structural-weight-characterisations}{%
\subsubsection{Statistical and Structural Weight
Characterisations}\label{statistical-and-structural-weight-characterisations}}

\label{sec:stat_struct_weight}

\hypertarget{magnitude}{%
\paragraph{Magnitude}\label{magnitude}}

\label{sec:magnitude}

A traditional measure for the ``importance'' of individual neurons in a
weight matrix is their magnitude, as exemplified by utilising weight
decay as a regulariser \cite{hanson1989comparing}. Similarly,
convolutional filters, considered as one unit, can be characterised by
their \(\ell_1\)-norms. Given that NN weights are typically randomly
initialised with small but non-zero values
\cite{glorot2010understanding, lecun1998efficient}, the presence of
many zeros or large values indicate significant changes during training.
We can compare these both across trained agents, and across the training
process (although change in magnitude may not correspond with a change
in task performance \cite{zhang2019all}).

\hypertarget{spectral-analysis}{%
\paragraph{Spectral Analysis}\label{spectral-analysis}}

\label{sec:spectral}

Convolutional filters are typically initialised pseudo-randomly, so that
there exists little or no spatial correlation within a single unit. We
hence propose using the 2D discrete power spectral density (PSD) as a
way of assessing the spatial organisation of convolutional filters, and
the power spectral entropy (PSE) as a measure of their complexity. Given
the mean-centred\footnote{Offsetting the mean exposes the
  \emph{relative} spatial structure.} 2D spatial-domain filter,
\(\mathbf{W}_{m, n}\), its corresponding spectral representation,
\(\mathbf{\hat{W}}_{u, v}\), can be calculated via the 2D discrete
Fourier transform of the original filter pattern (\(j=\sqrt{-1}\)):
\begin{align}
\mathbf{\hat{W}}_{u, v} = \sum_{m=0}^{M-1}\sum_{n=0}^{N-1}W_{m, n} \exp \left [ -\frac{j 2\pi}{MN}(u m + v n) \right ],
\end{align} \noindent and its PSD, \(\mathbf{S}_{u, v}\), from the
normalised squared amplitude of the spectrum: \begin{align}
\mathbf{S}_{u, v} = \frac{1}{U V} \left|\mathbf{\hat{W}}_{u, v}\right|^2,
\end{align} \noindent where \((m, n)\) are spatial indices, \((u, v)\)
are frequency indices, \((M, N)\) is the spatial extent of the filter,
and \((U, V)\) is the frequency extent of the filter.

When renormalised such that the sum of the PSD is 1, the PSD may be
thought of as a probability mass function over a dictionary of
components from a spatial Fourier transform. We can treat each location
\((u, v)\) in Fourier space as a symbol, and its corresponding value at
\(\mathbf{S}_{u, v}\) as the probability of that symbol appearing. The
PSE is then simply the Shannon entropy of this distribution, which we
use as a measure of spatial (dis)organisation. In our analysis
(Subsection \ref{sec:statistical_and_structural}), we include statistics
calculated over randomly initialised networks as a baseline. As the
initial weights for units are typically drawn independently from a
normal or uniform distribution, this leads to a fairly flat PSD with PSE
close to \(\log(MN)\)---an upper-bound on PSE.

One weakness of spectral analysis is that these measures will fail to
pick up strongly localised spatial features, as such filters would also
result in a roughly uniform PSD. In practice, global structure is still
useful to quantify, and matches well with human intuition (Figure
\ref{fig:entropy_vis}).

\begin{figure}
  \centering
  \begin{subfigure}{0.3\columnwidth}
    \includegraphics[width=\linewidth]{figures/chapter6/entropy_filter/bot_filts_dr}
    \subcaption{PSE = 2.63}
  \end{subfigure}
  \begin{subfigure}{0.3\columnwidth}
    \includegraphics[width=\linewidth]{figures/chapter6/entropy_filter/mid_filts_dr}
    \subcaption{PSE = 3.19}
  \end{subfigure}
  \begin{subfigure}{0.3\columnwidth}
    \includegraphics[width=\linewidth]{figures/chapter6/entropy_filter/top_filts_dr}
    \subcaption{PSE = 3.86}
  \end{subfigure}
  \caption{Convolutional filters from models trained with DR, with varying power spectral entropies, ranked from the lowest (a) to highest (c). The PSE value refers to the centre filter.}
  \label{fig:entropy_vis}
\end{figure}

Entropy as an information-theoretic measure has been used in DL in many
functions, from predicting neural network ensemble performance
\cite{hansen1990neural} to usage as a regulariser
\cite{khabou1999entropy} or pruning criteria \cite{luo2017entropy}
when applied to activations. Spectral entropy has been used as an input
feature for NNs
\cite{krkic1996eeg, misra2004spectral, srinivasan2005artificial, zheng1996digital},
but, to the best of our knowledge, not for quantifying aspects of the
network itself.

\hypertarget{unit-ablations}{%
\subsubsection{Unit Ablations}\label{unit-ablations}}

\label{sec:unit_abl}

Another way to characterise the importance of a single
neuron/convolutional filter is to remove it and observe how this affects
the performance of the NN: a large drop indicates that a particular unit
is by itself very important to the task at hand. More generally, rather
than only looking at performance, one might look for a large change in
the output. {While more sophisticated unit ablations have been applied in DRL}
\cite{meyes2020you}, {this has only been in the context of a control task with a low-dimensional symbolic state space}.

\hypertarget{layer-re-initialisation}{%
\subsubsection{Layer Re-initialisation}\label{layer-re-initialisation}}

\label{sec:layer_abl}

One can extend the concept of ablations to entire layers, and use this
to study the \emph{re-initialisation robustness} of trained networks
\cite{zhang2019all}. Typical neural network architectures, as used in
our work, are compositions of multiple parameterised layers, with
parameters \(\{\theta_1, \theta_2, \ldots, \theta_L\}\), where \(L\) is
the depth of the network. Using \(\theta_l^t\) to denote the set of
parameters of layer \(l \in [1, L]\) at training epoch \(t \in [1, T]\)
over a maximum of \(T\) epochs, we can study the evolution of each
layer's parameters over time---for example through the change in the
\(\ell_\infty\)- or \(\ell_2\)-norm of the set of parameters.

Zhang et al. \cite{zhang2019all} proposed re-initialisation robustness
as a measure of how important a layer's parameters are with respect to
task performance over the span of the optimisation procedure. After
training, for a given layer \(l\), re-initialisation robustness is
measured by replacing the parameters \(\theta_l^T\) with parameters
checkpointed from a previous timepoint \(t\), that is, setting
\(\theta_l^T \leftarrow \theta_l^t\), and then re-measuring task
performance. They observed that for common CNN architectures trained for
object classification, while the parameters of the latter layers of the
networks tended to change a lot by the \(\ell_\infty\)- and
\(\ell_2\)-norms, the same layers were robust to re-initialisation at
checkpoints early during the optimisation procedure, and even to the
initialisation at \(t = 0\). In the latter case, the parameters are
independent of the training data, which means that the effective number
of parameters is lower than the total number of parameters. Given that
the effective number of parameters is a better measure for model
complexity than total number, this potentially allows us to
differentiate between models with the same architecture. Unlike Zhang et
al. \cite{zhang2019all}, we use re-initialisation robustness to study
the effect of task complexity (training with and without DR, and with
and without proprioceptive inputs), but with networks of similar
capacity.

\hypertarget{recurrent-ablation}{%
\subsubsection{Recurrent Ablation}\label{recurrent-ablation}}

\label{sec:recurrent_abl}

When using recurrent units in the network architecture, we can test if
non-trivial recurrent dynamics are being used by forcing the hidden
state to be constant. If the performance of the agent degrades, then it
is somehow using the recurrent dynamics to perform the task---although
it is difficult to say what the exact ``strategy'' might be. However, if
the performance drop is zero or minimal, then the recurrency is not
being utilised. The constant values of the hidden states should be set
to the empirical average of the values during normal operation, as
naively setting all values to zero could cause a considerable shift in
the distribution of expected inputs---as the hypothesis is that the
network may have learned a constant offset, rather than completely
ignoring the hidden state. {While it is possible to investigate the hidden states of DRL agents over time, visualising and inspecting a high-dimensional time series can be difficult, even for domain experts}
\cite{jaunet2020drlviz}.

\hypertarget{entanglement}{%
\subsubsection{Entanglement}\label{entanglement}}

\label{sec:act_an}

Finally, we consider analysing the internal activations of trained
networks. One of the primary methods for examining activations is to
take the high-dimensional vectors and project them to a
lower-dimensional space (commonly \(\mathbb{R}^2\) for visualisation
purposes) using dimensionality reduction methods that try and preserve
the structure of the original data \cite{rauber2017visualizing}. Common
choices for visualising activations include both principal components
analysis (PCA; a linear projection)
\cite{elman1989representation, pearson1901liii} and t-distributed
stochastic neighbor embedding (t-SNE; a nonlinear projection)
\cite{hamel2010learning, maaten2008visualizing, mnih2015human}. {Prior work has used t-SNE to explore the state visitations of DRL agents, clustering states which are associated with similar actions}
\cite{mnih2015human, zahavy2016graying}.

While these works qualitatively examine the projections of the
activations for a single network, or compare them across trained
networks, we additionally introduce a method to use the projections
quantitatively. In supervised learning settings, one can examine class
overlap in the projected space \cite{rauber2017visualizing}. In our RL
setting there is no native concept of a ``class'', but we can instead
use activations taken under different OoD test scenarios (Subsection
\ref{sec:test_scenarios}) to see (beyond the generalisation performance)
how the internal representations of the trained networks vary under the
different scenarios. Specifically, we measure \emph{entanglement} (``how
close pairs of representations from the same class are, relative to
pairs of representations from different classes''
\cite{frosst2019analyzing}) using the soft nearest neighbour loss,
\(\mathcal{L}_{SNN}\) \cite{salakhutdinov2007learning}, defined over a
batch of size \(B\) with samples \(\mathbf{x}\) and classes \(y\) (where
in our case \(\mathbf{x}\) is a projected activation and \(y\) is a test
scenario) with temperature \(T\) (and using \(\delta_{i,j}\) as the
Kronecker-delta):

\begin{eqnarray}
\mathcal{L}_{SNN} = \frac{1}{B} \sum_{n=1}^B \left ( \log \left [ \sum_{b=1}^B (1-\delta_{b,n}) \cdot e^{-\frac{\lVert \mathbf{x}_n - \mathbf{x}_b \rVert_2^2}{T}} \right ] \right . \nonumber \\
\left . -\log \left [\sum_{a=1}^B (1-\delta_{a,n}) \cdot \delta_{y_a,y_n} \cdot e^{-\frac{\lVert \mathbf{x}_n - \mathbf{x}_a \rVert_2^2}{T}} \right ] \right )
\end{eqnarray}

In particular, if representations between different test scenarios are
highly entangled, this indicates that the network is largely invariant
to the factors of variation between the different scenarios. Considering
DR as a form of data augmentation, this is what we might expect of
networks trained with DR.

\hypertarget{experiments}{%
\section{Experiments}\label{experiments}}

\label{sec:experiments}

\hypertarget{environments}{%
\subsection{Environments}\label{environments}}

\label{sec:environments}

We base our experiments on the common setup of performing
target-reaching with visuomotor control. The tasks involve moving the
end effector of a robot arm to reach a randomly positioned target during
each episode, with visual (one RGB camera view) and sometimes
proprioceptive (joint positions, angles and velocities) input provided
to the agent. Unlike many DRL experiments where the position of the
joints and the target are explicitly provided \cite{plappert2018multi},
in our setup the agent must infer the position of the target, and
sometimes itself, purely through vision. Importantly, we use two robotic
arms---the Fetch Mobile Manipulator and the KINOVA JACO Assistive
robotic arm (pictured in Figure \ref{fig:robot_setups}; henceforth
referred to as Fetch and Jaco, respectively)---which have different
control schemes and different visual appearances. This leads to changes
in the relative importance of the visual and proprioceptive inputs,
which we explore in several of our experiments. These robot environments
are commonly used within the DRL literature
\cite{andrychowicz2017hindsight, gu2017deep, rusu2017sim} and,
therefore, adopting these in our experiments enables both comparison to
prior work and applicability to the DRL field.

\begin{figure}
  \centering
  \begin{subfigure}{0.49\columnwidth}
    \includegraphics[width=\linewidth]{figures/chapter6/fetch_sim.png}
    \subcaption{Fetch enviroment}
  \end{subfigure}
  \begin{subfigure}{0.49\columnwidth}
    \includegraphics[width=\linewidth]{figures/chapter6/fetch_view.png}
    \subcaption{Fetch camera view}
  \end{subfigure}
  \\
  \begin{subfigure}{0.49\columnwidth}
    \includegraphics[width=\linewidth]{figures/chapter6/jaco_sim.png}
    \subcaption{Jaco environment}
  \end{subfigure}
%  \begin{subfigure}{0.32\columnwidth}
%    \includegraphics[width=\linewidth]{jaco_side}
%    \subcaption{Jaco side view}
%  \end{subfigure}
  \begin{subfigure}{0.49\columnwidth}
    \includegraphics[width=\linewidth]{figures/chapter6/jaco_top.png}
    \subcaption{Jaco camera view}
  \end{subfigure}
  \caption{Fetch (a) and Jaco (c) environments, with associated camera views (b, d) that are provided as input to the agents.}
  \label{fig:robot_setups}
\end{figure}

The Fetch has a 7 degrees-of-freedom (DoF) arm, not including the
two-finger gripper. The original model and reaching task setup were
modified from the \texttt{FetchReach} task in OpenAI Gym
\cite{brockman2016openai, plappert2018multi} in order to provide an
additional camera feed for the agent (while also removing the
coordinates of the target from the input). The target can appear
anywhere on the 2D table surface. The agent has 3 sets of actions,
corresponding to position control of the end effector ({[}-5, 5{]} cm in
the x, y and z directions; gripper control is disabled).

The Jaco has been configured to be 6 DoF, with the 3 fingers disabled.
The target can appear anywhere within a 3D area to one side of the
robot's base. The agent has 6 sets of actions, corresponding to velocity
control of the arm joints ({[}-0.6, +0.6{]} rad/s). Due to the
difference in control schemes, 2D versus 3D target locations, and
homogeneous appearance of the Jaco, reaching tasks with the Jaco are
more challenging---particularly when proprioceptive input is not
provided to the agent. A summary of the different settings for the Fetch
and Jaco environments is provided in Table \ref{tbl:exp_setup}.

\begin{table}
  \caption{Summary of Fetch and Jaco experimental setups.}
  \label{tbl:exp_setup}
  \centering
  \begin{tabular}{c|cc}
    \toprule
    Setting & Fetch & Jaco \\
    \midrule
    Active (Total) DoF & 7 (8) & 6 (9) \\
    Target Range & $21 \times 31 \text{cm}^2$ & $40 \times 40 \times 40 \text{cm}^3$ \\
    Num. Test Targets & 80 & 250 \\
    Vision Input & $3 \times 64 \times 64$ & $3 \times 64 \times 64$ \\
    Proprioceptive Inputs & 30 & 18 \\
    Control Type & Position & Velocity \\
    Num. Actions & 3 & 6 \\
    Action Discretisation & 5 & 5 \\
    Control Frequency & 6.67Hz & 6.67Hz \\
    \bottomrule
  \end{tabular}
\end{table}

During training, target positions are sampled uniformly from within the
set range, with episodes terminating once the target is reached (within
10 cm of the target centre), or otherwise timing out in 100 timesteps.
The reward is sparse, with the only nonzero reward being +1 when the
target is reached. During testing, a fixed set of target positions,
covering a uniform grid over all possible target positions, are used; 80
positions in a 2D grid are used for Fetch, and 250 positions in a 3D
grid are used for Jaco. By using a deterministic policy and averaging
performance over the entire set of test target positions, we obtain an
empirical estimate of the probability of task success. Test episodes are
set to time out within 20 timesteps in order to minimise false positives
from the policy accidentally reaching the target.

We only randomise initial positions (for all agents) and visuals (for
some agents), but not dynamics, as this is still a sufficiently rich
task setup to explore. Henceforth we refer to agents trained with visual
randomisations as being under the DR condition, whereas agents trained
without are the standard (baseline) condition. Apart from the target, we
randomise the visuals of all other objects in the environment: the
robots, the table, the floor and the skybox. At the start of every
episode and at each timestep, we randomly alter the RGB colours,
textures and colour gradients of all surfaces (Figure
\ref{fig:dr_example}).

Importantly, there are several aspects that are not altered, as we also
want to test extrapolation to OoD scenarios (Subsection
\ref{sec:test_scenarios}). For example, one of the tests that we apply
to probe generalisation is to change a previously static
property---surface reflectivity, which is completely disabled during
training---and see how this affects the trained agents. All environments
were constructed in MuJoCo \cite{todorov2012mujoco}, a fast and
accurate physics simulator that is commonly used for DRL experiments.

\hypertarget{networks-and-training}{%
\subsection{Networks and Training}\label{networks-and-training}}

\label{sec:networks_training}

\begin{figure*}
  \centering
  \includegraphics[width=0.85\linewidth]{figures/chapter6/network.png}
  \caption{Actor-critic network architecture.}
  \label{fig:network}
\end{figure*}

We utilise the same basic actor-critic network architecture for each
experiment, based on the recurrent architecture used by Rusu et al.
\cite{rusu2017sim} for their Jaco experiments. The architecture has 2
convolutional layers, a fully-connected layer, a long short-term memory
(LSTM) layer \cite{gers2000learning, hochreiter1997long}, and a final
fully-connected layer for the policy and value outputs; rectified linear
units were used at the output of the convolutional layers and first
fully-connected layer. Proprioceptive inputs, when provided, were
concatenated with the outputs of the convolutional layers before being
input into the first fully-connected-layer. The policy,
\(\pi(\cdot; \theta)\), is a product of independent categorical
distributions, with one distribution per action dimension. Weights were
initialised using orthogonal weight initialisation
\cite{ilyas2018deep, saxe2014exact} and biases were set to zero. The
specifics of the architecture are detailed in Figure \ref{fig:network}.

During training, a stochastic policy
\(\mathbf{a} \sim \pi(\mathbf{a}|\mathbf{s}; \theta)\) is used and
trained with PPO with clip ratio \(\epsilon = 0.1\), GAE trace decay
\(\lambda = 0.95\) and discount \(\gamma = 0.99\). Each epoch of
training consists of 32 worker processes collecting 128 timesteps worth
of data each, then 4 PPO updates with a minibatch size of 1024. We train
for up to \(5 \times 10^3\) epochs, using the Adam optimiser
\cite{kingma2014adam} with learning rate \(= 2.5 \times 10^{-4}\),
\(\beta\)s \(= \{0.9, 0.999\}\), and \(\epsilon = 1 \times 10^{-5}\).
\(\mathcal{L}_{value}\) is weighted by 0.5 and \(\mathcal{L}_{entropy}\)
is weighted by 0.01. If the max \(\ell_2\)-norm of the gradients exceeds
0.5 they are rescaled to have a max \(\ell_2\)-norm of 0.5
\cite{pascanu2013difficulty}. During testing, the deterministic policy
\(\mathbf{a} = \argmax_\mathbf{a} \pi(\mathbf{a}|\mathbf{s}; \theta)\)
is used. Our training was implemented using PyTorch
\cite{paszke2019pytorch}. Training each model (each random seed) for
the full number of timesteps takes 1 day on a GTX 1080Ti; we trained
models with 5 different seeds for each of the 8 conditions. The overall
setup is detailed in Algorithm \ref{alg:overall_algo}.

\begin{algorithm}
\caption{PPO + DR training.}
\label{alg:overall_algo}
\begin{algorithmic}
\REQUIRE environment $E$, actor network $\pi(a|s;\theta)$, critic network $V_{\pi}(s;\theta)$, number of training epochs $N$, number of timesteps to collect data $T$, number of PPO updates $M$, on-policy transition memory $\mathcal{B}$
\FOR{epoch $= 1, 2, \ldots, N$}
  \STATE $\mathcal{B} \gets \varnothing$
  \FOR{$t = 1, 2, \ldots, T$}
    \IF{$E$ has terminated}
      \STATE Reset $E$ and receive DR \indent\indent state $s_0$
    \ENDIF
    \STATE Sample an action $a_t \sim \pi(a|s_t;\theta)$
    \STATE Execute $a_{t}$ in $E$ and receive reward $r_t$, DR next \indent\indent state $s_{t+1}$, and $terminal$
    \STATE Store the transition $\left(s_t, a_t, r_t, s_{t+1}, terminal \right)$ in $\mathcal{B}$
  \ENDFOR
  \FOR{PPO update $= 1, 2, \cdots, M$}
    \STATE Update $\pi(a|s;\theta)$ using gradient ascent on Eqn. \ref{eqn:l_clip} \indent\indent + Eqn. \ref{eqn:l_entropy} with data from $\mathcal{B}$
    \STATE Update $V_{\pi}(s;\theta)$ using gradient descent on Eqn. \ref{eqn:l_value} \indent\indent with data from $\mathcal{B}$
  \ENDFOR
\ENDFOR
\end{algorithmic}
\end{algorithm}

\hypertarget{domain-shift}{%
\subsection{Domain Shift}\label{domain-shift}}

\label{sec:domain_shift}

\begin{figure}
  \centering
  \hspace{0.025\textwidth}
  \begin{subfigure}{0.45\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/training_curves/jaco_prop.pdf}
    \caption{Proprioceptive inputs}
  \end{subfigure}
  \hfill
  \begin{subfigure}{0.45\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/training_curves/jaco_noprop.pdf}
    \caption{No proprioceptive inputs}
  \end{subfigure}
  \hspace{0.025\textwidth}
  \caption{Test performance of Jaco agents trained with DR and (a) with or (b) without proprioceptive inputs; the agents are tested against both standard and randomised visuals. Without proprioceptive inputs, the agents fail to fully deal with the domain gap between the randomised and standard visuals. Statistics (median and 95\% confidence interval) are calculated over all models (seeds) and test target locations.}
  \label{fig:domain_shift}
\end{figure}

Once agents are successfully trained on each of the different conditions
(Fetch/Jaco, DR/no DR, proprioceptive/no proprioceptive inputs), we can
perform further tests to see how they generalise. However, while the
agents achieve practically perfect test performance on the conditions
that they were trained under, the Jaco agents trained with DR but
without proprioceptive inputs fare worse when tested under the
simulator's standard visuals (Figure \ref{fig:domain_shift}),
demonstrating a drop in performance under domain shift. It is both
assumed and observed that domain shift occurs when transferring models
trained with DR to the more complex and noisy visuals of the real world,
but it is somewhat unexpected to see this happen when shifting to
simpler visuals, which are expected to be a subset of DR visuals---this
indicates that the agent is in some way overfitting to the DR visuals.
Because of this, it is not completely straightforward to compare
performance between different agents, but the change in performance of a
single agent over differing test conditions is still highly meaningful.

We also trained agents with visual DR where the visuals were only
randomised at the beginning of each episode, and kept fixed during.
These agents exhibited the same gap in performance between the standard
and randomised visuals, indicating that this is not an issue of temporal
consistency in the DR setup. Therefore, while agents trained with visual
DR may be invariant to the visual conditions observed during training,
this invariance may have limited OoD generalisation with respect to
backgrounds (Figure \ref{fig:strategies}).

\hypertarget{test-scenarios}{%
\subsection{Test Scenarios}\label{test-scenarios}}

\label{sec:test_scenarios}

In order to test how the agents generalise to different held-out
conditions, we constructed a suite of tests for the trained agents
(Figure \ref{fig:tests} for observations for Fetch under the different
conditions\footnote{Simulation environment parameters of the Mujoco can
  be referenced from \url{http://www.mujoco.org/book/XMLreference.html}.},
and Table \ref{tbl:tests} for the results):

\begin{figure*}
  \centering
  \begin{subfigure}{0.19\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/test_observations/standard}
    \caption{Standard}
  \end{subfigure}
  \begin{subfigure}{0.19\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/test_observations/colour_b}
    \caption{Colour}
  \end{subfigure}
  \begin{subfigure}{0.19\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/test_observations/shape}
    \caption{Shape}
  \end{subfigure}
  \begin{subfigure}{0.19\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/test_observations/illumination}
    \caption{Illumination (Lvl.)}
  \end{subfigure}
  \begin{subfigure}{0.19\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/test_observations/illumination_dir}
    \caption{Illumination (Dir.)}
  \end{subfigure}

  \begin{subfigure}{0.19\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/test_observations/noise}
    \caption{Noise}
  \end{subfigure}
  \begin{subfigure}{0.19\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/test_observations/reflection}
    \caption{Reflection}
  \end{subfigure}
  \begin{subfigure}{0.19\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/test_observations/translation}
    \caption{Translation}
  \end{subfigure}
  \begin{subfigure}{0.19\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/test_observations/invisible}
    \caption{Invisibility}
  \end{subfigure}
  \caption{Camera observations for Fetch under different test conditions.}
  \label{fig:tests}
\end{figure*}

\begin{description}
\item[Standard.] This is the standard evaluation procedure with the default simulator visuals, where the deterministic policy is applied to all test target positions and the performance is averaged (1.0 means that all targets were reached within 20 timesteps). 
\item[Colour.] This introduces a non-red sphere distractor object that is the same size and shape as the target. This tests the sensitivity of the policy to localising the target given another object of a different colour. We use yellow, blue, green, brown and purple; some of these colours have a red component (yellow, brown and purple), while others do not (blue and green).
\item[Shape.] This introduces a red distractor object that is the same width and colour as the target, but a different shape. We use a cube, ellipsoid, rectangle and diamond.
\item[Illumination (Lvl.).] This changes the diffuse colour of the main light. We use 5 illumination levels: 0.4 to 0.0 for Fetch, and 0.9 to 0.1 for Jaco.
\item[Illumination (Dir.).] This changes the location of the main light. We use 5 different directions for both robots. 
\item[Noise.] This adds Gaussian noise $\sim N(0, 0.25)$ to the visual observations. 
\item[Reflection.] This sets the table (for Fetch) or ground (for Jaco) to be reflective. This introduces reflections of the robot (and the target for Jaco) in the input.
\item[Translation.] This offsets the RGB camera. We use 5 locations: -20 to 20cm in the x direction for Jaco, and -20 to 20cm in the y direction for Fetch.
\item[Invisibility.] This makes the robot transparent; this is not a realistic alteration, but is instead used to test the importance of the visual inputs for self-localisation.
\end{description}

\begin{table*}
  \caption{Test performance of all models with local visual changes (distractors), global visual changes, and invisibility (visual self-localisation test). Colour and shape results are averaged over 9 different distractor locations, as well as all colours and shapes, respectively. Checkmarks and crosses indicate enabling/disabling DR and proprioceptive inputs (Prop.), respectively. Statistics are calculated over all models (seeds) and test target locations.}
  \label{tbl:tests}
  \resizebox{\linewidth}{!}{
  \begin{tabular}{c|cc|c|cc|cc|ccc|c}
    \toprule
    Robot & DR & Prop. & Standard & Colours & Shapes & Illumination (Lvl.) & Illumination (Dir.) & Noise & Reflection & Translation & Invisibility\\
    \midrule
    Fetch & \textcolor{red}{\xmark}   & \textcolor{red}{\xmark}   & 1.000$\pm 0.000$ & 0.993$\pm 0.004$ & 0.827$\pm 0.056$ & 0.694$\pm 0.049$ & 0.539$\pm 0.069$ & 0.980$\pm 0.006$ & 0.447$\pm 0.039$ & 0.270$\pm 0.079$ & 0.000$\pm 0.000$\\
    Fetch & \textcolor{red}{\xmark}   & \textcolor{green}{\cmark} & 1.000$\pm 0.000$ & 0.937$\pm 0.027$ & 0.524$\pm 0.087$ & 0.593$\pm 0.064$ & 0.491$\pm 0.066$ & 0.988$\pm 0.004$ & 0.570$\pm 0.078$ & 0.231$\pm 0.078$ & 0.000$\pm 0.000$\\
    Fetch & \textcolor{green}{\cmark} & \textcolor{red}{\xmark}   & 0.983$\pm 0.004$ & 0.974$\pm 0.004$ & 0.939$\pm 0.018$ & 0.957$\pm 0.008$ & 0.740$\pm 0.058$ & 0.985$\pm 0.007$ & 0.972$\pm 0.011$ & 0.480$\pm 0.064$ & 0.000$\pm 0.000$\\
    Fetch & \textcolor{green}{\cmark} & \textcolor{green}{\cmark} & 0.997$\pm 0.002$ & 0.997$\pm 0.001$ & 0.978$\pm 0.009$ & 0.992$\pm 0.002$ & 0.883$\pm 0.038$ & 0.970$\pm 0.008$ & 0.985$\pm 0.005$ & 0.536$\pm 0.063$ & 0.023$\pm 0.015$\\
    \hline
    Jaco  & \textcolor{red}{\xmark}   & \textcolor{red}{\xmark}   & 0.995$\pm 0.003$ & 0.772$\pm 0.058$ & 0.367$\pm 0.069$ & 0.966$\pm 0.012$ & 0.838$\pm 0.035$ & 0.635$\pm 0.028$ & 0.734$\pm 0.032$ & 0.556$\pm 0.063$ & 0.000$\pm 0.000$\\
    Jaco  & \textcolor{red}{\xmark}   & \textcolor{green}{\cmark} & 0.995$\pm 0.001$ & 0.808$\pm 0.044$ & 0.382$\pm 0.064$ & 0.876$\pm 0.032$ & 0.911$\pm 0.014$ & 0.478$\pm 0.059$ & 0.618$\pm 0.061$ & 0.688$\pm 0.048$ & 0.001$\pm 0.001$\\
    Jaco  & \textcolor{green}{\cmark} & \textcolor{red}{\xmark}   & 0.650$\pm 0.056$ & 0.652$\pm 0.022$ & 0.593$\pm 0.031$ & 0.643$\pm 0.027$ & 0.549$\pm 0.032$ & 0.575$\pm 0.040$ & 0.429$\pm 0.060$ & 0.310$\pm 0.041$ & 0.007$\pm 0.002$\\
    Jaco  & \textcolor{green}{\cmark} & \textcolor{green}{\cmark} & 0.991$\pm 0.004$ & 0.989$\pm 0.002$ & 0.843$\pm 0.061$ & 0.816$\pm 0.043$ & 0.882$\pm 0.031$ & 0.896$\pm 0.007$ & 0.946$\pm 0.006$ & 0.658$\pm 0.054$ & 0.916$\pm 0.022$\\
    \bottomrule
  \end{tabular}}
\end{table*}

\begin{table*}
  \caption{Test performance of all models with local visual changes (distractors). Checkmarks and crosses indicate enabling/disabling DR and proprioceptive inputs (Prop.), respectively. Statistics are calculated over all models (seeds), test target locations, and 9 distractor locations.}
  \label{tbl:diff_colours}
  \resizebox{\linewidth}{!}{
  \begin{tabular}{c|cc|ccccc|cccc}
    \toprule
    Robot & DR & Prop. & Yellow & Blue & Green & Brown & Purple & Cube & Ellipsoid & Rectangle & Diamond\\
    \midrule
    Fetch & \textcolor{red}{\xmark}   & \textcolor{red}{\xmark}   & 0.993$\pm 0.007$ & 1.000$\pm 0.000$ & 1.000$\pm 0.000$ & 0.978$\pm 0.017$ & 0.995$\pm 0.004$ & 0.775$\pm 0.085$ & 0.993$\pm 0.007$ & 0.960$\pm 0.036$ & 0.580$\pm 0.142$\\
    Fetch & \textcolor{red}{\xmark}   & \textcolor{green}{\cmark} & 0.875$\pm 0.088$ & 1.000$\pm 0.000$ & 1.000$\pm 0.000$ & 0.825$\pm 0.069$ & 0.983$\pm 0.006$ & 0.243$\pm 0.064$ & 0.980$\pm 0.006$ & 0.783$\pm 0.079$ & 0.090$\pm 0.048$\\
    Fetch & \textcolor{green}{\cmark} & \textcolor{red}{\xmark}   & 0.970$\pm 0.011$ & 0.980$\pm 0.006$ & 0.975$\pm 0.006$ & 0.973$\pm 0.010$ & 0.973$\pm 0.011$ & 0.913$\pm 0.042$ & 0.975$\pm 0.009$ & 0.965$\pm 0.013$ & 0.905$\pm 0.048$\\
    Fetch & \textcolor{green}{\cmark} & \textcolor{green}{\cmark} & 0.995$\pm 0.003$ & 0.998$\pm 0.002$ & 1.000$\pm 0.000$ & 0.998$\pm 0.002$ & 0.995$\pm 0.003$ & 0.963$\pm 0.020$ & 0.998$\pm 0.002$ & 0.995$\pm 0.003$ & 0.958$\pm 0.022$\\
    \hline
    Jaco  & \textcolor{red}{\xmark}   & \textcolor{red}{\xmark}   & 0.281$\pm 0.067$ & 0.871$\pm 0.036$ & 0.994$\pm 0.003$ & 0.730$\pm 0.095$ & 0.984$\pm 0.008$ & 0.274$\pm 0.077$ & 0.407$\pm 0.098$ & 0.759$\pm 0.066$ & 0.027$\pm 0.006$\\
    Jaco  & \textcolor{red}{\xmark}   & \textcolor{green}{\cmark} & 0.451$\pm 0.040$ & 0.941$\pm 0.032$ & 0.993$\pm 0.002$ & 0.706$\pm 0.064$ & 0.950$\pm 0.028$ & 0.258$\pm 0.044$ & 0.440$\pm 0.058$ & 0.784$\pm 0.044$ & 0.045$\pm 0.009$\\
    Jaco  & \textcolor{green}{\cmark} & \textcolor{red}{\xmark}   & 0.640$\pm 0.046$ & 0.655$\pm 0.055$ & 0.651$\pm 0.055$ & 0.668$\pm 0.046$ & 0.648$\pm 0.046$ & 0.636$\pm 0.040$ & 0.642$\pm 0.035$ & 0.666$\pm 0.047$ & 0.426$\pm 0.054$\\
    Jaco  & \textcolor{green}{\cmark} & \textcolor{green}{\cmark} & 0.987$\pm 0.005$ & 0.990$\pm 0.003$ & 0.987$\pm 0.003$ & 0.990$\pm 0.003$ & 0.990$\pm 0.003$ & 0.970$\pm 0.017$ & 0.970$\pm 0.011$ & 0.990$\pm 0.003$ & 0.441$\pm 0.128$\\
    \bottomrule
  \end{tabular}}
\end{table*}

\hypertarget{local-visual-changes}{%
\subsubsection{Local Visual Changes}\label{local-visual-changes}}

Noting that the baseline performance of the Jaco model trained with DR
but without proprioception is lower under standard visuals, across both
robots, DR confers robustness to both the colour and shape distractors
(Table \ref{tbl:tests}). However, there is not as consistent a pattern
between agents trained without DR. Table \ref{tbl:diff_colours} provides
an extended breakdown across the different colours and shapes, which
provides important information on what strategies the agents have
learned.

With Fetch, colour distractors have little effect on the agents, but the
shape distractor diminishes the performance of the non-DR agent trained
without proprioception somewhat, and the non-DR agent trained with
proprioception significantly. Given this, it seems that the latter agent
relies mainly on detecting \emph{pure} red (plus some shape information)
in order to locate the ball. As a result of self-localising based on
visual input alone, the former agent develops more sophisticated vision,
allowing the model to better distinguish both shapes and colours.

With Jaco, both non-DR agents suffer noticeable drops in performance in
the presence of distractors with a red component, whilst both DR agents
experience only a very small decrease in performance across all local
distractors (bar the diamond, which looks the most similar to a sphere,
especially at lower resolutions). While the non-DR agents also have
reduced success with the blue sphere distractor, it is less pronounced,
indicating that non-DR Jaco agents are primarily detecting large red
components as the target object.

In order to test that the location of the distractor does not also
influence the models' responses, we varied this and recorded the
corresponding success rates. The low standard deviations shown in Table
\ref{tbl:vary_distractor} indicate that the location only has a minimal
impact on the results.

\hypertarget{global-visual-changes}{%
\subsubsection{Global Visual Changes}\label{global-visual-changes}}

Referring to Table \ref{tbl:tests}, DR generally confers more
robustness, although this time the DR agents do exhibit noticeable drops
in performance across many of these tests.

Reducing the illumination levels drops the performance of all agents
monotonically with respect to dimness (Table \ref{tbl:diff_levels}),
although the Fetch agents trained with DR are the most robust.
Intriguingly, the Jaco agents trained without proprioception are more
robust with respect to this change, as compared to the agents trained
with. Their need to self-localise visually necessitates a more complex
visual system, whereas simpler visual processing may be thrown off by
the reduction in contrast or even simply the change in the pixel values
of the target. Given that the DR agents trained with proprioception tend
to be the most robust across most of the test conditions, this motivates
an additional consideration for training---when performing sensor fusion
within a model, the combination of information should be more resilient
to the loss or faulty functioning of any individual sensory input.

Changing the direction of the main illumination degrades the performance
of all agents. As before, Fetch agents trained with DR are more robust
than those trained without, but for Jaco agents the presence of
proprioceptive inputs are more important than training with DR. This
trend holds across different illumination directions (Table
\ref{tbl:diff_dirs}).

Additive Gaussian noise has very little effect on the Fetch agents, but
reduces the performance of the Jaco agents---by over 30\% for agents
trained without DR, but only by about 10\% for agents trained with DR.
Considering other factors are consistent across training conditions,
either the visual layout or difficulty of the task caused the Fetch
agents to be more robust to noise than the Jaco agents.

Making the table surface reflective throws off the Fetch agents trained
without DR, with an approximately 50\% drop in performance, but with DR
the agents are resilient to this change. The Jaco agents trained without
DR also incur a significant, yet smaller drop in performance. A likely
explanation for this difference is that the size of the robots relative
to the image differs, and the reflection of the Jaco arm simply changes
the input less. When given proprioceptive inputs, both the Fetch and the
Jaco agent trained with DR display similar levels of resilience.

Translating the camera causes a dramatic drop in performance in all
agents. DR confers some amount of resilience to this for the Fetch
agents. However, all Jaco agents are similarly affected, DR or not, and
their average success rates are higher than those of the Fetch agents.
This suggests that all Jaco agents manage to learn a degree of
translation invariance for their policies. One hypothesis for this is
that the requirement to reach a target in 3D confers a more
generalisable representation of space. Performance drops monotonically
with deviation from the original location (Table \ref{tbl:diff_trans}).
The drop is symmetric for the Fetch agents, but asymmetric for the Jaco
agents---a relative shift in the target towards the centre of the image
input is better than a shift away.

\hypertarget{visual-self-localisation}{%
\subsubsection{Visual
Self-localisation}\label{visual-self-localisation}}

\label{sec:vis_self_loc}

For nearly all agents, rendering the robot invisible drops performance
to zero. There are four non-zero performance scores, but three of these
are low enough to be attributable to chance. This test indicates that
perhaps either directly or indirectly the position of the robot is
inferred visually, although we cannot rule out that the drop in
performance is due to the domain shift that results from rendering the
arm invisible. The standout is the Jaco agent with proprioceptive inputs
and DR training, which only incurs a small drop in performance---this
agent is able to self-localise solely based on proprioceptive input.

\hypertarget{tests-summary}{%
\subsubsection{Tests Summary}\label{tests-summary}}

\begin{figure}
  \centering
  \begin{subfigure}{0.45\textwidth}
    \includegraphics[width=\linewidth]{figures/chapter6/dr_concept}
    \subcaption{Idealised theory}
  \end{subfigure}
  \hspace{0.05\textwidth}
  \begin{subfigure}{0.45\textwidth}
    \includegraphics[width=\linewidth]{figures/chapter6/dr_concept_results}
    \subcaption{Our experiments for testing OoD generalisation}
  \end{subfigure}
  \caption[Domain randomisation methodology.]{Domain randomisation methodology. (a) In theory, the range of simulation parameters should be varied in such a way as to successfully encompass visual/physical properties that would be encountered in the real world. (b) In practice, we limited the scope of variation to test generalisation to OoD inputs within simulation, and observed different levels of success depending on the type of change; depending on the training settings we even observe some level of overfitting. The local and global visual changes in our tests are designed to evaluate generalisation in a way that is reflective of the real world.}
  \label{fig:dr_concept}
\end{figure}

There is no single clear result from our evaluation of different setups
with different types of tests, beyond the general importance of sensor
fusion and DR to improve the ability for agents to generalise. The type
of DR used during training---randomising colours and textures---allows
generalisation to localised changes---distractor objects---but fails to
reliably improve generalisation across the more global changes, such as
illumination or translation (Figure \ref{fig:dr_concept}). This should
not come as a surprise given that our DR never changed the position of
the robot, nor the illumination of the target. The takeaway is that
``generalisation'' is more nuanced, and performing systematic tests can
help probe what strategies networks might be using to operate. Finding
failure cases for ``weaker'' agents can still be a useful exercise for
evaluating more robust agents, as it enables adversarial evaluation
\cite{uesato2018rigorous}, and can inform us about the design of DR.

\hypertarget{model-analysis}{%
\section{Model Analysis}\label{model-analysis}}

\label{sec:model_analysis}

The unit tests that we constructed can be used to evaluate the
performance of an arbitrary black box policy under differing conditions,
but we also have the ability to inspect the internals of our trained
agents. Although we cannot obtain a complete explanation for the learned
policies, we can still glean further information from both the learned
parameters and the sets of activations in the networks.

\hypertarget{saliency-maps-1}{%
\subsection{Saliency Maps}\label{saliency-maps-1}}

\label{sec:saliency_maps}

One of the first tests usually conducted is to examine saliency maps to
infer which aspects of the input influence the output of the agent. We
use the occlusion-based technique with average baseline, and focus on
distractors: we show saliency maps for both the standard test setup, and
with either the different colour (yellow) or different shape (cube)
distractors.

\begin{figure}
  \centering
  \begin{subfigure}{0.24\columnwidth}
    \includegraphics[width=\linewidth]{figures/chapter6/distractor_saliency_fetch_pro_off/standard_visual_std}
    \subcaption{Standard (no DR, no proprioception)}
  \end{subfigure}
  \begin{subfigure}{0.24\columnwidth}
    \includegraphics[width=\linewidth]{figures/chapter6/distractor_saliency_fetch_pro_off/color_visual_std}
    \subcaption{Colour (no DR, no proprioception)}
  \end{subfigure}
  \begin{subfigure}{0.24\columnwidth}
    \includegraphics[width=\linewidth]{figures/chapter6/distractor_saliency_fetch_pro_off/shape_visual_std}
    \subcaption{Shape (no DR, no proprioception)}
  \end{subfigure}
  \begin{subfigure}{0.24\columnwidth}
    \includegraphics[width=\linewidth]{figures/chapter6/distractor_saliency_fetch_pro_on/standard_sensor_std}
    \subcaption{Standard (no DR, proprioception)}
  \end{subfigure}

  \begin{subfigure}{0.24\columnwidth}
    \includegraphics[width=\linewidth]{figures/chapter6/distractor_saliency_fetch_pro_on/color_sensor_std}
    \subcaption{Colour (no DR, proprioception)}
  \end{subfigure}
  \begin{subfigure}{0.24\columnwidth}
    \includegraphics[width=\linewidth]{figures/chapter6/distractor_saliency_fetch_pro_on/shape_sensor_std}
    \subcaption{Shape (no DR, proprioception)}
  \end{subfigure}
  \begin{subfigure}{0.24\columnwidth}
    \includegraphics[width=\linewidth]{figures/chapter6/distractor_saliency_fetch_pro_off/standard_visual_random}
    \subcaption{Standard (DR, no proprioception)}
  \end{subfigure}
  \begin{subfigure}{0.24\columnwidth}
    \includegraphics[width=\linewidth]{figures/chapter6/distractor_saliency_fetch_pro_off/color_visual_random}
    \subcaption{Colour (DR, no proprioception)}
  \end{subfigure}

  \begin{subfigure}{0.24\columnwidth}
    \includegraphics[width=\linewidth]{figures/chapter6/distractor_saliency_fetch_pro_off/shape_visual_random}
    \subcaption{Shape (DR, no proprioception)}
  \end{subfigure}
  \begin{subfigure}{0.24\columnwidth}
    \includegraphics[width=\linewidth]{figures/chapter6/distractor_saliency_fetch_pro_on/standard_sensor_random}
    \subcaption{Standard (DR, proprioception)}
  \end{subfigure}
  \begin{subfigure}{0.24\columnwidth}
    \includegraphics[width=\linewidth]{figures/chapter6/distractor_saliency_fetch_pro_on/color_sensor_random}
    \subcaption{Colour (DR, proprioception)}
  \end{subfigure}
  \begin{subfigure}{0.24\columnwidth}
    \includegraphics[width=\linewidth]{figures/chapter6/distractor_saliency_fetch_pro_on/shape_sensor_random}
    \subcaption{Shape (DR, proprioception)}
  \end{subfigure}
  \caption{Occlusion-based saliency maps with Fetch models trained with (g-l) or without (a-f) DR and with (d-f, j-l) or without proprioception (a-c, g-i) in three different distractor conditions. The best Fetch model was used for each training condition.}
  \label{fig:saliency_fetch_distractor}
\end{figure}

\begin{figure*}
  \centering
  \begin{subfigure}{0.24\columnwidth}
    \includegraphics[width=\linewidth]{figures/chapter6/distractor_saliency_jaco_pro_off/standard_visual_std}
    \subcaption{Standard (no DR, no proprioception)}
  \end{subfigure}
  \begin{subfigure}{0.24\columnwidth}
    \includegraphics[width=\linewidth]{figures/chapter6/distractor_saliency_jaco_pro_off/color_visual_std}
    \subcaption{Colour (no DR, no proprioception)}
  \end{subfigure}
  \begin{subfigure}{0.24\columnwidth}
    \includegraphics[width=\linewidth]{figures/chapter6/distractor_saliency_jaco_pro_off/shape_visual_std}
    \subcaption{Shape (no DR, no proprioception)}
  \end{subfigure}
  \begin{subfigure}{0.24\columnwidth}
    \includegraphics[width=\linewidth]{figures/chapter6/distractor_saliency_jaco_pro_on/standard_sensor_std}
    \subcaption{Standard (no DR, proprioception)}
  \end{subfigure}

  \begin{subfigure}{0.24\columnwidth}
    \includegraphics[width=\linewidth]{figures/chapter6/distractor_saliency_jaco_pro_on/color_sensor_std}
    \subcaption{Colour (no DR, proprioception)}
  \end{subfigure}
  \begin{subfigure}{0.24\columnwidth}
    \includegraphics[width=\linewidth]{figures/chapter6/distractor_saliency_jaco_pro_on/shape_sensor_std}
    \subcaption{Shape (no DR, proprioception)}
  \end{subfigure}
  \begin{subfigure}{0.24\columnwidth}
    \includegraphics[width=\linewidth]{figures/chapter6/distractor_saliency_jaco_pro_off/standard_visual_random}
    \subcaption{Standard (DR, no proprioception)}
  \end{subfigure}
  \begin{subfigure}{0.24\columnwidth}
    \includegraphics[width=\linewidth]{figures/chapter6/distractor_saliency_jaco_pro_off/color_visual_random}
    \subcaption{Colour (DR, no proprioception)}
  \end{subfigure}

  \begin{subfigure}{0.24\columnwidth}
    \includegraphics[width=\linewidth]{figures/chapter6/distractor_saliency_jaco_pro_off/shape_visual_random}
    \subcaption{Shape (DR, no proprioception)}
  \end{subfigure}
  \begin{subfigure}{0.24\columnwidth}
    \includegraphics[width=\linewidth]{figures/chapter6/distractor_saliency_jaco_pro_on/standard_sensor_random}
    \subcaption{Standard (DR, proprioception)}
  \end{subfigure}
  \begin{subfigure}{0.24\columnwidth}
    \includegraphics[width=\linewidth]{figures/chapter6/distractor_saliency_jaco_pro_on/color_sensor_random}
    \subcaption{Colour (DR, proprioception)}
  \end{subfigure}
  \begin{subfigure}{0.24\columnwidth}
    \includegraphics[width=\linewidth]{figures/chapter6/distractor_saliency_jaco_pro_on/shape_sensor_random}
    \subcaption{Shape (DR, proprioception)}
  \end{subfigure}
  \caption{Occlusion-based saliency maps with Jaco models trained with (g-l) or without (a-f) DR and with (d-f, j-l) or without proprioception (a-c, g-i) in three different distractor conditions. The best Jaco model was used for each training condition.}
  \label{fig:saliency_jaco_distractor}
\end{figure*}

The saliency maps for the Fetch agents (Figure
\ref{fig:saliency_fetch_distractor}) differ between all models. Apart
from the model trained with DR and with proprioception (Figure
\ref{fig:saliency_fetch_distractor}j-l), all agents seem to use the
gripper to self-localise. Despite having access to clean proprioceptive
inputs, the Fetch agent trained without DR still pays attention to its
own body in the image (i.e., visual localisation strategy; Figure
\ref{fig:strategies})---so it is not necessarily the case that agents
will even utilise the inputs that we may expect. The Fetch agents
trained without DR show saliency on the distractors (Figure
\ref{fig:saliency_fetch_distractor}a-f), while the agents trained with
DR do not (with the exception of the model trained with DR and
proprioception on the shape distractor, as seen in Figure
\ref{fig:saliency_fetch_distractor}).

The saliency maps for the Jaco agents (Figure
\ref{fig:saliency_jaco_distractor}) are more homogeneous, with a large
amount of attention on the target, and little elsewhere. The saliency
for the agent trained without DR and without proprioception clearly
shows some attention around the base of the arm (Figure
\ref{fig:saliency_jaco_distractor}a-c) that would indicate visual
self-localisation. On an initial inspection, it may appear that there is
no saliency around the arm for the agent trained with DR and without
proprioception, although we know that in order to succeed it must be
relying on visual self-localisation. Indeed, there is saliency present
around the arm (Figure \ref{fig:saliency_jaco_distractor}g-i), but it is
difficult to perceive. This example indicates the subjective nature of
interpreting saliency maps, and hence why they should not be the sole
tool for analysis.

This recommendation is also borne out by the mismatch between the
saliency maps and performance. For the Fetch agents trained with DR, the
agent with proprioception shows saliency over the shape distractor
(Figure \ref{fig:saliency_fetch_distractor}l) in contrast to without
proprioception (Figure \ref{fig:saliency_fetch_distractor}i);
conversely, the performance drop is greater in the latter than the
former (Table \ref{tbl:diff_colours}). Similarly for the Jaco agents
trained without DR, the agent without proprioception shows a large
amount of saliency over the shape distractor (Figure
\ref{fig:saliency_jaco_distractor}c), while the agent with
proprioception demonstrates only a minimal amount of saliency (Figure
\ref{fig:saliency_jaco_distractor}f); however, they both have a similar
drop in performance (Table \ref{tbl:diff_colours}).

\hypertarget{activation-maximisation-1}{%
\subsection{Activation Maximisation}\label{activation-maximisation-1}}

\label{sec:activation_maximisation}

\begin{figure}
  \centering
  \begin{subfigure}{0.49\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/act_max/fetch_nodr_noprop_conv1}
    \caption{Convolution 1 (no DR, no proprioception)}
  \end{subfigure}
  \begin{subfigure}{0.49\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/act_max/fetch_nodr_prop_conv1}
    \caption{Convolution 1 (no DR, proprioception)}
  \end{subfigure}

  \begin{subfigure}{0.49\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/act_max/fetch_dr_noprop_conv1}
    \caption{Convolution 1 (DR, no proprioception)}
  \end{subfigure}
  \begin{subfigure}{0.49\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/act_max/fetch_dr_prop_conv1}
    \caption{Convolution 1 (DR, proprioception)}
  \end{subfigure}

  \begin{subfigure}{0.49\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/act_max/fetch_nodr_noprop_conv2}
    \caption{Convolution 2 (no DR, no proprioception)}
  \end{subfigure}
  \begin{subfigure}{0.49\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/act_max/fetch_nodr_prop_conv2}
    \caption{Convolution 2 (no DR, proprioception)}
  \end{subfigure}

  \begin{subfigure}{0.49\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/act_max/fetch_dr_noprop_conv2}
    \caption{Convolution 2 (DR, no proprioception)}
  \end{subfigure}
  \begin{subfigure}{0.49\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/act_max/fetch_dr_prop_conv2}
    \caption{Convolution 2 (DR, proprioception)}
  \end{subfigure}

  \begin{subfigure}{0.49\textwidth}
    \centering
    \includegraphics[width=0.63\textwidth]{figures/chapter6/act_max/fetch_nodr_noprop_value_policy}
    \caption{Value + Policy (no DR, no proprioception)}
  \end{subfigure}
  \begin{subfigure}{0.49\textwidth}
    \centering
    \includegraphics[width=0.63\textwidth]{figures/chapter6/act_max/fetch_nodr_prop_value_policy}
    \caption{Value + Policy (no DR, proprioception)}
  \end{subfigure}

  \begin{subfigure}{0.49\textwidth}
    \centering
    \includegraphics[width=0.63\textwidth]{figures/chapter6/act_max/fetch_dr_noprop_value_policy}
    \caption{Value + Policy (DR, no proprioception)}
  \end{subfigure}
  \begin{subfigure}{0.49\textwidth}
    \centering
    \includegraphics[width=0.63\textwidth]{figures/chapter6/act_max/fetch_dr_prop_value_policy}
    \caption{Value + Policy (DR, proprioception)}
  \end{subfigure}
  \caption{Activation maximisation for trained Fetch agents: first convolutional layer (a-d); second convolutional layer (e-h); value and policy outputs (i-l). The best Fetch model was used for each training condition. Proprioceptive inputs and hidden state for value and policy visualisations are set to zero. Agents trained without DR have many red filters (the colour of the target) in the second layer (e, f), while agents trained with DR have more structured oriented red-blue filters (g, h). In comparison, the Jaco task induces more structured filters even without DR (see Figure \ref{fig:act_max_jaco}).}
  \label{fig:act_max_fetch}
\end{figure}

\begin{figure}
  \centering
  \begin{subfigure}{0.49\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/act_max/jaco_nodr_noprop_conv1}
    \caption{Convolution 1 (no DR, no proprioception)}
  \end{subfigure}
  \begin{subfigure}{0.49\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/act_max/jaco_nodr_prop_conv1}
    \caption{Convolution 1 (no DR, proprioception)}
  \end{subfigure}

  \begin{subfigure}{0.49\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/act_max/jaco_dr_noprop_conv1}
    \caption{Convolution 1 (DR, no proprioception)}
  \end{subfigure}
  \begin{subfigure}{0.49\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/act_max/jaco_dr_prop_conv1}
    \caption{Convolution 1 (DR, proprioception)}
  \end{subfigure}

  \begin{subfigure}{0.49\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/act_max/jaco_nodr_noprop_conv2}
    \caption{Convolution 2 (no DR, no proprioception)}
  \end{subfigure}
  \begin{subfigure}{0.49\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/act_max/jaco_nodr_prop_conv2}
    \caption{Convolution 2 (no DR, proprioception)}
  \end{subfigure}

  \begin{subfigure}{0.49\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/act_max/jaco_dr_noprop_conv2}
    \caption{Convolution 2 (DR, no proprioception)}
  \end{subfigure}
  \begin{subfigure}{0.49\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/act_max/jaco_dr_prop_conv2}
    \caption{Convolution 2 (DR, proprioception)}
  \end{subfigure}

  \begin{subfigure}{0.49\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/act_max/jaco_nodr_noprop_value_policy}
    \caption{Value + Policy (no DR, no proprioception)}
  \end{subfigure}
  \begin{subfigure}{0.49\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/act_max/jaco_nodr_prop_value_policy}
    \caption{Value + Policy (no DR, proprioception)}
  \end{subfigure}

  \begin{subfigure}{0.49\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/act_max/jaco_dr_noprop_value_policy}
    \caption{Value + Policy (DR, no proprioception)}
  \end{subfigure}
  \begin{subfigure}{0.49\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/act_max/jaco_dr_prop_value_policy}
    \caption{Value + Policy (DR, proprioception)}
  \end{subfigure}
  \caption{Activation maximisation for trained Jaco agents: first convolutional layer (a-d); second convolutional layer (e-h); value and policy outputs (i-l). The best Jaco model was used for each training condition. Proprioceptive inputs and hidden state for value and policy visualisations are set to zero. All agents have colour-gradient filters in the second layer (e-h), indicating more visual complexity than needed for the Fetch task (Figure \ref{fig:act_max_fetch}).}
  \label{fig:act_max_jaco}
\end{figure}

In line with Such et al. \cite{such2018atari}, activation maximisation
applied to the first convolutional layer results in edge detectors, with
larger-scale spatial structure in the latter layers (Figure
\ref{fig:act_max_fetch} and Figure \ref{fig:act_max_jaco}). There are
several trends that apply to both the Fetch and Jaco agents. Firstly,
the agents trained without DR develop simpler, more colourful filters in
both layers. In contrast, the agents trained with DR develop more
edge-like detectors, with higher contrast, in their first convolutional
layers. In their second convolutional layers, the feature detectors
resemble the red target itself, surrounded by a complementary
blue-green. This style of detector is consistent across both the
DR-trained Fetch and Jaco agents, which suggests that it was not
developed in response to the green floor in the Jaco environment. One
noticeable difference between the second set of convolutional filters
between the Fetch (Figure \ref{fig:act_max_fetch}g,h and Jaco (Figure
\ref{fig:act_max_jaco}g,h) agents is that the latter also develop
positional sensitivity, indicating that target localisation may be more
difficult.

The Jaco agent trained without DR and without proprioceptive inputs has
convolutional filters that respond maximally to yellow (Figure
\ref{fig:act_max_jaco}b), and is also the most affected by the yellow
distractor, with performance dropping to 28\% (Table
\ref{tbl:diff_colours}). Therefore this agent has learned a
ball-localisation strategy that is not purely based on detecting red or
spherical objects (Figure \ref{fig:strategies}).

Finally, there is a more global, but largely uninterpretable structure
when maximising the value function or policy outputs (choosing the unit
that corresponds to the largest positive movement per action output).
For Fetch agents without DR, the visualisations are dominated by red
(the target colour), but with DR there is a wider spectrum of colours.
This trend is the same for the Jaco agents, although without DR and
without proprioceptive inputs the colours that maximise the value output
are purple and green (a constant hue shift on the usual red and blue).
The agents trained with DR but without proprioception have the most
plain activation maximisation images for the policy, perhaps suggesting
a more factorised control scheme. For the Jaco agent (Figure
\ref{fig:act_max_jaco}k), only the first and third actuators are
activated by strong visual inputs (given zeroes as the proprioceptive
inputs and hidden state), which correspond to the most important joints
for accomplishing this reaching task (the rotating base and the elbow).

As a reminder we note that activation maximisation may not (and is
practically unlikely to) converge to images within the training data
manifold \cite{mahendran2015understanding}---a disadvantage addressed
by the complementary technique of finding image patches within the
training data that maximally activate individual neurons
\cite{girshick2014rich}. {Alternatively, one could train a generative model on the state distribution and query this model to generate novel states of interest}
\cite{olson2021counterfactual, rupprecht2020finding}.

\hypertarget{statistical-and-structural-weight-characterisations-1}{%
\subsection{Statistical and Structural Weight
Characterisations}\label{statistical-and-structural-weight-characterisations-1}}

\label{sec:statistical_and_structural}

\begin{figure*}
  \centering
  \begin{subfigure}{0.24\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/weightdistributions/conv1_l1}
    \caption{$\ell_1$-norm (layer 1)}
  \end{subfigure}
  \begin{subfigure}{0.24\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/weightdistributions/conv2_l1}
    \caption{$\ell_1$-norm (layer 2)}
  \end{subfigure}
  \begin{subfigure}{0.24\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/weightdistributions/conv1_se}
    \caption{PSE (layer 1)}
  \end{subfigure}
  \begin{subfigure}{0.24\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/weightdistributions/conv2_se}
    \caption{PSE (layer 2)}
  \end{subfigure}
  \caption{Effect of DR on statistical and structural characterisiations of convolutional filters, using all filters from all models, along with models with randomly initialised weights. This effect is layer-dependent, with a large change in $\ell_1$-norm for layer 2, but not layer 1, and a relatively larger change in PSE for layer 1 as compared to layer 2.}
  \label{fig:norms_entropy}
\end{figure*}

We calculated statistical and structural weight characteristics over all
trained models (Fetch and Jaco, with/without proprioception,
with/without DR, 5 seeds), which allows us to average over 40 conditions
to examine the effects of DR. We analysed the norms (Subsection
\ref{sec:magnitude}) of all of the weights of the trained agents, and
could not find consistent trends across all layers. The most meaningful
characterisations were the \(\ell_1\)-norm and the power spectral
entropy, PSE, (Subsection \ref{sec:spectral}), applied to the
convolutional filters.

Figure \ref{fig:norms_entropy} shows a KDE of the \(\ell_1\)-norms and
PSEs of all of the 2D filters within the first and second convolutional
layers. For the \(\ell_1\)-norm, in layer 2 the distribution is skewed
towards higher values when the model is trained with DR. For the PSE, in
both layers, but particularly layer 1, the distribution is skewed
towards lower values when the model is trained with DR. Using the
nonparametric Kolmogorov-Smirnov (K-S) two-sided test between the two
distributions (DR versus non-DR), the \(p\)-value of the
\(\ell_1\)-norms is 0.014 (K-S statistic 0.072) for layer 1 and
\(\sim 0\) (K-S statistic 0.285) for layer 2, and the \(p\)-value of the
PSEs is \(5.71\times10^{-27}\) (K-S statistic 0.251) for layer 1 and
\(3.32\times10^{-9}\) (K-S statistic 0.044) for layer 2. Given the same
weight initialisation distributions across all models, this difference
indicates that DR causes a significant change in the final distribution
of weights, with both larger weights and greater spatial structure.

\hypertarget{unit-ablations-1}{%
\subsection{Unit Ablations}\label{unit-ablations-1}}

\label{sec:unit_ablations}

\begin{figure*}
  \centering
  \begin{subfigure}{0.49\linewidth}
    \includegraphics[width=\textwidth]{figures/chapter6/unitablation/conv1_ablations}
    \caption{Standard environment (layer 1)}
  \end{subfigure}
  \begin{subfigure}{0.49\linewidth}
    \includegraphics[width=\textwidth]{figures/chapter6/unitablation/conv1_ablations_noisy}
    \caption{Noisy environment (layer 1)}
  \end{subfigure}

  \begin{subfigure}{0.49\linewidth}
    \includegraphics[width=\textwidth]{figures/chapter6/unitablation/conv2_ablations}
    \caption{Standard environment (layer 2)}
  \end{subfigure}
  \begin{subfigure}{0.49\linewidth}
    \includegraphics[width=\textwidth]{figures/chapter6/unitablation/conv2_ablations_noisy}
    \caption{Noisy environment (layer 2)}
  \end{subfigure}
  \caption{Unit-wise ablation tests in two different visual test environments. Each point corresponds to one unit in layer 1 (a, b) or layer 2 (c, d), with the vertical bars representing baseline performance in the test environment. The training settings correspond to the Fetch (F) and Jaco (J) robots, whether additional proprioceptive inputs are available (Prop), and if DR was used. The best model was used for each training condition. Note that the Jaco agent trained with DR but without proprioception already has a lower base performance on the standard visuals than the other models (Table \ref{tbl:tests}).}
  \label{fig:ablations}
\end{figure*}

Given access to the trained models, unit ablations allow us to perform a
quantitative, white box analysis. To ablate units, we manually zero the
activations of one of the output channels in either the first or second
convolutional layers, iterating the process over every channel. We then
re-evaluate the modified agents for each of the 8 training settings,
using the agent with the best performance over all 5 seeds for each one
(noting that the performance of the best Jaco agent trained with DR and
without proprioception is significantly higher than the average, as
reported in Table \ref{tbl:tests}). These agents are tested on a single
\(x-y\) plane of the fixed test targets---the full 80 for Fetch, and 125
for Jaco---and both the standard visual and additive Gaussian noise test
scenarios (see Subsection \ref{sec:test_scenarios}), as the latter is
often used to mimic sensor noise in robotic learning tasks
\cite{jakobi1995noise}. The results of the ablations are presented in
Figure \ref{fig:ablations}.

We can make several observations from the plots in Figure
\ref{fig:ablations}. Firstly, the Fetch agents are barely affected by
unit ablations, whereas they have varying effects on the Jaco agents.
The higher variability for Jaco agents could be due to the increased
complexity of the Jaco task (both in terms of extracting relevant
information from the sensory inputs, and the difficulty of the
actuation).

Secondly, there is a greater spread of values in layer 1 ablations
(Figure \ref{fig:ablations}a,b) versus layer 2 (Figure
\ref{fig:ablations}c,d). In particular, there appear to be a few highly
important units in layer 1, resulting in highly skewed distributions. We
believe this supports what we observe in the activation maximisation
plots (Figure \ref{fig:act_max_fetch} and Figure
\ref{fig:act_max_jaco}), where there is a greater diversity in the layer
1 filters.

We observe a greater variability in the noisy environment (Figure
\ref{fig:ablations}b,d). Intriguingly, ablations can improve performance
beyond the baseline results, even for agents trained with DR---perhaps
indicating a sensitivity to high-frequency noise. While a thorough
discussion is beyond the scope of this work, we note that research on
corruption and adversarial robustness in supervised learning settings
could provide further insights on properties such as these
\cite{gilmer2019adversarial, hendrycks2019benchmarking}.

One of our original hypotheses was that DR might force the learned
representations to become more redundant---as quantified by reduced
variability under unit ablation---but the results do not support this.
Instead, the only clear outcome is that the baseline performance of the
agents trained with DR is simply higher than that of the agents trained
without DR in the noisy environment.

\hypertarget{layer-re-initialisation-1}{%
\subsection{Layer Re-initialisation}\label{layer-re-initialisation-1}}

\label{sec:layer_ablations}

Moving on from unit ablations, we now show the re-initialisation
robustness, as well as the change in \(\ell_\infty\)- and
\(\ell_2\)-norms of the parameters of my trained Fetch and Jaco agents
in Figures \ref{fig:fetch_robustness} and \ref{fig:jaco_robustness},
respectively. We use re-initialisation robustness to study the effect of
task complexity (training with and without DR, and with and without
proprioceptive inputs), but with networks of similar capacity. Our
results are mostly in line with Zhang et al.
\cite{zhang2019all}---despite continual changes in the weights during
training (as measured by weight norms), the latter layers of the network
are robust to re-initialisation after a few epochs of training, and in
the case of the Fetch agents, the policy layer is robust to
re-initialisation to the original set of weights. The agents trained
with DR are less robust to re-initialisation during
early-to-intermediate stages of training, implying that meaningful
changes in the learned representations occur for longer periods within
the entirety of training.

All agents require the second convolutional layer---where more
sophisticated target location occurs---to be trained for longer than the
first layer. Additionally, all agents with DR require the first
fully-connected layer (fc1) to be trained for longer than their
corresponding non-DR counterparts. This is most noticeable for the Jaco
agent trained with DR and proprioception (Figure
\ref{fig:jaco_robustness}j)---which is the only agent that can
self-localise in the absence of visual inputs (Subsection
\ref{sec:vis_self_loc}).

For nearly all agents, the recurrent layer is quite robust to
re-initialisation to the original set of weights (despite noticeable
changes in the weights as measured by both the \(\ell_\infty\)-
\(\ell_2\)-norms)---while this does not necessarily indicate that the
agents do not utilise information over time, it does imply that training
the recurrent connections is largely unnecessary for these tasks---a
hypothesis we test further in Subsection \ref{sec:recurrent_ablation}.

\begin{figure}
  \centering
  \begin{subfigure}{0.32\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/robustness/fetch/visual_std/error}
    \caption{Robustness (no DR, no proprioception)}
  \end{subfigure}
  \begin{subfigure}{0.32\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/robustness/fetch/visual_std/inf_dist}
    \caption{$\ell_\infty$-norm (no DR, no proprioception)}
  \end{subfigure}
  \begin{subfigure}{0.32\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/robustness/fetch/visual_std/l2_dist}
    \caption{$\ell_2$-norm (no DR, no proprioception)}
  \end{subfigure}
  \begin{subfigure}{0.32\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/robustness/fetch/sensor_std/error}
    \caption{Robustness (no DR, proprioception)}
  \end{subfigure}
  \begin{subfigure}{0.32\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/robustness/fetch/sensor_std/inf_dist}
    \caption{$\ell_\infty$-norm (no DR, proprioception)}
  \end{subfigure}
  \begin{subfigure}{0.32\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/robustness/fetch/sensor_std/l2_dist}
    \caption{$\ell_2$-norm (no DR, proprioception)}
  \end{subfigure}
  \begin{subfigure}{0.32\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/robustness/fetch/visual_random/error}
    \caption{Robustness (DR, no proprioception)}
  \end{subfigure}
  \begin{subfigure}{0.32\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/robustness/fetch/visual_random/inf_dist}
    \caption{$\ell_\infty$-norm (DR, no proprioception)}
  \end{subfigure}
  \begin{subfigure}{0.32\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/robustness/fetch/visual_random/l2_dist}
    \caption{$\ell_2$-norm (DR, no proprioception)}
  \end{subfigure}
  \begin{subfigure}{0.32\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/robustness/fetch/sensor_random/error}
    \caption{Robustness (DR, proprioception)}
  \end{subfigure}
  \begin{subfigure}{0.32\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/robustness/fetch/sensor_random/inf_dist}
    \caption{$\ell_\infty$-norm (DR, proprioception)}
  \end{subfigure}
  \begin{subfigure}{0.32\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/robustness/fetch/sensor_random/l2_dist}
    \caption{$\ell_2$-norm (DR, proprioception)}
  \end{subfigure}
  \caption{Re-initialisation robustness (1 for complete failure, and 0 for complete success), and change in $\ell_\infty$- and $\ell_2$-norm of parameters of Fetch agents trained with (g-l) and without (a-f) DR, and with (d-f, j-l) and without (a-c, g-i) proprioceptive inputs. Plots truncated to show detail during initial epochs. The best Fetch model was chosen for each training condition.}
  \label{fig:fetch_robustness}
\end{figure}

\begin{figure}
  \centering
  \begin{subfigure}{0.32\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/robustness/jaco/visual_std/error}
    \caption{Robustness (no DR, no proprioception)}
  \end{subfigure}
  \begin{subfigure}{0.32\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/robustness/jaco/visual_std/inf_dist}
    \caption{$\ell_\infty$-norm (no DR, no proprioception)}
  \end{subfigure}
  \begin{subfigure}{0.32\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/robustness/jaco/visual_std/l2_dist}
    \caption{$\ell_2$-norm (no DR, no proprioception)}
  \end{subfigure}
  \begin{subfigure}{0.32\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/robustness/jaco/sensor_std/error}
    \caption{Robustness (no DR, proprioception)}
  \end{subfigure}
  \begin{subfigure}{0.32\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/robustness/jaco/sensor_std/inf_dist}
    \caption{$\ell_\infty$-norm (no DR, proprioception)}
  \end{subfigure}
  \begin{subfigure}{0.32\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/robustness/jaco/sensor_std/l2_dist}
    \caption{$\ell_2$-norm (no DR, proprioception)}
  \end{subfigure}
  \begin{subfigure}{0.32\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/robustness/jaco/visual_random/error}
    \caption{Robustness (DR, no proprioception)}
  \end{subfigure}
  \begin{subfigure}{0.32\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/robustness/jaco/visual_random/inf_dist}
    \caption{$\ell_\infty$-norm (DR, no proprioception)}
  \end{subfigure}
  \begin{subfigure}{0.32\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/robustness/jaco/visual_random/l2_dist}
    \caption{$\ell_2$-norm (DR, no proprioception)}
  \end{subfigure}
  \begin{subfigure}{0.32\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/robustness/jaco/sensor_random/error}
    \caption{Robustness (DR, proprioception)}
  \end{subfigure}
  \begin{subfigure}{0.32\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/robustness/jaco/sensor_random/inf_dist}
    \caption{$\ell_\infty$-norm (DR, proprioception)}
  \end{subfigure}
  \begin{subfigure}{0.32\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/robustness/jaco/sensor_random/l2_dist}
    \caption{$\ell_2$-norm (DR, proprioception)}
  \end{subfigure}
  \caption{Re-initialisation robustness (1 indicates complete failure, and 0 indicates complete success), and change in $\ell_\infty$- and $\ell_2$-norm of parameters of Jaco agents trained with (g-l) and without (a-f) DR, and with (d-f, j-l) and without (a-c, g-i) proprioceptive inputs. Plots were truncated to show detail during initial epochs. The best Jaco model was chosen for each training condition. Note that the final failure rate of the best Jaco agent trained with DR and without proprioception on the standard environment is around 20\%. The re-initialisation robustness plot for this condition (g) indicates that all layers are necessary and that training continues to improve performance in the epochs depicted and beyond.}
  \label{fig:jaco_robustness}
\end{figure}

\hypertarget{recurrent-ablation-1}{%
\subsection{Recurrent Ablation}\label{recurrent-ablation-1}}

\label{sec:recurrent_ablation}

To test how useful the LSTM is, we set the hidden and cell states to
constant values and re-evaluated all models. Rather than naively zeroing
the hidden states, which may not be representative of the values during
rollouts, we instead use the empirical average values, as calculated
over the normal execution of the models in testing. Table
\ref{tbl:hidden_ablation} shows the results of this ablation---there is
a slight effect for agents trained without DR, but a significant effect
for agents trained with DR. This indicates that recurrent processing may
not be necessary for solving either robotic task without DR, but it is
useful when DR is active. In terms of the strategies \emph{learned} by
the agents (Figure \ref{fig:strategies}), memory is sufficient for the
agents trained without DR, but necessary for the agents trained with DR.

\begin{table}
  \caption{Test performance of all models with standard operation versus constant (empirical average) hidden states. Checkmarks and crosses indicate enabling/disabling DR and proprioceptive inputs (Prop.), respectively. Statistics are calculated over all models (seeds) and test target locations.}
  \label{tbl:hidden_ablation}
  \centering
  \begin{tabular}{c|cc|cc}
    \toprule
    Robot & DR & Prop. & Standard & Constant Hidden\\
    \midrule
    Fetch & \textcolor{red}{\xmark}   & \textcolor{red}{\xmark}   & 1.000$\pm 0.000$ & 0.988$\pm 0.016$\\
    Fetch & \textcolor{red}{\xmark}   & \textcolor{green}{\cmark} & 1.000$\pm 0.000$ & 0.990$\pm 0.012$\\
    Fetch & \textcolor{green}{\cmark} & \textcolor{red}{\xmark}   & 0.983$\pm 0.004$ & 0.658$\pm 0.106$\\
    Fetch & \textcolor{green}{\cmark} & \textcolor{green}{\cmark} & 0.997$\pm 0.002$ & 0.838$\pm 0.051$\\
    \hline
    Jaco  & \textcolor{red}{\xmark}   & \textcolor{red}{\xmark}   & 0.995$\pm 0.003$ & 0.919$\pm 0.032$\\
    Jaco  & \textcolor{red}{\xmark}   & \textcolor{green}{\cmark} & 0.995$\pm 0.001$ & 0.943$\pm 0.022$\\
    Jaco  & \textcolor{green}{\cmark} & \textcolor{red}{\xmark}   & 0.650$\pm 0.056$ & 0.422$\pm 0.083$\\
    Jaco  & \textcolor{green}{\cmark} & \textcolor{green}{\cmark} & 0.991$\pm 0.004$ & 0.746$\pm 0.060$\\
    \bottomrule
  \end{tabular}
\end{table}

\hypertarget{entanglement-1}{%
\subsection{Entanglement}\label{entanglement-1}}

\label{sec:activation_analysis}

\begin{figure*}
  \centering
  \begin{subfigure}{0.22\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/embeddings/jaco_no-DR_prop_conv1_PCA.png}
    \caption{Conv. 1 (PCA; no DR)}
  \end{subfigure}
  \begin{subfigure}{0.22\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/embeddings/jaco_no-DR_prop_conv2_PCA.png}
    \caption{Conv. 2 (PCA; no DR)}
  \end{subfigure}
  \begin{subfigure}{0.22\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/embeddings/jaco_no-DR_prop_fc1_PCA.png}
    \caption{FC (PCA; no DR)}
  \end{subfigure}
  \begin{subfigure}{0.22\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/embeddings/jaco_no-DR_prop_h_PCA.png}
    \caption{LSTM (PCA; no DR)}
  \end{subfigure}

  \begin{subfigure}{0.22\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/embeddings/jaco_DR_prop_conv1_PCA.png}
    \caption{Conv. 1 (PCA; DR)}
  \end{subfigure}
  \begin{subfigure}{0.22\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/embeddings/jaco_DR_prop_conv2_PCA.png}
    \caption{Conv. 2 (PCA; DR)}
  \end{subfigure}
  \begin{subfigure}{0.22\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/embeddings/jaco_DR_prop_fc1_PCA.png}
    \caption{FC (PCA; DR)}
  \end{subfigure}
  \begin{subfigure}{0.22\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/embeddings/jaco_DR_prop_h_PCA.png}
    \caption{LSTM (PCA; DR)}
  \end{subfigure}

  \begin{subfigure}{0.22\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/embeddings/jaco_no-DR_prop_conv1_tSNE.png}
    \caption{Conv. 1 (t-SNE; no DR)}
  \end{subfigure}
  \begin{subfigure}{0.22\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/embeddings/jaco_no-DR_prop_conv2_tSNE.png}
    \caption{Conv. 2 (t-SNE; no DR)}
  \end{subfigure}
  \begin{subfigure}{0.22\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/embeddings/jaco_no-DR_prop_fc1_tSNE.png}
    \caption{FC (t-SNE; no DR)}
  \end{subfigure}
  \begin{subfigure}{0.22\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/embeddings/jaco_no-DR_prop_h_tSNE.png}
    \caption{LSTM (t-SNE; no DR)}
  \end{subfigure}

  \begin{subfigure}{0.22\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/embeddings/jaco_DR_prop_conv1_tSNE.png}
    \caption{Conv. 1 (t-SNE; DR)}
  \end{subfigure}
  \begin{subfigure}{0.22\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/embeddings/jaco_DR_prop_conv2_tSNE.png}
    \caption{Conv. 2 (t-SNE; DR)}
  \end{subfigure}
  \begin{subfigure}{0.22\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/embeddings/jaco_DR_prop_fc1_tSNE.png}
    \caption{FC (t-SNE; DR)}
  \end{subfigure}
  \begin{subfigure}{0.22\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/embeddings/jaco_DR_prop_h_tSNE.png}
    \caption{LSTM (t-SNE; DR)}
  \end{subfigure}

  \begin{subfigure}{0.22\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/embeddings/jaco_no-DR_prop_conv1_UMAP.png}
    \caption{Conv. 1 (UMAP; no DR)}
  \end{subfigure}
  \begin{subfigure}{0.22\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/embeddings/jaco_no-DR_prop_conv2_UMAP.png}
    \caption{Conv. 2 (UMAP; no DR)}
  \end{subfigure}
  \begin{subfigure}{0.22\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/embeddings/jaco_no-DR_prop_fc1_UMAP.png}
    \caption{FC (UMAP; no DR)}
  \end{subfigure}
  \begin{subfigure}{0.22\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/embeddings/jaco_no-DR_prop_h_UMAP.png}
    \caption{LSTM (UMAP; no DR)}
  \end{subfigure}

  \begin{subfigure}{0.22\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/embeddings/jaco_DR_prop_conv1_UMAP.png}
    \caption{Conv. 1 (UMAP; DR)}
  \end{subfigure}
  \begin{subfigure}{0.22\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/embeddings/jaco_DR_prop_conv2_UMAP.png}
    \caption{Conv. 2 (UMAP; DR)}
  \end{subfigure}
  \begin{subfigure}{0.22\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/embeddings/jaco_DR_prop_fc1_UMAP.png}
    \caption{FC (UMAP; DR)}
  \end{subfigure}
  \begin{subfigure}{0.22\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter6/embeddings/jaco_DR_prop_h_UMAP.png}
    \caption{LSTM (UMAP; DR)}
  \end{subfigure}
  \caption{Embeddings for trained Jaco agents with proprioceptive inputs with and without DR. Test conditions that are entangled with the normal observations (orange) typically include changing the colour (dark blue) or shape (green) of the target, shifting the camera (light blue), and, for DR, adding reflections (yellow). Global changes---adding Gaussian noise (red) or changing the global lighting (purple)---are the least entangled with the normal observations.}
  \label{fig:act_embeddings}
\end{figure*}

\begin{table}
  \caption{Entanglement scores of different agents, for the first and second convolutional (conv.), fully-connected (FC) and LSTM layer, calculated over different testing conditions as classes (with $T = 0$). Checkmarks and crosses indicate enabling/disabling DR and proprioceptive inputs (Prop.), respectively.}
  \label{tbl:entanglement}
  \centering
  \resizebox{0.8\linewidth}{!}{
  \begin{tabular}{c|cc|cccc}
    \toprule
    Robot & DR & Prop. & 1\textsuperscript{st} Conv. & 2\textsuperscript{nd} Conv. & FC & LSTM\\
    \midrule
    Fetch & \textcolor{red}{\xmark}   & \textcolor{red}{\xmark}   & 0.11 & 0.30 & 0.56 & 0.68 \\
    Fetch & \textcolor{red}{\xmark}   & \textcolor{green}{\cmark} & 0.12 & 0.30 & 0.45 & 0.45 \\
    Fetch & \textcolor{green}{\cmark} & \textcolor{red}{\xmark}   & 0.23 & 0.38 & 0.62 & 0.92 \\
    Fetch & \textcolor{green}{\cmark} & \textcolor{green}{\cmark} & 0.24 & 0.41 & 0.58 & 1.15 \\
    \hline
    Jaco  & \textcolor{red}{\xmark}   & \textcolor{red}{\xmark}   & 0.14 & 0.29 & 0.52 & 0.68 \\
    Jaco  & \textcolor{red}{\xmark}   & \textcolor{green}{\cmark} & 0.11 & 0.08 & 0.43 & 0.66 \\
    Jaco  & \textcolor{green}{\cmark} & \textcolor{red}{\xmark}   & 0.41 & 0.37 & 0.55 & 0.73 \\
    Jaco  & \textcolor{green}{\cmark} & \textcolor{green}{\cmark} & 0.65 & 0.56 & 1.21 & 1.37 \\
    \bottomrule
  \end{tabular}}
\end{table}

Finally, we consider the quantitative analysis of activations from
different trained agents under the different training conditions. Table
\ref{tbl:entanglement} contains the entanglement scores
\cite{frosst2019analyzing} of the different trained agents, calculated
across the first 4 layers (not including the policy/value outputs); as
with the original work, we use a 2D t-SNE \cite{maaten2008visualizing}
embedding for the activations. There are two noticeable trends. Firstly,
the entanglement scores increase deeper into the network; this supports
the notion that the different testing conditions can result in very
different visual observations, but the difference between them
diminishes as they are further processed by the networks. Secondly, the
agents trained with DR have noticeably higher entanglement scores for
each layer as compared to their equivalents trained without DR. This
provides quantitative support for the hypothesis that DR makes agents
largely invariant to nuisance visual factors (as opposed to the agents
finding different strategies to cope with different visual conditions).

We can also qualitatively support these findings by visualising the same
activations in 2D (Figure \ref{fig:act_embeddings}). We use three common
embedding techniques in order to show different aspects of the data.
Firstly, we use PCA \cite{pearson1901liii}, which linearly embeds the
data into dimensions which explain the most variance in the original
data; as a result, linearly separable clusters have very different
global characteristics. Secondly, we use t-SNE
\cite{maaten2008visualizing}, which attempts to retain local structure
in the data by calculating pairwise similarities between datapoints and
creating a constrained graph layout in which distances in the original
high-dimensional and the low-dimensional projection are preserved as
much as possible. Thirdly, we use uniform manifold approximation and
projection (UMAP) \cite{mcinnes2018umap}, which operates similarly to
t-SNE at a high level, but better preserves global structure. Although
it is possible to tune t-SNE \cite{wattenberg2016use}, by default, UMAP
better shows relevant global structure.

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

\label{sec:discussion}

The main goal of this study was to understand the representations and
strategies (Figure \ref{fig:strategies}) learned by DRL agents in a
simulated robotics task. To do so, we examined 8 training configurations
simultaneously, resulting in novel insights on how the setup can
influence what agents learn and how well they generalise.

One of the main axes of variation was the presence of DR. In line with
prior work, DR improves performance across a wider distribution of
testing conditions. In particular, our implementation of DR, which
varied colours and textures, allowed generalisation to scenarios with
``local'' perturbations, but was more variable when more global changes
were made to the setup; overall, agents trained with DR were nearly
always more robust than agents trained without (Subsection
\ref{sec:test_scenarios}). Adding DR to a task makes it more challenging
to solve, in terms of sample complexity, although under the current
experimental setup the models do not appear to require additional
architectural depth, as all agents\footnote{Except for the Jaco agent
  trained with DR and without proprioception, which has lower final
  performance under standard visuals.} are robust to re-initialisation
of the final (policy) layer (Subsection \ref{sec:layer_ablations}). The
application of entanglement \cite{frosst2019analyzing}, with respect to
visual perturbations, shows that throughout the network the
representations that are learned appear to be more invariant to these
changes in the visuals, as the embeddings of representations from the
different conditions have higher overlap (Subsection
\ref{sec:activation_analysis}).

At the lower levels of the networks, DR results in significant changes
in the \(\ell_1\)-norms of the convolutional filters (Subsection
\ref{sec:statistical_and_structural}), with more sophisticated feature
detectors (Subsection \ref{sec:activation_maximisation}). Supporting
this, visualising the saliency maps of the agents shows that DR agents
have more focused attention on task-specific features, such as the arm
or ball (Subsection \ref{sec:saliency_maps}). Counter to initial
expectations, we did not find that DR reduced the variability of
performance under convolutional filter ablations---the agents merely
have better baseline performance (Subsection \ref{sec:unit_ablations}).
Deeper within the networks, we found that DR caused the agents to
utilise the recurrent dynamics of the LSTM, whilst the agents trained
without DR were hardly impacted by keeping their recurrent state
constant (Subsection \ref{sec:recurrent_ablation}).

While we observe these general trends, it is notable that some of the
results are not \emph{a priori} as obvious. For example, even when
provided with proprioceptive inputs, the Fetch agent trained without DR
still uses its visual inputs for self-localisation (Subsection
\ref{sec:saliency_maps}), although the addition of DR removes this
observed effect. We believe that the relative simplicity of the Fetch
reaching task---including both sensing and actuation---leads to less
pronounced effects with DR (Subsection \ref{sec:test_scenarios}). The
most unexpected finding was that the performance of the Jaco agent
trained with DR and without proprioception dropped when shifting from DR
visuals to the standard simulator visuals, demonstrating that DR can
overfit (Subsection \ref{sec:domain_shift}). With proprioception the gap
disappears, which supports the idea that the form of input can have a
significant effect on generalisation in agents
\cite{hill2019emergent}---meriting further investigation. While we
focused on investigating 8 configurations in detail, orthogonal factors
of variation, such as in the RL algorithm, or network architecture,
could lead to other insights. {Furthermore, examining the states encountered/policies learned during training can be more informative than simply looking at agents post-training}
\cite{sequeira2020interestingness}.

This work has focused on understanding the effects of DR, but also has a
dual purpose, which is to inform research in an opposite sense: in
situations where DR is expensive or even infeasible, what approaches can
we take to improve generalisation in sim2real transfer? If certain
characteristics are positively correlated with DR training, explicitly
enforcing them---without requiring the DR pipeline---may also lead to
improved generalisation. For instance, Cobbe et al.
\cite{cobbe2018quantifying} showed that standard regularisation
techniques improve generalisation to a limited extent. Similarly, Pinto
et al. \cite{pinto2017robust} showed that adversarial training could
improve the robustness of DRL policies. In line with this, enforcing
greater spatial structure in the convolutional filters (Subsection
\ref{sec:spectral}), which is higher in agents trained with DR
(Subsection \ref{sec:statistical_and_structural}), could be used as a
novel regularisation objective.

A broader goal of these experiments was to assess the suitability of
interpretability methods within the context of DRL. Beyond noticing
limitations as discussed in previous works
\cite{kindermans2016investigating, mahendran2015understanding}, there
is a larger positive outcome from using a wide suite of interpretability
techniques. Firstly, when used together they can cross-check the
validity of each other's results. For example, supposedly ``dead'' units
in the Jaco model with DR and proprioceptive inputs do in fact worsen
performance when ablated (Subsection \ref{sec:activation_maximisation}).
Additionally, although the LSTM layer within DR agents are robust to
re-initialisation at early stages of training (Subsection
\ref{sec:layer_ablations}), the recurrent ablations show that the agents
depend heavily on recurrent processing (Subsection
\ref{sec:recurrent_ablation}). Secondly, the complementary answers these
techniques provide lead to a better understanding of the model as a
whole. For instance, unit ablations (Subsection
\ref{sec:unit_ablations}) can be related to diversity in activation
maximisation (Subsection \ref{sec:activation_maximisation}), and
entanglement (Subsection \ref{sec:activation_analysis}) can explain the
generalisation of agents trained with DR (Subsection
\ref{sec:test_scenarios}).

To conclude, we provide some recommendations for any practitioner aiming
to study DRL agents:

\begin{itemize}
\item Use agents trained with test and control conditions. While it may be possible to interpret results absolutely, it is more reliable to interpret results relatively. An example is inferring what convolutional filters, as visualised using activation maximisation, are selective for.
\item Do not assume that results generalise. While the Jaco agent trained with DR and without proprioception obeys some trends, it remains an outlier in many other aspects.
\item Do not expect clear results. While some methods, such as entanglement, uncovered clear trends, others, such as unit ablations, did not. This does not mean that unit ablations are generally useless---one can imagine that if several units were highly important without DR and no units are individually important with DR, we would have found a significant trend using this method.
\item If possible, use a range of interpretability methods. Uncertainty about results from one method may be resolved by results from another method.
\item No single interpretability method is more useful than another---each type of method reveals different, complementary pieces of information.
\item Finally, use interpretability methods before making claims about the strategies used by agents. While one may assume that an agent that performs well at a task on average is using an intelligent strategy, in all likelihood it may be using heuristics that fail to generalise \cite{ruderman2019uncovering}.
\end{itemize}

### 2022-08-21 conclusion/conclusion.tex
update the final version

@@
 \section{Summary of Thesis Achievements}
-In this thesis, we have proposed several approaches to improve the sample efficiency of deep reinforcement learning (DRL) in tasks with sparse and delayed rewards.
+% In this thesis, we have proposed several approaches to improve the sample efficiency of deep reinforcement learning (DRL) in tasks with sparse and delayed rewards.
+Moreover, we provide a comprehensive quantitative and qualitative analysis on the agents trained with and without domain randomisation (DR) to visualise and explain why DR can improve the robustness of the agents against unseen scenarios.
+In this thesis, in order to improve the sample efficiency of deep reinforcement learning (DRL) in the environments with sparse or delayed rewards, three novel DRL methods are proposed.
+In these methods, we propose 1) a novel self-imitation learning loss term that allows the agent to overcome hard exploration problems by using hindsight experiences, 2) a novel sampling strategy that increases the chance to sample valuable transitions for the training and 3) a diversity-based intrinsic reward that can provide auxiliary dense feedback to the agent for helping the exploration in the environments with delayed or sparse rewards.
+From extensive experiments, these methods achieve better performance than other baseline approaches when using the same number of samples for training, which demonstrates that our methods indeed improve the sample efficiency of DRL algorithms.
 Moreover, we provide a comprehensive quantitative and qualitative analysis on the agents trained with and without domain randomisation (DR) to visualise and explain why DR can improve the robustness of the agents against unseen scenarios.
 
-Firstly, in \textbf{Chapter}~\ref{ch:esil} and \textbf{Chapter}~\ref{ch:dtgsh}, our methods aim to improve the learning efficiency of goal-oriented RL in continuous control tasks with sparse rewards.
-In \textbf{Chapter}~\ref{ch:esil}, to overcome the difficulty that the agent rarely achieves positive rewards from the environment, hindsight goal relabelling is leveraged to modify failed trajectories into successful ones.
-Then, these modified trajectories are used to accelerate the training/exploration in the initial stages of training via a self-imitation learning scheme.
-In addition, a trajectory selection module and an adaptive weight are proposed to remove redundant hindsight experiences and maintain the balance between the exploration and exploiting.
-In \textbf{Chapter}~\ref{ch:dtgsh}, we argue that when using hindsight goal relabelling, not all agent's experiences contribute equally to the training, and the naive uniform sampling may lead to inefficient learning.
-In order to achieve valuable experience, we first sample trajectories according to the diversity of the goal states as modelled by determinantal point processes (DPPs).
-Then, transitions with diverse goal states are selected from the trajectories by using $k$-DPPs.
-Extensive experiments in these two chapters demonstrate the effectiveness of our methods and show that our methods achieve state-of-the-art results on several simulated robot manipulation tasks.
-
-Secondly, in \textbf{Chapter}~\ref{ch:daim}, diversity is again used via a mechanism of intrinsic motivation, a diversity-augmented intrinsic motivation is proposed to encourage the agent keep exploring novel states in the environment.
-This intrinsic reward is formulated based on the diversity of adjacent states, measured under the framework of DPPs.
-Furthermore, a bi-level optimisation is used to guarantee that when increasing the intrinsic reward to encourage the agent to explore novel states, the extrinsic returns are still maximised.
-Experimental results show that this method outperforms other baselines on the MuJoCo continuous control tasks with standard and delayed reward settings, and it accelerates exploration in the initial stage of training in the arcade learning environments.
-
-Finally, in \textbf{Chapter}~\ref{ch:drl_interp}, we investigate the agents trained with and without DR to interpret why the agents trained with DR are more robust to unseen scenarios.
+Firstly, in \textbf{Chapter}~\ref{ch:drl_interp}, we investigate the agents trained with and without DR to interpret why the agents trained with DR are more robust to unseen scenarios.
 In this work, we built two simulated robot environments that provide both visual and proprioceptive inputs.
@@
 We further propose an improved occlusion-based saliency map method to provide better attention maps of the agents at each timestep.
-\td{From results, we show that the DR agents have better generalisation ability to local visual changes, and also exhibit a degree of robustness to global visual changes.
+From results, we show that the DR agents have better generalisation ability to local visual changes, and also exhibit a degree of robustness to global visual changes.
 The saliency maps of the DR agents focus more on the target object, while non-DR agents show attention to the distractors.
 In the layer re-initialisation, DR agents also show significant changes in the parameters, which leads to more robust features.
-The recurrent ablations also reflect the LSTM is only necessary when DR is activated.} In the end, some recommendations for applying interpretability methods to DRL agents are also suggested.
+The recurrent ablations also reflect that LSTM is only necessary when DR is activated.
+In the end, some recommendations for applying interpretability methods to DRL agents are also suggested.
 
+Secondly, in \textbf{Chapter}~\ref{ch:esil} and \textbf{Chapter}~\ref{ch:dtgsh}, research aims to improve the sample efficiency of goal-oriented RL in continuous control tasks with sparse rewards.
+In \textbf{Chapter}~\ref{ch:esil}, a novel self-imitation learning loss is proposed to overcome the difficulty that the agent rarely achieves positive rewards from the environment, where hindsight goal relabelling is leveraged to modify failed trajectories into successful ones.
+Then, these modified trajectories are used to accelerate the training/exploration in the initial stages of training via a self-imitation learning scheme.
+In addition, a trajectory selection module and an adaptive weight are proposed to remove redundant hindsight experiences and maintain the balance between the exploration and exploiting.
+In \textbf{Chapter}~\ref{ch:dtgsh}, it can be argued that when using hindsight goal relabelling, not all agent's experiences contribute equally to the training, and the naive uniform sampling may lead to inefficient learning.
+In order to achieve valuable experience, a more efficient sampling strategy is introduced.
+We first sample trajectories according to the diversity of the goal states as modelled by determinantal point processes (DPPs).
+Then, transitions with diverse goal states are selected from the trajectories by using $k$-DPPs.
+Extensive experiments in these two chapters demonstrate the effectiveness of our methods and show that our methods achieve state-of-the-art results on several simulated robot manipulation tasks.
+
+Finally, in \textbf{Chapter}~\ref{ch:daim}, diversity is again used via a mechanism of intrinsic motivation, a diversity-augmented intrinsic motivation is proposed to encourage the agent keep exploring novel states in the environment.
+This intrinsic reward is formulated based on the diversity of adjacent states, measured under the framework of DPPs.
+Furthermore, a bi-level optimisation is used to guarantee that when increasing the intrinsic reward to encourage the agent to explore novel states, the extrinsic returns are still maximised.
+Experimental results show that this method outperforms other baselines on the MuJoCo continuous control tasks with standard and delayed reward settings, and it accelerates exploration in the initial stage of training in the arcade learning environments.
 %\section{Applications}
@@
 \subsection{Limitations and Extensions}
+In \textbf{Chapter}~\ref{ch:drl_interp}, the current work focuses solely on the analysis of visual domain randomisation.
+In the future, we will also analysis the trained agents with dynamic DR (e.g., changing the frictional coefficient.).
+Furthermore, instead of using a simple reach task as the test environment, we will further develop more complicated tasks for analysing, such as a stacking cube task or a pick and place task, which are more in line with real-world applications.
+Another direction for the future is to apply the related interpretability methods to our proposed methods in other chapters and then compare the results with the PPO algorithm, because the learned features and policies may have changed and thus some interesting findings may be obtained to explain why these methods can improve the performance of vanilla PPO algorithm.
+
 In \textbf{Chapter}~\ref{ch:esil}, the proposed episodic self-imitation learning (ESIL) allows on-policy DRL algorithms (e.g., PPO) and hindsight experience replay (HER) to be combined and can be used to solve continuous control tasks with the sparse rewards.
@@
 Nair \textit{et al.}~\cite{nair2018overcoming} have already done relevant work, where they collect expert demonstrations for imitation learning.
-In stead of manually collecting demonstrations, \td{a potential improvement is to utilise} past good trajectories for self-imitation learning in the off-policy DRL paradigm.
+Instead of manually collecting demonstrations, a potential improvement is to utilise past good trajectories for self-imitation learning in the off-policy DRL paradigm.
 
@@
 In \textbf{Chapter}~\ref{ch:daim}, the proposed diversity-augmented intrinsic motivation (DAIM) achieves limited performance in the arcade learning environment (ALE), and it is also unable to address hard exploration tasks, such as Montezuma's Revenge.
-Therefore, \td{tackling such challenging hard exploration tasks is a potential direction for improvement.}
+Therefore, tackling such challenging hard exploration tasks is a potential direction for improvement.
+One possible solution is to combine the proposed self-imitation loss in Chapter~\ref{ch:esil} with the diversity-based intrinsic reward to enhance the exploration in the above challenging hard exploration tasks.
 
-In \textbf{Chapter}~\ref{ch:drl_interp}, the current work focuses solely on the analysis of visual domain randomisation.
-In the future, we will also analysis the trained agents with dynamic DR (e.g., changing the frictional coefficient.).
-Furthermore, instead of using a simple reach task as the test environment, we will further develop more complicated tasks for analysing, such as a stacking cube task or a pick and place task, which are more in line with real-world applications.
-
-In addition, although there exist plentiful methods to improve sample efficiency in DRL.
+In addition, although there exists plentiful methods to improve sample efficiency in DRL.
 However, sample efficiency in DRL is still a challenging problem.
@@
 Extensive experiments also show that C-learning outperforms other baselines in various goal-oriented tasks.
-Therefore, C-learning proposes a new way to solve goal-oriented RL, and it deserves further investigation.
+Therefore, C-learning provides a new way to solve goal-oriented RL, and it deserves further investigation.
 
-Another way to improve sample efficiency in DRL is to reuse prior knowledge from previously unseen tasks.
+Another way to improve sample efficiency in DRL is to reuse prior knowledge to solve previously unseen tasks.
 This kind of method can differ from imitation learning methods~\cite{bojarski2016end,xu2017end,torabi2018behavioral} which require collecting demonstrations from the same task.
-Some related methods like meta-learning based RL~\cite{duan2016rl,wang2017learning,finn2017model} and meta-imitation learning~\cite{duan2017one,finn2017one,huang2019continuous} try to leverage prior knowledge from unseen tasks to accelerate the learning of new tasks.
-Another interesting method called prior accelerated reinforcement (PARROT)~\cite{singh2020parrot}, first pretrains a flow-based generative model that maps the latent variables, which are conditioned on the current state, to an action by using expert demonstrations from the previous unseen tasks.
+Some related methods like meta-learning based RL~\cite{duan2016rl,wang2017learning,finn2017model} and meta-imitation learning~\cite{duan2017one,finn2017one,huang2019continuous} try to leverage prior knowledge to accelerate the learning of new tasks.
+Another interesting method called prior accelerated reinforcement (PARROT)~\cite{singh2020parrot}, first pretrains a flow-based generative model that maps the latent variables, which are conditioned on the current state, to an action by using expert demonstrations.
 Subsequently, this pretrained generative model can be used as a behaviour prior to facilitate the exploration in the new tasks.
@@
   \end{subfigure}\hfill
-  \caption{Examples of different types of axon images.
+  \caption[Examples of different types of axon images.]{Examples of different types of axon images.
 (a) is a real microscopy image, (b) is the basic 2D synthetic axon image, (c) is the 2D synthetic axon image generated by CapsPix2Pix~\cite{bass2018image}, and (d) is the 3D synthetic axon image.}
@@
 One potential application of DRL is to track the elongated structure in biomedical/microscopy images, such as axons and blood vessels.
+
 Automatically tracking elongated structures is a challenging problem in the field of biomedical imaging.
-\td{In contrast to segmentation, tracking also allows the establishment of correspondences and state estimation to be performed.} Some existing algorithms, such as Kalman filtering, particle filtering, or semi-heuristic connectivity algorithms, require tuning parameters manually for different datasets.
+In contrast to segmentation, tracking also allows the establishment of correspondences and state estimation to be performed.
+Some existing algorithms, such as Kalman filtering, particle filtering, or semi-heuristic connectivity algorithms, require tuning parameters manually for different datasets.
 To alleviate the need for hand-engineering trackers for different biomedical image datasets, we
@@
 We have proposed two DRL trackers~\cite{dai2019deep,balaram2019maximum} which are able to solve 2D axon tracking and achieve comparable results with state-of-the-art axon tracking method -- Vaa3D~\cite{peng2010v3d}.
-However, the above DRL trackers \td{do not yet} scale to the 3D axon tracking task (see Figure~\ref{subfig:3d_synthetic}), because they lack effective exploration in the environment.
+However, the above DRL trackers do not yet scale to the 3D axon tracking task (see Figure~\ref{subfig:3d_synthetic}), because they lack effective exploration in the environment.
 In the future, we will apply different techniques (e.g., intrinsic motivation) to overcome the exploration problems in the 3D tracking task.

### 2022-08-21 introduction/introduction.tex
update the final version

@@
 \chapter{Introduction}
-
 % TODO List:
 % Focus on two parts: sample efficiency (sparse reward setting and general exploration strategy), interperbility.
-
 % chang the title of subsection 1.2 -> general exploration strategies in DRL (-> intrinsic motivation... -> we propose a method based on intrinisc).
-
 % Finally, we discuss the interperbility stuff.
-
 \label{chapter1}
@@
 In recent years, with the success of deep learning (DL) and rapid iteration of computer hardware (e.g., Graphic Processing Unit (GPU)), deep neural networks have significantly improved the state-of-the-art in computer vision, such as for object classification~\cite{krizhevsky2012imagenet} and detection~\cite{ren2015faster}.
-\td{Incorporating deep neural networks with RL is a common approach}, which is capable of scaling to previously unsolved problems with high-dimensional inputs (e.g., images), and this learning paradigm is known as deep reinforcement learning (DRL).
-\td{Some other methods~\cite{guo2016k} reduce the dimension of inputs by adopting clustering algorithms.}
+Incorporating deep neural networks into RL is, increasingly, a popular approach, which is capable of scaling to previously unsolved problems with high-dimensional inputs (e.g., images), and this learning paradigm is known as deep reinforcement learning (DRL).
+Some other methods~\cite{guo2016k} reduce the dimension of inputs by adopting clustering algorithms.
 
@@
 For example, in 2015, DeepMind proposed the deep Q-network (DQN)~\cite{mnih2015human} which enabled the agent to outperform previous RL algorithms on Atari 2600 games and achieve comparable results to professional human players by only receiving the image of a game screen as the inputs.
-In 2016, AlphaGo~\cite{silver2016mastering} defeated the world-famous Go player Mr.
-Lee Sedol.
+In 2016, AlphaGo~\cite{silver2016mastering} defeated the world-famous Go player Lee Sedol.
 In 2019, DRL also successfully addressed one of the most challenging real-time strategy (RTS) games - StarCraft II~\cite{vinyals2019grandmaster} and multiplayer online battle arena (MOBA) games - DOTA2~\cite{berner2019dota} which exhibited the potential of using DRL to tackle scientific and real-world problems.
 In addition to the milestone achievements above, DRL is also utilised as a solution in some other scenarios, such as dexterous robot manipulation~\cite{andrychowicz2020learning}, autonomous driving~\cite{kiran2021deep}, and urban traffic signal control~\cite{wu2020multi}.
-However, some existing challenges may become obstacles to impede people from using RL in real-world applications.
+However, some remaining challenges may become obstacles to impede people from using RL in real-world applications.
 
@@
 Like DL, DRL also requires large amounts of interaction data to train the agent, which can lead to unacceptable costs in real-world applications.
-For example, even for training agents in a video game, such as StarCraft II, \td{16 Tensor Processing Units (TPUs) over 14 days were required} to complete the training (approximately 200 years of \td{human rate} StarCraft II play)~\cite{vinyals2019grandmaster}.
+For example, even for training agents in a video game, such as StarCraft II, 16 Tensor Processing Units (TPUs) over 14 days were required to complete the training (approximately 200 years of human-rate StarCraft II play)~\cite{vinyals2019grandmaster}.
 Therefore, improving the sample efficiency of DRL algorithms is an important research direction for the community.
 
-In this thesis, we mainly focus on overcoming some problems that lead to sample inefficiency in DRL.
-The proposed methods and investigation are based on the following three aspects: 1) solving DRL with a sparse reward setting, 2) improving general exploration strategies in DRL, and 3) generalisation and interpretability of DRL.
+In this thesis, we focus mainly on overcoming some problems that lead to sample inefficiency in DRL.
+Furthermore, we also do some preliminary work on the analysis of trained DRL agents.
+The proposed methods and investigation are based on the following three aspects: 1) solving DRL in environments with a sparse reward setting, 2) improving general exploration strategies in DRL, and 3) analysing DRL agents in the simulated environments.
+\newpage
 
@@
 In RL, the goal or skills that the agent is expected to learn can be formalised into a specific reward function~\cite{sutton2018reinforcement}.
-The purpose of reward shaping is to design a task-specific reward function towards known optimal behaviours, and this usually incorporates domain knowledge and hand-crafted design.
+The purpose of \textit{reward shaping} is to design a task-specific reward function that guides the agent towards known optimal behaviours, and this usually incorporates domain knowledge and hand-crafted design.
 For example, in a locomotion task, a two-legged robot is required to move forward, and the reward function can be designed as~\cite{lee2018composing}:
@@
 \mathcal{R}(s_{t}, a_{t}) = & \lambda_{\text{vel}}\cdot |v_{\text{x}} - v_{\text{target}}| + \lambda_{\text{alive}} - \lambda_{\text{height}}\cdot |1.1 - \min (1.1, \Delta h)| \\
-        & + \lambda_{\text{angle}}\cdot \cos (\theta) \lambda_{\text{foot}}(v_{\text{left\_foot}} - v_{\text{right\_foot}}), \nonumber
+        & + \lambda_{\text{angle}}\cdot \cos (\theta) -  \lambda_{\text{foot}}\cdot(v_{\text{left\_foot}} + v_{\text{right\_foot}}), \nonumber
 \label{eq:designed_reward_function}
 \end{align}
-where $v_{x}$, $v_{\text{left\_foot}}$ and $v_{\text{right\_foot}}$ are the forward velocity, right foot angular velocity, left foot angular velocity.
+where $v_{x}$, $v_{\text{left\_foot}}$ and $v_{\text{right\_foot}}$ are forward velocity, right foot angular velocity, left foot angular velocity.
 $\Delta h$ is the height between the foot and the torso.
@@
 $\lambda_{*}$ are the hyperparameters for the task.
-Therefore, even for simple moving forward behaviour, a complicated reward function needs to be designed, and this reward function cannot be generalised across different tasks and morphology of the robot.
+Therefore, even for simple forward walking behaviour, a complicated reward function needs to be designed, and this reward function cannot be easily generalised across different tasks and morphology of the robot.
 Furthermore, ill-designed reward functions usually lead to unexpected behaviors~\cite{riedmiller2018learning} and limit exploration.
@@
 Details about the goal-oriented RL and sparse reward function are described in Chapter~\ref{ch:background}.
+\newpage
 
@@
 The trade-off between \textit{exploration} and \textit{exploitation} is a challenging topic in DRL.
-In order to achieve sufficient rewards from the environment, an agent needs to exploit previously learned actions that can lead to higher rewards; however, it also needs to explore in the environment to acquire better action selection in the future.
-Without enough exploration, the agent might be stuck at local-minima or even fail to solve tasks.
+In order to achieve sufficient rewards from the environment, an agent can exploit previously learned actions that can lead to higher rewards; however, it also needs to explore the environment and its associated action space to acquire better action selection in the future.
+Without enough exploration, the agent might be stuck at local-minima performing sub-optimally, or even fail to solve particular tasks.
 
-The classic exploration strategies adopted in DRL include: $\epsilon$-greedy~\cite{mnih2015human,wang2016dueling}, entropy regularisation term~\cite{mnih2016asynchronous,schulman2017proximal,schulman2015trust}, and noise-based exploration~\cite{lillicrap2015continuous,plappert2018parameter,fortunato2018noisy}.
+The classic exploration strategies adopted in DRL include: $\epsilon$-greedy~\cite{mnih2015human,wang2016dueling}, entropy regularisation~\cite{mnih2016asynchronous,schulman2017proximal,schulman2015trust}, and noise-based exploration~\cite{lillicrap2015continuous,plappert2018parameter,fortunato2018noisy}.
 For $\epsilon$-greedy, the agent selects optimal actions with probability of $1-\epsilon$ and selects random actions with probability of $\epsilon$; for entropy regularisation, to encourage the agent to select diverse actions in the environment, an additional entropy term $\mathcal{H}(\pi(a_{t}|s_{t}))$ is used to increase the uncertainty of the policy $\pi$; for noise-based exploration, the noise (e.g., Gaussian noise) is added to the action space or parameter space to improve the exploration.
-However, these classic exploration strategies are relatively simple and can only provide limited contributions when the complexity of the environment is increased.
+However, these classic exploration strategies are relatively simple and can only provide limited contributions when the complexity of the environment increases.
 
@@
 More specifically, without the guidance of task-specific reward signals, intrinsic motivation can still keep organisms exploring and emerging broad behaviours or skills based on curiosity.
-Although such emerged skills may not be able to solve one specific task directly, they can be used as low-level skills to address various high-level tasks and thereby improve the learning efficiency.
+Although such emerged skills may not be able to solve one specific task directly, they can be used as low-level skills to address various high-level tasks and thus improve the learning efficiency.
 
@@
 
-\section{Generalisation and Interpretability}
-Generate a policy that can be generalised to unseen environments is also essential when deploying RL models in real-world applications~\cite{kirk2021survey}.
+\section{Analysing DRL Agents}
+% Generate a policy that can be generalised to unseen environments is also essential when deploying RL models in real-world applications~\cite{kirk2021survey}.
 In the DRL community, instead of using a physical robot, a robot simulator is often selected to collect samples during training.
@@
 
-Interpretability is also another challenge in DRL.
-Although DRL has achieved impressive performance in many tasks.
-However, they are often treated as black boxes during inference.
+Understanding how DRL agents behave in an environment is also another challenge in DRL.
+Although DRL has achieved impressive performance in many tasks, agents are often treated as black boxes.
 When applying DRL models in real-world tasks where security and reliability are important (e.g., autonomous driving and robot manipulation), it is necessary to understand how DRL models make decisions, in order to prevent potential risks during deployment.
@@
 \section{Contributions and Outline}
-The contributions and outline of this thesis are organized as follows:
+The contributions and outline of this thesis are organised as follows:
 % make a list to summarize the contribution
@@
 
+    \item In \textbf{Chapter}~\ref{ch:drl_interp}, we investigate trained DRL agents with various types of interpretability methods.
+Large amounts of quantitative and qualitative results are used to analyse the agent trained with and without domain randomisation (DR) to understand the effect of DR.
+This work was done in collaboration with Dr.
+Kai Arulkumaran, I was responsible for most of conceptual design, implementation and most of experiments.
+
     \item In \textbf{Chapter}~\ref{ch:esil}, we propose the episodic self-imitation learning (ESIL) algorithm, in order to improve the sample efficiency of RL in continuous control tasks within the sparse reward setting.
 The proposed method utilises episodic modified trajectories to update the policy.
-In addition, an auxiliary trajectory selection module and an adaptive weight are leveraged to remove undesired trajectories and balance the contribution between the RL loss term and self-imitation learning loss term.
+In addition, an auxiliary trajectory selection module and an adaptive weight are leveraged to remove undesired trajectories and balance the contribution between the RL loss term and the self-imitation learning loss term.
 Extensive experiments show that ESIL achieves better performance than other baselines in four robot simulation tasks.
+In this work, I was responsible for conceptual design, implementation and experiments.
 
@@
 From experiments, our method outperforms other baselines in all benchmark tasks.
+In this work, I was responsible for conceptual design, implementation and experiments.
 
-    \item In \textbf{Chapter}~\ref{ch:daim}, we design a diversity-based intrinsic reward to encourage the agent explore novel states in the tasks with sparse or delayed reward settings.
+    \item In \textbf{Chapter}~\ref{ch:daim}, we design a diversity-based intrinsic reward to encourage the agent to explore novel states in tasks with sparse or delayed reward settings.
 The key idea in our work is to use determinantal point processes (DPPs) to model the discrepancy between states as an intrinsic reward.
-Furthermore, A bi-level optimisation method is utilised to combine with the proposed diversity-based measure, so that the intrinsic reward will increase the extrinsic reward while increasing diversity.
-From extensive experiments, our method achieves state-of-the-art performance in the MuJoCo locomotion environment within standard and delayed reward settings, and it also improves the learning efficiency at the initial stage of training in the arcade learning environment (ALE).
-
-    \item In \textbf{Chapter}~\ref{ch:drl_interp}, we investigate the trained DRL agents with various types of interpretability methods.
-Large amounts of quantitative and qualitative results are used to analyse the agent trained with and without domain randomisation (DR) to understand the effect of DR.
+Furthermore, a bi-level optimisation method is utilised to combine with the proposed diversity-based measure, so that the intrinsic reward will increase the extrinsic reward while increasing diversity.
+From extensive experiments, our method achieves state-of-the-art performance in the MuJoCo locomotion environment within standard and delayed reward settings, and also improves learning efficiency at the initial stage of training in the arcade learning environment (ALE).
+In this work, I was responsible for conceptual design, implementation and experiments.
 
@@
     \item In \textbf{Chapter}~\ref{ch:conclusions}, we briefly summarise the contributions of this thesis, and introduce future works and potential applications.
+In the preliminary work of axon tracking, I was responsible for the implementation and experiments.
 \end{enumerate}
@@
 \begin{enumerate}
+	\item \underline{\textbf{Tianhong Dai}$^{*}$}, Kai Arulkumaran$^{*}$, Tamara Gerbert, Samyakh Tukra, Feryal Behbahani, and Anil Anthony Bharath, ``Analysing Deep Reinforcement Learning Agents Trained with Domain Randomisation", Neurocomputing, vol.493, pp.143--165, 2022.
+[\textbf{Chapter}~\ref{ch:drl_interp}]
 	\item \underline{\textbf{Tianhong Dai}}, Hengyan Liu, Anil Anthony Bharath, ``Episodic Self-Imitation Learning with Hindsight", Electronics, Special Issue on Deep Reinforcement Learning: Methods and Applications, vol.9, pp.1742, 2020.
@@
 [\textbf{Chapter}~\ref{ch:daim}]
-	\item \underline{\textbf{Tianhong Dai}$^{*}$}, Kai Arulkumaran$^{*}$, Tamara Gerbert, Samyakh Tukra, Feryal Behbahani, and Anil Anthony Bharath, ``Analysing Deep Reinforcement Learning Agents Trained with Domain Randomisation", Neurocomputing, vol.493, pp.143--165, 2022.
-[\textbf{Chapter}~\ref{ch:drl_interp}]
 	\item \underline{\textbf{Tianhong Dai}}, Magda Dubois, Kai Arulkumaran, Jonathan Campbell, Cher Bass, Benjamin Billot, Fatmatulzehra Uslu, Vincenzo de Paola, Claudia Clopath, and Anil Anthony Bharath, ``Deep Reinforcement Learning for Subpixel Neural Tracking", International Conference on Medical Imaging with Deep Learning, pp.130--150, 2019 (Spotlight).
@@
 % List extra publications
-\td{In addition to the above publications, the following publications that I authored/co-authored during my Ph.D. study also contribute to deep reinforcement learning and its related applications.}
+In addition to the above publications, the following publications that I co-authored during my Ph.D. study also contribute to the research of deep reinforcement learning and its related applications.
 \begin{enumerate}
